### 2014-09-02 article.Rmd
Article template

---
title: "Gender Article"
author: "Cameron Blevins and Lincoln Mullen"
date: "September 2, 2014"
output: html_document
---

```{r}
library(gender)
```

### 2014-09-03 setup.R
Update dependencies

@@ -3,4 +3,4 @@
 install.packages(c("devtools", "knitr", "packrat"), dependencies = TRUE)
 library(devtools)
 install_github("lmullen/gender-data-pkg")
-install_github("ropensci/gender", branch = "develop") # development version
+install_github("ropensci/gender", ref = "develop") # development version

### 2014-09-03 article.Rmd
Update dependencies
Import article from Google Docs

@@
 ---
-title: "Gender Article"
+title: "Jane, John … Leslie? Algorithmic Gender Detection"
 author: "Cameron Blevins and Lincoln Mullen"
-date: "September 2, 2014"
 output: html_document
@@
 library(gender)
+library(dplyr)
 ```
+
+### Abstract
+
+This paper combines two trends in the humanities: the rise of gender as an analytical lens of inquiry and the proliferation of large, machine-readable datasets.
+It outlines a new method for algorithmically inferring the gender of a name by incorporating historical data to better reflect how naming practices change over time.
+This method ensures a higher level of accuracy with a precise statement of its reliability in inferring gender while tailoring it for humanistic study.
+Predicting gender from names requires a fundamentally historical method, which is demonstrated in this article.
+The article reports on the “gender” package for R which implements this method.
+The article demonstrates the method’s utility by applying it to a case study examining gender representation in the historical profession.
+
+### INTRODUCTION. The "Leslie" Problem
+
+As researchers grapple with larger and larger datasets they are increasingly using algorithms not just to extract information from existing data, but to infer additional information from a dataset.
+[find example – network analysis, maybe http://kieranhealy.org/blog/archives/2013/06/09/using-metadata-to-find-paul-revere/] Many of these datasets are made up of records of individual people, such as company payrolls or military rosters.
+Some might contain associated information like a person’s home address, age, or rank.
+Other datasets contain little more than a personal name.
+What kind of information can researchers infer from just a name?
+Perhaps most fundamentally, a person’s gender can often be inferred from just their first name.
+“Jane Fay” is almost certainly a woman, while her brother “John Fay” is almost certainly a man.
+A straightforward computer program can infer their respective genders with relative ease.
+
+The problem arises from the fact that the link between gender and naming practices, like language itself, is not static.
+What about Jane and John Fay’s long-lost third sibling, Leslie?
+If Leslie was born in the past sixty years, chances are good that she would be their sister.
+But if the three siblings were born in the early twentieth century, chances are good that Leslie would be their brother.
+That is because the gendered naming convention surrounding “Leslie” had completely flipped over the course of one hundred years.
+92% of the babies born in the United States in 1900 who were named “Leslie” were male, while 96% of those identically named babies born in the year 2000 were female.
+A similar pattern holds true for any number of American names: Madison, Morgan, Sydney, Kendall.
+The historical change in the prevailing gender of a name tends to follow set patterns.
+In general, names tend to shift from being predominantly male to being predominantly female; the shift almost never happens in the other direction.
+The reason that the shift tends to be from male to female depends on the cause of the shift.
+In general names that move from male to female are also shifting from being unpopular to popular names.For example the name Madison changed from male to female suddenly after the 1985 because the name became extremely popular for naming baby girls, whereas before it had been an infrequently used male name.
+In other words, since the late twentieth century US at any rate, girls are more likely than boys to be receive a trendy name, and once a trendy name is associated with females it is unlikely to be used for males.
+The few exceptions of names which go from being female to male, such as Jan, have a different causal mechanism: it seems likely that Jan as an English name is female but that Jan as a Scandinavian (Finnish?
+Swedish?) name is male.
+Because the name is relatively unpopular, changing patterns of immigration can explain the change.
+
+[CITE: literature on baby names generally]
+
+[GRAPHIC chart of changes in names – from README]
+[GRAPHIC showing change in popularity of names like Madison]
+
+For researchers working on contemporary subjects the “Leslie” problem is not an especially pressing one.
+An education policy analyst studying the demographics of different urban school districts, for instance, can turn to a variety of tools that use current databases of names.
+(http://genderize.io/, http://genderchecker.com/about-us.aspx) These tools will most likely identify the gender of Leslie Fay correctly.
+But researchers studying historical subjects need to take into account the dynamic relationship between gender and naming practices.
+If that same researcher wanted to compare the demographics of contemporary urban schools with those from the early twentieth century, Leslie Fay’s gender would get misidentified.
+And the “Leslie” problem is not just one for historians studying the past.
+Anyone studying a longer timeframe will also encounter this problem: as of 2012, the average American has a lifespan of nearly seventy-nine years – more than enough time for naming practices to change quite dramatically.
+(cite World Bank).
+
+Our solution to the “Leslie” problem is to create a library that explicitly takes into account changes in naming practices.
+The program incorporates this temporal dimension using historical databases to more accurately infer whether “Leslie” was male or female depending on when he or she lived.
+It not only achieves better accuracy in inferring gender, but it also provides the relative likelihood of a person’s gender for a given range of years: 52% of the babies named “Leslie” born in 1950 were girls and 48% were boys.
+Rather than a black box spitting out a stark FEMALE or MALE designation, the program computes a measurement for each gender’s relative “share” of a given name.
+This allows researchers to conduct more nuanced analysis and to determine for themselves what proportion is acceptable for making a prediction of male or female.
+
+We recognize that gender is a fluid concept, both today and in the past, and that this method glosses over the complexities of historical definitions of gender and of the relationship between biological sex and gender.
+In some sense this is necessary since the large data sets which make this kind of analysis possible were created by state agencies, and for gender (as for race) state agencies have created a limited number of categories which have tremendous definitional power.
+[CITE e.g.
+Margot Canady, The Straight State; works on e.g.
+Virginia’s definition of race.
+Shane Landrum on birth certificates.
+James C.
+Scott on Seeing Like a State or is that obvious to everyone by now?] Nevertheless, this method is suitable for two reasons.
+First, researchers are likely to use this method in the aggregate rather than for individuals.
+The assumption that a name implies male or female (and that those are possible categories to be inferred) has a higher validity when applied to large groups of people than it does when applied to individuals.
+Second, this methods is neutral as to what the categories male or female actually mean in any given historical context.
+Researchers are free---indeed, obligated---to make that determination for themselves, and this package provides access to the underlying data set so that a technologically competent researcher can interrogate every assumption through the process.
+This method does not argue against the complexities of historical understandings of gender, but as with any project it does require researchers who use the method to think through how the data was gathered, what it implies, and how it reflects social and cultural practices.
+
+The remainder of the article is divided into two sections.
+First, we present the methodology of our approach by comparing it to existing methods and detailing the strengths and weaknesses of the program.
+Second, we apply the method to a case study in order to demonstrate how humanities scholars might use it by analyzing gender representation within the historical profession.
+
+PART I. The Method
+
+Existing Methods:
+
+The historical method to be described below may be compared to a method available in the Natural Language Toolkit for Python.
+The NLTK provides many tools for analyzing natural language, including two lists of male and female names created by Mark Kantrowitz with assistance from Bill Ross.
+This list includes 7579 [DOUBLE CHECK] unique names.
+Using the Kantrowitz names corpus in NLTK, one could look up a name like Jane or John and find out that it is male or female.
+[NOTE: The Kantrowitz corpus is also implemented in the gender package for R for comparison, though its use is not recommended for actual research.]
+
+[SAMPLE OUTPUT in Python from NLTK names corpus converted to an array of tuples.]
+
+[('Aaron', 'male'), 
+('Abbey', 'male'), 
+('Abbie', 'male'), 
+('Abbot', 'male')
+('Zorine', 'female'), 
+('Zsa Zsa', 'female'), 
+('Zsazsa', 'female'), 
+('Zulema', 'female'), 
+('Zuzana', 'female')]
+
+[CITE: Link to original name corpus: http://www.cs.cmu.edu/afs/cs/project/ai-repository/ai/areas/nlp/corpora/names/0.html]
+
+The most significant problem with the Kantrowitz names corpus and thus the NLTK implementation is that it assumes that names are timeless.
+As pointed out above, this makes the corpus unusable for historical purposes.
+Furthermore the corpus includes some oddities which make it less suitable.
+Some names like “Bill” are overwhelmingly male, but the corpus includes them as both male and female.
+And because the Kantrowitz corpus contains only 7,579 (a mere 8.2% of the 91,320 names released by the Social Security Administration for the US alone) there are many names that it cannot identify.
+[NOTE: The SSA data includes only names used more than 5 times in a given year, so the “long tail” of actual number of names in the US is considerably larger.
+Link to SSA data.] In this chart, for example, the “unidentified” category grows rapidly after 1965.
+While our Social Security method relies primarily on children born in the US but also people who become naturalized citizens (double check) whereas as census data for example also includes immigrants, nevertheless it does far better at identifying non-European names since it contains over an order of magnitude more names.
+
+[CITE existing literature on gender classification]
+
+We mention the Kantrowitz name corpus as implemented in NLTK because the Natural Language Toolkit is rightly regarded as influential for scholarship.
+It’s flaws are also emblematic of the problems with most other implementations of gender prediction algorithms.
+The Genderize website and API is, for example, a much more sophisticated implementation of a gender prediction.
+Besides predicting male or female for gender, it also reports the proportion of male or female names, along with a count of the number of uses of the name that it is basing its prediction on.
+Genderize will also permit the user to customization a prediction for different countries: an extremely important feature.
+Genderize includes “database contains 142848 distinct names across 77 countries and 85 languages.” Genderize is unsuitable for historical work, however, because it is based only on contemporary data.
+According to the documentation for its API, “It utilizes big datasets of information, from user profiles across major social networks.” [SOURCE: http://genderize.io] It would be anachronistic to apply these datasets to the past, and Genderize provides no functionality to filter results chronologically as it does geographically.
+[NOTE: A wrapper to the genderize API is included in the gender package.
+There are also wrappers available in Ruby and Python.
+There are a number of other website which provide similar functionality based on apparently similar datasets, none of which permit historical variation.)
+
+Surprisingly, CRAN does not include any packages for doing this in R, though some packages do include functions for this.
+Python has no existing package (double check) besides the NLTK and Genderize.Thus the two of the most commonly used languages for data analysis, particularly within the digital humanities community, have no satisfactory package that could easily solve the problem of discovering the gender of names, though options for geocoding, for example, abound.
+
+The Solution:
+
+The possibilities for encoding gender from names depends on two things.
+First, it depends on having a suitable (and suitably large) data set for the time period and region under study.
+Unsurprisingly such data sets are almost always gathered in the first instance by states, though the compilation and digitization may be undertaken by scholarly researchers.
+Second, it depends on a suitable algorithms for estimating the proportion of male and female names for a given year, or range of years since often a person cannot be associated with an exact date.
+It is important that the algorithm take into account any biases in the data to make correct predictions.
+
+Data:
+
+Besides including the
+
+The Social Security Administration provides annual data on the names and genders of babies stretching back to the year 1880.
+This is database encompases 91,320 unique names.
+For comparison, [example of other databases and their size].
+What makes it even more useful, however, is its temporal dimension.
+This allows a researcher to figure out not just the likelihood that “Leslie” is a male or female name, but the likelihood that “Leslie” is a male or female name for a specific range of years, or for a specific year.
+
+Using for example the SSA data for a few names (once it has been munged into a more useful form).
+The columns male and female record the number of people of that gender (sex?) born in that year who received Social Security cards.
+
+name         year    female     male
+---------  ------  --------  -------
+jane         1890       372        0
+jane         1990       771        8
+john         1890        56     8502
+john         1990        81    29066
+madison      1890         0       14
+madison      1990      1407      145
+
+Furthermore, we can calculate the proportions of male and female and on that basis guess the gender for a given year.
+
+name         year    female     male    proportion_male    proportion_female  gender  
+---------  ------  --------  -------  -----------------  -------------------  --------
+jane         1890       372        0             0.0000               1.0000  female  
+jane         1990       771        8             0.0103               0.9897  female  
+john         1890        56     8502             0.9935               0.0065  male    
+john         1990        81    29066             0.9972               0.0028  male    
+madison      1890         0       14             1.0000               0.0000  male    
+madison      1990      1407      145             0.0934               0.9066  female
+ADD LESLIE AND ANOTHER EXAMPLES WE USE
+
+Furthermore, one can specify a range of years and calculate the proportion for that time period.
+
+Given this data set, it is simple to perform perform a merge (a join, in SQL terms) on either a name or a combination of name and year to link this data to a date set.
+
+Our program is best understood through an example.
+Let’s say a historian is working with records from a public school system and wants to know what percentage of its total student body each year was male or female.
+This historian can tell the program the approximate age range of its target population - in this case, ages 4 to 18.
+Our program then iterates through all of the names attempting to identify their gender based on a) the date of the data, and b) the age range specified by the user.
+Sidney Dubner, for instance, was enrolled as a student in the year 1918.
+Given the age range of 4 to 18 specified by the user, the program guesses that Sidney Dubner was born between 1900 and 1914.
+It then looks up every birth of a child named “Sidney” recorded by the Social Security Administration between 1900 and 1914 and calculates that 94.7% of these babies named “Sidney” were male.
+Chances are good that this was a male student.
+But what if the year was 1998 instead of 1918?
+The chances that “Sidney” was a boy falls to 66.4%.
+Fast-forward another ten years to 2008, and suddenly it’s more likely that “Sidney” is in fact a girl, not a boy (69.3% female to 30.7% male).
+
+Reference Implementations/Software:
+
+Description of Lincoln’s R package [link].
+
+Implements both the Kantrowitz method and this new method. Extensible through the addition of additional data sets as they become available.
+Includes cleaned up, tidied (in the Hadley Wickham sense) versions the Kantrowitz data sets and the SSA national and state data.
+(Link to gender repo.) These allow researchers to load the package and use the data directly, in addition to accessing through methods.
+Furthermore
+Allows for encoding of individual names by year, of data sets where each name is associated with a year, and data sets where there are names of years and a range of years is provided.
+Eventually will include ability to specify by
+Samples of code:
+
+find_gender(“madison”) # returns “female” and probability
+find_gender(“madison”, 1990) # returns “female” and probability
+find_gender(“madison”, 1890) # returns “male” and probability
+
+year  name     
+------  ---------
+ 1890  john     
+ 1990  john     
+ 1990  john     
+ 1890  jane     
+ 1990  jane     
+ 1890  madison  
+ 2012  madison
+For data sets where there is no year available, you can specify the range.
+find_gender(test_data, years = c(1880, 1940)) # returns male, male, male, male, female, female
+
+For data sets where there is a year available, say a year of birth, then you can specify the column that stores that data, and each person will be looked up individually.
+find_gender(test_data, year_column = T) # returns male, female, male, male, female, female
+
+In cases where the data set includes non-birth date, and because those dates vary significantly, one wishes to operate on the basis of date, one is advised to split up the data into cohorts, apply the gender checking methods, then recombine the split up data sets.
+(This is akin to Wickhams Split Apply Combine method: citation to article.) In this example of the dissertations data sets, I have split the data set into people whose dissertations were before 1950 and those who were after.
+(In actual practice, I would use more fine grained cohorts, and a more sophisticated method for splitting and applying, such as the `ddply` command in `plyr`.)
+
+library(dplyr)
+dissertations_before_1950 <- dissertations %.%
+ filter(year < 1950)
+dissertations_after_1950 <- dissertations %.%
+ filter(year >= 1950)
+dissertations_before_1950 <- dissertations_before_1950 %.%
+ find_gender(years = c(1880,1940)) # reasonable guess as to ages
+dissertations_after_1950 <- dissertations_after_1950 %.%
+ find_gender(years = c(1930,1990)) # reasonable guess as to ages
+dissertations_combined < rbind(dissertations_before_1950, dissertations_after_1950)
+
+Limitations:
+
+On the methodological front, perhaps the program’s most glaring limitation is its weakness in inferring the gender of non-American names.
+Once again, this is a reflection of where the source data originated: U.S.
+government agencies.
+More recent datasets exist for non-American names, but we have been unable to find or incorporate information on how these names have changed over time.
+The program is extensible, however, meaning that if someone does locate a suitably temporal dataset of non-American names they can implement it into our program.
+
+Our program uses source data that only goes back as far as 1880, making it a far more effective method for twentieth-century analysis.
+Additionally, our program is not as effective at identifying highly uncommon names.
+The annual reference data it uses from the Social Security Administration only includes baby names that occurred at least five times in a given year.
+If a name was never given to five or more babies over the course of any year from 1880 to the present, our program will not be able to infer its gender.
+[more details – see below?]
+[- SSA restricts its data to only names with at least 5 occurrences (we’re still missing a bit of the long tail).
+Not a big problem because we still have more names, and more obscure names, than any other data set.
+More of a problem when localizing to the state, which starts in 1910 (double check) and because it includes only >5 occurrences in much smaller jurisdictions, more of the long tail is cut off.]
+
+PART II. Measuring Gatekeeping in the Historical Profession
+
+In a 2005 report for the American Historical Association, Elizabeth Lunbeck acknowledged “a sea change in the [historical] profession with respect to gender” before going on to describe the painful limitations of this sea change for female historians: ongoing personal discrimination, lower salaries, and barriers to securing high-ranking positions.
+Lunbeck’s report drew in part on a survey of 362 female historians that produced a rich source of responses detailing the deep and multi-faceted challenges facing women in the profession.
+What follows is a quantitative supplement to Lunbeck’s analysis that uses our program to analyze gender representation amongst historians across a much larger scale and a much longer time period.
+http://www.historians.org/Documents/About%20AHA%20and%20Membership/CWH-Report_5.20.05.pdf
+
+Our analysis focuses on one of the bedrocks of the historical profession: scholarly research.
+We begin with the history dissertation, often the defining scholarly output of a historian’s early career.
+The completion of a dissertation marks a standardized moment of transition out of the “training” phase of the historical profession.
+Data supplied by ProQuest provided roughly [80,000] PhD(?) dissertations completed since [1950].
+Identifying the gender of their authors gives a sense for the number of women and men who are completing PhD-level training in history each year.
+Our program uses what year a historian wrote their dissertation to estimate a period for when they might have been born: a historian who completed their dissertation in 1980 was likely born between ____ and _____.
+Using this temporal information, we are better able to infer their gender and chart how the larger representation of women and men changes year-by-year.
+
+[chart – raw numbers]
+
+Another way to examine this trend is to look at the proportion of male and female dissertation authors over time, which smoothes out changes in the absolute number of dissertations produced.
+The proportion of dissertations written by women has steadily increased over the past half-century, a change that began in the late 1960s and continued through the early 2000’s.
+Since that point, the proportion of dissertations written by women has largely plateaued at a few percentage points below the proportion written by men.
+Female historians have achieved something approaching parity with male historians in terms of how many women and women complete dissertations each year.
+
+[chart - proportion]
+
+But what happens after the dissertation?
+It is, after all, only the first major stage of a historian’s research output.
+The “coin of the realm” for many historians remains the publication of a scholarly monograph to be read and evaluated by peers.
+This process often takes the form of book reviews published in academic journals.
+One of the leading journals in the historical profession is the American Historical Review, the flagship journal of the American Historical Association.
+Along with a few dozen articles The AHR publishes five issues each year that contain, in aggregate, roughly one thousand books reviews covering (in its own words) “every major field of historical study.” The AHR is not only one of the widest-ranging journals in the profession, it is also the oldest; the journal has been publishing continuously since 1895.
+The range, scope, and length of the AHR makes it an ideal source to analyze the historical profession on a large scale.
+
+We take the AHR as a rough proxy for gatekeeping within the larger historical profession.
+When a book appears in the AHR it serves as a signal that other historians in positions of power have “approved” it for consideration.
+This signal is not necessarily correlated with the quality of the book.
+Even if it garners a negative review, the fact that it appears in the journal at all is a measure of the fact that the profession’s gatekeepers have deemed it important enough to review.
+On the flip side, a book that does not appear in the AHR may have gone on to have a significant impact on the profession.
+It is precisely this professional gatekeeping dimension that makes the AHR useful to study in the context of gender representation.
+
+Scraping the table-of-contents of every issue of the American Historical Review results in a dataset of close to 60,000 books reviewed by the journal since it began publication in 1895.
+Our program then inferred the gender of the authors of these books, which we could then use to plot the proportion of female and male authors over time.
+The temporal trajectory of gender representations roughly resembles that of history dissertations: women began making inroads in the late 1960s and have made steady gains over the past four decades.
+By the twenty-first century the proportion of female authors reviewed in the AHR had more than tripled.
+
+[graphic]
+
+But a closer look shows important differences between the proportion of women writing history dissertations and the proportion of women appearing in the AHR.
+Although both have trended upwards since the 1970s the slope is a lot steeper for dissertations than it is for the AHR.
+Female historians have very nearly closed the gaps in terms of newly completed dissertations, but the glass ceiling remains stubbornly low in terms of what happens from that point onwards.
+In book reviews published in the AHR male authors continue to outnumber female authors by a factor of nearly 2 to 1.
+The long-term gains made by female doctoral students do not carry over in the pages of the historical profession’s leading journal.
+
+[chart comparing dissertations with AHR]
+
+The path from a dissertation to an AHR book review is not an automatic one.
+Not all dissertations are published as books, not all books are submitted to the American Historical Review, and not all book submissions are accepted by the AHR editors for review.
+As of August 2014 the journal’s website reminds readers that “the AHR receives over 3,000 books a year; we have the resources to publish at most 1,000 reviews a year (approximately 200 per issue).” (http://www.indiana.edu/~ahrweb/book_review_guide.html - retrieved August 25 2014).
+Fewer and fewer female historians make it through each successive step.
+Establishing a single causal reason for this pattern is all but impossible, and instead a murky concoction of interrelated factors likely contributes to the gap.
+For one, a 2010 survey by the American Historical Association revealed the female historians dedicated substantially more time to child and elder care than their male colleagues, leaving less time for research.
+This gender gap is more likely to impact the production of a book than a dissertation, as graduate students are generally younger and less likely to be juggling child-rearing responsibilities while writing than their older peers.
+More insidiously, women continue to face insidious gender discrimination.
+
+We then used our program to infer the gender of not just book authors reviewed by the AHR, but the gender of the reviewers themselves.
+The story is much the same for reviewers as it is for authors: more than than twice as many men as women appear as reviewers in the journal.
+But gender inflects book reviews in less direct ways by shaping who writes reviews of which book authors.
+About three times as many men write reviews of male-authored books as do women.
+In the case of female-authored books the ratio of female-to-male reviewers is much closer to 50/50.
+In short, women are much more likely to write reviews of other women.
+And while men still write the majority of reviews of female-authored books, they tend to gravitate more towards male authors – who are, of course, already over-represented in the AHR.
+
+[chart: male authored books]
+[chart: female authored books]
+
+The American Historical Association also categorizes their reviews under different historical sub-fields.
+Although the journal’s taxonomy has changed quite a bit over time, it did allow us to refine our analysis.
+Were there more reviews of female authors in certain sub-fields rather than others, and has that changed over the past forty years?
+Caribbean/Latin American history has had something approaching equal representation for the past decade-and-a-half.
+In both African history and Ancient/Medieval history female historians made some quite dramatic gains during the late-nineties and aughts.
+The guiltiest parties, however, are also the two subject categories that publish the most book reviews in the AHR: Modern/Early Modern Europe and the United States/Canada.
+Both sub-fields have made progress over the four decades but still hover at around two-thirds male.
+Gender representation is not evenly distributed across the profession.
+
+[chart – sub-fields]
+
+One 2013 analysis of 2,500 recent history Ph.D.’s found that “gender played little role in employment patterns across particular professions and industries.” (http://historians.org/Documents/Many_Careers_of_History_PhDs_Final.pdf) Our own analysis reveals a more complicated reality.
+The flagship journal of professional historians continues to publish twice as many reviews of books by female authors as male authors.
+This disparity is put into even sharper relief when set against the relative parity achieved by women in producing history dissertations.
+Women might be getting hired at comparable rates to men, but discriminatory gatekeeping remains alive and well in the historical profession.
+Our program allows us to quantitatively measure the scope of the problem on a large scale, how it’s changed over time, and what remains to be done.
+
+Conclusion
+
+Our contributions:
+
+-     Broadly: gender studies + big data
+-   	Temporal dimension makes it better suited to fit the needs of humanities research than contemporary approaches – both more accurate (historical) and more nuanced (supplying the gendered proportion of a name for different periods)
+-   	Extensible – non US-names, older names. (future work)
+
+The fundamental points are that we have a much larger corpus of names.
+More importantly the ability to vary the look up based on a specific date of birth or a specified range of dates.
+Making this much more useful for historians, but also for any research in any field, given that the normal human lifespan present in any data set means that names can vary as much as a century, and so everyone needs to be using a fundamentally historical method.
+(Take that, sociologists.)
+
+Future work:
+
+Both the Python script and the R package are easily extensible to allow additional data sets, say from other countries.
+(Are there other data sets, say for Great Britain, http://data.gov.uk/data/search?q=names ??) And eventually perhaps US census data, which would solve the 19th century problem.
+
+Works Cited:

### 2014-09-03 article.Rmd
Added plot

@@
 ```{r}
+library(genderdata)
 library(gender)
 library(dplyr)
+library(ggplot2)
 ```
@@
 [GRAPHIC chart of changes in names – from README]
+
 [GRAPHIC showing change in popularity of names like Madison]
+
+```{r}
+plot_names <-  c("madison",
+                 "keanu",
+                 "addison",
+                 "jan",
+                 "sydney",
+                 "morgan",
+                 "leslie")
+
+ssa_national %>%
+  filter(name %in% plot_names) %>%
+  mutate(proportion_female = female / (female + male)) %>%
+ggplot(aes(x = year, y = proportion_female, color = name)) +
+  geom_line() +
+  ggtitle("Changes in selected names")
+
+```
 

### 2014-09-06 article.Rmd
Revise abstract and title

@@
 ---
-title: "Jane, John … Leslie? Algorithmic Gender Detection"
+title: "Jane, John … Leslie? A Historical Method for Algorithmic Gender Detection"
 author: "Cameron Blevins and Lincoln Mullen"
@@
 
-```{r}
+```{r, echo=FALSE, message=FALSE}
 library(genderdata)
@@
 
-This paper combines two trends in the humanities: the rise of gender as an analytical lens of inquiry and the proliferation of large, machine-readable datasets.
-It outlines a new method for algorithmically inferring the gender of a name by incorporating historical data to better reflect how naming practices change over time.
-This method ensures a higher level of accuracy with a precise statement of its reliability in inferring gender while tailoring it for humanistic study.
-Predicting gender from names requires a fundamentally historical method, which is demonstrated in this article.
-The article reports on the “gender” package for R which implements this method.
-The article demonstrates the method’s utility by applying it to a case study examining gender representation in the historical profession.
+This paper addresses two trends in the humanities: the rise of gender as an analytical tool and the proliferation of large, machine-readable datasets.
+It outlines a new method for algorithmically inferring the gender of a name from historical data in order to take into account how naming practices change over time and space.
+In contrast to existing methods of predicting gender, this method provides a higher level of accuracy and a precise statement of the likelihood that a name is a given gender.
+The article reports on the [Gender](https://github.com/ropensci/gender) package for the [R programming language](http://www.r-project.org/) which implements this method.
+The article demonstrates the method's utility by applying it to a case study in which we study gender in the history profession by inferring the gender of authors of history dissertations and of authors and reviewers in the *American Historical Review*.
 

### 2014-09-06 article.Rmd
Revised through selected names

@@
 
-```{r, echo=FALSE, message=FALSE}
+```{r load-libraries, include=FALSE}
 library(genderdata)
@@
 library(ggplot2)
+library(mullenMisc)
+library(knitr)
+
+# TRUE to show code and warnings; FALSE to remove them for publication
+display_code_blocks(FALSE)
 ```
@@
 In contrast to existing methods of predicting gender, this method provides a higher level of accuracy and a precise statement of the likelihood that a name is a given gender.
-The article reports on the [Gender](https://github.com/ropensci/gender) package for the [R programming language](http://www.r-project.org/) which implements this method.
+The article describes the methodology as implemented in the [Gender](https://github.com/ropensci/gender) package for the [R programming language](http://www.r-project.org/).
 The article demonstrates the method's utility by applying it to a case study in which we study gender in the history profession by inferring the gender of authors of history dissertations and of authors and reviewers in the *American Historical Review*.
 
-### INTRODUCTION. The "Leslie" Problem
+### Introduction: The "Leslie" Problem
 
-As researchers grapple with larger and larger datasets they are increasingly using algorithms not just to extract information from existing data, but to infer additional information from a dataset.
-[find example – network analysis, maybe http://kieranhealy.org/blog/archives/2013/06/09/using-metadata-to-find-paul-revere/] Many of these datasets are made up of records of individual people, such as company payrolls or military rosters.
-Some might contain associated information like a person’s home address, age, or rank.
+As historians and other humanities researchers grapple with larger and larger datasets, they are increasingly using algorithms not just to extract information from existing data but to infer additional information from the data that is available.^[CITE: find example – network analysis, maybe http://kieranhealy.org/blog/archives/2013/06/09/using-metadata-to-find-paul-revere/] Datasets of interest to historians---for example membership lists, company payrolls, military rosters, passenger bookings, records of correspondence, and lists of published works---might included records with the names of individual people.
+Some datasets might contain associated information like a person's address, age, or rank.
 Other datasets contain little more than a personal name.
+Given large but sparse datasets, researchers have to use their ingenuity to derive additional kinds of information.
 What kind of information can researchers infer from just a name?
-Perhaps most fundamentally, a person’s gender can often be inferred from just their first name.
-“Jane Fay” is almost certainly a woman, while her brother “John Fay” is almost certainly a man.
+Most obviously first names often imply a person's gender.
+"Jane Fay" is almost certainly a woman, while her brother "John Fay" is almost certainly a man.
 A straightforward computer program can infer their respective genders with relative ease.
@@
 The problem arises from the fact that the link between gender and naming practices, like language itself, is not static.
-What about Jane and John Fay’s long-lost third sibling, Leslie?
-If Leslie was born in the past sixty years, chances are good that she would be their sister.
+What about Jane and John Fay's sibling, Leslie?
+If Leslie was born in the past sixty years, chances are good that Leslie would be their sister.
 But if the three siblings were born in the early twentieth century, chances are good that Leslie would be their brother.
-That is because the gendered naming convention surrounding “Leslie” had completely flipped over the course of one hundred years.
+That is because the gendered naming convention surrounding "Leslie" completely changed over the course of one hundred years.
 92% of the babies born in the United States in 1900 who were named “Leslie” were male, while 96% of those identically named babies born in the year 2000 were female.
-A similar pattern holds true for any number of American names: Madison, Morgan, Sydney, Kendall.
-The historical change in the prevailing gender of a name tends to follow set patterns.
-In general, names tend to shift from being predominantly male to being predominantly female; the shift almost never happens in the other direction.
-The reason that the shift tends to be from male to female depends on the cause of the shift.
-In general names that move from male to female are also shifting from being unpopular to popular names.For example the name Madison changed from male to female suddenly after the 1985 because the name became extremely popular for naming baby girls, whereas before it had been an infrequently used male name.
+
+```{r leslie}
+# We cannot simply use the Social Security figures (even though they probably
+# don't present a particular problem for the name Leslie) because later we will
+# argue that the gender ratios for the SSA figures before 1930 are skewed. So we
+# are going to splice the IPUMS data from before 1930 and the SSA data from
+# after 1930.
+
+ipums_leslie <- genderdata::ipums_usa %>%
+  filter(name == "leslie", year <= 1930)
+
+ssa_leslie <- genderdata::ssa_national %>%
+  filter(name == "leslie", year > 1930)
+
+leslie_df <- rbind(ipums_leslie, ssa_leslie) %>%
+  mutate(prop_female = female / (female + male))
+
+leslie_plot <- leslie_df %>% 
+  ggplot(aes(x = year, y = prop_female)) +
+  geom_smooth() + 
+  geom_point() +
+  ggtitle("Proportion of female uses of the name 'Leslie,' 1820-2012") +
+  xlab(NULL) + ylab(NULL) + ylim(0, 1) +
+  geom_hline(yintercept = 0.5) + 
+  annotate("text", x = 1900, y = 0.45, label = "Male") +
+  annotate("text", x = 1900, y = 0.55, label = "Female")
+
+leslie_plot %>%
+  gg_attribution("Data: IPUMS USA; Social Security Administration") %>%
+  print()
+```
+
+Many other names as used in the United States, such as Madison, Morgan, Sydney, and Kendall, have also changed genders.
+Such changes in the prevailing gender of a name usually follow set patterns.
+In general, names tend to shift from being predominantly male to being predominantly female; the shift rarely happens in the other direction.
+In general, names that move from male to female also change from being unpopular to popular names.
+For example the name Madison changed from male to female suddenly after the 1985 because the name became extremely popular for naming baby girls, whereas before it had been an infrequently used male name.
 In other words, since the late twentieth century US at any rate, girls are more likely than boys to be receive a trendy name, and once a trendy name is associated with females it is unlikely to be used for males.
-The few exceptions of names which go from being female to male, such as Jan, have a different causal mechanism: it seems likely that Jan as an English name is female but that Jan as a Scandinavian (Finnish?
-Swedish?) name is male.
-Because the name is relatively unpopular, changing patterns of immigration can explain the change.
-
-[CITE: literature on baby names generally]
-
-[GRAPHIC chart of changes in names – from README]
-
-[GRAPHIC showing change in popularity of names like Madison]
+The few exceptions of names which go from being female to male, such as Jan, have a different causal mechanism.
+It seems likely that Jan as an English name is female but that Jan as a Scandinavian name is male.
+Because the name has remained relatively uncommong, other causes besides changing gender preferences can explain the change.
 
@@
                  "morgan",
-                 "leslie")
+                 "leslie",
+                 "kendall")
 
-ssa_national %>%
+selected_names_plot <- ssa_national %>%
   filter(name %in% plot_names) %>%
-  mutate(proportion_female = female / (female + male)) %>%
-ggplot(aes(x = year, y = proportion_female, color = name)) +
-  geom_line() +
-  ggtitle("Changes in selected names")
+  mutate(prop_female = female / (female + male)) %>%
+ggplot(aes(x = year, y = prop_female, color = name)) +
+  geom_smooth(se = FALSE) +
+  ggtitle("Changes in selected names, 1930-2012") +
+  xlim(1930, 2012) + 
+  xlab(NULL) + ylab(NULL) + ylim(0, 1) +
+  geom_hline(yintercept = 0.5) +
+  annotate("text", x = 1965, y = 0.45, label = "Male") +
+  annotate("text", x = 1965, y = 0.55, label = "Female")
 
+selected_names_plot %>%
+  gg_attribution("Data: Social Security Administration") %>%
+  print()
 ```

### 2014-09-06 article.Rmd
Paragraph on Genderize.io

@@
 
-For researchers working on contemporary subjects the “Leslie” problem is not an especially pressing one.
-An education policy analyst studying the demographics of different urban school districts, for instance, can turn to a variety of tools that use current databases of names.
-(http://genderize.io/, http://genderchecker.com/about-us.aspx) These tools will most likely identify the gender of Leslie Fay correctly.
+For researchers working on contemporary subjects the "Leslie" problem is not an especially pressing one.
+An education policy analyst studying the demographics of different urban school districts, for instance, can turn to a variety of tools that use current databases of names, and these tools will likely identify the gender of Leslie Fay correctly.
+For example, the [Genderize.io](http://genderize.io/) online service determines the gender of first names "from user profiles across major social networks." Checking the service for the name Leslie returns a prediction that the name is female, along with an estimate of what proportion of the uses among members of unnamed social media sites is female and male.^[In this example the Genderize.io service has been accessed through the Gender package described below.]
+
+```{r genderize, echo=TRUE, results='markup'}
+gender("Leslie", method = "genderize")
+```
+
 But researchers studying historical subjects need to take into account the dynamic relationship between gender and naming practices.

### 2014-09-07 article.Rmd
Revised rest of the introduction

@@
 
-But researchers studying historical subjects need to take into account the dynamic relationship between gender and naming practices.
-If that same researcher wanted to compare the demographics of contemporary urban schools with those from the early twentieth century, Leslie Fay’s gender would get misidentified.
-And the “Leslie” problem is not just one for historians studying the past.
-Anyone studying a longer timeframe will also encounter this problem: as of 2012, the average American has a lifespan of nearly seventy-nine years – more than enough time for naming practices to change quite dramatically.
-(cite World Bank).
+Researchers studying historical subjects, on the other hand, need to take into account changes in naming practices.
+If that same researcher wanted to compare the demographics of contemporary urban schools with those from the early twentieth century, contemporary datasets would misidentify Leslie Fay's gender.
+The "Leslie" problem is not just for researchers who think of themselves as historians.
+Anyone studying a timeframe longer than a few years, or anyone studying a group whose characteristics do not match the probably unknowable characteristics of the groups that comprise social media datasets will also encounter this problem.
+As of 2012, the average American has a lifespan of nearly seventy-nine years–--more than enough time for naming practices to change quite dramatically.^[CITATION] Predicting gender from first names therefore requires a fundamentally historical method.
 
-Our solution to the “Leslie” problem is to create a library that explicitly takes into account changes in naming practices.
-The program incorporates this temporal dimension using historical databases to more accurately infer whether “Leslie” was male or female depending on when he or she lived.
-It not only achieves better accuracy in inferring gender, but it also provides the relative likelihood of a person’s gender for a given range of years: 52% of the babies named “Leslie” born in 1950 were girls and 48% were boys.
-Rather than a black box spitting out a stark FEMALE or MALE designation, the program computes a measurement for each gender’s relative “share” of a given name.
+Our solution to the "Leslie" problem is to create a software package that combines a predictive algorithm with a range of historical dataset suitable to various times and regions.
+The algorithm calculates the proportion of male or female uses of a name in a given birth year or range of years.
+It thus can provide not only a prediction of the gender of a name, but also an estimate of how likely that prediction is.
+For example 52 percent of the babies named Leslie born in the United States in 1950 were girls and 48% were boys.
 This allows researchers to conduct more nuanced analysis and to determine for themselves what proportion is acceptable for making a prediction of male or female.
@@
 We recognize that gender is a fluid concept, both today and in the past, and that this method glosses over the complexities of historical definitions of gender and of the relationship between biological sex and gender.
-In some sense this is necessary since the large data sets which make this kind of analysis possible were created by state agencies, and for gender (as for race) state agencies have created a limited number of categories which have tremendous definitional power.
-[CITE e.g.
+In some sense this is unavoidable since the large datasets which make this kind of analysis possible for the past were created by state agencies.
+For the category of gender (as for race) states have offered a limited number of boxes to check, and those boxes extert a tremendous power to define.^[CITE e.g.
 Margot Canady, The Straight State; works on e.g.
@@
 First, researchers are likely to use this method in the aggregate rather than for individuals.
-The assumption that a name implies male or female (and that those are possible categories to be inferred) has a higher validity when applied to large groups of people than it does when applied to individuals.
-Second, this methods is neutral as to what the categories male or female actually mean in any given historical context.
-Researchers are free---indeed, obligated---to make that determination for themselves, and this package provides access to the underlying data set so that a technologically competent researcher can interrogate every assumption through the process.
+The assumption that a name implies male or female (and that those are possible categories to be inferred) is more valid when applied to large groups of people than it does when applied to individuals.
+Second, this method is neutral as to what the categories male or female actually mean in any given historical context.
+Researchers are free---indeed, obligated---to make that determination for themselves, and this package provides access to the underlying dataset so that researchers can interrogate every assumption we have made.
 This method does not argue against the complexities of historical understandings of gender, but as with any project it does require researchers who use the method to think through how the data was gathered, what it implies, and how it reflects social and cultural practices.
 
-The remainder of the article is divided into two sections.
-First, we present the methodology of our approach by comparing it to existing methods and detailing the strengths and weaknesses of the program.
-Second, we apply the method to a case study in order to demonstrate how humanities scholars might use it by analyzing gender representation within the historical profession.
+The remainder of this article is divided into two sections.
+First, we describe our method in more detail, compare it to existing methods, and explain how to use the method.
+Second, we apply the method to a case study of the historical profession in order to demonstrate the usefulness of the method.
 

### 2014-09-07 article.Rmd
Fixed headings
Revised through problem with existing methods

@@
 library(knitr)
+library(scales)
 
@@
 display_code_blocks(FALSE)
+
+# Add a thousands separator
+knit_hooks$set(inline = function(x) {
+  prettyNum(x, big.mark=",")
+})
+
 ```
@@
 
-```{r}
+```{r plot_names_that_change}
 plot_names <-  c("madison",
@@
 
-PART I. The Method
+### I. The Method
 
-Existing Methods:
+```{r calc_unique_names}
+# Numbers of unique names in different data sets, for reference below
+unique_names <- function(dataset) {
+  dataset$name %>% unique %>% length
+  }
 
-The historical method to be described below may be compared to a method available in the Natural Language Toolkit for Python.
-The NLTK provides many tools for analyzing natural language, including two lists of male and female names created by Mark Kantrowitz with assistance from Bill Ross.
-This list includes 7579 [DOUBLE CHECK] unique names.
-Using the Kantrowitz names corpus in NLTK, one could look up a name like Jane or John and find out that it is male or female.
-[NOTE: The Kantrowitz corpus is also implemented in the gender package for R for comparison, though its use is not recommended for actual research.]
+kantrowitz_unique <- unique_names(genderdata::kantrowitz)
+ssa_unique <- unique_names(genderdata::ssa_national)
+ipums_unique <- unique_names(genderdata::ipums_usa)
+```
 
-[SAMPLE OUTPUT in Python from NLTK names corpus converted to an array of tuples.]
+Our historical method for inferring gender from names may be compared to a method available in the [Natural Language Toolkit](http://www.nltk.org/) for Python.^[Steven Bird, Edward Loper, and Ewan Klein, *Natural Language Processing with Python* (O’Reilly, 2009), ch.
+2.
+Chapter 6 of the NLTK book also describes a method of predicting gender from the last letter of the name, but this example is given for the sake of explaining supervised classification and not as an end in itself, so we do not have reason to address this ahistorical method.] The NLTK is an influential software package for scholarship because it provides an extraordinary range of tools for analyzing natural language.
+Included in the NLTK are two lists of male and female names created by the computer scientist Mark Kantrowitz with assistance from Bill Ross.^[The [original lists of names](http://www.cs.cmu.edu/afs/cs/project/ai-repository/ai/areas/nlp/corpora/names/0.html) are available at the website of the computer science department at Carnegie Mellon University, and the NLTK implementation is described in its [documentation](http://www.nltk.org/howto/corpus.html).] These lists include `r kantrowitz_unique` unique names.
+Using the Kantrowitz names corpus in the NLTK, one could look up a name like Jane or John and find out that it is male or female.
+The Kantrowitz corpuse provides the list of names.^[The examples below is drawn from our Gender package which implements the Kantrowitz dataset for comparsion's sake, though its use is not recommended for actual research.
+The Gender package combines the Kantrowitz corpus's two lists into a single list, identifying names that are ambiguous because they appear on both lists as `"either"`.]
 
-[('Aaron', 'male'), 
-('Abbey', 'male'), 
-('Abbie', 'male'), 
-('Abbot', 'male')
-('Zorine', 'female'), 
-('Zsa Zsa', 'female'), 
-('Zsazsa', 'female'), 
-('Zulema', 'female'), 
-('Zuzana', 'female')]
+```{r kantrowitz list, echo=TRUE, results='markup'}
+genderdata::kantrowitz
+```
 
-[CITE: Link to original name corpus: http://www.cs.cmu.edu/afs/cs/project/ai-repository/ai/areas/nlp/corpora/names/0.html]
+One can then easily write a function which looks up the gender of a given name.
+
+```{r abby-kantrowitz, echo=TRUE, results='markup'}
+gender("abby", method = "kantrowitz")
+```
+
+```{r compare-kantrowitz-ssa, cache=TRUE, eval=FALSE}
+# Calculate the error rate of the Kantrowitz data set. This code is very slow, 
+# so I have disabled evaluating and hardcoded the value into the text for now.
+# We can re-run this for the final version of the article.
+cf_kantrowitz_ssa <- data.frame(name = unique(genderdata::kantrowitz$name),
+                                stringsAsFactors = "FALSE")
+
+# For just the names in the Kantrowitz corpus, predict the gender with both the Kantrowitz and the SSA method.
+cf_kantrowitz_ssa <- cf_kantrowitz_ssa %>%
+  group_by(name) %>%
+  do(kantrowitz = gender(name = .$name, method = "kantrowitz")$gender,
+     ssa = gender(name = .$name, method = "ssa")$gender)
+
+cf_results <- as.character(cf_kantrowitz_ssa$kantrowitz) ==
+  as.character(cf_kantrowitz_ssa$ssa)
+
+# Error rate is the number of mismatches to the total number of names
+kantrowitz_error_rate <- 1 -(sum(cf_results) / length(cf_results))
+```
+
+```{r}
+# Hardcode the error rate for now. This value is based on running the code above
+# locally.
+kantrowitz_error_rate <- 0
+```
 
@@
 As pointed out above, this makes the corpus unusable for historical purposes.
-Furthermore the corpus includes some oddities which make it less suitable.
-Some names like “Bill” are overwhelmingly male, but the corpus includes them as both male and female.
-And because the Kantrowitz corpus contains only 7,579 (a mere 8.2% of the 91,320 names released by the Social Security Administration for the US alone) there are many names that it cannot identify.
-[NOTE: The SSA data includes only names used more than 5 times in a given year, so the “long tail” of actual number of names in the US is considerably larger.
-Link to SSA data.] In this chart, for example, the “unidentified” category grows rapidly after 1965.
-While our Social Security method relies primarily on children born in the US but also people who become naturalized citizens (double check) whereas as census data for example also includes immigrants, nevertheless it does far better at identifying non-European names since it contains over an order of magnitude more names.
-
-[CITE existing literature on gender classification]
+Furthermore the corpus includes some oddities which make it less than suitable.
+Some names like Abby are overwhelmingly female and some like Bill are overwhelmingly male, but the corpus includes them as both male and female.
+The Kantrowitz corpus contains only `r kantrowitz_unique`, a mere `r percent(kantrowitz_unique / ssa_unique)` of the `r ssa_unique` names in the Social Security Administration dataset and `r percent(kantrowitz_unique / ipums_unique)` of the `r ipums_unique` names in the census records provided by IPUMS USA.
+There are therefore many names that it cannot identify.
+Assuming for the moment that our method provides more accurate results, we estimate that `r percent(kantrowitz_error_rate)` percent of the names in the Kantrowitz corpuse are misclassified or classified as ambigious when a prediction could be made.
 
 We mention the Kantrowitz name corpus as implemented in NLTK because the Natural Language Toolkit is rightly regarded as influential for scholarship.
-It’s flaws are also emblematic of the problems with most other implementations of gender prediction algorithms.
-The Genderize website and API is, for example, a much more sophisticated implementation of a gender prediction.
+Its flaws are also typical of the problems with most other implementations of gender prediction algorithms.
+The Genderize.io API is, for example, a more sophisticated implementation of gender prediction than the NLTK algorithm.
 Besides predicting male or female for gender, it also reports the proportion of male or female names, along with a count of the number of uses of the name that it is basing its prediction on.
 Genderize will also permit the user to customization a prediction for different countries: an extremely important feature.
-Genderize includes “database contains 142848 distinct names across 77 countries and 85 languages.” Genderize is unsuitable for historical work, however, because it is based only on contemporary data.
-According to the documentation for its API, “It utilizes big datasets of information, from user profiles across major social networks.” [SOURCE: http://genderize.io] It would be anachronistic to apply these datasets to the past, and Genderize provides no functionality to filter results chronologically as it does geographically.
-[NOTE: A wrapper to the genderize API is included in the gender package.
-There are also wrappers available in Ruby and Python.
-There are a number of other website which provide similar functionality based on apparently similar datasets, none of which permit historical variation.)
+Genderize.io reports that its "database contains 142848 distinct names across 77 countries and 85 languages." Genderize is unsuitable for historical work, however, because it is based only on contemporary data.
+According to the documentation for its API, "It utilizes big datasets of information, from user profiles across major social networks."^[Quotations are taken from the [Genderize.io website](http://genderize.io), with correspondence from the developer behind the site for confirmation.] It would be anachronistic to apply these datasets to the past, and Genderize.io provides no functionality to filter results chronologically as it does geographically.
+In addition, exactly what the dataset comprises and how it was gathered is unclear, which keeps scholars from interrogating the value of the source.
 
-Surprisingly, CRAN does not include any packages for doing this in R, though some packages do include functions for this.
-Python has no existing package (double check) besides the NLTK and Genderize.Thus the two of the most commonly used languages for data analysis, particularly within the digital humanities community, have no satisfactory package that could easily solve the problem of discovering the gender of names, though options for geocoding, for example, abound.
+R and Python are the most commonly used languages for data analysis among digital humanists.
+Python's existing packages for gender prediction all implement a method similar to the NLTK or Genderize.io.^[Python library names] To our knowledge, [CRAN](http://cran.rstudio.com/) (the central repository for R packages) does not include any packages dedicated to gender prediction.
+Though options abound for geocoding, for example, there is no satisfactory existing method of gender prediction for historical and humanistic research.
 
@@
 
-PART II. Measuring Gatekeeping in the Historical Profession
+### II. Measuring Gatekeeping in the Historical Profession
 
@@
 
-Conclusion
+### Conclusion
 
@@
 
-Works Cited:
+----
+
+Things cut to be moved to more suitable places:
+
+[NOTE: The SSA data includes only names used more than 5 times in a given year, so the "long tail" of actual number of names in the US is considerably larger. Link to SSA data.]

### 2014-09-07 article.Rmd
Hardcoded Kantrowitz error rate

@@
 # locally.
-kantrowitz_error_rate <- 0
+kantrowitz_error_rate <- 0.2325766
 ```
@@
 
-The Solution:
+To that end we have created the [Gender](https://github.com/ropensci/gender) package for R which contains the , and an associated [genderdata](https://github.com/lmullen/gender-data-pkg) package which contains various historical datasets.
 
 The possibilities for encoding gender from names depends on two things.
-First, it depends on having a suitable (and suitably large) data set for the time period and region under study.
-Unsurprisingly such data sets are almost always gathered in the first instance by states, though the compilation and digitization may be undertaken by scholarly researchers.
-Second, it depends on a suitable algorithms for estimating the proportion of male and female names for a given year, or range of years since often a person cannot be associated with an exact date.
-It is important that the algorithm take into account any biases in the data to make correct predictions.
-
-Data:
+First, it depends on having a suitable (and suitably large) dataset for the time period and region under study.
+Unsurprisingly such datasets are almost always gathered in the first instance by states, though the compilation and digitization may be undertaken by scholarly researchers.
+Second, it depends on a suitable algorithm for estimating the proportion of male and female names for a given year, or range of years since often a person cannot be associated with an exact date.
+It is especially important that the algorithm take into account any biases in the data to make correct predictions.
 

### 2014-09-07 article.Rmd
Description of project history

@@
 There are therefore many names that it cannot identify.
-Assuming for the moment that our method provides more accurate results, we estimate that `r percent(kantrowitz_error_rate)` percent of the names in the Kantrowitz corpuse are misclassified or classified as ambigious when a prediction could be made.
+Assuming for the moment that our method provides more accurate results, we estimate that `r percent(kantrowitz_error_rate)` percent of the names in the Kantrowitz corpuse are misclassified or classified as ambigious when a valid prediction could be made.
 
@@
 
-To that end we have created the [Gender](https://github.com/ropensci/gender) package for R which contains the , and an associated [genderdata](https://github.com/lmullen/gender-data-pkg) package which contains various historical datasets.
+To that end we have created the [Gender](https://github.com/ropensci/gender) package for R which contains the predictive algorithm, and an associated [genderdata](https://github.com/lmullen/gender-data-pkg) package which contains various historical datasets.
+This R implementation is based on an [earlier Python implementation](https://github.com/cblevins/Gender-ID-By-Time) by [Cameron Blevins](http://www.cameronblevins.org/) and [Bridget Baird](http://www.conncoll.edu/directories/emeritus-faculty/bridget-baird/).
 
@@
 It is especially important that the algorithm take into account any biases in the data to make correct predictions.
-
-Besides including the
+Development of the R package has had two primary aims.
+The first is to abstract the predictive algorithm to the simplest possible form so that it is usable for a wide range of historical problems rather than depending on the format of any particular data set.
+The second has been to provide as many datasets as possible in order to localize the predictions to particular times and places.
+To that end, [Ben Schmidt](http://benschmidt.org/) contributed a dataset of US Census Data and accompanying code, as well as providing a significant correction to predictions based on Social Security Administration data.
 

### 2014-09-08 article.Rmd
Overview of data; mention rOpenSci

@@
 This R implementation is based on an [earlier Python implementation](https://github.com/cblevins/Gender-ID-By-Time) by [Cameron Blevins](http://www.cameronblevins.org/) and [Bridget Baird](http://www.conncoll.edu/directories/emeritus-faculty/bridget-baird/).
+The Gender package is affiliated with [rOpenSci](http://ropensci.org/), an initiative that supports reproducible research and open data for scientists using R.
+The rOpenSci advisory board has provided code review and publicity for the project.
 
@@
 To that end, [Ben Schmidt](http://benschmidt.org/) contributed a dataset of US Census Data and accompanying code, as well as providing a significant correction to predictions based on Social Security Administration data.
+
+ The Gender package currently includes two datasets which make it suitable for use for United States history since the first federal census in 1790.^[The authors of the package are both American historians, which explains why the initial release of the package focuses on that country.
+We have plans, however, to extend the underlying data to six other European and North American countries included in the [North Atlantic Population Project](https://www.nappdata.org/napp/) for the nineteenth century.
+We also welcome contributions of datasets for other regions and times from other scholars.] The first dataset contains names of applicants for Social Security and available from Data.gov.
+The second dataset contains names gathered in the decennial censuses and available from the IPUMS-USA (Integrated Public Use Microdata Series) service from the Minnesota Population Center at the University of Minnesota.
 

### 2014-09-08 article.Rmd
Describe SSA dataset

@@
 
-The Social Security Administration provides annual data on the names and genders of babies stretching back to the year 1880.
-This is database encompases 91,320 unique names.
-For comparison, [example of other databases and their size].
-What makes it even more useful, however, is its temporal dimension.
-This allows a researcher to figure out not just the likelihood that “Leslie” is a male or female name, but the likelihood that “Leslie” is a male or female name for a specific range of years, or for a specific year.
+The Social Security Administration "[Baby Names](http://catalog.data.gov/dataset/baby-names-from-social-security-card-applications-national-level-data)" dataset was created as a result of the Social Security Act of 1935 during the New Deal.
+^[Social Security Administration, "[Baby Names from Social Security Card Applications-National Level Data](http://catalog.data.gov/dataset/baby-names-from-social-security-card-applications-national-level-data)," Data.gov dataset ID US-GOV-SSA-338 (accessed August 31, 2014).
+The Social Security Administration also provides a related dataset which contains the names of applicants broken down by state and year since 1910.
+Social Security Administration, "[Baby Names from Social Security Card Applications-Data by State and District of Columbia](http://catalog.data.gov/dataset/baby-names-from-social-security-card-applications-data-by-state-and-district-of-)," Data.gov dataset ID US-GOV-SSA-339 (accessed August 31, 2014).] This dataset contains a list of first names with the number of male and female uses of the name in every year beginning with 1880 and ending with 2012.
+The SSA has included any name which was used more than five times in a given year; otherwise the dataset is a 100 percent sample.
+The description "baby names" provided by the SSA is a serious misnomer.
+When Social Security first became available during the New Deal, its first beneficiaries were adults past or near retirment age.
+The dataset goes back to 1880, the birth year for a 55 year-old adult when Social Security was enacted.
+Even after 1935, registration at birth for Social Security was not mandatory until 1986.
+As we will demonstrate below, the way in which the data was gathered requires an adjustment to our predictions of gender.^[CITE history of SSA.
+On the history of birth registration see Shane Landrum, "The State's Big Family Bible: Birth Certificates, Personal Identity, and Citizenship in the United States, 1840--1950" (PhD dissertation, Brandeis University, 2014).
+It is well-known among American historians that Social Security was written to exclude farmworkers with the aim of excluding African Americans.
+In examining the dataset we have not found any reason to believe that this policy noticably skewed the prediction of gender from names.]
 
-Using for example the SSA data for a few names (once it has been munged into a more useful form).
-The columns male and female record the number of people of that gender (sex?) born in that year who received Social Security cards.
-
-name         year    female     male
----------  ------  --------  -------
-jane         1890       372        0
-jane         1990       771        8
-john         1890        56     8502
-john         1990        81    29066
-madison      1890         0       14
-madison      1990      1407      145
-
-Furthermore, we can calculate the proportions of male and female and on that basis guess the gender for a given year.
-
-name         year    female     male    proportion_male    proportion_female  gender  
----------  ------  --------  -------  -----------------  -------------------  --------
-jane         1890       372        0             0.0000               1.0000  female  
-jane         1990       771        8             0.0103               0.9897  female  
-john         1890        56     8502             0.9935               0.0065  male    
-john         1990        81    29066             0.9972               0.0028  male    
-madison      1890         0       14             1.0000               0.0000  male    
-madison      1990      1407      145             0.0934               0.9066  female
-ADD LESLIE AND ANOTHER EXAMPLES WE USE
-
-Furthermore, one can specify a range of years and calculate the proportion for that time period.
-
-Given this data set, it is simple to perform perform a merge (a join, in SQL terms) on either a name or a combination of name and year to link this data to a date set.
+TODO IPUMS
 

### 2014-09-08 article.Rmd
Deleted extract
minor editing

@@
 The article describes the methodology as implemented in the [Gender](https://github.com/ropensci/gender) package for the [R programming language](http://www.r-project.org/).
-The article demonstrates the method's utility by applying it to a case study in which we study gender in the history profession by inferring the gender of authors of history dissertations and of authors and reviewers in the *American Historical Review*.
+It goes on to demonstrate the method's utility by applying it to a case study in which we study gender in the history profession by inferring the gender of authors of history dissertations and of authors and reviewers in the *American Historical Review*.
 
@@
 
-Many other names as used in the United States, such as Madison, Morgan, Sydney, and Kendall, have also changed genders.
-Such changes in the prevailing gender of a name usually follow set patterns.
-In general, names tend to shift from being predominantly male to being predominantly female; the shift rarely happens in the other direction.
-In general, names that move from male to female also change from being unpopular to popular names.
-For example the name Madison changed from male to female suddenly after the 1985 because the name became extremely popular for naming baby girls, whereas before it had been an infrequently used male name.
-In other words, since the late twentieth century US at any rate, girls are more likely than boys to be receive a trendy name, and once a trendy name is associated with females it is unlikely to be used for males.
-The few exceptions of names which go from being female to male, such as Jan, have a different causal mechanism.
+Many other American names, such as Madison, Morgan, Sydney, and Kendall, have also changed gender over the past century.
+Such changes usually follow set patterns.
+In general, names tend to shift from being predominantly male to being predominantly female; the shift rarely moves in the other direction.
+In general, names that transition from male to female also became increasingly popular over time.
+For example, the name "Madison" changed from male to female suddenly after 1985.
+Whereas before it had been an infrequently used male name, after 1985 it skyrocketed in popularity, and by the early 2000s it was one of the ten most frequent names for baby girls.
+Broadly speaking, since the late twentieth century Americans girls are more likely than American boys to receive a "trendy" name, and once that name is associated with females it is unlikely to be given to male babies.
+The few exceptions of names which go from being female to male (such as "Jan") have a different causal mechanism.
 It seems likely that Jan as an English name is female but that Jan as a Scandinavian name is male.
-Because the name has remained relatively uncommong, other causes besides changing gender preferences can explain the change.
+Because the name has remained relatively uncommon, other causes besides changing gender preferences can explain the change.
+**this last bit about "Jan" might just muddle things - unless we have a clear explanation or hypothesis as to what does explain the change, not sure it's worth highlighting**
 
@@
 (Are there other data sets, say for Great Britain, http://data.gov.uk/data/search?q=names ??) And eventually perhaps US census data, which would solve the 19th century problem.
-
-----
-
-Things cut to be moved to more suitable places:
-
-[NOTE: The SSA data includes only names used more than 5 times in a given year, so the "long tail" of actual number of names in the US is considerably larger. Link to SSA data.]

### 2014-09-09 article.Rmd
Tweaks and questions
Minor editing, a few citations, and posed a couple of questions.

@@
 For example 52 percent of the babies named Leslie born in the United States in 1950 were girls and 48% were boys.
-This allows researchers to conduct more nuanced analysis and to determine for themselves what proportion is acceptable for making a prediction of male or female.
+This allows researchers to conduct more nuanced analysis and to determine for themselves what proportion is acceptable for making a prediction of gender.
 
 We recognize that gender is a fluid concept, both today and in the past, and that this method glosses over the complexities of historical definitions of gender and of the relationship between biological sex and gender.
-In some sense this is unavoidable since the large datasets which make this kind of analysis possible for the past were created by state agencies.
-For the category of gender (as for race) states have offered a limited number of boxes to check, and those boxes extert a tremendous power to define.^[CITE e.g.
-Margot Canady, The Straight State; works on e.g.
-Virginia’s definition of race.
-Shane Landrum on birth certificates.
-James C.
-Scott on Seeing Like a State or is that obvious to everyone by now?] Nevertheless, this method is suitable for two reasons.
+In some sense this is unavoidable since the large historical datasets which make this kind of analysis possible were created by state agencies.
+For the category of gender (as for race) states have offered a limited number of boxes to check, and those boxes extert a tremendous power to define state subjects.^[CITE One of the best-known theorists on gender is Judith Butler, *Gender Trouble* and Butler, *Bodies that Matter*; For the problems with the word "identity" in the humanities and social sciences, see Rogers Brubaker and Frederick Cooper, "Beyond 'Identity'"; Several historians have written about the American state and its drive to impose identities onto subjects: Margot Canady, The Straight State;  Peggy Pascoe, *What Comes Naturally*; Shane Landrum on birth certificates.] Nevertheless, this method is suitable for two reasons.
 First, researchers are likely to use this method in the aggregate rather than for individuals.
 The assumption that a name implies male or female (and that those are possible categories to be inferred) is more valid when applied to large groups of people than it does when applied to individuals.
-Second, this method is neutral as to what the categories male or female actually mean in any given historical context.
-Researchers are free---indeed, obligated---to make that determination for themselves, and this package provides access to the underlying dataset so that researchers can interrogate every assumption we have made.
+**I think this first line of reasoning is dangerous.
+This leaves us open to critiques of "big data" in general from humanists and theorists: that the process of aggregation is actually really dangerous and that assuming that individual identity and experience can be subsumed under the generalities of the crowd, etc.
+etc.
+I think the second reason is strong enough on its own.** Second, this method is neutral as to what the categories male or female actually mean in any given historical context.
+Researchers are free---indeed, obligated---to make that determination for themselves, and this package provides access to the underlying dataset so that researchers can interrogate every assumption that comes with them.
 This method does not argue against the complexities of historical understandings of gender, but as with any project it does require researchers who use the method to think through how the data was gathered, what it implies, and how it reflects social and cultural practices.
@@
 The remainder of this article is divided into two sections.
-First, we describe our method in more detail, compare it to existing methods, and explain how to use the method.
-Second, we apply the method to a case study of the historical profession in order to demonstrate the usefulness of the method.
+First, we describe our method in more detail, compare it to existing methods, and explain how to use it.
+Second, we apply the method to a case study of gatekeeping in the historical profession in order to demonstrate its utility.
 
@@
 
-Our historical method for inferring gender from names may be compared to a method available in the [Natural Language Toolkit](http://www.nltk.org/) for Python.^[Steven Bird, Edward Loper, and Ewan Klein, *Natural Language Processing with Python* (O’Reilly, 2009), ch.
+Why do scholars need a historical method for inferring gender?
+Let's start with a comparison.
+One existing method for predicting gender is available in the [Natural Language Toolkit](http://www.nltk.org/) for Python.^[Steven Bird, Edward Loper, and Ewan Klein, *Natural Language Processing with Python* (O’Reilly, 2009), ch.
 2.
-Chapter 6 of the NLTK book also describes a method of predicting gender from the last letter of the name, but this example is given for the sake of explaining supervised classification and not as an end in itself, so we do not have reason to address this ahistorical method.] The NLTK is an influential software package for scholarship because it provides an extraordinary range of tools for analyzing natural language.
-Included in the NLTK are two lists of male and female names created by the computer scientist Mark Kantrowitz with assistance from Bill Ross.^[The [original lists of names](http://www.cs.cmu.edu/afs/cs/project/ai-repository/ai/areas/nlp/corpora/names/0.html) are available at the website of the computer science department at Carnegie Mellon University, and the NLTK implementation is described in its [documentation](http://www.nltk.org/howto/corpus.html).] These lists include `r kantrowitz_unique` unique names.
+Chapter 6 of the NLTK book also describes a method of predicting gender from the last letter of the name, but this example is given for the sake of explaining supervised classification and not as an end in itself, so we do not have reason to address this a historical method.] The NLTK is an influential software package for scholarship because it provides an extraordinary range of tools for analyzing natural language.
+**I think we could develop this even more, if we have information about it.
+My impression is that it's the "industry standard" for a lot of folks, but it would be nice to give some evidence to then show why our intervention matters, and that the Kantrowitz method isn't just a straw man** Included in the NLTK are two lists of male and female names created by the computer scientist Mark Kantrowitz with assistance from Bill Ross.^[The [original lists of names](http://www.cs.cmu.edu/afs/cs/project/ai-repository/ai/areas/nlp/corpora/names/0.html) are available at the website of the computer science department at Carnegie Mellon University, and the NLTK implementation is described in its [documentation](http://www.nltk.org/howto/corpus.html).] These lists include `r kantrowitz_unique` unique names.
 Using the Kantrowitz names corpus in the NLTK, one could look up a name like Jane or John and find out that it is male or female.
-The Kantrowitz corpuse provides the list of names.^[The examples below is drawn from our Gender package which implements the Kantrowitz dataset for comparsion's sake, though its use is not recommended for actual research.
+The Kantrowitz corpus provides the list of names.^[The examples below is drawn from our Gender package which implements the Kantrowitz dataset for comparsion's sake, though its use is not recommended for actual research.
 The Gender package combines the Kantrowitz corpus's two lists into a single list, identifying names that are ambiguous because they appear on both lists as `"either"`.]
@@
 The most significant problem with the Kantrowitz names corpus and thus the NLTK implementation is that it assumes that names are timeless.
-As pointed out above, this makes the corpus unusable for historical purposes.
-Furthermore the corpus includes some oddities which make it less than suitable.
+As pointed out above, this makes the corpus problematic for historical purposes.
+Furthermore the Kantrowitz corpus includes other oddities which make it less suitable.
 Some names like Abby are overwhelmingly female and some like Bill are overwhelmingly male, but the corpus includes them as both male and female.
-The Kantrowitz corpus contains only `r kantrowitz_unique`, a mere `r percent(kantrowitz_unique / ssa_unique)` of the `r ssa_unique` names in the Social Security Administration dataset and `r percent(kantrowitz_unique / ipums_unique)` of the `r ipums_unique` names in the census records provided by IPUMS USA.
+The Kantrowitz corpus contains only `r kantrowitz_unique` names, a mere `r percent(kantrowitz_unique / ssa_unique)` of the `r ssa_unique` names in the Social Security Administration dataset and `r percent(kantrowitz_unique / ipums_unique)` of the `r ipums_unique` names in the census records provided by IPUMS USA.
 There are therefore many names that it cannot identify.
-Assuming for the moment that our method provides more accurate results, we estimate that `r percent(kantrowitz_error_rate)` percent of the names in the Kantrowitz corpuse are misclassified or classified as ambigious when a valid prediction could be made.
+Assuming for the moment that our method provides more accurate results, we estimate that `r percent(kantrowitz_error_rate)` percent of the names in the Kantrowitz corpus are misclassified or classified as ambigious when a valid prediction could be made.
+**This isn't immediately obvious to me how this error rate was derived.
+Do we know why or how Kantrowitz is classifying things as ambiguous?
+Maybe he's just using a much higher, and theoretically more rigorous, cut-off (say 90%) than our 50% cut-off.**
 
@@
 Besides predicting male or female for gender, it also reports the proportion of male or female names, along with a count of the number of uses of the name that it is basing its prediction on.
-Genderize will also permit the user to customization a prediction for different countries: an extremely important feature.
+Genderize will also permit the user to customize a prediction for different countries, which is an extremely important feature.
 Genderize.io reports that its "database contains 142848 distinct names across 77 countries and 85 languages." Genderize is unsuitable for historical work, however, because it is based only on contemporary data.
 According to the documentation for its API, "It utilizes big datasets of information, from user profiles across major social networks."^[Quotations are taken from the [Genderize.io website](http://genderize.io), with correspondence from the developer behind the site for confirmation.] It would be anachronistic to apply these datasets to the past, and Genderize.io provides no functionality to filter results chronologically as it does geographically.
-In addition, exactly what the dataset comprises and how it was gathered is unclear, which keeps scholars from interrogating the value of the source.
+In addition, Genderize.io does not make clear exactly what comprises the dataset and how it was gathered, which keeps scholars from interrogating the value of the source.
 
-R and Python are the most commonly used languages for data analysis among digital humanists.
-Python's existing packages for gender prediction all implement a method similar to the NLTK or Genderize.io.^[Python library names] To our knowledge, [CRAN](http://cran.rstudio.com/) (the central repository for R packages) does not include any packages dedicated to gender prediction.
-Though options abound for geocoding, for example, there is no satisfactory existing method of gender prediction for historical and humanistic research.
+R and Python are the most commonly used languages for data analysis among digital humanists.^[**I completely agree, but do we have any evidence for this?
+Surveys or anything?**] Python's existing packages for gender prediction all implement a method similar to the NLTK or Genderize.io.^[Python library names] To our knowledge, [CRAN](http://cran.rstudio.com/) (the central repository for R packages) does not include any packages dedicated to gender prediction.
+In two of the most popular data analysis languages in the digital humanities there is currently no satisfactory existing method of gender prediction for historical and humanistic research.
 
-To that end we have created the [Gender](https://github.com/ropensci/gender) package for R which contains the predictive algorithm, and an associated [genderdata](https://github.com/lmullen/gender-data-pkg) package which contains various historical datasets.
+To that end we have created the [Gender](https://github.com/ropensci/gender) package for R which includes both the predictive algorithm and an associated [genderdata](https://github.com/lmullen/gender-data-pkg) package containing various historical datasets.
 This R implementation is based on an [earlier Python implementation](https://github.com/cblevins/Gender-ID-By-Time) by [Cameron Blevins](http://www.cameronblevins.org/) and [Bridget Baird](http://www.conncoll.edu/directories/emeritus-faculty/bridget-baird/).
@@
 
-The possibilities for encoding gender from names depends on two things.
-First, it depends on having a suitable (and suitably large) dataset for the time period and region under study.
-Unsurprisingly such datasets are almost always gathered in the first instance by states, though the compilation and digitization may be undertaken by scholarly researchers.
-Second, it depends on a suitable algorithm for estimating the proportion of male and female names for a given year, or range of years since often a person cannot be associated with an exact date.
-It is especially important that the algorithm take into account any biases in the data to make correct predictions.
+The possibilities for inferring gender from names depends on two things.
+First, it needs a suitable (and suitably large) dataset for the time period and region under study.
+Unsurprisingly such datasets are almost always gathered in the first instance by government entities, though the compilation and digitization may be undertaken by scholarly researchers.
+Second, it depends on a suitable algorithm for estimating the proportion of male and female names for a given year or range of years.
+since often a person cannot be associated with an exact date.
+It is especially important that the algorithm take into account any biases in the data to formulate more accurate predictions.
 Development of the R package has had two primary aims.
@@
 
- The Gender package currently includes two datasets which make it suitable for use for United States history since the first federal census in 1790.^[The authors of the package are both American historians, which explains why the initial release of the package focuses on that country.
+The Gender package currently includes two datasets which make it suitable for studying the United States from the first federal census in 1790 onwards.^[The authors of the package are both American historians, which explains why the initial release of the package focuses on that country.
 We have plans, however, to extend the underlying data to six other European and North American countries included in the [North Atlantic Population Project](https://www.nappdata.org/napp/) for the nineteenth century.
 We also welcome contributions of datasets for other regions and times from other scholars.] The first dataset contains names of applicants for Social Security and available from Data.gov.
-The second dataset contains names gathered in the decennial censuses and available from the IPUMS-USA (Integrated Public Use Microdata Series) service from the Minnesota Population Center at the University of Minnesota.
+The second dataset contains names gathered in the decennial censuses and available from the IPUMS-USA (Integrated Public Use Microdata Series) service from the Minnesota Population Center at the University of Minnesota.^[**citation?**]
 
@@
 Social Security Administration, "[Baby Names from Social Security Card Applications-Data by State and District of Columbia](http://catalog.data.gov/dataset/baby-names-from-social-security-card-applications-data-by-state-and-district-of-)," Data.gov dataset ID US-GOV-SSA-339 (accessed August 31, 2014).] This dataset contains a list of first names with the number of male and female uses of the name in every year beginning with 1880 and ending with 2012.
-The SSA has included any name which was used more than five times in a given year; otherwise the dataset is a 100 percent sample.
+The SSA list includes any name which was used more than five times in a given year, thereby capturing all but the most infrequently used names from each year.
 The description "baby names" provided by the SSA is a serious misnomer.
@@
 The dataset goes back to 1880, the birth year for a 55 year-old adult when Social Security was enacted.
-Even after 1935, registration at birth for Social Security was not mandatory until 1986.
+**Okay, so that means we need to tweak some stuff where we use the word "baby" earlier, yes?** Even after 1935, registration at birth for Social Security was not mandatory until 1986.
 As we will demonstrate below, the way in which the data was gathered requires an adjustment to our predictions of gender.^[CITE history of SSA.
 On the history of birth registration see Shane Landrum, "The State's Big Family Bible: Birth Certificates, Personal Identity, and Citizenship in the United States, 1840--1950" (PhD dissertation, Brandeis University, 2014).
-It is well-known among American historians that Social Security was written to exclude farmworkers with the aim of excluding African Americans.
+It is well-known among American historians that Social Security was written to exclude farmworkers as a means of also excluding African Americans.
 In examining the dataset we have not found any reason to believe that this policy noticably skewed the prediction of gender from names.]
@@
 Lunbeck’s report drew in part on a survey of 362 female historians that produced a rich source of responses detailing the deep and multi-faceted challenges facing women in the profession.
-What follows is a quantitative supplement to Lunbeck’s analysis that uses our program to analyze gender representation amongst historians across a much larger scale and a much longer time period.
-http://www.historians.org/Documents/About%20AHA%20and%20Membership/CWH-Report_5.20.05.pdf
+What follows is a quantitative supplement to Lunbeck’s analysis that uses our program to analyze gender representation amongst historians across a much larger scale and a much longer time period.^[http://www.historians.org/Documents/About%20AHA%20and%20Membership/CWH-Report_5.20.05.pdf]
 

### 2014-09-09 article.Rmd
Finished description of data

@@
 
-TODO IPUMS
+The IPUMS-USA dataset, contributed by Benjamin Schmidt, contains records from the United States decennial census from 1790 to 1930.
+This dataset contains the birth year and numbers of males and females under the age of 62 for all the years in that range.
+This data has been aggregated by IPUMS at the University of Minnesota and is released as a sample of the total census data.
+Unlike the SSA dataset, which includes 100% of every name reported to the Social Security Administration, the IPUMS data contains 5% or 10% samples.
+Because the `gender()` function relies on proportions of uses of names, rather than raw counts of people with the names, the sampling does not diminish the reliability of the function's predicitons.^[Steven Ruggles, J.
+Trent Alexander, Katie Genadek, Ronald Goeken, Matthew B.
+Schroeder, and Matthew Sobek.
+Integrated Public Use Microdata Series: Version 5.0 [Machine-readable database].
+Minneapolis, MN: Minnesota Population Center [producer and distributor], 2010.]
+
+These two datasets (and any future datasets to be added to the package) thus contain a simple table of data. The columns
+
+^[The format of this data departs from the more generally desirable "tidy data" framework proposed by Hadley Wickham for performance reasons.
+By keeping the data in a "wider" format instead of a tidy format as defined by Wickham, the function is able to make its predictions faster.
+Hadley Wickham, "[Tidy Data](http://vita.had.co.nz/papers/tidy-data.html)," *Journal of Statistical Software* (forthcoming).] The columns contain the `name`, `year`, and number of `female` and `male` uses of a name.
+Each row contains a combination for a name and a year.
+
+```{r ssa_national, echo=TRUE, results='markup'}
+genderdata::ssa_national
+```
+
+```{r ipums_usa, echo=TRUE, results='markup'}
+genderdata::ipums_usa
+```
 

### 2014-09-09 article.Rmd
Delete old explanation; intro to new

@@
 
-Our program is best understood through an example.
-Let’s say a historian is working with records from a public school system and wants to know what percentage of its total student body each year was male or female.
-This historian can tell the program the approximate age range of its target population - in this case, ages 4 to 18.
-Our program then iterates through all of the names attempting to identify their gender based on a) the date of the data, and b) the age range specified by the user.
-Sidney Dubner, for instance, was enrolled as a student in the year 1918.
-Given the age range of 4 to 18 specified by the user, the program guesses that Sidney Dubner was born between 1900 and 1914.
-It then looks up every birth of a child named “Sidney” recorded by the Social Security Administration between 1900 and 1914 and calculates that 94.7% of these babies named “Sidney” were male.
-Chances are good that this was a male student.
-But what if the year was 1998 instead of 1918?
-The chances that “Sidney” was a boy falls to 66.4%.
-Fast-forward another ten years to 2008, and suddenly it’s more likely that “Sidney” is in fact a girl, not a boy (69.3% female to 30.7% male).
-
-Reference Implementations/Software:
-
-Description of Lincoln’s R package [link].
-
-Implements both the Kantrowitz method and this new method. Extensible through the addition of additional data sets as they become available.
-Includes cleaned up, tidied (in the Hadley Wickham sense) versions the Kantrowitz data sets and the SSA national and state data.
-(Link to gender repo.) These allow researchers to load the package and use the data directly, in addition to accessing through methods.
-Furthermore
-Allows for encoding of individual names by year, of data sets where each name is associated with a year, and data sets where there are names of years and a range of years is provided.
-Eventually will include ability to specify by
-Samples of code:
-
-find_gender(“madison”) # returns “female” and probability
-find_gender(“madison”, 1990) # returns “female” and probability
-find_gender(“madison”, 1890) # returns “male” and probability
-
-year  name     
-------  ---------
- 1890  john     
- 1990  john     
- 1990  john     
- 1890  jane     
- 1990  jane     
- 1890  madison  
- 2012  madison
-For data sets where there is no year available, you can specify the range.
-find_gender(test_data, years = c(1880, 1940)) # returns male, male, male, male, female, female
-
-For data sets where there is a year available, say a year of birth, then you can specify the column that stores that data, and each person will be looked up individually.
-find_gender(test_data, year_column = T) # returns male, female, male, male, female, female
-
-In cases where the data set includes non-birth date, and because those dates vary significantly, one wishes to operate on the basis of date, one is advised to split up the data into cohorts, apply the gender checking methods, then recombine the split up data sets.
-(This is akin to Wickhams Split Apply Combine method: citation to article.) In this example of the dissertations data sets, I have split the data set into people whose dissertations were before 1950 and those who were after.
-(In actual practice, I would use more fine grained cohorts, and a more sophisticated method for splitting and applying, such as the `ddply` command in `plyr`.)
-
-library(dplyr)
-dissertations_before_1950 <- dissertations %.%
- filter(year < 1950)
-dissertations_after_1950 <- dissertations %.%
- filter(year >= 1950)
-dissertations_before_1950 <- dissertations_before_1950 %.%
- find_gender(years = c(1880,1940)) # reasonable guess as to ages
-dissertations_after_1950 <- dissertations_after_1950 %.%
- find_gender(years = c(1930,1990)) # reasonable guess as to ages
-dissertations_combined < rbind(dissertations_before_1950, dissertations_after_1950)
+Our method for predicting gender is best understood through a series of examples.
+First we will demonstrate a simplified version of the inner workings of the function for predicting a single name.
+This example will demonstrate fundamentally how the function works.
+Next we will demonstrate how to use the function provided in the gender package to solve a simple historical problem.
 

### 2014-09-09 article.Rmd
Explanation of inner workings

@@
 This example will demonstrate fundamentally how the function works.
-Next we will demonstrate how to use the function provided in the gender package to solve a simple historical problem.
+Next we will demonstrate how to use the function provided in the gender package to predict gender from a small sample dataset.
+
+The method for predicting gender from a name using the provided datasets is fundamentally simple.
+Let's begin by assuming that using the Social Security dataset we want to predict the gender of someone named Sidney who was born in 1935.
+Because the dataset contains a list of names for each year, we can simply lookup the row for Sidney in 1935.
+Using the [dplyr](https://github.com/hadley/dplyr) package for R, this can be expressed in a simple grammar of filtering by name and year.
+
+```{r, echo=TRUE, results='markup'}
+genderdata::ssa_national %>%
+  filter(name == "sidney", year == 1935)
+```
+
+Thus according to the Social Security Administration, there were 974 boys and 93 females named Sidney born in 1935.
+We can add another command to calculate the proportion of females and males ("mutate" in dplyr's parlance).
+
+```{r, echo=TRUE, results='markup'}
+genderdata::ssa_national %>%
+  filter(name == "sidney", year == 1935) %>%
+  mutate(proportion_female = female / (male + female),
+         proportion_male   = 1 - proportion_female)
+```
+
+In other words, there is an approximately `r gender("sidney", year = 1935)$proportion_male %>% percent()` percent chance that a person born in 1935 named Sidney was male.
+(In 2012, for comparison, there was an approximately `r gender("sidney", year = 2012)$proportion_female %>% percent()` percent chance that a person born named Sidney was female.)
+
+The method is only slightly more complex if we do not know the exact year when someone was born, as is often the case for historical data.
+Suppose we know that Sidney was born in the 1930s but cannot identify the exact year of his or her birth.
+Using the same method as above we can look up the name for all of those years.
+
+```{r, echo=TRUE, results='markup'}
+genderdata::ssa_national %>%
+  filter(name == "sidney", year >= 1930 & year <= 1939) 
+```
+
+Next we can sum up the male and female columns ("summarize" in dplyr's vocabulary) and calculate the proportions of female and male uses.
+
+```{r, echo=TRUE, results='markup'}
+genderdata::ssa_national %>%
+  filter(name == "sidney", year >= 1930 & year <= 1939) %>%
+  group_by(name) %>%
+  summarize(female = sum(female),
+            male = sum(male)) %>%
+  mutate(proportion_female = female / (male + female),
+         proportion_male   = 1 - proportion_female)
+```
+
+In other words, for the decade of the 1930s, we can calculate that there is a `r gender("sidney", year = c(1930, 1939))$proportion_male %>% percent()` percent chance tha person named Sidney was male.
+This is roughly the same as the probability that the name was male in just 1935, but by looking at the underlying counts on which it is based, we can see this proportion was calculated from ten years instead of one.
 

### 2014-09-09 article.Rmd
Explain skewed gender ratios

@@
 
+As previously mentioned, the history of how the Social Security Administration collected the data effects its validity.
+Specifically, because the data extends back to 1880 but the first applications were gathered after 1935, the sex ratios in the dataset are skewed, as demonstrated in this plot.
+
+```{r}
+genderdata::ssa_national %>%
+  group_by(year) %>%
+  summarize(ratio = sum(female) /sum(female + male)) %>%
+  ggplot(aes(x = year, y = ratio)) +
+  geom_line() +
+  ggtitle("Percentage of SSA dataset that is female") +
+  xlab(NULL) + ylab(NULL)
+```
+
+It is extremely improbable that nearly 70 percent of the people born in 1900 were female.
+Exactly why the dataset has this bias is unclear, though we speculate that it is because the applicants for Social Security in the early years of the program were approaching retirement age, which was set much closer to the average life expectancy in 1935 than it is today.
+Since women tend to live longer than men, they were overrepresented in Social Security applications.
+
+The solution to this problem is two-fold.
+First, we recommend that researchers use the IPUMS-USA dataset to make predictions for years from 1790 to 1930 (which avoids the "bubble" in the SSA data) and that they use the SSA dataset for years since 1930.^[The IPUMS-USA dataset includes only apparently random fluctuations in the gender ratio in any given year.] Second, we have built in a correction to the SSA dataset when using the `gender()` function.
+If we assume that the secondary sex ratio (that is, the ratio of male to female births) in any given year does not deviate from `0.5` (that is, equality), it is possible to calculate a correction factor for each year or range of years to even out the dataset.
+We apply this correction factor automatically when using the SSA dataset.
+
 Limitations:

### 2014-09-10 article.Rmd
Finished first cut at methodological section

@@
 Thus according to the Social Security Administration, there were 974 boys and 93 females named Sidney born in 1935.
-We can add another command to calculate the proportion of females and males ("mutate" in dplyr's parlance).
+We can add another command to calculate the proportion of females and males ("mutate" in dplyr's vocabulary).
 
@@
 
-Limitations:
+These internal workings are wrapped by the prediction function so that researchers can focus on the choice of datasets and the range of years for making their predictions.
+Let's use the following sample dataset, say, a membership role, as an example for how to use the function.^[See the package vignette for more details: `vignette("predicting-gender", package = "gender")`.]
 
-On the methodological front, perhaps the program’s most glaring limitation is its weakness in inferring the gender of non-American names.
-Once again, this is a reflection of where the source data originated: U.S.
-government agencies.
-More recent datasets exist for non-American names, but we have been unable to find or incorporate information on how these names have changed over time.
-The program is extensible, however, meaning that if someone does locate a suitably temporal dataset of non-American names they can implement it into our program.
+```{r}
+membership <- data.frame(name = c("Madison", "Madison", "Morgan",
+                                  "Morgan", "Jan", "Jan"),
+                         birth_year = c(1934, 1990, 1948, 1970, 1965, 1998),
+                         stringsAsFactors = FALSE) %>% tbl_df()
+```
 
-Our program uses source data that only goes back as far as 1880, making it a far more effective method for twentieth-century analysis.
-Additionally, our program is not as effective at identifying highly uncommon names.
-The annual reference data it uses from the Social Security Administration only includes baby names that occurred at least five times in a given year.
-If a name was never given to five or more babies over the course of any year from 1880 to the present, our program will not be able to infer its gender.
-[more details – see below?]
-[- SSA restricts its data to only names with at least 5 occurrences (we’re still missing a bit of the long tail).
-Not a big problem because we still have more names, and more obscure names, than any other data set.
-More of a problem when localizing to the state, which starts in 1910 (double check) and because it includes only >5 occurrences in much smaller jurisdictions, more of the long tail is cut off.]
+```{r, echo=TRUE, results='markup'}
+membership
+```
+
+The simplest way to use the package would be to call the `gender()` function, specifying the method, in this case the Social Security Administration dataset since all the names are after 1930, and convert the results to a data frame (one of R's native data structures).
+
+```{r, echo=TRUE, results='markup'}
+gender(membership$name, method = "ssa") %>% do.call(rbind.data.frame, .)
+```
+
+This most basic call would be suitable if we did not know the year of birth. But we can also specify that in the `gender()` function by using this call:
+
+```{r, echo=TRUE, results='markup'}
+Map(gender, membership$name, years = membership$birth_year) %>% 
+  do.call(rbind.data.frame, .)
+```
+
+By taking into account the year of birth, we find that four of our six names were likely male, whereas we might otherwise have predicted that all six were female.
+We also now know the approximate likelyhood that our predictions are correct: at least 80% for all of thse predictions.
 

### 2014-09-10 article.Rmd
Removed outdated stuff

@@
 
-In a 2005 report for the American Historical Association, Elizabeth Lunbeck acknowledged “a sea change in the [historical] profession with respect to gender” before going on to describe the painful limitations of this sea change for female historians: ongoing personal discrimination, lower salaries, and barriers to securing high-ranking positions.
+In a 2005 report for the American Historical Association, Elizabeth Lunbeck acknowledged "a sea change in the [historical] profession with respect to gender" before going on to describe the painful limitations of this sea change for female historians: ongoing personal discrimination, lower salaries, and barriers to securing high-ranking positions.
 Lunbeck’s report drew in part on a survey of 362 female historians that produced a rich source of responses detailing the deep and multi-faceted challenges facing women in the profession.
-What follows is a quantitative supplement to Lunbeck’s analysis that uses our program to analyze gender representation amongst historians across a much larger scale and a much longer time period.^[http://www.historians.org/Documents/About%20AHA%20and%20Membership/CWH-Report_5.20.05.pdf]
+What follows is a quantitative supplement to Lunbeck's analysis that uses our program to analyze gender representation amongst historians across a much larger scale and a much longer time period.^[http://www.historians.org/Documents/About%20AHA%20and%20Membership/CWH-Report_5.20.05.pdf]
 
 Our analysis focuses on one of the bedrocks of the historical profession: scholarly research.
-We begin with the history dissertation, often the defining scholarly output of a historian’s early career.
-The completion of a dissertation marks a standardized moment of transition out of the “training” phase of the historical profession.
+We begin with the history dissertation, often the defining scholarly output of a historian's early career.
+The completion of a dissertation marks a standardized moment of transition out of the "training" phase of the historical profession.
 Data supplied by ProQuest provided roughly [80,000] PhD(?) dissertations completed since [1950].
@@
 (Take that, sociologists.)
-
-Future work:
-
-Both the Python script and the R package are easily extensible to allow additional data sets, say from other countries.
-(Are there other data sets, say for Great Britain, http://data.gov.uk/data/search?q=names ??) And eventually perhaps US census data, which would solve the 19th century problem.

### 2014-09-11 R/load-dissertations.R
Import load script for dissertations

@@ -1,0 +1,334 @@
+# load2.r 
+# This script loads the data and cleans it. It is a copy of `load.r`, which was
+# used for posts 1-5, but modified to clean the data more rigorously. It is also
+# better optimized. I have two load scripts so that results from the earlier
+# posts are preserved unmodified.
+
+library(dplyr)
+library(stringr)
+
+# Read the historical history dissertations data
+raw <- read.csv("data/history-df.csv", stringsAsFactors=F)
+
+# Turn import into a tbl_df to make it easier to work with dplyr
+raw <- tbl_df(raw)
+
+# Turn empty strings into NAs
+raw[raw == ""] <- NA
+
+# Figure out the universities outside of the U.S. and Canada to drop
+# universities <- raw %.%
+#   group_by(university) %.%
+#   summarise(count = length(id)) %.%
+#   arrange(university)
+# 
+# write.csv(universities, "temp/universities.csv")
+
+# Universities to delete
+universities_to_delete <- c(
+  "Adelphi University, School of Social Work.",
+  "Adelphi University, The Institute of Advanced Psychological Studies.",
+  "Adler School of Professional Psychology.",
+  "Ahmadu Bello University (Nigeria).",
+  "Allegheny University of Health Sciences.",
+  "American Conservatory of Music.",
+  "American School of Professional Psychology - Rosebridge.",
+  "American Conservatory of Music.",
+  "American School of Professional Psychology - Rosebridge.",
+  "Bar-Ilan University (Israel).",
+  "Bogazici Universitesi (Turkey).",
+  "Boston University School of Education.",
+  "Brandeis University, The Heller School for Social Policy and Management.",
+  "Brandenburgische Technische Universitaet Cottbus (Germany).",
+  "Bryn Mawr College, Graduate School of Social Work and Social Research.",
+  "Cairo University (Egypt).",
+  "Caribbean Graduate School of Theology (Jamaica).",
+  "Case Western Reserve University (Health Sciences).",
+  "Centro de Estudios Avanzados de Puerto Rico y el Caribe (Puerto Rico).",
+  "Christian-Albrechts Universitaet zu Kiel (Germany).",
+  "Chuo Daigaku (Japan).",
+  "Columbia University Teachers College under the Faculty of Union Theological Seminary in the City of New York.",
+  "Duquesne University School of Nursing.",
+  "Eberhard Karls Universitaet Tuebingen (Germany).",
[…]

### 2014-09-12 article.Rmd
Encode gender for AHR

@@
 library(knitr)
+library(stringr)
 library(scales)
@@
 # TRUE to show code and warnings; FALSE to remove them for publication
-display_code_blocks(FALSE)
+display_code_blocks(TRUE)
 
@@
 In a 2005 report for the American Historical Association, Elizabeth Lunbeck acknowledged "a sea change in the [historical] profession with respect to gender" before going on to describe the painful limitations of this sea change for female historians: ongoing personal discrimination, lower salaries, and barriers to securing high-ranking positions.
-Lunbeck’s report drew in part on a survey of 362 female historians that produced a rich source of responses detailing the deep and multi-faceted challenges facing women in the profession.
+Lunbeck's report drew in part on a survey of 362 female historians that produced a rich source of responses detailing the deep and multi-faceted challenges facing women in the profession.
 What follows is a quantitative supplement to Lunbeck's analysis that uses our program to analyze gender representation amongst historians across a much larger scale and a much longer time period.^[http://www.historians.org/Documents/About%20AHA%20and%20Membership/CWH-Report_5.20.05.pdf]
@@
 It is, after all, only the first major stage of a historian’s research output.
-The “coin of the realm” for many historians remains the publication of a scholarly monograph to be read and evaluated by peers.
+The "coin of the realm" for many historians remains the publication of a scholarly monograph to be read and evaluated by peers.
 This process often takes the form of book reviews published in academic journals.
-One of the leading journals in the historical profession is the American Historical Review, the flagship journal of the American Historical Association.
-Along with a few dozen articles The AHR publishes five issues each year that contain, in aggregate, roughly one thousand books reviews covering (in its own words) “every major field of historical study.” The AHR is not only one of the widest-ranging journals in the profession, it is also the oldest; the journal has been publishing continuously since 1895.
+One of the leading journals in the historical profession is the *American Historical Review*, the flagship journal of the American Historical Association.
+Along with a few dozen articles, the AHR publishes five issues each year that contain, in aggregate, roughly one thousand books reviews covering (in its own words) "every major field of historical study." The AHR is not only one of the widest-ranging journals in the profession, it is also the oldest; the journal has been publishing continuously since 1895.
 The range, scope, and length of the AHR makes it an ideal source to analyze the historical profession on a large scale.
@@
 We take the AHR as a rough proxy for gatekeeping within the larger historical profession.
-When a book appears in the AHR it serves as a signal that other historians in positions of power have “approved” it for consideration.
+When a book appears in the AHR it serves as a signal that other historians in positions of power have "approved" it for consideration.
 This signal is not necessarily correlated with the quality of the book.
-Even if it garners a negative review, the fact that it appears in the journal at all is a measure of the fact that the profession’s gatekeepers have deemed it important enough to review.
+Even if it garners a negative review, the fact that it appears in the journal at all is a measure of the fact that the profession's gatekeepers have deemed it important enough to review.
 On the flip side, a book that does not appear in the AHR may have gone on to have a significant impact on the profession.
@@
 
-[graphic]
+```{r gender-book-authors}
+ahr <- read.csv("data/ahr.utf8.csv", stringsAsFactors = FALSE) %>%
+  tbl_df()
+
+# For a number of entries the first name column contains first and middle names,
+# so we're going to parse out just the first names.
+ahr <- ahr %>%
+  mutate(first_name = word(Author_First))
+
+# We need to guess at the range of ages for book authors. 25 to 75 seems
+# sensible enough.
+min_age <- 25
+max_age <- 75
+
+# Now we will create columns for the minimum year and the maximum year for
+# guessing gender.
+ahr <- ahr %>%
+  mutate(min_year = issueDate - max_age,
+         max_year = issueDate - min_age)
+
+# Now we need to decide which method to use when predicting the gender. In 
+# general we want to use IPUMS before 1930 and SSA after 1930, but sometimes 
+# we'll have to compromise. There are several kinds of obvious cases:
+# 
+# - If max_age < 1930, use IPUMS
+# 
+# - If min_age > 1930, use SSA
+# 
+# The tricky cases are when the terminal values overlap the ranges of one of the
+# methods.
+# 
+# - If min_age > 1880 and max_age > 1930, use SSA
+# 
+# If our range of ages was greater than 50, we might have a person who we 
+# guessed was born before 1880 or after 1930. But by keeping our range smaller
+# than 50 (the difference between 1930 and 1880) we can avoid that problem.
+
+# That does indeed match all cases. So now we can write a function to return the
+# proper method for a given range of years.
+pick_method <- function(min_age, max_age) {
+  ifelse(max_age <= 1930, "ipums",
+         ifelse(min_age > 1930, "ssa",
+         ifelse(min_age > 1880 & max_age > 1930, "ssa", NA)))
+}
+
+# Now assign a method to the AHR dataset
+ahr <- ahr %>%
+  mutate(method = pick_method(min_year, max_year))
+
+pass_through <- function(x) { return(x) }
+
+apply_gender <- function(df) {
+  gender(df$first_name, method = df$method, 
+         years = c(df$min_year, df$max_year))
+  }
+
+ahr_gender <- ahr %>% 
+  group_by(first_name, vol_issue_ID, Author_Full, Reviewer_First,
+           Reviewer_Full, category) %>%
+  do(gender_author = apply_gender(.))
+
+# Finally we are ready to predict the gender. Because we have to combine years
+# we're going to do this as a for loop, which is ugly.
+# ahr_gender.l <- vector(mode = "list", nrow(ahr))
+
+# for(i in 1:nrow(ahr)) {
+#   row <- ahr[i, ]
+#   result <- gender(row$first_name, method = row$method,
+#                    years = c(row$min_year, row$max_year))
+#   ahr_gender.l[[i]] <- result
+# }
+
+# ahr_gender.df <- ahr_gender.l %>% do.call(rbind.data.frame, .)
+```
 

### 2014-09-12 article.Rmd
Revised conclusion

@@
 Not all dissertations are published as books, not all books are submitted to the American Historical Review, and not all book submissions are accepted by the AHR editors for review.
-As of August 2014 the journal’s website reminds readers that “the AHR receives over 3,000 books a year; we have the resources to publish at most 1,000 reviews a year (approximately 200 per issue).” (http://www.indiana.edu/~ahrweb/book_review_guide.html - retrieved August 25 2014).
+As of August 2014 the journal’s website reminds readers that "the AHR receives over 3,000 books a year; we have the resources to publish at most 1,000 reviews a year (approximately 200 per issue)."^[http://www.indiana.edu/~ahrweb/book_review_guide.html - retrieved August 25 2014].
 Fewer and fewer female historians make it through each successive step.
@@
 In both African history and Ancient/Medieval history female historians made some quite dramatic gains during the late-nineties and aughts.
-The guiltiest parties, however, are also the two subject categories that publish the most book reviews in the AHR: Modern/Early Modern Europe and the United States/Canada.
+The most uneven parties, however, are also the two subject categories that publish the most book reviews in the AHR: Modern/Early Modern Europe and the United States/Canada.
 Both sub-fields have made progress over the four decades but still hover at around two-thirds male.
@@
 
-One 2013 analysis of 2,500 recent history Ph.D.’s found that “gender played little role in employment patterns across particular professions and industries.” (http://historians.org/Documents/Many_Careers_of_History_PhDs_Final.pdf) Our own analysis reveals a more complicated reality.
+One 2013 analysis of 2,500 recent history Ph.D.'s found that "gender played little role in employment patterns across particular professions and industries."^[http://historians.org/Documents/Many_Careers_of_History_PhDs_Final.pdf] Our own analysis reveals a more complicated reality.
 The flagship journal of professional historians continues to publish twice as many reviews of books by female authors as male authors.
@@
 Women might be getting hired at comparable rates to men, but discriminatory gatekeeping remains alive and well in the historical profession.
-Our program allows us to quantitatively measure the scope of the problem on a large scale, how it’s changed over time, and what remains to be done.
+Our program allows us to quantitatively measure the scope of the problem on a large scale, how it's changed over time, and what remains to be done.
 
@@
 
-Our contributions:
-
--     Broadly: gender studies + big data
--   	Temporal dimension makes it better suited to fit the needs of humanities research than contemporary approaches – both more accurate (historical) and more nuanced (supplying the gendered proportion of a name for different periods)
--   	Extensible – non US-names, older names. (future work)
-
-The fundamental points are that we have a much larger corpus of names.
-More importantly the ability to vary the look up based on a specific date of birth or a specified range of dates.
-Making this much more useful for historians, but also for any research in any field, given that the normal human lifespan present in any data set means that names can vary as much as a century, and so everyone needs to be using a fundamentally historical method.
-(Take that, sociologists.)
+In conclusion, our algorith for predicting gender from first names combined with historical datasets in the Gender package provides a powerful new method for historians interested in the study of gender.
+The digital turn has made available very large datasets, yet in the midst of abundance historians always have to deal with the problem of scarcity in the sources.
+Our methods uses the abundant datasets to produce more information from scarce datasets.
+The temporal dimension of our work matches a historical question to a historical method, providing much improved over simple, anachronistic lookup methods.
+This method is extensible to any time or place for which there is a suitable dataset of names from which to make the predictions.
+As we have demonstrated in our study of gatekeeping in the American historical profession, our method can reveal important details about gender even from datasets which do not otherwise include that information, and it can do so over long historical periods.

### 2014-09-12 article.Rmd
Placeholder images

@@
 # TRUE to show code and warnings; FALSE to remove them for publication
-display_code_blocks(TRUE)
+display_code_blocks(FALSE)
 
@@
 Whereas before it had been an infrequently used male name, after 1985 it skyrocketed in popularity, and by the early 2000s it was one of the ten most frequent names for baby girls.
-Broadly speaking, since the late twentieth century Americans girls are more likely than American boys to receive a "trendy" name, and once that name is associated with females it is unlikely to be given to male babies.
-The few exceptions of names which go from being female to male (such as "Jan") have a different causal mechanism.
+Broadly speaking, since the late twentieth century Americans girls are more likely than American boys to receive a "trendy" name, and once that name is associated with females it is unlikely to be given to male babies.^[The few exceptions of names which go from being female to male (such as "Jan") have a different causal mechanism.
 It seems likely that Jan as an English name is female but that Jan as a Scandinavian name is male.
-Because the name has remained relatively uncommon, other causes besides changing gender preferences can explain the change.
-**this last bit about "Jan" might just muddle things - unless we have a clear explanation or hypothesis as to what does explain the change, not sure it's worth highlighting**
+Because the name has remained relatively uncommon, other causes besides changing gender preferences can explain the change.]
 
@@
 
-[chart – raw numbers]
+![Number of dissertations by gender](img/plot-gender.png)
 
@@
 
-[chart - proportion]
+![Proportion of dissertations for each gender](img/dissertations-by-gender-proportion.png)
 
@@
 
-```{r gender-book-authors}
+![Gender of authors of books reviewed in the AHR](img/book-authors.jpg)
+
+```{r gender-book-authors, eval=FALSE}
 ahr <- read.csv("data/ahr.utf8.csv", stringsAsFactors = FALSE) %>%
@@
 In book reviews published in the AHR male authors continue to outnumber female authors by a factor of nearly 2 to 1.
-The long-term gains made by female doctoral students do not carry over in the pages of the historical profession’s leading journal.
+The long-term gains made by female doctoral students do not carry over in the pages of the historical profession's leading journal.
 
-[chart comparing dissertations with AHR]
+![Gap between female authorship of dissertations and books](img/book-dissertation-gap.jpg)
 
@@
 
-[chart: male authored books]
-[chart: female authored books]
+![Gender of reviewers of books authored by women](img/female_authors_withreviewers.jpg)
 
@@
 
-[chart – sub-fields]
+![Gender of reviewers by category of review](img/categories_gender_bytime.jpg)
 

### 2014-09-12 article.Rmd
Implemented most of Cameron's suggestions

@@
 That is because the gendered naming convention surrounding "Leslie" completely changed over the course of one hundred years.
-92% of the babies born in the United States in 1900 who were named “Leslie” were male, while 96% of those identically named babies born in the year 2000 were female.
+92% of the babies born in the United States in 1900 who were named "Leslie" were male, while 96% of those identically named babies born in the year 2000 were female.
 
@@
 
-Many other American names, such as Madison, Morgan, Sydney, and Kendall, have also changed gender over the past century.
+Many other names as used in the United States, such as Madison, Morgan, Sydney, and Kendall, have also changed gender over the past century.
 Such changes usually follow set patterns.
@@
 For example, the name "Madison" changed from male to female suddenly after 1985.
-Whereas before it had been an infrequently used male name, after 1985 it skyrocketed in popularity, and by the early 2000s it was one of the ten most frequent names for baby girls.
+Whereas before it had been an infrequently used male name, after 1985 it skyrocketed in popularity, and by the early 2000s it was one of the ten most frequently used names for baby girls.
 Broadly speaking, since the late twentieth century Americans girls are more likely than American boys to receive a "trendy" name, and once that name is associated with females it is unlikely to be given to male babies.^[The few exceptions of names which go from being female to male (such as "Jan") have a different causal mechanism.
@@
 The "Leslie" problem is not just for researchers who think of themselves as historians.
-Anyone studying a timeframe longer than a few years, or anyone studying a group whose characteristics do not match the probably unknowable characteristics of the groups that comprise social media datasets will also encounter this problem.
-As of 2012, the average American has a lifespan of nearly seventy-nine years–--more than enough time for naming practices to change quite dramatically.^[CITATION] Predicting gender from first names therefore requires a fundamentally historical method.
+Anyone studying a period longer than a few years, or anyone studying a group whose characteristics do not match the probably unknowable characteristics of the groups that comprise social media datasets will also encounter this problem.
+As of 2012, the average American has a lifespan of nearly seventy-nine years–--more than enough time for naming practices to change quite dramatically.^[World Bank, *World Bank Open Data*, "Life expectancy at birth, total (years)," <http://data.worldbank.org/indicator/SP.DYN.LE00.IN> (accessed September 12, 2014).] Predicting gender from first names therefore requires a fundamentally historical method.
 
-Our solution to the "Leslie" problem is to create a software package that combines a predictive algorithm with a range of historical dataset suitable to various times and regions.
+Our solution to the "Leslie" problem is to create a software package that combines a predictive algorithm with a several historical datasets suitable to various times and regions.
 The algorithm calculates the proportion of male or female uses of a name in a given birth year or range of years.
 It thus can provide not only a prediction of the gender of a name, but also an estimate of how likely that prediction is.
-For example 52 percent of the babies named Leslie born in the United States in 1950 were girls and 48% were boys.
+For example 52 percent of the babies named Leslie born in the United States in 1950 were girls and 48 percent were boys.
 This allows researchers to conduct more nuanced analysis and to determine for themselves what proportion is acceptable for making a prediction of gender.
@@
 In some sense this is unavoidable since the large historical datasets which make this kind of analysis possible were created by state agencies.
-For the category of gender (as for race) states have offered a limited number of boxes to check, and those boxes extert a tremendous power to define state subjects.^[CITE One of the best-known theorists on gender is Judith Butler, *Gender Trouble* and Butler, *Bodies that Matter*; For the problems with the word "identity" in the humanities and social sciences, see Rogers Brubaker and Frederick Cooper, "Beyond 'Identity'"; Several historians have written about the American state and its drive to impose identities onto subjects: Margot Canady, The Straight State;  Peggy Pascoe, *What Comes Naturally*; Shane Landrum on birth certificates.] Nevertheless, this method is suitable for two reasons.
-First, researchers are likely to use this method in the aggregate rather than for individuals.
-The assumption that a name implies male or female (and that those are possible categories to be inferred) is more valid when applied to large groups of people than it does when applied to individuals.
-**I think this first line of reasoning is dangerous.
-This leaves us open to critiques of "big data" in general from humanists and theorists: that the process of aggregation is actually really dangerous and that assuming that individual identity and experience can be subsumed under the generalities of the crowd, etc.
-etc.
-I think the second reason is strong enough on its own.** Second, this method is neutral as to what the categories male or female actually mean in any given historical context.
+For the category of gender (as for race) states have offered a limited number of boxes to check, and those boxes extert a tremendous power to define state subjects.^[One of the best-known theorists on gender is Judith Butler, *Gender Trouble* and Butler, *Bodies that Matter*; For the problems with the word "identity" in the humanities and social sciences, see Rogers Brubaker and Frederick Cooper, "Beyond 'Identity'"; Several historians have written about the American state and its drive to impose identities onto subjects: Margot Canady, The Straight State;  Peggy Pascoe, *What Comes Naturally*; Shane Landrum on birth certificates.] Nevertheless, this method is neutral as to what the categories male or female actually mean in any given historical context.
 Researchers are free---indeed, obligated---to make that determination for themselves, and this package provides access to the underlying dataset so that researchers can interrogate every assumption that comes with them.
@@
 2.
-Chapter 6 of the NLTK book also describes a method of predicting gender from the last letter of the name, but this example is given for the sake of explaining supervised classification and not as an end in itself, so we do not have reason to address this a historical method.] The NLTK is an influential software package for scholarship because it provides an extraordinary range of tools for analyzing natural language.
-**I think we could develop this even more, if we have information about it.
-My impression is that it's the "industry standard" for a lot of folks, but it would be nice to give some evidence to then show why our intervention matters, and that the Kantrowitz method isn't just a straw man** Included in the NLTK are two lists of male and female names created by the computer scientist Mark Kantrowitz with assistance from Bill Ross.^[The [original lists of names](http://www.cs.cmu.edu/afs/cs/project/ai-repository/ai/areas/nlp/corpora/names/0.html) are available at the website of the computer science department at Carnegie Mellon University, and the NLTK implementation is described in its [documentation](http://www.nltk.org/howto/corpus.html).] These lists include `r kantrowitz_unique` unique names.
+Chapter 6 of the NLTK book also describes a method of predicting gender from the last letter of the name, but this example is given for the sake of explaining supervised classification and not as an end in itself, so we do not have reason to address this a historical method.] The NLTK is an influential software package for scholarship because it provides an extraordinary range of tools for analyzing natural language.^[The NLTK book (the most common way to cite the software) has been cited 812 times according to [Google Scholar](http://scholar.google.com/scholar?cites=3735380965977150091&as_sdt=5,47&sciodt=0,47&hl=en).
+Because in the humanities (unlike in other fields) the practice of citing even software which determines the outcome of an analysis is regrettably rare, even nonexistent, such metrics are difficult to determine.
+Yet searching of journals such as *Digital Humanities Quarterly* and the *Journal of Digital Humanities* reveals that natural language processing, for which the NLTK is the foremost implementation, is one of the key methodological categories of the digital humanities.] Included in the NLTK are two lists of male and female names created by the computer scientist Mark Kantrowitz with assistance from Bill Ross.^[The [original lists of names](http://www.cs.cmu.edu/afs/cs/project/ai-repository/ai/areas/nlp/corpora/names/0.html) are available at the website of the computer science department at Carnegie Mellon University, and the NLTK implementation is described in its [documentation](http://www.nltk.org/howto/corpus.html).] These lists include `r kantrowitz_unique` unique names.
 Using the Kantrowitz names corpus in the NLTK, one could look up a name like Jane or John and find out that it is male or female.
@@
 
-R and Python are the most commonly used languages for data analysis among digital humanists.^[**I completely agree, but do we have any evidence for this?
-Surveys or anything?**] Python's existing packages for gender prediction all implement a method similar to the NLTK or Genderize.io.^[Python library names] To our knowledge, [CRAN](http://cran.rstudio.com/) (the central repository for R packages) does not include any packages dedicated to gender prediction.
+R and Python are the most commonly used languages for data analysis among digital humanists.^[Surveys and analysis of useage point out the growth of R and the continuing popularity of Python among "data science" generally.
+See, for example, Robert A.
+Muenchen, "The Popularity of Data Analysis Software," *r4stats.com*, <http://r4stats.com/articles/popularity/> (accessed September 12, 2014); Gregory Piatetsky, "Top Languages for analytics, data mining, data science," *KDN Nuggets*, <http://www.kdnuggets.com/2013/08/languages-for-analytics-data-mining-data-science.html> (accessed September 12, 2014).] Python's existing packages for gender prediction all implement a method similar to the NLTK or Genderize.io.^[Python library names] To our knowledge, [CRAN](http://cran.rstudio.com/) (the central repository for R packages) does not include any packages dedicated to gender prediction.
 In two of the most popular data analysis languages in the digital humanities there is currently no satisfactory existing method of gender prediction for historical and humanistic research.
@@
 The Gender package is affiliated with [rOpenSci](http://ropensci.org/), an initiative that supports reproducible research and open data for scientists using R.
-The rOpenSci advisory board has provided code review and publicity for the project.
+The rOpenSci advisory board has provided code review and publicity for the project.^[The gender package depends especially on R Core Team, *R: A language and environment for statistical computing*.
+R Foundation for Statistical Computing, Vienna, Austria (2014).
+<http://www.R-project.org/>;  Hadley Wickham and Romain Francois, *dplyr: dplyr: a grammar of data manipulation*.
+R package version 0.2 (2014): <http://CRAN.R-project.org/package=dplyr>.
+The sources of the data are cited below.]
 
@@
 We also welcome contributions of datasets for other regions and times from other scholars.] The first dataset contains names of applicants for Social Security and available from Data.gov.
-The second dataset contains names gathered in the decennial censuses and available from the IPUMS-USA (Integrated Public Use Microdata Series) service from the Minnesota Population Center at the University of Minnesota.^[**citation?**]
+The second dataset contains names gathered in the decennial censuses and available from the IPUMS-USA (Integrated Public Use Microdata Series) service from the Minnesota Population Center at the University of Minnesota.
 
@@
 The dataset goes back to 1880, the birth year for a 55 year-old adult when Social Security was enacted.
-**Okay, so that means we need to tweak some stuff where we use the word "baby" earlier, yes?** Even after 1935, registration at birth for Social Security was not mandatory until 1986.
+Even after 1935, registration at birth for Social Security was not mandatory until 1986.
 As we will demonstrate below, the way in which the data was gathered requires an adjustment to our predictions of gender.^[CITE history of SSA.
@@
 The completion of a dissertation marks a standardized moment of transition out of the "training" phase of the historical profession.
-Data supplied by ProQuest provided roughly [80,000] PhD(?) dissertations completed since [1950].
-Identifying the gender of their authors gives a sense for the number of women and men who are completing PhD-level training in history each year.
-Our program uses what year a historian wrote their dissertation to estimate a period for when they might have been born: a historian who completed their dissertation in 1980 was likely born between ____ and _____.
+Data supplied by ProQuest provided roughly [80,000] PhD(?) dissertations completed since [1950].^[ProQuest's dataset does not permit us to limit dissertations just to dissertations produced in history departments.
+These dissertations include dissertations *about* history.
+For a justification of this approach, see a [series of blog posts](http://lincolnmullen.com/research/history-dissertations/) exploring the dataset.
+The American Historical Association's dataset of dissertations relies on reporting from departments and does not have the chronological range of the ProQuest data.
+Nevertheless, applying the same method to the AHA's dataset does not provide any reason for thinking that the proportion authorship by gender over time is substantially different when limited to just history department dissertations.
+American Historical Association, *Directory of History Dissertations*, https://secure.historians.org/pubs/dissertations/.
+Data export provided to author by Liz Townsend, March 3, 2014.] Identifying the gender of their authors gives a sense for the number of women and men who are completing PhD-level training in history each year.
+Our program uses the year a historian wrote his or her dissertation to estimate a period for when they might have been born: a historian who completed their dissertation in 1980 was likely born between ____ and _____.
 Using this temporal information, we are better able to infer their gender and chart how the larger representation of women and men changes year-by-year.
@@
 Another way to examine this trend is to look at the proportion of male and female dissertation authors over time, which smoothes out changes in the absolute number of dissertations produced.
-The proportion of dissertations written by women has steadily increased over the past half-century, a change that began in the late 1960s and continued through the early 2000’s.
+The proportion of dissertations written by women has steadily increased over the past half-century, a change that began in the late 1960s and continued through the early 2000's.
 Since that point, the proportion of dissertations written by women has largely plateaued at a few percentage points below the proportion written by men.

### 2014-09-15 article.Rmd
Adding citations and editing

@@
 The article describes the methodology as implemented in the [Gender](https://github.com/ropensci/gender) package for the [R programming language](http://www.r-project.org/).
-It goes on to demonstrate the method's utility by applying it to a case study in which we study gender in the history profession by inferring the gender of authors of history dissertations and of authors and reviewers in the *American Historical Review*.
+It goes on to apply the method to a case study in which we study gender and gatekeeping in the history profession by inferring the gender of authors of history dissertations and of authors and reviewers in the *American Historical Review*.
 
@@
 
-As historians and other humanities researchers grapple with larger and larger datasets, they are increasingly using algorithms not just to extract information from existing data but to infer additional information from the data that is available.^[CITE: find example – network analysis, maybe http://kieranhealy.org/blog/archives/2013/06/09/using-metadata-to-find-paul-revere/] Datasets of interest to historians---for example membership lists, company payrolls, military rosters, passenger bookings, records of correspondence, and lists of published works---might included records with the names of individual people.
-Some datasets might contain associated information like a person's address, age, or rank.
-Other datasets contain little more than a personal name.
+As humanities researchers grapple with larger and larger datasets they will increasingly be able to not just extract existing information from that data but to also infer additional information from it \[Goldstone and Underwood 2014]\.
+^[The availability of large-scale datasets dealing with personal information raises important ethical issues for digital humanists, particularly in light of the revelations about domestic spying at the NSA brought forth by Edward Snowden in 2013.
+\[Weingart 2014]\ Kieran Healy has used network analysis to demonstrate how researchers can infer information from datasets while also touching on the troubling ethical issue of surveillance \[Kiely 2013]\.] Many of these datasets deal with lists of people: company payrolls, military rosters, passenger bookings, records of correspondence, or lists of published works.
+Some of these records might contain associated information such a person's address, age, or rank, while others contain little more than a list of personal names.
 Given large but sparse datasets, researchers have to use their ingenuity to derive additional kinds of information.
-What kind of information can researchers infer from just a name?
-Most obviously first names often imply a person's gender.
+What can researchers infer from just a name?
+On a basic level, first names often imply a person's gender.
 "Jane Fay" is almost certainly a woman, while her brother "John Fay" is almost certainly a man.
@@
 
-The problem arises from the fact that the link between gender and naming practices, like language itself, is not static.
+The problem with inferring gender from names is that the link between gender and naming practices, like language itself, is not static.
 What about Jane and John Fay's sibling, Leslie?
@@
 But if the three siblings were born in the early twentieth century, chances are good that Leslie would be their brother.
-That is because the gendered naming convention surrounding "Leslie" completely changed over the course of one hundred years.
+That is because the gendered naming convention surrounding "Leslie reversed course completely over the course of one hundred years.
 92% of the babies born in the United States in 1900 who were named "Leslie" were male, while 96% of those identically named babies born in the year 2000 were female.
@@
 
-Many other names as used in the United States, such as Madison, Morgan, Sydney, and Kendall, have also changed gender over the past century.
+Other names used in the United States such as Madison, Morgan, Sydney, and Kendall, have also "flipped" their gender over the past century.
 Such changes usually follow set patterns.
@@
 In general, names that transition from male to female also became increasingly popular over time.
-For example, the name "Madison" changed from male to female suddenly after 1985.
-Whereas before it had been an infrequently used male name, after 1985 it skyrocketed in popularity, and by the early 2000s it was one of the ten most frequently used names for baby girls.
-Broadly speaking, since the late twentieth century Americans girls are more likely than American boys to receive a "trendy" name, and once that name is associated with females it is unlikely to be given to male babies.^[The few exceptions of names which go from being female to male (such as "Jan") have a different causal mechanism.
+For example, the name "Madison" changed from male to female quite suddenly after 1985.
+Whereas before it had been an infrequently used male name, after 1985 it skyrocketed in popularity and by the early 2000s was one of the ten most frequently used names for baby girls.
+Broadly speaking, in the late twentieth century Americans girls became increasingly more likely than American boys to receive a "trendy" name, and once that name was associated with women it was unlikely to be given to male babies.
+These names present a moving target for researchers hoping to use them to infer additional information about the people under study.
+^[The few exceptions of names which flip from being female to male, such as "Jan", have a different causal mechanism.
 It seems likely that Jan as an English name is female but that Jan as a Scandinavian name is male.
@@
 
-For researchers working on contemporary subjects the "Leslie" problem is not an especially pressing one.
+For those working on contemporary subjects the "Leslie" problem is not an especially pressing one.
 An education policy analyst studying the demographics of different urban school districts, for instance, can turn to a variety of tools that use current databases of names, and these tools will likely identify the gender of Leslie Fay correctly.
-For example, the [Genderize.io](http://genderize.io/) online service determines the gender of first names "from user profiles across major social networks." Checking the service for the name Leslie returns a prediction that the name is female, along with an estimate of what proportion of the uses among members of unnamed social media sites is female and male.^[In this example the Genderize.io service has been accessed through the Gender package described below.]
+One online service, [Genderize.io](http://genderize.io/), determines the gender of first names "from user profiles across major social networks." Checking the service for the name Leslie returns a prediction that the name is female, along with an estimate of male/female proportions of the uses among members of unnamed social media sites.^[In this example the Genderize.io service has been accessed through the Gender package described below.]
 
@@
 
-Researchers studying historical subjects, on the other hand, need to take into account changes in naming practices.
-If that same researcher wanted to compare the demographics of contemporary urban schools with those from the early twentieth century, contemporary datasets would misidentify Leslie Fay's gender.
+But researchers studying a longer timespan need to take into account the changing nature of naming practices.
+If that same education researcher wanted to compare the demographics of contemporary urban schools with those from the early twentieth century, contemporary tools like Genderize.io would misidentify Leslie Fay's gender in historical datasets.
 The "Leslie" problem is not just for researchers who think of themselves as historians.
-Anyone studying a period longer than a few years, or anyone studying a group whose characteristics do not match the probably unknowable characteristics of the groups that comprise social media datasets will also encounter this problem.
-As of 2012, the average American has a lifespan of nearly seventy-nine years–--more than enough time for naming practices to change quite dramatically.^[World Bank, *World Bank Open Data*, "Life expectancy at birth, total (years)," <http://data.worldbank.org/indicator/SP.DYN.LE00.IN> (accessed September 12, 2014).] Predicting gender from first names therefore requires a fundamentally historical method.
+Anyone studying a period longer than a few years, or anyone studying a group whose demographics do not match those of a contemporary tool like Genderize.io will also encounter this problem.
+As of 2012, the average American has a lifespan of nearly seventy-nine years–--more than enough time for naming practices to change quite dramatically between their birth and death.^[World Bank, *World Bank Open Data*, "Life expectancy at birth, total (years)," <http://data.worldbank.org/indicator/SP.DYN.LE00.IN> (accessed September 12, 2014).] Predicting gender from first names therefore requires a fundamentally historical method.
 
@@
 As we have demonstrated in our study of gatekeeping in the American historical profession, our method can reveal important details about gender even from datasets which do not otherwise include that information, and it can do so over long historical periods.
+
+### References
+
+Goldstone and Underwood 2014.
+Goldstone, Andrew, and Ted Underwood.
+"The Quiet Transformations of Literary Studies: What Thirteen Thousand Scholars Could Tell Us." *New Literary History* (Forthcoming 2014).
+
+Healy 2013. Healy, Kieran. "Using Metadata to Find Paul Revere." *kieranhealy.org*. 9 June 2013, http://kieranhealy.org/blog/archives/2013/06/09/using-metadata-to-find-paul-revere/
+
+Weingart 2014. Weingart, Scott. "The Moral Role of DH in a Data-Driven World." *the scottbot irregular*. 9 September 2014, http://www.scottbot.net/HIAL/?p=40944

### 2014-09-15 article.Rmd
Citations

@@
 Anyone studying a period longer than a few years, or anyone studying a group whose demographics do not match those of a contemporary tool like Genderize.io will also encounter this problem.
-As of 2012, the average American has a lifespan of nearly seventy-nine years–--more than enough time for naming practices to change quite dramatically between their birth and death.^[World Bank, *World Bank Open Data*, "Life expectancy at birth, total (years)," <http://data.worldbank.org/indicator/SP.DYN.LE00.IN> (accessed September 12, 2014).] Predicting gender from first names therefore requires a fundamentally historical method.
+As of 2012, the average American has a lifespan of nearly seventy-nine years–--more than enough time for naming practices to change quite dramatically between their birth and death [World Bank 2014].
+Predicting gender from first names therefore requires a fundamentally historical method.
 
-Our solution to the "Leslie" problem is to create a software package that combines a predictive algorithm with a several historical datasets suitable to various times and regions.
+Our solution to the "Leslie" problem is to create a software package that combines a predictive algorithm with several historical datasets suitable to various times and regions.
 The algorithm calculates the proportion of male or female uses of a name in a given birth year or range of years.
-It thus can provide not only a prediction of the gender of a name, but also an estimate of how likely that prediction is.
-For example 52 percent of the babies named Leslie born in the United States in 1950 were girls and 48 percent were boys.
-This allows researchers to conduct more nuanced analysis and to determine for themselves what proportion is acceptable for making a prediction of gender.
+It thus can provide not only a prediction of the gender of a name, but also a measure of that prediction's likelihood.
+For example, our program predicts that a person named Leslie born in 1950 is female, but only just barely: 52 percent of the babies named Leslie born in the United States in 1950 were girls and 48 percent were boys.
+Including these proportion figures allows researchers to conduct more nuanced analysis and to determine for themselves what proportion is acceptable predicting the gender of names.
 
-We recognize that gender is a fluid concept, both today and in the past, and that this method glosses over the complexities of historical definitions of gender and of the relationship between biological sex and gender.
+We recognize that gender is a fluid concept, both today and in the past, and that this method glosses over the complexities of historical definitions of gender and of the relationship between biological sex and gender \[Butler 1990, Butler 1993]\.
 In some sense this is unavoidable since the large historical datasets which make this kind of analysis possible were created by state agencies.
-For the category of gender (as for race) states have offered a limited number of boxes to check, and those boxes extert a tremendous power to define state subjects.^[One of the best-known theorists on gender is Judith Butler, *Gender Trouble* and Butler, *Bodies that Matter*; For the problems with the word "identity" in the humanities and social sciences, see Rogers Brubaker and Frederick Cooper, "Beyond 'Identity'"; Several historians have written about the American state and its drive to impose identities onto subjects: Margot Canady, The Straight State;  Peggy Pascoe, *What Comes Naturally*; Shane Landrum on birth certificates.] Nevertheless, this method is neutral as to what the categories male or female actually mean in any given historical context.
+For the category of gender (as for race) states have offered a limited number of boxes to check, and those boxes extert a tremendous power to define state subjects \[Pascoe 2009, Canaday 2011, Landrum 2014]\.^[The word "identity" is a loaded one in the humanities and social sciences \[Brubaker and Cooper 2000]\.
+] Nevertheless, this method is neutral as to what the categories male or female actually mean in any given historical context.
 Researchers are free---indeed, obligated---to make that determination for themselves, and this package provides access to the underlying dataset so that researchers can interrogate every assumption that comes with them.
@@
 Let's start with a comparison.
-One existing method for predicting gender is available in the [Natural Language Toolkit](http://www.nltk.org/) for Python.^[Steven Bird, Edward Loper, and Ewan Klein, *Natural Language Processing with Python* (O’Reilly, 2009), ch.
-2.
-Chapter 6 of the NLTK book also describes a method of predicting gender from the last letter of the name, but this example is given for the sake of explaining supervised classification and not as an end in itself, so we do not have reason to address this a historical method.] The NLTK is an influential software package for scholarship because it provides an extraordinary range of tools for analyzing natural language.^[The NLTK book (the most common way to cite the software) has been cited 812 times according to [Google Scholar](http://scholar.google.com/scholar?cites=3735380965977150091&as_sdt=5,47&sciodt=0,47&hl=en).
-Because in the humanities (unlike in other fields) the practice of citing even software which determines the outcome of an analysis is regrettably rare, even nonexistent, such metrics are difficult to determine.
-Yet searching of journals such as *Digital Humanities Quarterly* and the *Journal of Digital Humanities* reveals that natural language processing, for which the NLTK is the foremost implementation, is one of the key methodological categories of the digital humanities.] Included in the NLTK are two lists of male and female names created by the computer scientist Mark Kantrowitz with assistance from Bill Ross.^[The [original lists of names](http://www.cs.cmu.edu/afs/cs/project/ai-repository/ai/areas/nlp/corpora/names/0.html) are available at the website of the computer science department at Carnegie Mellon University, and the NLTK implementation is described in its [documentation](http://www.nltk.org/howto/corpus.html).] These lists include `r kantrowitz_unique` unique names.
+One existing method for predicting gender is available in the [Natural Language Toolkit](http://www.nltk.org/) for Python \[Bird et al.
+2009, Chapter 2]\.^[Chapter 6 of \[Bird et al.
+2009] also describes a method of predicting gender from the last letter of the name, but this example is given for the sake of explaining supervised classification and not as an end in itself, so we do not have reason to address this a historical method.] The NLTK is an influential software package for scholarship because it provides an extraordinary range of tools for analyzing natural language.^[\[Bird et al.
+2009]\ has been cited 812 times according to [Google Scholar](http://scholar.google.com/scholar?cites=3735380965977150091&as_sdt=5,47&sciodt=0,47&hl=en).
+Because in the humanities (unlike in other fields) the practice of citing software (even software which determines the outcome of an analysis) is regrettably rare, even nonexistent, such metrics are difficult to determine.
+Yet searching journals such as *Digital Humanities Quarterly* and the *Journal of Digital Humanities* reveals that natural language processing, for which the NLTK is the foremost implementation, is one of the key methodological categories of the digital humanities.] Included in the NLTK are two lists of male and female names created by the computer scientist Mark Kantrowitz with assistance from Bill Ross.^[The [original lists of names](http://www.cs.cmu.edu/afs/cs/project/ai-repository/ai/areas/nlp/corpora/names/0.html) are available at the website of the computer science department at Carnegie Mellon University, and the NLTK implementation is described in its [documentation](http://www.nltk.org/howto/corpus.html).] These lists include `r kantrowitz_unique` unique names.
 Using the Kantrowitz names corpus in the NLTK, one could look up a name like Jane or John and find out that it is male or female.
@@
 Genderize.io reports that its "database contains 142848 distinct names across 77 countries and 85 languages." Genderize is unsuitable for historical work, however, because it is based only on contemporary data.
-According to the documentation for its API, "It utilizes big datasets of information, from user profiles across major social networks."^[Quotations are taken from the [Genderize.io website](http://genderize.io), with correspondence from the developer behind the site for confirmation.] It would be anachronistic to apply these datasets to the past, and Genderize.io provides no functionality to filter results chronologically as it does geographically.
-In addition, Genderize.io does not make clear exactly what comprises the dataset and how it was gathered, which keeps scholars from interrogating the value of the source.
+According to the documentation for its API, "It utilizes big datasets of information, from user profiles across major social networks." It would be anachronistic to apply these datasets to the past, and Genderize.io provides no functionality to filter results chronologically as it does geographically.
+In addition, Genderize.io does not make clear exactly what comprises the dataset and how it was gathered, which keeps scholars from interrogating the value of the source \[Genderize.io 2014-]\.^[Quotations are taken from the [Genderize.io website](http://genderize.io), with correspondence from the developer behind the site for confirmation.]
 
-R and Python are the most commonly used languages for data analysis among digital humanists.^[Surveys and analysis of useage point out the growth of R and the continuing popularity of Python among "data science" generally.
-See, for example, Robert A.
-Muenchen, "The Popularity of Data Analysis Software," *r4stats.com*, <http://r4stats.com/articles/popularity/> (accessed September 12, 2014); Gregory Piatetsky, "Top Languages for analytics, data mining, data science," *KDN Nuggets*, <http://www.kdnuggets.com/2013/08/languages-for-analytics-data-mining-data-science.html> (accessed September 12, 2014).] Python's existing packages for gender prediction all implement a method similar to the NLTK or Genderize.io.^[Python library names] To our knowledge, [CRAN](http://cran.rstudio.com/) (the central repository for R packages) does not include any packages dedicated to gender prediction.
+R and Python are two of the most commonly used languages for data analysis \[Piatesky 2014, Muenchen 2014-]\.^[Surveys and analysis of useage point out the growth of R and the continuing popularity of Python among "data science" generally.
+See, for example, ] Python's existing packages for gender prediction all implement a method similar to the NLTK or Genderize.io.^[Python library names] To our knowledge, [CRAN](http://cran.rstudio.com/) (the central repository for R packages) does not include any packages dedicated to gender prediction.
 In two of the most popular data analysis languages in the digital humanities there is currently no satisfactory existing method of gender prediction for historical and humanistic research.
@@
 
+Bird et al. 2009. Bird, Steve, Edward Loper, and Ewan Klein, *Natural Language Processing with Python* (O’Reilly, 2009)
+
+Brubaker and Cooper 2000. Brubaker, Rogers, and Frederick Cooper. “Beyond ‘identity.’” Theory and Society 29.1 (2000): 1–47.
+
+Butler 1989. Butler, Judith. *Gender Trouble: Feminism and the Subversion of Identity*. New York: Routledge, 1990.
+
+Butler 1993. Butler, Judith. *Bodies That Matter: On the Discursive Limits of Sex*. New York: Routledge, 1993.
+
+Canaday 2011. Canaday, Margot. *The Straight State: Sexuality and Citizenship in Twentieth-Century America*. Princeton, N.J.: Princeton University Press, 2011.
+
+Genderize.io 2014-. *Genderize.io*. Accessed 15 September 2014, http://genderize.io/.
+
 Goldstone and Underwood 2014.
@@
 
+Muenchen 2014-. Muenchen, Robert A. "The Popularity of Data Analysis Software," *r4stats.com*. Accessed 12 September 2014, http://r4stats.com/articles/popularity/.
+
+Pascoe 2009. Pascoe, Peggy. *What Comes Naturally: Miscegenation Law and the Making of Race in America*. Oxford University Press, USA, 2009.
+
+Piatetsky 2013.
+Piatetsky, Gregory.
+"Top Languages for analytics, data mining, data science," *KDN Nuggets*.
+27 August 2013, http://www.kdnuggets.com/2013/08/languages-for-analytics-data-mining-data-science.html.
+
 Weingart 2014. Weingart, Scott. "The Moral Role of DH in a Data-Driven World." *the scottbot irregular*. 9 September 2014, http://www.scottbot.net/HIAL/?p=40944
+
+World Bank 2014-. World Bank. "Life expectancy at birth, total (years)." *World Bank Open Data*. Accessed 12 September 2014, http://data.worldbank.org/indicator/SP.DYN.LE00.IN.

### 2014-09-15 article.Rmd
More citations

@@
 
-R and Python are two of the most commonly used languages for data analysis \[Piatesky 2014, Muenchen 2014-]\.^[Surveys and analysis of useage point out the growth of R and the continuing popularity of Python among "data science" generally.
-See, for example, ] Python's existing packages for gender prediction all implement a method similar to the NLTK or Genderize.io.^[Python library names] To our knowledge, [CRAN](http://cran.rstudio.com/) (the central repository for R packages) does not include any packages dedicated to gender prediction.
-In two of the most popular data analysis languages in the digital humanities there is currently no satisfactory existing method of gender prediction for historical and humanistic research.
+R and Python are two of the most commonly used languages for data analysis \[Piatesky 2014, Muenchen 2014-]\.^[Surveys and analysis of useage point out the growth of R and the continuing popularity of Python among "data science" generally.] Python's existing packages for gender prediction all implement a method similar to the NLTK or Genderize.io.^[**Python library names**] To our knowledge, [CRAN](http://cran.rstudio.com/) (the central repository for R packages) does not include any packages dedicated to gender prediction.
+In two of the most popular data analysis languages used in the digital humanities there is currently no satisfactory existing method of gender prediction for historical and humanistic research.
 
@@
 The Gender package is affiliated with [rOpenSci](http://ropensci.org/), an initiative that supports reproducible research and open data for scientists using R.
-The rOpenSci advisory board has provided code review and publicity for the project.^[The gender package depends especially on R Core Team, *R: A language and environment for statistical computing*.
-R Foundation for Statistical Computing, Vienna, Austria (2014).
-<http://www.R-project.org/>;  Hadley Wickham and Romain Francois, *dplyr: dplyr: a grammar of data manipulation*.
-R package version 0.2 (2014): <http://CRAN.R-project.org/package=dplyr>.
-The sources of the data are cited below.]
+The rOpenSci advisory board has provided code review and publicity for the project.^[The gender package depends especially on \[R Core Team 2014-, Wickham and Francois 2014]\ ]
 
@@
 Unsurprisingly such datasets are almost always gathered in the first instance by government entities, though the compilation and digitization may be undertaken by scholarly researchers.
-Second, it depends on a suitable algorithm for estimating the proportion of male and female names for a given year or range of years.
-since often a person cannot be associated with an exact date.
+Second, it depends on a suitable algorithm for estimating the proportion of male and female names for a given year or range of years, since often a person cannot be associated with an exact date.
 It is especially important that the algorithm take into account any biases in the data to formulate more accurate predictions.
@@
 
-The Social Security Administration "[Baby Names](http://catalog.data.gov/dataset/baby-names-from-social-security-card-applications-national-level-data)" dataset was created as a result of the Social Security Act of 1935 during the New Deal.
-^[Social Security Administration, "[Baby Names from Social Security Card Applications-National Level Data](http://catalog.data.gov/dataset/baby-names-from-social-security-card-applications-national-level-data)," Data.gov dataset ID US-GOV-SSA-338 (accessed August 31, 2014).
-The Social Security Administration also provides a related dataset which contains the names of applicants broken down by state and year since 1910.
-Social Security Administration, "[Baby Names from Social Security Card Applications-Data by State and District of Columbia](http://catalog.data.gov/dataset/baby-names-from-social-security-card-applications-data-by-state-and-district-of-)," Data.gov dataset ID US-GOV-SSA-339 (accessed August 31, 2014).] This dataset contains a list of first names with the number of male and female uses of the name in every year beginning with 1880 and ending with 2012.
+The Social Security Administration (SSA) "[Baby Names](http://catalog.data.gov/dataset/baby-names-from-social-security-card-applications-national-level-data)" dataset was created as a result of the Social Security Act of 1935 during the New Deal.
+^[The Social Security Administration also provides a related dataset which contains the names of applicants broken down by state and year since 1910 \[Social Security Administration 2014-b]\.] This dataset contains a list of first names with the number of male and female uses of the name in every year beginning with 1880 and ending with 2012.
 The SSA list includes any name which was used more than five times in a given year, thereby capturing all but the most infrequently used names from each year.
@@
 Even after 1935, registration at birth for Social Security was not mandatory until 1986.
-As we will demonstrate below, the way in which the data was gathered requires an adjustment to our predictions of gender.^[CITE history of SSA.
-On the history of birth registration see Shane Landrum, "The State's Big Family Bible: Birth Certificates, Personal Identity, and Citizenship in the United States, 1840--1950" (PhD dissertation, Brandeis University, 2014).
-It is well-known among American historians that Social Security was written to exclude farmworkers as a means of also excluding African Americans.
+As we will demonstrate below, the way in which the data was gathered requires an adjustment to our predictions of gender.^[Shane Landrum has studie dthe history of birth registration in the United States \[Landrum 2014]\.
+It is well-known among American historians that the Social Security Act was written to exclude farmworkers as a means of also excluding African Americans.
 In examining the dataset we have not found any reason to believe that this policy noticably skewed the prediction of gender from names.]
@@
 
-In conclusion, our algorith for predicting gender from first names combined with historical datasets in the Gender package provides a powerful new method for historians interested in the study of gender.
+In conclusion, our algorithm for predicting gender from first names combined with historical datasets in the Gender package provides a powerful new method for historians interested in the study of gender.
 The digital turn has made available very large datasets, yet in the midst of abundance historians always have to deal with the problem of scarcity in the sources.
@@
 
+Landrum, in progress.
+Landrum, Shane.
+"The State's Big Family Bible: Birth Certificates, Personal Identity, and Citizenship in the United States, 1840--1950." PhD dissertation, Brandeis University, in progress.
+
 Muenchen 2014-. Muenchen, Robert A. "The Popularity of Data Analysis Software," *r4stats.com*. Accessed 12 September 2014, http://r4stats.com/articles/popularity/.
@@
 
+R Core Team 2014-. R Core Team, *R: A language and environment for statistical computing*. R Foundation for Statistical Computing, Vienna, Austria, 2014. <http://www.R-project.org/>
+
+Social Security Administration 2014-a.
+Social Security Administration, "Baby Names from Social Security Card Applications-National Level Data".
+Data.gov dataset ID US-GOV-SSA-338.
+Accessed 31 August 2014, http://catalog.data.gov/dataset/baby-names-from-social-security-card-applications-national-level-data
+
+Social Security Administration 2014-b.
+Social Security Administration, "Baby Names from Social Security Card Applications-Data by State and District of Columbia." Data.gov dataset ID US-GOV-SSA-339.
+Accessed 31 August 2014, http://catalog.data.gov/dataset/baby-names-from-social-security-card-applications-data-by-state-and-district-of-
+
 Weingart 2014. Weingart, Scott. "The Moral Role of DH in a Data-Driven World." *the scottbot irregular*. 9 September 2014, http://www.scottbot.net/HIAL/?p=40944
 
+Wickham and Francois 2014-. Wickham, Hadley and Romain Francois, *dplyr: dplyr: a grammar of data manipulation*. R package version 0.2 (2014): <http://CRAN.R-project.org/package=dplyr>.
+
 World Bank 2014-. World Bank. "Life expectancy at birth, total (years)." *World Bank Open Data*. Accessed 12 September 2014, http://data.worldbank.org/indicator/SP.DYN.LE00.IN.

### 2014-09-15 article.Rmd
Citations

@@
 Anyone studying a period longer than a few years, or anyone studying a group whose demographics do not match those of a contemporary tool like Genderize.io will also encounter this problem.
-As of 2012, the average American has a lifespan of nearly seventy-nine years–--more than enough time for naming practices to change quite dramatically between their birth and death [World Bank 2014].
+As of 2012, the average American has a lifespan of nearly seventy-nine years–--more than enough time for naming practices to change quite dramatically between their birth and death \[World Bank 2014]\.
 Predicting gender from first names therefore requires a fundamentally historical method.
@@
 Even after 1935, registration at birth for Social Security was not mandatory until 1986.
-As we will demonstrate below, the way in which the data was gathered requires an adjustment to our predictions of gender.^[Shane Landrum has studie dthe history of birth registration in the United States \[Landrum 2014]\.
+As we will demonstrate below, the way in which the data was gathered requires an adjustment to our predictions of gender.^[Shane Landrum has studied the history of birth registration in the United States \[Landrum 2014]\.
 It is well-known among American historians that the Social Security Act was written to exclude farmworkers as a means of also excluding African Americans.
@@
 Unlike the SSA dataset, which includes 100% of every name reported to the Social Security Administration, the IPUMS data contains 5% or 10% samples.
-Because the `gender()` function relies on proportions of uses of names, rather than raw counts of people with the names, the sampling does not diminish the reliability of the function's predicitons.^[Steven Ruggles, J.
-Trent Alexander, Katie Genadek, Ronald Goeken, Matthew B.
-Schroeder, and Matthew Sobek.
-Integrated Public Use Microdata Series: Version 5.0 [Machine-readable database].
-Minneapolis, MN: Minnesota Population Center [producer and distributor], 2010.]
+Because the `gender()` function relies on proportions of uses of names, rather than raw counts of people with the names, the sampling does not diminish the reliability of the function's predicitons \[Ruggles et al.
+2010]\.
 
-These two datasets (and any future datasets to be added to the package) thus contain a simple table of data. The columns
-
+These two datasets (and any future datasets to be added to the package) thus contain a simple table of data.
 ^[The format of this data departs from the more generally desirable "tidy data" framework proposed by Hadley Wickham for performance reasons.
-By keeping the data in a "wider" format instead of a tidy format as defined by Wickham, the function is able to make its predictions faster.
-Hadley Wickham, "[Tidy Data](http://vita.had.co.nz/papers/tidy-data.html)," *Journal of Statistical Software* (forthcoming).] The columns contain the `name`, `year`, and number of `female` and `male` uses of a name.
+By keeping the data in a "wider" format instead of a tidy format as defined by Wickham, the function is able to make its predictions faster \[Wickham forthcoming]\.] The columns contain the `name`, `year`, and number of `female` and `male` uses of a name.
 Each row contains a combination for a name and a year.
@@
 
+Ruggles et al.
+2010.
+Ruggles, Steven, J.
+Trent Alexander, Katie Genadek, Ronald Goeken, Matthew B.
+Schroeder, and Matthew Sobek.
+Integrated Public Use Microdata Series: Version 5.0 \[Machine-readable database]\.
+Minneapolis, MN: Minnesota Population Center \[producer and distributor]\, 2010.
+
 Social Security Administration 2014-a.
@@
 
+Wickham, forthcoming. Hadley Wickham, "Tidy Data." *Journal of Statistical Software* (forthcoming), http://vita.had.co.nz/papers/tidy-data.html.
+
 Wickham and Francois 2014-. Wickham, Hadley and Romain Francois, *dplyr: dplyr: a grammar of data manipulation*. R package version 0.2 (2014): <http://CRAN.R-project.org/package=dplyr>.

### 2014-09-16 article.Rmd
Editing - see comments in bold

@@
 For example, our program predicts that a person named Leslie born in 1950 is female, but only just barely: 52 percent of the babies named Leslie born in the United States in 1950 were girls and 48 percent were boys.
-Including these proportion figures allows researchers to conduct more nuanced analysis and to determine for themselves what proportion is acceptable predicting the gender of names.
+Including these proportion figures allows researchers to conduct more nuanced analysis and to determine for themselves what proportion is acceptable for predicting the gender of names.
 
@@
 In some sense this is unavoidable since the large historical datasets which make this kind of analysis possible were created by state agencies.
-For the category of gender (as for race) states have offered a limited number of boxes to check, and those boxes extert a tremendous power to define state subjects \[Pascoe 2009, Canaday 2011, Landrum 2014]\.^[The word "identity" is a loaded one in the humanities and social sciences \[Brubaker and Cooper 2000]\.
-] Nevertheless, this method is neutral as to what the categories male or female actually mean in any given historical context.
+For the category of gender (as for race) states have offered a limited number of boxes to check, and those boxes extert a tremendous power to define state subjects \[Brubaker and Cooper 2000, 14-17, Pascoe 2009, Canaday 2011, Landrum 2014]\.
+This method does not presuppose that the gender binary imposed by governments and their wider societies reflect how an individual identifies oneself and is neutral as to what the categories "male" or "female" actually mean in any given historical context.
 Researchers are free---indeed, obligated---to make that determination for themselves, and this package provides access to the underlying dataset so that researchers can interrogate every assumption that comes with them.
@@
 Some names like Abby are overwhelmingly female and some like Bill are overwhelmingly male, but the corpus includes them as both male and female.
-The Kantrowitz corpus contains only `r kantrowitz_unique` names, a mere `r percent(kantrowitz_unique / ssa_unique)` of the `r ssa_unique` names in the Social Security Administration dataset and `r percent(kantrowitz_unique / ipums_unique)` of the `r ipums_unique` names in the census records provided by IPUMS USA.
+The Kantrowitz corpus contains only `r kantrowitz_unique` names, a mere `r percent(kantrowitz_unique / ssa_unique)` of the `r ssa_unique` names in the Social Security Administration dataset and `r percent(kantrowitz_unique / ipums_unique)` of the `r ipums_unique` names in the census records provided by Integrated Public Use Microdata (IPUMS) USA.
 There are therefore many names that it cannot identify.
@@
 The IPUMS-USA dataset, contributed by Benjamin Schmidt, contains records from the United States decennial census from 1790 to 1930.
-This dataset contains the birth year and numbers of males and females under the age of 62 for all the years in that range.
+This dataset includes the birth year and numbers of males and females under the age of 62 for all the years in that range.
 This data has been aggregated by IPUMS at the University of Minnesota and is released as a sample of the total census data.
-Unlike the SSA dataset, which includes 100% of every name reported to the Social Security Administration, the IPUMS data contains 5% or 10% samples.
+Unlike the SSA dataset, which includes 100% of every name reported to the Social Security Administration **not quite - only names that occur more than five times, right?**, the IPUMS data contains 5% or 10% samples.
 Because the `gender()` function relies on proportions of uses of names, rather than raw counts of people with the names, the sampling does not diminish the reliability of the function's predicitons \[Ruggles et al.
@@
 ^[The format of this data departs from the more generally desirable "tidy data" framework proposed by Hadley Wickham for performance reasons.
-By keeping the data in a "wider" format instead of a tidy format as defined by Wickham, the function is able to make its predictions faster \[Wickham forthcoming]\.] The columns contain the `name`, `year`, and number of `female` and `male` uses of a name.
-Each row contains a combination for a name and a year.
+By keeping the data in a "wider" format instead of a tidy format as defined by Wickham, the function is able to make its predictions faster \[Wickham forthcoming]\.] The columns contain the `name`, `year`, and number of `female` and `male` uses of that name in a particular year.
 
@@
 
+**unfortunatley the first ten isn't intuitive since it's just "!" rather than names - this is going to confuse readers, I think**
 ```{r ipums_usa, echo=TRUE, results='markup'}
@@
 Our method for predicting gender is best understood through a series of examples.
-First we will demonstrate a simplified version of the inner workings of the function for predicting a single name.
-This example will demonstrate fundamentally how the function works.
-Next we will demonstrate how to use the function provided in the gender package to predict gender from a small sample dataset.
+First we will use it to predict the gender of a single name in order to demonstrate a simplified version of the inner workings of the function.
+We will then apply it to a small sample dataset to show how a researcher might use it in practice.
 
-The method for predicting gender from a name using the provided datasets is fundamentally simple.
-Let's begin by assuming that using the Social Security dataset we want to predict the gender of someone named Sidney who was born in 1935.
+**Example #1: A Sample Name**
+
+The method for predicting gender from a name using the package's datasets is simple.
+Let's begin by assuming that we want to predict the gender of someone named Sidney who was born in 1935 using the Social Security Administration dataset.
 Because the dataset contains a list of names for each year, we can simply lookup the row for Sidney in 1935.
@@
 
-Thus according to the Social Security Administration, there were 974 boys and 93 females named Sidney born in 1935.
-We can add another command to calculate the proportion of females and males ("mutate" in dplyr's vocabulary).
+Thus, according to the Social Security Administration, there were 974 boys and 93 girls named Sidney born in 1935.
+We can add another command to calculate the *proportion* of females and males ("mutate" in dplyr's vocabulary) rather than raw numbers.
 
@@
 In other words, there is an approximately `r gender("sidney", year = 1935)$proportion_male %>% percent()` percent chance that a person born in 1935 named Sidney was male.
-(In 2012, for comparison, there was an approximately `r gender("sidney", year = 2012)$proportion_female %>% percent()` percent chance that a person born named Sidney was female.)
+In 2012, for comparison, there was an approximately `r gender("sidney", year = 2012)$proportion_female %>% percent()` percent chance that a person born named Sidney was female.
 
@@
 
-Next we can sum up the male and female columns ("summarize" in dplyr's vocabulary) and calculate the proportions of female and male uses.
+Next we can sum up the male and female columns ("summarize" in dplyr's vocabulary) and calculate the proportions of female and male uses of "Sideny" during that decade.
 
@@
 
-In other words, for the decade of the 1930s, we can calculate that there is a `r gender("sidney", year = c(1930, 1939))$proportion_male %>% percent()` percent chance tha person named Sidney was male.
-This is roughly the same as the probability that the name was male in just 1935, but by looking at the underlying counts on which it is based, we can see this proportion was calculated from ten years instead of one.
+In other words, for the decade of the 1930s, we can calculate that there is a `r gender("sidney", year = c(1930, 1939))$proportion_male %>% percent()` percent chance that person named Sidney was male.
+This is roughly the same as the probability we calculated above for just 1935, but our method also returns the figures it used to calculate those probabilities: 1,067 total instances of "Sidney" in 1935 versus 10,110 instances for the decade as a whole.
+
+**Example #2: A Sample Dataset**
+
+The method's real utility stems from being able to process larger datasets than a single name.
+Let's use, say, a list of editors from a college newspaper to illustrate how a researcher might apply it to their own data.
+The package's prediction function allows researchers to choose which reference datasets they would like to use and the range of years for making their predictions.^[See the package vignette for more details: `vignette("predicting-gender", package = "gender")`.]
+
+```{r}
+editors <- data.frame(name = c("Madison", "Madison", "Morgan",
+                                  "Morgan", "Jan", "Jan"),
+                         birth_year = c(1934, 1990, 1948, 1970, 1965, 1998),
+                         stringsAsFactors = FALSE) %>% tbl_df()
+```
+
+```{r, echo=TRUE, results='markup'}
+editors
+```
+
+The simplest way to use the package would be to call the `gender()` function, specifying the method, in this case the Social Security Administration dataset since all the names are after 1930, and convert the results to a data frame (one of R's native data structures).
+
+```{r, echo=TRUE, results='markup'}
+gender(editors$name, method = "ssa") %>% do.call(rbind.data.frame, .)
+```
+
+This basic call of the `gender()` function would be suitable if we did not know the birth year for each editor. But we can also specify that in the `gender()` function by using this call:
+
+```{r, echo=TRUE, results='markup'}
+Map(gender, editors$name, years = editors$birth_year) %>% 
+  do.call(rbind.data.frame, .)
+```
+
+By taking into account the year of birth, we find that four of our six names were likely male, whereas we might otherwise have predicted that all six were female.
+We also now know the approximate likelihood that our predictions are correct: at least 80% for all of thse predictions.
+
+**I think it might be worth expanding this Example #2 section by a paragraph to show how you can dynamically incorporate periods of time rather than single birth years.
+A single birth year might make humanists uncomfortable since it implies a precision that a lot of people aren't going to have (like in the case of our dissertations and AHR authors).
+I changed the example to be a list of editors of a college paper.
+That way we can easily use an age range of 18-22 to demonstrate how a researcher might implement this.**
+
+**I moved this next caveat section about the SSA data because it felt kind of out of place sandwiched between the two examples.
+I'm still not sure it fits here, though.
+I'm trying to figrue out if there's a way to integrate it into the paper that isn't going to be quite so diversionary for the reader.**
 
@@
 We apply this correction factor automatically when using the SSA dataset.
-
-These internal workings are wrapped by the prediction function so that researchers can focus on the choice of datasets and the range of years for making their predictions.
-Let's use the following sample dataset, say, a membership role, as an example for how to use the function.^[See the package vignette for more details: `vignette("predicting-gender", package = "gender")`.]
-
-```{r}
-membership <- data.frame(name = c("Madison", "Madison", "Morgan",
-                                  "Morgan", "Jan", "Jan"),
-                         birth_year = c(1934, 1990, 1948, 1970, 1965, 1998),
-                         stringsAsFactors = FALSE) %>% tbl_df()
-```
-
-```{r, echo=TRUE, results='markup'}
-membership
-```
-
-The simplest way to use the package would be to call the `gender()` function, specifying the method, in this case the Social Security Administration dataset since all the names are after 1930, and convert the results to a data frame (one of R's native data structures).
-
-```{r, echo=TRUE, results='markup'}
-gender(membership$name, method = "ssa") %>% do.call(rbind.data.frame, .)
-```
-
-This most basic call would be suitable if we did not know the year of birth. But we can also specify that in the `gender()` function by using this call:
-
-```{r, echo=TRUE, results='markup'}
-Map(gender, membership$name, years = membership$birth_year) %>% 
-  do.call(rbind.data.frame, .)
-```
-
-By taking into account the year of birth, we find that four of our six names were likely male, whereas we might otherwise have predicted that all six were female.
-We also now know the approximate likelyhood that our predictions are correct: at least 80% for all of thse predictions.
 
@@
 
-Brubaker and Cooper 2000. Brubaker, Rogers, and Frederick Cooper. “Beyond ‘identity.’” Theory and Society 29.1 (2000): 1–47.
+Brubaker and Cooper 2000. Brubaker, Rogers, and Frederick Cooper. “Beyond ‘identity.’” *Theory and Society* 29.1 (2000): 1–47.
 

### 2014-09-16 article.Rmd
Break loading and predicting into separate scripts

@@
 
-```{r gender-book-authors, eval=FALSE}
-ahr <- read.csv("data/ahr.utf8.csv", stringsAsFactors = FALSE) %>%
-  tbl_df()
-
-# For a number of entries the first name column contains first and middle names,
-# so we're going to parse out just the first names.
-ahr <- ahr %>%
-  mutate(first_name = word(Author_First))
-
-# We need to guess at the range of ages for book authors. 25 to 75 seems
-# sensible enough.
-min_age <- 25
-max_age <- 75
-
-# Now we will create columns for the minimum year and the maximum year for
-# guessing gender.
-ahr <- ahr %>%
-  mutate(min_year = issueDate - max_age,
-         max_year = issueDate - min_age)
-
-# Now we need to decide which method to use when predicting the gender. In 
-# general we want to use IPUMS before 1930 and SSA after 1930, but sometimes 
-# we'll have to compromise. There are several kinds of obvious cases:
-# 
-# - If max_age < 1930, use IPUMS
-# 
-# - If min_age > 1930, use SSA
-# 
-# The tricky cases are when the terminal values overlap the ranges of one of the
-# methods.
-# 
-# - If min_age > 1880 and max_age > 1930, use SSA
-# 
-# If our range of ages was greater than 50, we might have a person who we 
-# guessed was born before 1880 or after 1930. But by keeping our range smaller
-# than 50 (the difference between 1930 and 1880) we can avoid that problem.
-
-# That does indeed match all cases. So now we can write a function to return the
-# proper method for a given range of years.
-pick_method <- function(min_age, max_age) {
-  ifelse(max_age <= 1930, "ipums",
-         ifelse(min_age > 1930, "ssa",
-         ifelse(min_age > 1880 & max_age > 1930, "ssa", NA)))
-}
-
-# Now assign a method to the AHR dataset
-ahr <- ahr %>%
-  mutate(method = pick_method(min_year, max_year))
-
-pass_through <- function(x) { return(x) }
-
-apply_gender <- function(df) {
-  gender(df$first_name, method = df$method, 
-         years = c(df$min_year, df$max_year))
-  }
-
-ahr_gender <- ahr %>% 
-  group_by(first_name, vol_issue_ID, Author_Full, Reviewer_First,
-           Reviewer_Full, category) %>%
-  do(gender_author = apply_gender(.))
-
-# Finally we are ready to predict the gender. Because we have to combine years
-# we're going to do this as a for loop, which is ugly.
-# ahr_gender.l <- vector(mode = "list", nrow(ahr))
-
-# for(i in 1:nrow(ahr)) {
-#   row <- ahr[i, ]
-#   result <- gender(row$first_name, method = row$method,
-#                    years = c(row$min_year, row$max_year))
-#   ahr_gender.l[[i]] <- result
-# }
-
-# ahr_gender.df <- ahr_gender.l %>% do.call(rbind.data.frame, .)
-```
-
 But a closer look shows important differences between the proportion of women writing history dissertations and the proportion of women appearing in the AHR.

### 2014-09-16 R/load-data.R

@@ -1,0 +1,17 @@
+# Load the data and do any cleaning
+
+library(dplyr)
+library(stringr)
+
+# The dissertations data has already been cleaned.
+# See https://github.com/lmullen/dissertations-data
+diss <- read.csv("data/h_diss2.csv", stringsAsFactors = FALSE) %>%
+  tbl_df()
+
+# The AHR data has been converted to UTF8. For some entries the first name
+# column contains first and middle names, so we're going to parse out just the
+# first names.
+ahr <- read.csv("data/ahr.utf8.csv", stringsAsFactors = FALSE) %>%
+  tbl_df() %>%
+  mutate(author_first_name = word(Author_First),
+         reviewer_first_name = word(Reviewer_First))

### 2014-09-16 R/load-dissertations.R

@@ -1,334 +1,0 @@
-# load2.r 
-# This script loads the data and cleans it. It is a copy of `load.r`, which was
-# used for posts 1-5, but modified to clean the data more rigorously. It is also
-# better optimized. I have two load scripts so that results from the earlier
-# posts are preserved unmodified.
-
-library(dplyr)
-library(stringr)
-
-# Read the historical history dissertations data
-raw <- read.csv("data/history-df.csv", stringsAsFactors=F)
-
-# Turn import into a tbl_df to make it easier to work with dplyr
-raw <- tbl_df(raw)
-
-# Turn empty strings into NAs
-raw[raw == ""] <- NA
-
-# Figure out the universities outside of the U.S. and Canada to drop
-# universities <- raw %.%
-#   group_by(university) %.%
-#   summarise(count = length(id)) %.%
-#   arrange(university)
-# 
-# write.csv(universities, "temp/universities.csv")
-
-# Universities to delete
-universities_to_delete <- c(
-  "Adelphi University, School of Social Work.",
-  "Adelphi University, The Institute of Advanced Psychological Studies.",
-  "Adler School of Professional Psychology.",
-  "Ahmadu Bello University (Nigeria).",
-  "Allegheny University of Health Sciences.",
-  "American Conservatory of Music.",
-  "American School of Professional Psychology - Rosebridge.",
-  "American Conservatory of Music.",
-  "American School of Professional Psychology - Rosebridge.",
-  "Bar-Ilan University (Israel).",
-  "Bogazici Universitesi (Turkey).",
-  "Boston University School of Education.",
-  "Brandeis University, The Heller School for Social Policy and Management.",
-  "Brandenburgische Technische Universitaet Cottbus (Germany).",
-  "Bryn Mawr College, Graduate School of Social Work and Social Research.",
-  "Cairo University (Egypt).",
-  "Caribbean Graduate School of Theology (Jamaica).",
-  "Case Western Reserve University (Health Sciences).",
-  "Centro de Estudios Avanzados de Puerto Rico y el Caribe (Puerto Rico).",
-  "Christian-Albrechts Universitaet zu Kiel (Germany).",
-  "Chuo Daigaku (Japan).",
-  "Columbia University Teachers College under the Faculty of Union Theological Seminary in the City of New York.",
-  "Duquesne University School of Nursing.",
-  "Eberhard Karls Universitaet Tuebingen (Germany).",
[…]

### 2014-09-16 R/predict-gender.R

@@ -1,0 +1,64 @@
+# We need to guess at the range of ages for book authors. 25 to 75 seems
+# sensible enough.
+min_age <- 25
+max_age <- 75
+
+# Now we will create columns for the minimum year and the maximum year for
+# guessing gender.
+ahr <- ahr %>%
+  mutate(min_year = issueDate - max_age,
+         max_year = issueDate - min_age)
+
+# Now we need to decide which method to use when predicting the gender. In
+# general we want to use IPUMS before 1930 and SSA after 1930, but sometimes
+# we'll have to compromise. There are several kinds of obvious cases:
+#
+# - If max_age < 1930, use IPUMS
+#
+# - If min_age > 1930, use SSA
+#
+# The tricky cases are when the terminal values overlap the ranges of one of the
+# methods.
+#
+# - If min_age > 1880 and max_age > 1930, use SSA
+#
+# If our range of ages was greater than 50, we might have a person who we
+# guessed was born before 1880 or after 1930. But by keeping our range smaller
+# than 50 (the difference between 1930 and 1880) we can avoid that problem.
+
+# That does indeed match all cases. So now we can write a function to return the
+# proper method for a given range of years.
+pick_method <- function(min_age, max_age) {
+  ifelse(max_age <= 1930, "ipums",
+         ifelse(min_age > 1930, "ssa",
+         ifelse(min_age > 1880 & max_age > 1930, "ssa", NA)))
+}
+
+# Now assign a method to the AHR dataset
+ahr <- ahr %>%
+  mutate(method = pick_method(min_year, max_year))
+
+pass_through <- function(x) { return(x) }
+
+apply_gender <- function(df) {
+  gender(df$first_name, method = df$method,
+         years = c(df$min_year, df$max_year))
+  }
+
+ahr_gender <- ahr %>%
+  group_by(first_name, vol_issue_ID, Author_Full, Reviewer_First,
+           Reviewer_Full, category) %>%
+  do(gender_author = apply_gender(.))
+
+# Finally we are ready to predict the gender. Because we have to combine years
+# we're going to do this as a for loop, which is ugly.
+# ahr_gender.l <- vector(mode = "list", nrow(ahr))
+
+# for(i in 1:nrow(ahr)) {
+#   row <- ahr[i, ]
+#   result <- gender(row$first_name, method = row$method,
+#                    years = c(row$min_year, row$max_year))
[…]

### 2014-09-16 R/load-data.R
Update data loading

@@ -3,10 +3,20 @@
 library(dplyr)
 library(stringr)
 
+extract_first_name_from_diss <- function(name) {
+  name %>%
+    tolower() %>%
+    str_extract(", [a-z'-]{2,}") %>%
+    str_extract("[a-z'-]{2,}") %>%
+    str_trim()
+}
+
 # The dissertations data has already been cleaned.
 # See https://github.com/lmullen/dissertations-data
+# But it needs a new column for the author first name.
 diss <- read.csv("data/h_diss2.csv", stringsAsFactors = FALSE) %>%
-  tbl_df()
+  tbl_df() %>%
+  mutate(author_first_name = extract_first_name_from_diss(author))
 
 # The AHR data has been converted to UTF8. For some entries the first name
 # column contains first and middle names, so we're going to parse out just the

### 2014-09-16 R/predict-gender.R
Script to predict genders faster

@@ -1,13 +1,29 @@
-# We need to guess at the range of ages for book authors. 25 to 75 seems
-# sensible enough.
-min_age <- 25
-max_age <- 75
+library(gender)
+library(dplyr)
+library(tidyr)
+
+# The gender function is not particular fast. Okay, it is downright slow. So
+# it's important that we predict gender as few times as possible, then cache the
+# result. The best way to run the function as few times as possible. So if there
+# are two "charles" authors in the same year, then we can create a summarized
+# data frame of just the unique combinations of names, and then merge the
+# results back into the origional data frame.
+
+# We'll start with just the dissertations. This gives us about 35K observations
+# instead of about 80K.
+diss_gender <- diss %>%
+  group_by(author_first_name, year) %>%
+  summarize(n = n())
+
+# We need to provide a range of years in which we think the person was born.
+min_dissertator_age <- 25
+max_dissertator_age <- 45
 
 # Now we will create columns for the minimum year and the maximum year for
 # guessing gender.
-ahr <- ahr %>%
-  mutate(min_year = issueDate - max_age,
-         max_year = issueDate - min_age)
+diss_gender <- diss_gender %>%
+  mutate(min_year = year - max_dissertator_age,
+         max_year = year - min_dissertator_age)
 
 # Now we need to decide which method to use when predicting the gender. In
 # general we want to use IPUMS before 1930 and SSA after 1930, but sometimes
@@ -34,31 +50,62 @@ pick_method <- function(min_age, max_age) {
          ifelse(min_age > 1880 & max_age > 1930, "ssa", NA)))
 }
 
-# Now assign a method to the AHR dataset
-ahr <- ahr %>%
+# Now assign a method to the diss_gender dataset
+diss_gender <- diss_gender %>%
   mutate(method = pick_method(min_year, max_year))
 
-pass_through <- function(x) { return(x) }
-
 apply_gender <- function(df) {
-  gender(df$first_name, method = df$method,
+  gender(df$author_first_name, method = df$method,
          years = c(df$min_year, df$max_year))
-  }
+}
 
+# Now do the prediction, caching the results
+diss_results_fn <- "data/diss-gender-results.rda"
+if(!file.exists(diss_results_fn)) {
+  diss_gender_results <- diss_gender %>%
+      rowwise() %>%
+      do(gender_author = apply_gender(.))
[…]

### 2014-09-17 article.Rmd
Works Cited

@@
 
-Landrum, in progress.
+Landrum 2014.
 Landrum, Shane.
-"The State's Big Family Bible: Birth Certificates, Personal Identity, and Citizenship in the United States, 1840--1950." PhD dissertation, Brandeis University, in progress.
+"The State's Big Family Bible: Birth Certificates, Personal Identity, and Citizenship in the United States, 1840--1950." PhD dissertation, Brandeis University, August 2014.
 

### 2014-09-18 R/predict-gender.R
Extract values from predictions

@@ -109,3 +109,17 @@ if(!file.exists(ahr_results_fn)) {
 } else {
   load(ahr_results_fn)
 }
+
+# The previous operation returns a data frame of lists, so lets get just the
+# lists, then turn them into a data frame. Also, since we neglected to save the
+# year of the dissertation or book, which we'll need for the merge, we have to
+# recalculate that from our assumptions about dissertator or author age.
+diss_gender_to_merge <- diss_gender_results$gender_author %>%
+  do.call(rbind.data.frame, .) %>%
+  mutate(year = year_min + max_dissertator_age) %>%
+  tbl_df()
+
+ahr_gender_to_merge <- ahr_gender_results$gender_author %>%
+  do.call(rbind.data.frame, .) %>%
+  mutate(year = year_min + max_author_age) %>%
+  tbl_df()

### 2014-09-18 R/predict-gender.R
Merge and save the predicted data frames

@@ -123,3 +123,30 @@ ahr_gender_to_merge <- ahr_gender_results$gender_author %>%
   do.call(rbind.data.frame, .) %>%
   mutate(year = year_min + max_author_age) %>%
   tbl_df()
+
+# We need to merge the dissertations and AHR datasets to get the genders in
+# those data sets. We actually need to do this twice for the AHR, because we
+# have both the reviewer name and the author name to predict. We'll also select just the minimum data we need to do our analysis. Then we'll cache the results.
+diss_predicted <- diss %>%
+  merge(diss_gender_to_merge,
+        by.x = c("author_first_name", "year"),
+        by.y = c("name", "year")) %>%
+  tbl_df() %>%
+  select(author, title, gender_author = gender, year)
+
+ahr_predicted <- ahr %>%
+  merge(ahr_gender_to_merge,
+        by.x = c("author_first_name", "issueDate"),
+        by.y = c("name", "year")) %>%
+  tbl_df() %>%
+  mutate(gender_author = gender) %>%
+  merge(ahr_gender_to_merge,
+        by.x = c("reviewer_first_name", "issueDate"),
+        by.y = c("name", "year")) %>%
+  mutate(gender_reviewer = gender.y) %>%
+  select(author = Author_Full, category, title, reviewer = Reviewer_Full,
+         year = issueDate, gender_author, gender_reviewer)
+
+# Save the files
+save(diss_predicted, file = "data/diss-predicted.rda", compress = TRUE)
+save(ahr_predicted, file = "data/ahr-predicted.rda", compress = TRUE)

### 2014-09-18 article.Rmd
Remove escape sequence that is breaking syntax highlighting

--- 2014-09-17 article.Rmd
@@
 
-As humanities researchers grapple with larger and larger datasets they will increasingly be able to not just extract existing information from that data but to also infer additional information from it \[Goldstone and Underwood 2014]\.
+As humanities researchers grapple with larger and larger datasets they will increasingly be able to not just extract existing information from that data but to also infer additional information from it [Goldstone and Underwood 2014].
 ^[The availability of large-scale datasets dealing with personal information raises important ethical issues for digital humanists, particularly in light of the revelations about domestic spying at the NSA brought forth by Edward Snowden in 2013.
-\[Weingart 2014]\ Kieran Healy has used network analysis to demonstrate how researchers can infer information from datasets while also touching on the troubling ethical issue of surveillance \[Kiely 2013]\.] Many of these datasets deal with lists of people: company payrolls, military rosters, passenger bookings, records of correspondence, or lists of published works.
+[Weingart 2014] Kieran Healy has used network analysis to demonstrate how researchers can infer information from datasets while also touching on the troubling ethical issue of surveillance [Kiely 2013].] Many of these datasets deal with lists of people: company payrolls, military rosters, passenger bookings, records of correspondence, or lists of published works.
 Some of these records might contain associated information such a person's address, age, or rank, while others contain little more than a list of personal names.
@@
 Anyone studying a period longer than a few years, or anyone studying a group whose demographics do not match those of a contemporary tool like Genderize.io will also encounter this problem.
-As of 2012, the average American has a lifespan of nearly seventy-nine years–--more than enough time for naming practices to change quite dramatically between their birth and death \[World Bank 2014]\.
+As of 2012, the average American has a lifespan of nearly seventy-nine years–--more than enough time for naming practices to change quite dramatically between their birth and death [World Bank 2014].
 Predicting gender from first names therefore requires a fundamentally historical method.
@@
 
-We recognize that gender is a fluid concept, both today and in the past, and that this method glosses over the complexities of historical definitions of gender and of the relationship between biological sex and gender \[Butler 1990, Butler 1993]\.
+We recognize that gender is a fluid concept, both today and in the past, and that this method glosses over the complexities of historical definitions of gender and of the relationship between biological sex and gender [Butler 1990, Butler 1993].
 In some sense this is unavoidable since the large historical datasets which make this kind of analysis possible were created by state agencies.
-For the category of gender (as for race) states have offered a limited number of boxes to check, and those boxes extert a tremendous power to define state subjects \[Brubaker and Cooper 2000, 14-17, Pascoe 2009, Canaday 2011, Landrum 2014]\.
+For the category of gender (as for race) states have offered a limited number of boxes to check, and those boxes extert a tremendous power to define state subjects [Brubaker and Cooper 2000, 14-17, Pascoe 2009, Canaday 2011, Landrum 2014].
 This method does not presuppose that the gender binary imposed by governments and their wider societies reflect how an individual identifies oneself and is neutral as to what the categories "male" or "female" actually mean in any given historical context.
@@
 Let's start with a comparison.
-One existing method for predicting gender is available in the [Natural Language Toolkit](http://www.nltk.org/) for Python \[Bird et al.
-2009, Chapter 2]\.^[Chapter 6 of \[Bird et al.
-2009] also describes a method of predicting gender from the last letter of the name, but this example is given for the sake of explaining supervised classification and not as an end in itself, so we do not have reason to address this a historical method.] The NLTK is an influential software package for scholarship because it provides an extraordinary range of tools for analyzing natural language.^[\[Bird et al.
-2009]\ has been cited 812 times according to [Google Scholar](http://scholar.google.com/scholar?cites=3735380965977150091&as_sdt=5,47&sciodt=0,47&hl=en).
+One existing method for predicting gender is available in the [Natural Language Toolkit](http://www.nltk.org/) for Python [Bird et al.
+2009, Chapter 2].^[Chapter 6 of [Bird et al.
+2009] also describes a method of predicting gender from the last letter of the name, but this example is given for the sake of explaining supervised classification and not as an end in itself, so we do not have reason to address this a historical method.] The NLTK is an influential software package for scholarship because it provides an extraordinary range of tools for analyzing natural language.^[[Bird et al.
+2009] has been cited 812 times according to [Google Scholar](http://scholar.google.com/scholar?cites=3735380965977150091&as_sdt=5,47&sciodt=0,47&hl=en).
 Because in the humanities (unlike in other fields) the practice of citing software (even software which determines the outcome of an analysis) is regrettably rare, even nonexistent, such metrics are difficult to determine.
@@
 According to the documentation for its API, "It utilizes big datasets of information, from user profiles across major social networks." It would be anachronistic to apply these datasets to the past, and Genderize.io provides no functionality to filter results chronologically as it does geographically.
-In addition, Genderize.io does not make clear exactly what comprises the dataset and how it was gathered, which keeps scholars from interrogating the value of the source \[Genderize.io 2014-]\.^[Quotations are taken from the [Genderize.io website](http://genderize.io), with correspondence from the developer behind the site for confirmation.]
+In addition, Genderize.io does not make clear exactly what comprises the dataset and how it was gathered, which keeps scholars from interrogating the value of the source [Genderize.io 2014-].^[Quotations are taken from the [Genderize.io website](http://genderize.io), with correspondence from the developer behind the site for confirmation.]
 
-R and Python are two of the most commonly used languages for data analysis \[Piatesky 2014, Muenchen 2014-]\.^[Surveys and analysis of useage point out the growth of R and the continuing popularity of Python among "data science" generally.] Python's existing packages for gender prediction all implement a method similar to the NLTK or Genderize.io.^[**Python library names**] To our knowledge, [CRAN](http://cran.rstudio.com/) (the central repository for R packages) does not include any packages dedicated to gender prediction.
+R and Python are two of the most commonly used languages for data analysis [Piatesky 2014, Muenchen 2014-].^[Surveys and analysis of useage point out the growth of R and the continuing popularity of Python among "data science" generally.] Python's existing packages for gender prediction all implement a method similar to the NLTK or Genderize.io.^[**Python library names**] To our knowledge, [CRAN](http://cran.rstudio.com/) (the central repository for R packages) does not include any packages dedicated to gender prediction.
 In two of the most popular data analysis languages used in the digital humanities there is currently no satisfactory existing method of gender prediction for historical and humanistic research.
@@
 The Gender package is affiliated with [rOpenSci](http://ropensci.org/), an initiative that supports reproducible research and open data for scientists using R.
-The rOpenSci advisory board has provided code review and publicity for the project.^[The gender package depends especially on \[R Core Team 2014-, Wickham and Francois 2014]\ ]
+The rOpenSci advisory board has provided code review and publicity for the project.^[The gender package depends especially on [R Core Team 2014-, Wickham and Francois 2014] ]
 
@@
 The Social Security Administration (SSA) "[Baby Names](http://catalog.data.gov/dataset/baby-names-from-social-security-card-applications-national-level-data)" dataset was created as a result of the Social Security Act of 1935 during the New Deal.
-^[The Social Security Administration also provides a related dataset which contains the names of applicants broken down by state and year since 1910 \[Social Security Administration 2014-b]\.] This dataset contains a list of first names with the number of male and female uses of the name in every year beginning with 1880 and ending with 2012.
+^[The Social Security Administration also provides a related dataset which contains the names of applicants broken down by state and year since 1910 [Social Security Administration 2014-b].] This dataset contains a list of first names with the number of male and female uses of the name in every year beginning with 1880 and ending with 2012.
 The SSA list includes any name which was used more than five times in a given year, thereby capturing all but the most infrequently used names from each year.
@@
 Even after 1935, registration at birth for Social Security was not mandatory until 1986.
-As we will demonstrate below, the way in which the data was gathered requires an adjustment to our predictions of gender.^[Shane Landrum has studied the history of birth registration in the United States \[Landrum 2014]\.
+As we will demonstrate below, the way in which the data was gathered requires an adjustment to our predictions of gender.^[Shane Landrum has studied the history of birth registration in the United States [Landrum 2014].
 It is well-known among American historians that the Social Security Act was written to exclude farmworkers as a means of also excluding African Americans.
@@
 Unlike the SSA dataset, which includes 100% of every name reported to the Social Security Administration **not quite - only names that occur more than five times, right?**, the IPUMS data contains 5% or 10% samples.
-Because the `gender()` function relies on proportions of uses of names, rather than raw counts of people with the names, the sampling does not diminish the reliability of the function's predicitons \[Ruggles et al.
-2010]\.
+Because the `gender()` function relies on proportions of uses of names, rather than raw counts of people with the names, the sampling does not diminish the reliability of the function's predicitons [Ruggles et al.
+2010].
 
@@
 ^[The format of this data departs from the more generally desirable "tidy data" framework proposed by Hadley Wickham for performance reasons.
-By keeping the data in a "wider" format instead of a tidy format as defined by Wickham, the function is able to make its predictions faster \[Wickham forthcoming]\.] The columns contain the `name`, `year`, and number of `female` and `male` uses of that name in a particular year.
+By keeping the data in a "wider" format instead of a tidy format as defined by Wickham, the function is able to make its predictions faster [Wickham forthcoming].] The columns contain the `name`, `year`, and number of `female` and `male` uses of that name in a particular year.
 
@@
 Schroeder, and Matthew Sobek.
-Integrated Public Use Microdata Series: Version 5.0 \[Machine-readable database]\.
-Minneapolis, MN: Minnesota Population Center \[producer and distributor]\, 2010.
+Integrated Public Use Microdata Series: Version 5.0 [Machine-readable database].
+Minneapolis, MN: Minnesota Population Center [producer and distributor], 2010.
 

### 2014-09-18 article.Rmd
Load the cached data
Remove sample IPUMS data

@@
 
+# Load the data frames with the predicted gender, if they exist. Otherwise run
+# the analysis.
+if(file.exists("data/diss-predicted.rda") &
+     file.exists("data/ahr-predicted.rda")) {
+  load("data/diss-predicted.rda")
+  load("data/ahr-predicted.rda")
+} else {
+  source("R/load-data.R")
+  source("predict-gender.R")
+}
 ```
@@
 genderdata::ssa_national
-```
-
-**unfortunatley the first ten isn't intuitive since it's just "!" rather than names - this is going to confuse readers, I think**
-```{r ipums_usa, echo=TRUE, results='markup'}
-genderdata::ipums_usa
 ```

### 2014-09-18 article.Rmd
Plot of dissertation counts by gender

@@
 
-![Number of dissertations by gender](img/plot-gender.png)
+```{r}
+short_scale <- scale_x_continuous(breaks = seq(1950, 2010, 10),
+                                  labels = seq(1950, 2010, 10),
+                                  limits = c(1950, 2012))
+
+diss_predicted %>%
+  ggplot(aes(x = year, fill = gender_author)) + 
+  geom_bar(binwidth = 1) +
+  ggtitle("Dissertations about History by Gender, 1950-2012") +
+  xlab(NULL) + ylab(NULL) +
+  short_scale + theme(legend.position = "bottom")
+```
 

### 2014-09-18 article.Rmd
Proportion of dissertations

--- 2014-09-18 article.Rmd
@@
 library(scales)
+library(tidyr)
 
@@
 
-![Number of dissertations by gender](img/plot-gender.png)
+```{r}
+short_scale <- scale_x_continuous(breaks = seq(1950, 2010, 10),
+                                  labels = seq(1950, 2010, 10),
+                                  limits = c(1950, 2012))
+
+diss_predicted %>%
+  ggplot(aes(x = year, fill = gender_author)) + 
+  geom_bar(binwidth = 1) +
+  ggtitle("Dissertations about History by Gender, 1950-2012") +
+  xlab(NULL) + ylab(NULL) +
+  short_scale + theme(legend.position = "bottom") +
+  scale_fill_discrete(name = "",
+                      labels = c("Female", "Male"))
+```
 
@@
 
-![Proportion of dissertations for each gender](img/dissertations-by-gender-proportion.png)
+```{r}
+diss_predicted %>%
+  select(year, gender_author) %>%
+  filter(year >= 1950) %>%
+  group_by(year, gender_author) %>%
+  summarize(n = n()) %>%
+  spread(gender_author, n) %>% 
+  mutate(female_proportion = female / (female + male),
+         male_proportion = male / (female + male)) %>% 
+  select(female_proportion, male_proportion, year) %>%
+  gather(gender, proportion, -year) %>% 
+  ggplot(aes(x = year, y = proportion, fill = gender)) + 
+  geom_bar(binwidth = 1, stat = "identity") +
+  ggtitle("Proportion of Dissertations about History by Gender, 1950-2012") +
+  xlab(NULL) + ylab(NULL) +
+  short_scale + theme(legend.position = "bottom") +
+  geom_hline(yintercept = 0.5) +
+  scale_fill_discrete(name = "",
+                      labels = c("Female", "Male"))
+```
 

### 2014-09-18 R/load-data.R
Remove factors

@@ -1,5 +1,5 @@
 # Load the data and do any cleaning
-
+options(stringsAsFactors = FALSE)
 library(dplyr)
 library(stringr)
 

### 2014-09-18 R/predict-gender.R

@@ -1,3 +1,4 @@
+options(stringsAsFactors = FALSE)
 library(gender)
 library(dplyr)
 library(tidyr)

### 2014-09-18 article.Rmd
AHR first chart, chart theme

--- 2014-09-18 article.Rmd
@@
 ```{r load-libraries, include=FALSE}
+options(stringsAsFactors = FALSE)
 library(genderdata)
@@
 }
+
+diss_predicted$gender_author <- diss_predicted$gender_author %>% as.character()
+ahr_predicted$gender_author <- ahr_predicted$gender_author %>% as.character()
+ahr_predicted$gender_reviewer <- ahr_predicted$gender_reviewer %>% as.character()
+
+diss_predicted[is.na(diss_predicted)] <- "unknown"
+ahr_predicted[is.na(ahr_predicted)] <- "unknown"
+
+# Specify article wide chart options
+short_scale <- scale_x_continuous(breaks = seq(1950, 2010, 10),
+                                  labels = seq(1950, 2010, 10),
+                                  limits = c(1950, 2012))
+
+article_theme <- theme(legend.title = element_blank())
 ```
@@
 
-![Number of dissertations by gender](img/plot-gender.png)
+**Do we even need this chart, or is the chart of proportions sufficient?**
+
+```{r}
+diss_predicted %>%
+  ggplot(aes(x = year, fill = gender_author)) + 
+  geom_bar(binwidth = 1) +
+  ggtitle("Dissertations about History by Gender, 1950-2012") +
+  xlab(NULL) + ylab(NULL) +
+  short_scale + theme(legend.position = "bottom") +
+  scale_fill_discrete(name = "",
+                      labels = c("Female", "Male")) +
+  article_theme
+```
 
@@
 
-![Proportion of dissertations for each gender](img/dissertations-by-gender-proportion.png)
+```{r}
+diss_predicted %>%
+  select(year, gender_author) %>%
+  filter(year >= 1950) %>%
+  group_by(year, gender_author) %>% 
+  summarize(n = n()) %>% 
+  spread(gender_author, n, fill = 0) %>% 
+  mutate(female_proportion = female / (female + male + unknown),
+         male_proportion = male / (female + male + unknown),
+         unknown_proportion = unknown / (female + male + unknown)) %>% 
+  select(female_proportion, male_proportion, unknown_proportion, year) %>%
+  gather(gender, proportion, -year) %>% 
+  ggplot(aes(x = year, y = proportion, fill = gender)) +
+  geom_bar(binwidth = 1, stat = "identity") +
+  ggtitle("Proportion of Dissertations about History by Gender, 1950-2012") +
+  xlab(NULL) + ylab(NULL) +
+  short_scale + theme(legend.position = "bottom") +
+  geom_hline(yintercept = 0.5) +
+  scale_fill_discrete(labels = c("Female", "Male", "Unknown")) +
+  article_theme
+```
 
@@
 
-![Gender of authors of books reviewed in the AHR](img/book-authors.jpg)
+```{r}
+ahr_predicted %>%
+  select(year, gender_author) %>%  
+  group_by(year, gender_author) %>% 
+  summarize(n = n()) %>% 
+  spread(gender_author, n, fill = 0) %>%
+  mutate(female_proportion = female / (female + male + unknown),
+         male_proportion = male / (female + male + unknown),
+         unknown_proportion = unknown / (female + male + unknown)) %>%  
+  select(female_proportion, male_proportion, unknown_proportion, year) %>%
+  gather(gender, proportion, -year) %>% 
+  ggplot(aes(x = year, y = proportion, fill = gender)) + 
+  geom_bar(binwidth = 1, stat = "identity") +
+  ggtitle("Authors of Books Reviewed in the AHR by Gender") +
+  xlab(NULL) + ylab(NULL) +
+  theme(legend.position = "bottom") +
+  scale_fill_discrete(labels = c("Female", "Male")) +
+  article_theme
+```
 

### 2014-09-18 article.Rmd
Plot of the gap

@@
 ```{r}
-diss_predicted %>%
+diss_proportions <- diss_predicted %>%
   select(year, gender_author) %>%
@@
   select(female_proportion, male_proportion, unknown_proportion, year) %>%
-  gather(gender, proportion, -year) %>% 
+  gather(gender, proportion, -year)
+
+diss_proportions  %>%
   ggplot(aes(x = year, y = proportion, fill = gender)) +
@@
 ```{r}
-ahr_predicted %>%
+ahr_proportions <- ahr_predicted %>%
   select(year, gender_author) %>%  
@@
   select(female_proportion, male_proportion, unknown_proportion, year) %>%
-  gather(gender, proportion, -year) %>% 
+  gather(gender, proportion, -year)
+
+ahr_proportions %>%
   ggplot(aes(x = year, y = proportion, fill = gender)) + 
@@
   theme(legend.position = "bottom") +
-  scale_fill_discrete(labels = c("Female", "Male")) +
+  scale_fill_discrete(labels = c("Female", "Male", "Unknown")) +
   article_theme
@@
 ![Gap between female authorship of dissertations and books](img/book-dissertation-gap.jpg)
+
+```{r}
+diss_proportions$type <- "dissertation"
+ahr_proportions$type <- "book"
+
+combined_proportions <- rbind(diss_proportions, ahr_proportions)
+
+combined_proportions %>%
+  filter(gender != "unknown_proportion") %>%
+  ggplot(aes(x = year, y = proportion, color = gender, size = type)) + 
+  geom_line() +
+  short_scale +
+  scale_size_discrete(name = "Type",
+                      labels = c("Book", "Dissertation"), range = c(2,1)) +
+  scale_color_discrete(name = "Gender",
+                       labels = c("Female", "Male")) +
+  xlab(NULL) + ylab("Proportion of authors by gender") +
+  ggtitle("The gap between between authorship of dissertations and books reviewed in the AHR ")
+```
 

### 2014-09-18 article.Rmd
Remove static image
Make charts wider and more consistent

@@
 })
+
+knitr::opts_chunk$set(fig.width = 8, fig.height = 5)
 
@@
   theme(legend.position = "bottom") +
+  geom_hline(yintercept = 0.5) +
   scale_fill_discrete(labels = c("Female", "Male", "Unknown")) +
@@
 The long-term gains made by female doctoral students do not carry over in the pages of the historical profession's leading journal.
-
-![Gap between female authorship of dissertations and books](img/book-dissertation-gap.jpg)
 
@@
   short_scale +
+  geom_hline(yintercept = 0.5) +
   scale_size_discrete(name = "Type",
@@
   xlab(NULL) + ylab("Proportion of authors by gender") +
-  ggtitle("The gap between between authorship of dissertations and books reviewed in the AHR ")
+  ggtitle("The gap between the genders of book and dissertation authors")
 ```

### 2014-09-18 article.Rmd
Faceted by subject category

@@
   short_scale +
-  geom_hline(yintercept = 0.5) +
   scale_size_discrete(name = "Type",
@@
 
-![Gender of reviewers of books authored by women](img/female_authors_withreviewers.jpg)
+```{r}
+ahr_predicted %>%
+  filter(gender_author == "female") %>%
+  select(year, gender_reviewer) %>%  
+  group_by(year, gender_reviewer) %>% 
+  summarize(n = n()) %>% 
+  spread(gender_reviewer, n, fill = 0) %>%
+  mutate(female_proportion = female / (female + male + unknown),
+         male_proportion = male / (female + male + unknown),
+         unknown_proportion = unknown / (female + male + unknown)) %>%  
+  select(female_proportion, male_proportion, unknown_proportion, year) %>%
+  gather(gender, proportion, -year) %>%
+  ggplot(aes(x = year, y = proportion, fill = gender)) + 
+  geom_bar(binwidth = 1, stat = "identity") +
+  ggtitle("Gender of reviewers of female book authors in AHR") +
+  xlab(NULL) + ylab(NULL) +
+  theme(legend.position = "bottom") +
+  geom_hline(yintercept = 0.5) +
+  scale_fill_discrete(labels = c("Female", "Male", "Unknown")) +
+  article_theme +
+  short_scale
+```
 
@@
 ![Gender of reviewers by category of review](img/categories_gender_bytime.jpg)
+
+**Do you have the mapping you use from the AHR's 68 categories to the 8 or so you used? With that, I just have one or two more lines to facet this plot.**
+
+```{r}
+ahr_predicted %>%
+  filter(gender_author == "female") %>%
+  select(year, gender_reviewer, category) %>% 
+  group_by(year, gender_reviewer, category) %>%  
+  summarize(n = n()) %>%
+  spread(gender_reviewer, n, fill = 0) %>% 
+  mutate(female_proportion = female / (female + male + unknown),
+         male_proportion = male / (female + male + unknown),
+         unknown_proportion = unknown / (female + male + unknown)) %>%  
+  select(female_proportion, male_proportion, unknown_proportion,
+         year, category) %>% 
+  gather(gender, proportion, -year, -category) -> temp
+ggplot(aes(x = year, y = proportion, fill = gender)) + 
+  geom_bar(binwidth = 1, stat = "identity") +
+  ggtitle("Gender of reviewers of female book authors in AHR") +
+  xlab(NULL) + ylab(NULL) +
+  theme(legend.position = "bottom") +
+  geom_hline(yintercept = 0.5) +
+  scale_fill_discrete(labels = c("Female", "Male", "Unknown")) +
+  article_theme +
+  short_scale +
+```
 

### 2014-09-18 article.Rmd
Color Palette

--- 2014-09-18 article.Rmd
@@
 library(tidyr)
+library(ggthemes)
+library(RColorBrewer)
 
@@
 article_theme <- theme(legend.title = element_blank())
+
+article_palette <- "Dark2"
 ```
@@
   short_scale + theme(legend.position = "bottom") +
-  scale_fill_discrete(name = "",
+  scale_fill_brewer(palette = article_palette, name = "",
                       labels = c("Female", "Male")) +
@@
   geom_hline(yintercept = 0.5) +
-  scale_fill_discrete(labels = c("Female", "Male", "Unknown")) +
+  scale_fill_brewer(palette = article_palette,
+                    labels = c("Female", "Male", "Unknown")) +
   article_theme
@@
   geom_hline(yintercept = 0.5) +
-  scale_fill_discrete(labels = c("Female", "Male", "Unknown")) +
+  scale_fill_brewer(palette = article_palette,
+                      labels = c("Female", "Male", "Unknown")) +
   article_theme
@@
   short_scale +
-  geom_hline(yintercept = 0.5) +
   scale_size_discrete(name = "Type",
                       labels = c("Book", "Dissertation"), range = c(2,1)) +
-  scale_color_discrete(name = "Gender",
+  scale_color_brewer(palette = article_palette, name = "Gender",
                        labels = c("Female", "Male")) +
@@
 
-![Gender of reviewers of books authored by women](img/female_authors_withreviewers.jpg)
+```{r}
+ahr_predicted %>%
+  filter(gender_author == "female") %>%
+  select(year, gender_reviewer) %>%  
+  group_by(year, gender_reviewer) %>% 
+  summarize(n = n()) %>% 
+  spread(gender_reviewer, n, fill = 0) %>%
+  mutate(female_proportion = female / (female + male + unknown),
+         male_proportion = male / (female + male + unknown),
+         unknown_proportion = unknown / (female + male + unknown)) %>%  
+  select(female_proportion, male_proportion, unknown_proportion, year) %>%
+  gather(gender, proportion, -year) %>%
+  ggplot(aes(x = year, y = proportion, fill = gender)) + 
+  geom_bar(binwidth = 1, stat = "identity") +
+  ggtitle("Gender of reviewers of female book authors in AHR") +
+  xlab(NULL) + ylab(NULL) +
+  theme(legend.position = "bottom") +
+  geom_hline(yintercept = 0.5) +
+  scale_fill_brewer(palette = article_palette, 
+                    labels = c("Female", "Male", "Unknown")) +
+  article_theme +
+  short_scale
+```
 
@@
 
-![Gender of reviewers by category of review](img/categories_gender_bytime.jpg)
+**Cameron: Do you have the mapping you use from the AHR's 68 categories to the 8 or so you used?
+With that, I just have one or two more lines to facet this plot.
+(It doesn't look right because it's not faceted yet.)**
+
+```{r}
+ahr_predicted %>%
+  filter(gender_author == "female") %>%
+  select(year, gender_reviewer, category) %>% 
+  group_by(year, gender_reviewer, category) %>%  
+  summarize(n = n()) %>%
+  spread(gender_reviewer, n, fill = 0) %>% 
+  mutate(female_proportion = female / (female + male + unknown),
+         male_proportion = male / (female + male + unknown),
+         unknown_proportion = unknown / (female + male + unknown)) %>%  
+  select(female_proportion, male_proportion, unknown_proportion,
+         year, category) %>% 
+  gather(gender, proportion, -year, -category) %>%
+ggplot(aes(x = year, y = proportion, fill = gender)) + 
+  geom_bar(binwidth = 1, stat = "identity") +
+  ggtitle("Gender of reviewers of female book authors in AHR") +
+  xlab(NULL) + ylab(NULL) +
+  theme(legend.position = "bottom") +
+  geom_hline(yintercept = 0.5) +
+  scale_fill_brewer(palette = article_palette,
+                    labels = c("Female", "Male", "Unknown")) +
+  article_theme +
+  short_scale
+```
 

### 2014-09-19 article.Rmd
Minor change to abstract

@@
 It outlines a new method for algorithmically inferring the gender of a name from historical data in order to take into account how naming practices change over time and space.
-In contrast to existing methods of predicting gender, this method provides a higher level of accuracy and a precise statement of the likelihood that a name is a given gender.
+In contrast to existing methods of predicting gender, this method provides a higher level of accuracy and a precise statement of the likelihood that a name was a given gender.
 The article describes the methodology as implemented in the [Gender](https://github.com/ropensci/gender) package for the [R programming language](http://www.r-project.org/).

### 2014-09-19 article.Rmd
Stylistic changes to the introduction

@@
 ^[The availability of large-scale datasets dealing with personal information raises important ethical issues for digital humanists, particularly in light of the revelations about domestic spying at the NSA brought forth by Edward Snowden in 2013.
-[Weingart 2014] Kieran Healy has used network analysis to demonstrate how researchers can infer information from datasets while also touching on the troubling ethical issue of surveillance [Kiely 2013].] Many of these datasets deal with lists of people: company payrolls, military rosters, passenger bookings, records of correspondence, or lists of published works.
+[Weingart 2014] Kieran Healy has used network analysis to demonstrate how researchers can infer information from datasets while also touching on the troubling ethical issue of surveillance [Healy 2013].] Many of these datasets deal with lists of people: company payrolls, military rosters, passenger bookings, records of correspondence, or lists of published works.
 Some of these records might contain associated information such a person's address, age, or rank, while others contain little more than a list of personal names.
@@
 "Jane Fay" is almost certainly a woman, while her brother "John Fay" is almost certainly a man.
-A straightforward computer program can infer their respective genders with relative ease.
+Given a dataset matching names to genders, a straightforward computer program can easily infer their respective genders.
 
@@
 But if the three siblings were born in the early twentieth century, chances are good that Leslie would be their brother.
-That is because the gendered naming convention surrounding "Leslie reversed course completely over the course of one hundred years.
-92% of the babies born in the United States in 1900 who were named "Leslie" were male, while 96% of those identically named babies born in the year 2000 were female.
+That is because the conventional gender for the name Leslie switched over the course of one hundred years.
+In 1900 some 92% of the babies born in the United States who were named Leslie were male, while in 2000 about 96% of the Leslies born in that year were female.
 
-```{r leslie}
+```{r plot-leslie}
 # We cannot simply use the Social Security figures (even though they probably

### 2014-09-19 article.Rmd
Use theme_few() from ggthemes

@@
 
-article_theme <- theme(legend.title = element_blank())
+# I like 
+article_theme <- theme(legend.title = element_blank()) + theme_few()
 
@@
   geom_point() +
-  ggtitle("Proportion of female uses of the name 'Leslie,' 1820-2012") +
+  ggtitle("Proportion of female uses of the name Leslie, 1820-2012") +
   xlab(NULL) + ylab(NULL) + ylim(0, 1) +
@@
   annotate("text", x = 1900, y = 0.45, label = "Male") +
-  annotate("text", x = 1900, y = 0.55, label = "Female")
+  annotate("text", x = 1900, y = 0.55, label = "Female") +
+  article_theme
 
@@
   annotate("text", x = 1965, y = 0.45, label = "Male") +
-  annotate("text", x = 1965, y = 0.55, label = "Female")
+  annotate("text", x = 1965, y = 0.55, label = "Female") +
+  article_theme
 
@@
   ggtitle("Percentage of SSA dataset that is female") +
-  xlab(NULL) + ylab(NULL)
+  xlab(NULL) + ylab(NULL) +
+  article_theme
 ```

### 2014-09-19 article.Rmd
Made first two plots' labels more precise

@@
 
-As humanities researchers grapple with larger and larger datasets they will increasingly be able to not just extract existing information from that data but to also infer additional information from it [Goldstone and Underwood 2014].
-^[The availability of large-scale datasets dealing with personal information raises important ethical issues for digital humanists, particularly in light of the revelations about domestic spying at the NSA brought forth by Edward Snowden in 2013.
+As humanities researchers grapple with larger and larger datasets they will increasingly be able to not just extract existing information from that data but to also infer additional information from it [Goldstone and Underwood 2014].^[The availability of large-scale datasets dealing with personal information raises important ethical issues for digital humanists, particularly in light of the revelations about domestic spying at the NSA brought forth by Edward Snowden in 2013.
 [Weingart 2014] Kieran Healy has used network analysis to demonstrate how researchers can infer information from datasets while also touching on the troubling ethical issue of surveillance [Healy 2013].] Many of these datasets deal with lists of people: company payrolls, military rosters, passenger bookings, records of correspondence, or lists of published works.
@@
   geom_point() +
-  ggtitle("Proportion of female uses of the name Leslie, 1820-2012") +
-  xlab(NULL) + ylab(NULL) + ylim(0, 1) +
+  ggtitle("Gender of the name Leslie, 1820-2012") +
+  xlab(NULL) + ylab("Proportion of female uses") + ylim(0, 1) +
   geom_hline(yintercept = 0.5) + 
@@
 
-Other names used in the United States such as Madison, Morgan, Sydney, and Kendall, have also "flipped" their gender over the past century.
+Other names used in the United States such as Madison, Morgan, Sydney, and Kendall, have also flipped their gender over the past century.
 Such changes usually follow set patterns.
-In general, names tend to shift from being predominantly male to being predominantly female; the shift rarely moves in the other direction.
+In general, names tend to shift from being predominantly male to being predominantly female; the shift rarely happens in the other direction.
 In general, names that transition from male to female also became increasingly popular over time.
@@
 Whereas before it had been an infrequently used male name, after 1985 it skyrocketed in popularity and by the early 2000s was one of the ten most frequently used names for baby girls.
-Broadly speaking, in the late twentieth century Americans girls became increasingly more likely than American boys to receive a "trendy" name, and once that name was associated with women it was unlikely to be given to male babies.
+Broadly speaking, in the late twentieth century Americans girls became increasingly more likely than American boys to receive a "trendy" name, and once that name was associated with females it was unlikely to be given to male babies.
 These names present a moving target for researchers hoping to use them to infer additional information about the people under study.
-^[The few exceptions of names which flip from being female to male, such as "Jan", have a different causal mechanism.
-It seems likely that Jan as an English name is female but that Jan as a Scandinavian name is male.
+In addition, the prevalance of changes in the gender of names in the last half of the twentieth century implies that gendered lists of names drawn from contemporary data are anachronistic and so of little use for earlier periods.^[The few exceptions of names which flip from being female to male, such as "Jan", have a different causal mechanism.
+We speculate that Jan as an English name is female but that Jan as a Scandinavian name is male.
 Because the name has remained relatively uncommon, other causes besides changing gender preferences can explain the change.]
 
-```{r plot_names_that_change}
+```{r plot-names-that-change}
 plot_names <-  c("madison",
@@
   geom_smooth(se = FALSE) +
-  ggtitle("Changes in selected names, 1930-2012") +
+  ggtitle("Changes in gender of selected names, 1930-2012") +
   xlim(1930, 2012) + 
-  xlab(NULL) + ylab(NULL) + ylim(0, 1) +
+  xlab(NULL) + ylab("Proportion of female uses") + ylim(0, 1) +
   geom_hline(yintercept = 0.5) +
@@
   annotate("text", x = 1965, y = 0.55, label = "Female") +
-  article_theme
+  article_theme +
+  scale_color_brewer(palette = article_palette, name = "")
 

### 2014-09-19 article.Rmd
More stylistic changes to introduction

@@
 An education policy analyst studying the demographics of different urban school districts, for instance, can turn to a variety of tools that use current databases of names, and these tools will likely identify the gender of Leslie Fay correctly.
-One online service, [Genderize.io](http://genderize.io/), determines the gender of first names "from user profiles across major social networks." Checking the service for the name Leslie returns a prediction that the name is female, along with an estimate of male/female proportions of the uses among members of unnamed social media sites.^[In this example the Genderize.io service has been accessed through the Gender package described below.]
+One online service, [Genderize.io](http://genderize.io/), determines the gender of first names "from user profiles across major social networks." Checking the service for the name Leslie returns a prediction that the name is female, along with an estimate of male/female proportions of the uses among members of unnamed social media sites.^[In this example the Genderize.io service has been accessed through an API wrapper provided in our Gender package described below.]
 
@@
 The "Leslie" problem is not just for researchers who think of themselves as historians.
-Anyone studying a period longer than a few years, or anyone studying a group whose demographics do not match those of a contemporary tool like Genderize.io will also encounter this problem.
-As of 2012, the average American has a lifespan of nearly seventy-nine years–--more than enough time for naming practices to change quite dramatically between their birth and death [World Bank 2014].
+Anyone studying a period longer than a few years, or anyone studying a group whose demographics do not match the groups used by a contemporary tool like Genderize.io will also encounter this problem.
+As of 2012, average Americans have a lifespan of nearly seventy-nine years---more than enough time for naming practices to change quite dramatically between their births and deaths [World Bank 2014].
 Predicting gender from first names therefore requires a fundamentally historical method.
@@
 For example, our program predicts that a person named Leslie born in 1950 is female, but only just barely: 52 percent of the babies named Leslie born in the United States in 1950 were girls and 48 percent were boys.
-Including these proportion figures allows researchers to conduct more nuanced analysis and to determine for themselves what proportion is acceptable for predicting the gender of names.
+Including these probability figures allows researchers to conduct more nuanced analysis and to determine for themselves what level of certainty is acceptable for predicting the gender of names.
 
@@
 This method does not presuppose that the gender binary imposed by governments and their wider societies reflect how an individual identifies oneself and is neutral as to what the categories "male" or "female" actually mean in any given historical context.
-Researchers are free---indeed, obligated---to make that determination for themselves, and this package provides access to the underlying dataset so that researchers can interrogate every assumption that comes with them.
+Researchers are free---indeed, obligated---to make that determination for themselves, and this package provides access to the underlying dataset so that researchers can interrogate every assumption that comes with it.
 This method does not argue against the complexities of historical understandings of gender, but as with any project it does require researchers who use the method to think through how the data was gathered, what it implies, and how it reflects social and cultural practices.

### 2014-09-19 article.Rmd
Stylistic changes
Stylistic changes to the section on Kantrowitz

@@
 First, we describe our method in more detail, compare it to existing methods, and explain how to use it.
-Second, we apply the method to a case study of gatekeeping in the historical profession in order to demonstrate its utility.
+Second, we apply the method to a case study of gatekeeping in the historical profession in order to demonstrate its usefulness.
 
@@
 Let's start with a comparison.
-One existing method for predicting gender is available in the [Natural Language Toolkit](http://www.nltk.org/) for Python [Bird et al.
+One existing method for predicting gender is available in the [Natural Language Toolkit](http://www.nltk.org/) for the Python programming language [Bird et al.
 2009, Chapter 2].^[Chapter 6 of [Bird et al.
-2009] also describes a method of predicting gender from the last letter of the name, but this example is given for the sake of explaining supervised classification and not as an end in itself, so we do not have reason to address this a historical method.] The NLTK is an influential software package for scholarship because it provides an extraordinary range of tools for analyzing natural language.^[[Bird et al.
+2009] also describes a method of predicting gender from the last letter of the name, but this example is given for the sake of explaining supervised classification and not as an end in itself, so we do not have reason to address this as a historical method.] The NLTK is an influential software package for scholarship because it provides an extraordinary range of tools for analyzing natural language.^[[Bird et al.
 2009] has been cited 812 times according to [Google Scholar](http://scholar.google.com/scholar?cites=3735380965977150091&as_sdt=5,47&sciodt=0,47&hl=en).
-Because in the humanities (unlike in other fields) the practice of citing software (even software which determines the outcome of an analysis) is regrettably rare, even nonexistent, such metrics are difficult to determine.
-Yet searching journals such as *Digital Humanities Quarterly* and the *Journal of Digital Humanities* reveals that natural language processing, for which the NLTK is the foremost implementation, is one of the key methodological categories of the digital humanities.] Included in the NLTK are two lists of male and female names created by the computer scientist Mark Kantrowitz with assistance from Bill Ross.^[The [original lists of names](http://www.cs.cmu.edu/afs/cs/project/ai-repository/ai/areas/nlp/corpora/names/0.html) are available at the website of the computer science department at Carnegie Mellon University, and the NLTK implementation is described in its [documentation](http://www.nltk.org/howto/corpus.html).] These lists include `r kantrowitz_unique` unique names.
-Using the Kantrowitz names corpus in the NLTK, one could look up a name like Jane or John and find out that it is male or female.
-The Kantrowitz corpus provides the list of names.^[The examples below is drawn from our Gender package which implements the Kantrowitz dataset for comparsion's sake, though its use is not recommended for actual research.
+Because in the humanities (unlike in other fields) the practice of citing software essential to performing scholarly analysis is regrettably rare, even nonexistent, such metrics are difficult to determine.
+Yet searching journals such as *Digital Humanities Quarterly* and the *Journal of Digital Humanities* reveals that natural language processing, for which the NLTK is the foremost implementation, is one of the key methodologies of the digital humanities.] Included in the NLTK are two lists of male and female names created by the computer scientist Mark Kantrowitz with assistance from Bill Ross.^[The [original lists of names](http://www.cs.cmu.edu/afs/cs/project/ai-repository/ai/areas/nlp/corpora/names/0.html) are available at the website of the computer science department at Carnegie Mellon University, and the NLTK implementation is described in its [documentation](http://www.nltk.org/howto/corpus.html).] These lists include `r kantrowitz_unique` unique names.
+Using the Kantrowitz names corpus in the NLTK, one could look up a name like Jane or John and find out whether it is male or female.
+
+The Kantrowitz corpus provides the list of names.^[The examples below is drawn from our Gender package which implements the Kantrowitz dataset for sake of comparsion, though its use is not recommended for actual research.
 The Gender package combines the Kantrowitz corpus's two lists into a single list, identifying names that are ambiguous because they appear on both lists as `"either"`.]
@@
 As pointed out above, this makes the corpus problematic for historical purposes.
-Furthermore the Kantrowitz corpus includes other oddities which make it less suitable.
-Some names like Abby are overwhelmingly female and some like Bill are overwhelmingly male, but the corpus includes them as both male and female.
-The Kantrowitz corpus contains only `r kantrowitz_unique` names, a mere `r percent(kantrowitz_unique / ssa_unique)` of the `r ssa_unique` names in the Social Security Administration dataset and `r percent(kantrowitz_unique / ipums_unique)` of the `r ipums_unique` names in the census records provided by Integrated Public Use Microdata (IPUMS) USA.
+Furthermore the Kantrowitz corpus includes other oddities which make it less useful for research.
+Some names such as Abby are overwhelmingly female and some such as Bill are overwhelmingly male, but the corpus includes them as both male and female.
+The Kantrowitz corpus contains only `r kantrowitz_unique` unique names, a mere `r percent(kantrowitz_unique / ssa_unique)` of the `r ssa_unique` unique names in a dataset provided by the Social Security Administration and `r percent(kantrowitz_unique / ipums_unique)` of the `r ipums_unique` unique names in the census records provided in the Integrated Public Use Microdata Series (IPUMS) USA dataset.
 There are therefore many names that it cannot identify.

### 2014-09-19 article.Rmd
Minor corrections; list Python packages

@@
 cf_kantrowitz_ssa <- data.frame(name = unique(genderdata::kantrowitz$name),
-                                stringsAsFactors = "FALSE")
+                                stringsAsFactors = FALSE)
 
@@
 We mention the Kantrowitz name corpus as implemented in NLTK because the Natural Language Toolkit is rightly regarded as influential for scholarship.
-Its flaws are also typical of the problems with most other implementations of gender prediction algorithms.
+Its flaws for predicting gender, which are a minor part of the software's total functionality, are also typical of the problems with most other implementations of gender prediction algorithms.
 The Genderize.io API is, for example, a more sophisticated implementation of gender prediction than the NLTK algorithm.
 Besides predicting male or female for gender, it also reports the proportion of male or female names, along with a count of the number of uses of the name that it is basing its prediction on.
-Genderize will also permit the user to customize a prediction for different countries, which is an extremely important feature.
+Genderize will also permit the user to customize a prediction for different countries, which is an important feature.
 Genderize.io reports that its "database contains 142848 distinct names across 77 countries and 85 languages." Genderize is unsuitable for historical work, however, because it is based only on contemporary data.
 According to the documentation for its API, "It utilizes big datasets of information, from user profiles across major social networks." It would be anachronistic to apply these datasets to the past, and Genderize.io provides no functionality to filter results chronologically as it does geographically.
-In addition, Genderize.io does not make clear exactly what comprises the dataset and how it was gathered, which keeps scholars from interrogating the value of the source [Genderize.io 2014-].^[Quotations are taken from the [Genderize.io website](http://genderize.io), with correspondence from the developer behind the site for confirmation.]
+In addition, Genderize.io does not make clear exactly what comprises the dataset and how it was gathered, which keeps scholars from interrogating the value of the source [Genderize.io 2014-].^[Quotations are taken from the [Genderize.io website](http://genderize.io), with correspondence from the site's developer for confirmation.]
 
-R and Python are two of the most commonly used languages for data analysis [Piatesky 2014, Muenchen 2014-].^[Surveys and analysis of useage point out the growth of R and the continuing popularity of Python among "data science" generally.] Python's existing packages for gender prediction all implement a method similar to the NLTK or Genderize.io.^[**Python library names**] To our knowledge, [CRAN](http://cran.rstudio.com/) (the central repository for R packages) does not include any packages dedicated to gender prediction.
-In two of the most popular data analysis languages used in the digital humanities there is currently no satisfactory existing method of gender prediction for historical and humanistic research.
+R and Python are two of the most commonly used languages for data analysis [Piatesky 2014, Muenchen 2014-].^[Surveys and analysis of useage point out the growth of R and the continuing popularity of Python for "data science" generally.] Python's existing packages for gender prediction all implement a method similar to the NLTK or Genderize.io.^[These Python libraries include [estimate.gender](https://pypi.python.org/pypi/estimate.gender/0.4), [genderize](https://github.com/SteelPangolin/genderize), and [gender.py](https://github.com/block8437/gender.py)] To our knowledge, [CRAN](http://cran.rstudio.com/) (the central repository for R packages) does not include any packages dedicated to gender prediction.
+Thus two of the most popular data analysis languages used in the digital humanities currently have no satisfactory method for predicting gender from names for historical and humanistic research.
 
 To that end we have created the [Gender](https://github.com/ropensci/gender) package for R which includes both the predictive algorithm and an associated [genderdata](https://github.com/lmullen/gender-data-pkg) package containing various historical datasets.
+Both of these packages can be extended to include other time periods and regions.
 This R implementation is based on an [earlier Python implementation](https://github.com/cblevins/Gender-ID-By-Time) by [Cameron Blevins](http://www.cameronblevins.org/) and [Bridget Baird](http://www.conncoll.edu/directories/emeritus-faculty/bridget-baird/).
 The Gender package is affiliated with [rOpenSci](http://ropensci.org/), an initiative that supports reproducible research and open data for scientists using R.
-The rOpenSci advisory board has provided code review and publicity for the project.^[The gender package depends especially on [R Core Team 2014-, Wickham and Francois 2014] ]
+The rOpenSci advisory board has provided code review and publicity for the project.^[The gender package depends especially on [R Core Team 2014-, Wickham and Francois 2014].]
 

### 2014-09-19 article.Rmd
Minor corrections

@@
 To that end we have created the [Gender](https://github.com/ropensci/gender) package for R which includes both the predictive algorithm and an associated [genderdata](https://github.com/lmullen/gender-data-pkg) package containing various historical datasets.
-Both of these packages can be extended to include other time periods and regions.
 This R implementation is based on an [earlier Python implementation](https://github.com/cblevins/Gender-ID-By-Time) by [Cameron Blevins](http://www.cameronblevins.org/) and [Bridget Baird](http://www.conncoll.edu/directories/emeritus-faculty/bridget-baird/).
@@
 First, it needs a suitable (and suitably large) dataset for the time period and region under study.
-Unsurprisingly such datasets are almost always gathered in the first instance by government entities, though the compilation and digitization may be undertaken by scholarly researchers.
+Unsurprisingly such datasets are almost always gathered in the first instance by governments, though their compilation and digitization may be undertaken by scholarly researchers.
 Second, it depends on a suitable algorithm for estimating the proportion of male and female names for a given year or range of years, since often a person cannot be associated with an exact date.
@@
 The second has been to provide as many datasets as possible in order to localize the predictions to particular times and places.
-To that end, [Ben Schmidt](http://benschmidt.org/) contributed a dataset of US Census Data and accompanying code, as well as providing a significant correction to predictions based on Social Security Administration data.
+To that end, [Ben Schmidt](http://benschmidt.org/) contributed a dataset of US Census Data and accompanying code, as well as provided a significant correction to predictions based on Social Security Administration data.
 
-The Gender package currently includes two datasets which make it suitable for studying the United States from the first federal census in 1790 onwards.^[The authors of the package are both American historians, which explains why the initial release of the package focuses on that country.
+The Gender package currently uses two datasets which make it suitable for studying the United States from the first federal census in 1790 onwards.^[The authors of the package are both American historians, which explains why the initial release of the package focuses on that country.
 We have plans, however, to extend the underlying data to six other European and North American countries included in the [North Atlantic Population Project](https://www.nappdata.org/napp/) for the nineteenth century.
-We also welcome contributions of datasets for other regions and times from other scholars.] The first dataset contains names of applicants for Social Security and available from Data.gov.
+We also welcome contributions of datasets for other regions and times from other scholars.] The first dataset contains names of applicants for Social Security and available from [Data.gov](https://www.data.gov/).
 The second dataset contains names gathered in the decennial censuses and available from the IPUMS-USA (Integrated Public Use Microdata Series) service from the Minnesota Population Center at the University of Minnesota.
@@
 The description "baby names" provided by the SSA is a serious misnomer.
-When Social Security first became available during the New Deal, its first beneficiaries were adults past or near retirment age.
+When Social Security became available during the New Deal, its first beneficiaries were adults past or near retirment age.
 The dataset goes back to 1880, the birth year for a 55 year-old adult when Social Security was enacted.

### 2014-09-19 article.Rmd
More precise explanation of sampling

@@
 This data has been aggregated by IPUMS at the University of Minnesota and is released as a sample of the total census data.
-Unlike the SSA dataset, which includes 100% of every name reported to the Social Security Administration **not quite - only names that occur more than five times, right?**, the IPUMS data contains 5% or 10% samples.
+Unlike the SSA dataset, which includes a 100% sample for every reported to the Social Security Administration and used more than five times, the IPUMS data contains 5% or 10% samples.
 Because the `gender()` function relies on proportions of uses of names, rather than raw counts of people with the names, the sampling does not diminish the reliability of the function's predicitons [Ruggles et al.

### 2014-09-19 article.Rmd
Typo

@@
 Unlike the SSA dataset, which includes a 100% sample for every reported to the Social Security Administration and used more than five times, the IPUMS data contains 5% or 10% samples.
-Because the `gender()` function relies on proportions of uses of names, rather than raw counts of people with the names, the sampling does not diminish the reliability of the function's predicitons [Ruggles et al.
+Because the `gender()` function relies on proportions of uses of names, rather than raw counts of people with the names, the sampling does not diminish the reliability of the function's predictions [Ruggles et al.
 2010].

### 2014-09-19 article.Rmd
Wickham Tidy Data article is out now

@@
 ^[The format of this data departs from the more generally desirable "tidy data" framework proposed by Hadley Wickham for performance reasons.
-By keeping the data in a "wider" format instead of a tidy format as defined by Wickham, the function is able to make its predictions faster [Wickham forthcoming].] The columns contain the `name`, `year`, and number of `female` and `male` uses of that name in a particular year.
+By keeping the data in a "wider" format instead of a tidy format as defined by Wickham, the function is able to make its predictions faster [Wickham 2014b].] The columns contain the `name`, `year`, and number of `female` and `male` uses of that name in a particular year.
 
@@
 
-Bird et al. 2009. Bird, Steve, Edward Loper, and Ewan Klein, *Natural Language Processing with Python* (O’Reilly, 2009)
+Bird et al. 2009. Bird, Steve, Edward Loper, and Ewan Klein, *Natural Language Processing with Python* (O’Reilly, 2009).
 
@@
 
-Wickham, forthcoming. Hadley Wickham, "Tidy Data." *Journal of Statistical Software* (forthcoming), http://vita.had.co.nz/papers/tidy-data.html.
+Wickham 2014. Hadley Wickham, "Tidy Data." *Journal of Statistical Software*, 50 (2014): http://www.jstatsoft.org/v59/i10.
 
-Wickham and Francois 2014-. Wickham, Hadley and Romain Francois, *dplyr: dplyr: a grammar of data manipulation*. R package version 0.2 (2014): <http://CRAN.R-project.org/package=dplyr>.
+Wickham and Francois 2014. Wickham, Hadley and Romain Francois, *dplyr: dplyr: a grammar of data manipulation*. R package version 0.2 (2014): <http://CRAN.R-project.org/package=dplyr>.
 

### 2014-09-19 article.Rmd
Trying to comply with some bizarre citation style

@@
 
-Bird et al. 2009. Bird, Steve, Edward Loper, and Ewan Klein, *Natural Language Processing with Python* (O’Reilly, 2009).
+[Bird et al. 2009] Bird, Steve, Edward Loper, and Ewan Klein. *Natural Language Processing with Python* (O’Reilly, 2009).
 
-Brubaker and Cooper 2000. Brubaker, Rogers, and Frederick Cooper. “Beyond ‘identity.’” *Theory and Society* 29.1 (2000): 1–47.
+[Brubaker and Cooper 2000] Brubaker, Rogers, and Frederick Cooper. "Beyond 'identity.'" *Theory and Society* 29.1 (2000): 1–47.
 
-Butler 1989. Butler, Judith. *Gender Trouble: Feminism and the Subversion of Identity*. New York: Routledge, 1990.
+[Butler 1989] Butler, Judith. *Gender Trouble: Feminism and the Subversion of Identity*. New York: Routledge, 1990.
 
-Butler 1993. Butler, Judith. *Bodies That Matter: On the Discursive Limits of Sex*. New York: Routledge, 1993.
+[Butler 1993] Butler, Judith. *Bodies That Matter: On the Discursive Limits of Sex*. New York: Routledge, 1993.
 
-Canaday 2011. Canaday, Margot. *The Straight State: Sexuality and Citizenship in Twentieth-Century America*. Princeton, N.J.: Princeton University Press, 2011.
+[Canaday 2011] Canaday, Margot. *The Straight State: Sexuality and Citizenship in Twentieth-Century America*. Princeton, N.J.: Princeton University Press, 2011.
 
-Genderize.io 2014-. *Genderize.io*. Accessed 15 September 2014, http://genderize.io/.
+[Genderize.io 2014-] *Genderize.io*. Accessed 15 September 2014, http://genderize.io/.
 
-Goldstone and Underwood 2014.
-Goldstone, Andrew, and Ted Underwood.
+[Goldstone and Underwood 2014] Goldstone, Andrew, and Ted Underwood.
 "The Quiet Transformations of Literary Studies: What Thirteen Thousand Scholars Could Tell Us." *New Literary History* (Forthcoming 2014).
 
-Healy 2013. Healy, Kieran. "Using Metadata to Find Paul Revere." *kieranhealy.org*. 9 June 2013, http://kieranhealy.org/blog/archives/2013/06/09/using-metadata-to-find-paul-revere/
+[Healy 2013] Healy, Kieran. "Using Metadata to Find Paul Revere." *kieranhealy.org*. 9 June 2013, http://kieranhealy.org/blog/archives/2013/06/09/using-metadata-to-find-paul-revere/
 
-Landrum 2014.
-Landrum, Shane.
+[Landrum 2014] Landrum, Shane.
 "The State's Big Family Bible: Birth Certificates, Personal Identity, and Citizenship in the United States, 1840--1950." PhD dissertation, Brandeis University, August 2014.
 
-Muenchen 2014-. Muenchen, Robert A. "The Popularity of Data Analysis Software," *r4stats.com*. Accessed 12 September 2014, http://r4stats.com/articles/popularity/.
+[Muenchen 2014-] Muenchen, Robert A. "The Popularity of Data Analysis Software," *r4stats.com*. Accessed 12 September 2014, http://r4stats.com/articles/popularity/.
 
-Pascoe 2009. Pascoe, Peggy. *What Comes Naturally: Miscegenation Law and the Making of Race in America*. Oxford University Press, USA, 2009.
+[Pascoe 2009] Pascoe, Peggy. *What Comes Naturally: Miscegenation Law and the Making of Race in America*. Oxford University Press, USA, 2009.
 
-Piatetsky 2013.
-Piatetsky, Gregory.
+[Piatetsky 2013] Piatetsky, Gregory.
 "Top Languages for analytics, data mining, data science," *KDN Nuggets*.
@@
 
-R Core Team 2014-. R Core Team, *R: A language and environment for statistical computing*. R Foundation for Statistical Computing, Vienna, Austria, 2014. <http://www.R-project.org/>
+[R Core Team 2014-] R Core Team. *R: A language and environment for statistical computing*. R Foundation for Statistical Computing, Vienna, Austria, 2014. <http://www.R-project.org/>
 
-Ruggles et al.
-2010.
-Ruggles, Steven, J.
+[Ruggles et al.
+2010] Ruggles, Steven, J.
 Trent Alexander, Katie Genadek, Ronald Goeken, Matthew B.
@@
 
-Social Security Administration 2014-a.
-Social Security Administration, "Baby Names from Social Security Card Applications-National Level Data".
+[Social Security Administration 2014-a] Social Security Administration.
+"Baby Names from Social Security Card Applications-National Level Data".
 Data.gov dataset ID US-GOV-SSA-338.
@@
 
-Social Security Administration 2014-b.
-Social Security Administration, "Baby Names from Social Security Card Applications-Data by State and District of Columbia." Data.gov dataset ID US-GOV-SSA-339.
+[Social Security Administration 2014-b] Social Security Administration.
+"Baby Names from Social Security Card Applications-Data by State and District of Columbia." Data.gov dataset ID US-GOV-SSA-339.
 Accessed 31 August 2014, http://catalog.data.gov/dataset/baby-names-from-social-security-card-applications-data-by-state-and-district-of-
 
-Weingart 2014. Weingart, Scott. "The Moral Role of DH in a Data-Driven World." *the scottbot irregular*. 9 September 2014, http://www.scottbot.net/HIAL/?p=40944
+[Weingart 2014] Weingart, Scott. "The Moral Role of DH in a Data-Driven World." *the scottbot irregular*. 9 September 2014, http://www.scottbot.net/HIAL/?p=40944
 
-Wickham 2014. Hadley Wickham, "Tidy Data." *Journal of Statistical Software*, 50 (2014): http://www.jstatsoft.org/v59/i10.
+[Wickham 2014] Wickham, Hadley. "Tidy Data." *Journal of Statistical Software*, 50 (2014): http://www.jstatsoft.org/v59/i10.
 
-Wickham and Francois 2014. Wickham, Hadley and Romain Francois, *dplyr: dplyr: a grammar of data manipulation*. R package version 0.2 (2014): <http://CRAN.R-project.org/package=dplyr>.
+[Wickham and Francois 2014] Wickham, Hadley, and Romain Francois. *dplyr: dplyr: a grammar of data manipulation*. R package version 0.2 (2014): <http://CRAN.R-project.org/package=dplyr>.
 
-World Bank 2014-. World Bank. "Life expectancy at birth, total (years)." *World Bank Open Data*. Accessed 12 September 2014, http://data.worldbank.org/indicator/SP.DYN.LE00.IN.
+[World Bank 2014-] World Bank. "Life expectancy at birth, total (years)." *World Bank Open Data*. Accessed 12 September 2014, http://data.worldbank.org/indicator/SP.DYN.LE00.IN.

### 2014-09-19 article.Rmd
Minor changes to first example

@@
 The second has been to provide as many datasets as possible in order to localize the predictions to particular times and places.
-To that end, [Ben Schmidt](http://benschmidt.org/) contributed a dataset of US Census Data and accompanying code, as well as provided a significant correction to predictions based on Social Security Administration data.
+To that end, [Benjamin Schmidt](http://benschmidt.org/) contributed a dataset of US Census Data and accompanying code, as well as provided a significant correction to predictions based on Social Security Administration data.
 
@@
 
-These two datasets (and any future datasets to be added to the package) thus contain a simple table of data.
-^[The format of this data departs from the more generally desirable "tidy data" framework proposed by Hadley Wickham for performance reasons.
+These two datasets (and any future datasets to be added to the package) thus contain a simple table of data.^[The format of this data departs from the more generally desirable "tidy data" framework proposed by Hadley Wickham for performance reasons.
 By keeping the data in a "wider" format instead of a tidy format as defined by Wickham, the function is able to make its predictions faster [Wickham 2014b].] The columns contain the `name`, `year`, and number of `female` and `male` uses of that name in a particular year.
@@
 Because the dataset contains a list of names for each year, we can simply lookup the row for Sidney in 1935.
-Using the [dplyr](https://github.com/hadley/dplyr) package for R, this can be expressed in a simple grammar of filtering by name and year.
+Using the [dplyr](https://github.com/hadley/dplyr) package for R, which provided a grammar for data manipulation, this can be expressed with the action "filter":
 
@@
 
-Next we can sum up the male and female columns ("summarize" in dplyr's vocabulary) and calculate the proportions of female and male uses of "Sideny" during that decade.
+Next we can sum up the male and female columns ("summarize" in dplyr's vocabulary) and calculate the proportions of female and male uses of "Sidney" during that decade.
 
@@
 In other words, for the decade of the 1930s, we can calculate that there is a `r gender("sidney", year = c(1930, 1939))$proportion_male %>% percent()` percent chance that person named Sidney was male.
-This is roughly the same as the probability we calculated above for just 1935, but our method also returns the figures it used to calculate those probabilities: 1,067 total instances of "Sidney" in 1935 versus 10,110 instances for the decade as a whole.
+This is roughly the same as the probability we calculated above for just 1935, but our method also returns the figures it used to calculate those probabilities: 1,067 instances of "Sidney" in 1935 versus 10,110 total instances for the decade as a whole.
 

### 2014-09-19 article.Rmd
Added paragraph about birth year prediction

@@
 By taking into account the year of birth, we find that four of our six names were likely male, whereas we might otherwise have predicted that all six were female.
-We also now know the approximate likelihood that our predictions are correct: at least 80% for all of thse predictions.
+We also now know the approximate likelihood that our predictions are correct: at least 80% for all of these predictions.
 
-**I think it might be worth expanding this Example #2 section by a paragraph to show how you can dynamically incorporate periods of time rather than single birth years.
-A single birth year might make humanists uncomfortable since it implies a precision that a lot of people aren't going to have (like in the case of our dissertations and AHR authors).
-I changed the example to be a list of editors of a college paper.
-That way we can easily use an age range of 18-22 to demonstrate how a researcher might implement this.**
+It is also possible to use a range of years for a dataset like this.
+If our list of editors contained the year in which the person served on the newspaper rather than the birth year, we could make a reasonable assumption that their ages were likely to be between 18 and 24.
+We could then calculate a minimum and a maximum year of birth for each, and run the prediction function on that range for each person.
+The exact code to accomplish these types of analysis can be found in the Gender package's vignette.
 
@@
   xlab(NULL) + ylab("Proportion of authors by gender") +
-  ggtitle("The gap between the genders of book and dissertation authors")
+  ggtitle("The gap between the genders of book and dissertation authors") + 
+  article_theme +
+  geom_hline(yintercept = 0.5)
 ```

### 2014-09-19 article.Rmd
Note on SSA data

@@
 I'm still not sure it fits here, though.
-I'm trying to figrue out if there's a way to integrate it into the paper that isn't going to be quite so diversionary for the reader.**
+I'm trying to figrue out if there's a way to integrate it into the paper that isn't going to be quite so diversionary for the reader.
+LAM: okay, wherever you think is best.**
 

### 2014-09-19 article.Rmd
Lunbeck citation

@@
 In a 2005 report for the American Historical Association, Elizabeth Lunbeck acknowledged "a sea change in the [historical] profession with respect to gender" before going on to describe the painful limitations of this sea change for female historians: ongoing personal discrimination, lower salaries, and barriers to securing high-ranking positions.
-Lunbeck's report drew in part on a survey of 362 female historians that produced a rich source of responses detailing the deep and multi-faceted challenges facing women in the profession.
-What follows is a quantitative supplement to Lunbeck's analysis that uses our program to analyze gender representation amongst historians across a much larger scale and a much longer time period.^[http://www.historians.org/Documents/About%20AHA%20and%20Membership/CWH-Report_5.20.05.pdf]
+Lunbeck's report drew in part on a survey of 362 female historians that produced a rich source of responses detailing the deep and multi-faceted challenges facing women in the profession [Lunbeck 2005].
+What follows is a quantitative supplement to Lunbeck's analysis that uses our program to analyze gender representation amongst historians across a much larger scale and a much longer time period.
 
@@
 
+[Lunbeck 2005] Lunbeck, Elizabeth, et al.
+The Status of Women in the Historical Profession.
+American Historical Association Committee on Women Historians (2005): http://www.historians.org/Documents/About%20AHA%20and%20Membership/CWH-Report_5.20.05.pdf.
+
 [Muenchen 2014-] Muenchen, Robert A. "The Popularity of Data Analysis Software," *r4stats.com*. Accessed 12 September 2014, http://r4stats.com/articles/popularity/.

### 2014-09-19 article.Rmd
Filled in some details

@@
 The completion of a dissertation marks a standardized moment of transition out of the "training" phase of the historical profession.
-Data supplied by ProQuest provided roughly [80,000] PhD(?) dissertations completed since [1950].^[ProQuest's dataset does not permit us to limit dissertations just to dissertations produced in history departments.
+Data supplied by ProQuest provided roughly eighty thousand PhD dissertations completed in the United States.^[ProQuest's dataset does not permit us to limit dissertations just to dissertations produced in history departments.
 These dissertations include dissertations *about* history.
@@
 Data export provided to author by Liz Townsend, March 3, 2014.] Identifying the gender of their authors gives a sense for the number of women and men who are completing PhD-level training in history each year.
-Our program uses the year a historian wrote his or her dissertation to estimate a period for when they might have been born: a historian who completed their dissertation in 1980 was likely born between ____ and _____.
+Our program uses the year a historian wrote his or her dissertation to estimate a period for when they might have been born: we have assumed that a historian was between 25 and 45 years old when completing the dissertation.
 Using this temporal information, we are better able to infer their gender and chart how the larger representation of women and men changes year-by-year.
-
-**Do we even need this chart, or is the chart of proportions sufficient?**
 
@@
 Another way to examine this trend is to look at the proportion of male and female dissertation authors over time, which smoothes out changes in the absolute number of dissertations produced.
-The proportion of dissertations written by women has steadily increased over the past half-century, a change that began in the late 1960s and continued through the early 2000's.
+The proportion of dissertations written by women has steadily increased over the past half-century, a change that began in the late 1960s and continued through the early 2000s.
 Since that point, the proportion of dissertations written by women has largely plateaued at a few percentage points below the proportion written by men.

### 2014-09-20 article.Rmd
Delete accepted note

@@
 
-**I moved this next caveat section about the SSA data because it felt kind of out of place sandwiched between the two examples.
-I'm still not sure it fits here, though.
-I'm trying to figrue out if there's a way to integrate it into the paper that isn't going to be quite so diversionary for the reader.
-LAM: okay, wherever you think is best.**
-
 As previously mentioned, the history of how the Social Security Administration collected the data effects its validity.

### 2014-09-20 article.Rmd
Minor stylistic changes

@@
 The completion of a dissertation marks a standardized moment of transition out of the "training" phase of the historical profession.
-Data supplied by ProQuest provided roughly eighty thousand PhD dissertations completed in the United States.^[ProQuest's dataset does not permit us to limit dissertations just to dissertations produced in history departments.
+To study dissertations about history, we used data supplied by ProQuest for roughly eighty thousand PhD dissertations completed in the United States.^[ProQuest's dataset does not permit us to limit dissertations just to dissertations produced in history departments.
 These dissertations include dissertations *about* history.
@@
 The American Historical Association's dataset of dissertations relies on reporting from departments and does not have the chronological range of the ProQuest data.
-Nevertheless, applying the same method to the AHA's dataset does not provide any reason for thinking that the proportion authorship by gender over time is substantially different when limited to just history department dissertations.
+Nevertheless, applying the same method to the AHA's dataset does not provide any reason for thinking that the proportion of authorship by gender over time is substantially different when limited to just history department dissertations.
 American Historical Association, *Directory of History Dissertations*, https://secure.historians.org/pubs/dissertations/.
-Data export provided to author by Liz Townsend, March 3, 2014.] Identifying the gender of their authors gives a sense for the number of women and men who are completing PhD-level training in history each year.
+Data export provided to author by Liz Townsend, March 3, 2014.] Identifying the gender of their authors identifies the approximate number of women and men who are completing PhD-level training in history each year.
 Our program uses the year a historian wrote his or her dissertation to estimate a period for when they might have been born: we have assumed that a historian was between 25 and 45 years old when completing the dissertation.
@@
 One of the leading journals in the historical profession is the *American Historical Review*, the flagship journal of the American Historical Association.
-Along with a few dozen articles, the AHR publishes five issues each year that contain, in aggregate, roughly one thousand books reviews covering (in its own words) "every major field of historical study." The AHR is not only one of the widest-ranging journals in the profession, it is also the oldest; the journal has been publishing continuously since 1895.
+Along with a few dozen articles published in its five issues each year, the AHR publishes roughly one thousand books reviews per year covering (in its own words) "every major field of historical study." The AHR is not only one of the widest-ranging journals in the profession, it is also the oldest; the journal has been publishing continuously since 1895.
 The range, scope, and length of the AHR makes it an ideal source to analyze the historical profession on a large scale.

### 2014-09-20 article.Rmd
Sentence capitalization for figure titles

@@
   geom_bar(binwidth = 1) +
-  ggtitle("Dissertations about History by Gender, 1950-2012") +
+  ggtitle("Dissertations about history by gender, 1950-2012") +
   xlab(NULL) + ylab(NULL) +
@@
   geom_bar(binwidth = 1, stat = "identity") +
-  ggtitle("Proportion of Dissertations about History by Gender, 1950-2012") +
+  ggtitle("Proportion of dissertations about history by gender, 1950-2012") +
   xlab(NULL) + ylab(NULL) +
@@
   geom_bar(binwidth = 1, stat = "identity") +
-  ggtitle("Authors of Books Reviewed in the AHR by Gender") +
+  ggtitle("Authors of books reviewed in the AHR by gender") +
   xlab(NULL) + ylab(NULL) +
@@
   geom_bar(binwidth = 1, stat = "identity") +
-  ggtitle("Gender of reviewers of female book authors in AHR") +
+  ggtitle("Gender of reviewers of female book authors in AHR by category") +
   xlab(NULL) + ylab(NULL) +

### 2014-09-20 article.Rmd
Don't run gg_atribution

@@
 
-leslie_plot <- leslie_df %>%
+leslie_df %>%
   ggplot(aes(x = year, y = prop_female)) +
@@
   article_theme
-
-leslie_plot %>%
-  gg_attribution("Data: IPUMS USA; Social Security Administration") %>%
-  print()
 ```
@@
 
-selected_names_plot <- ssa_national %>%
+ssa_national %>%
   filter(name %in% plot_names) %>%
@@
   scale_color_brewer(palette = article_palette, name = "")
-
-selected_names_plot %>%
-  gg_attribution("Data: Social Security Administration") %>%
-  print()
 ```

### 2014-09-20 article.Rmd
Specify precise error rates for the Kantrowitz corpus

@@
 
-```{r compare-kantrowitz-ssa, cache=TRUE, eval=FALSE}
+```{r compare-kantrowitz-ssa}
 # Calculate the error rate of the Kantrowitz data set. This code is very slow, 
@@
 # We can re-run this for the final version of the article.
+if (file.exists("data/cf-kantrowitz-ssa.rda")) {
+  load("data/cf-kantrowitz-ssa.rda")
+  } else {
   cf_kantrowitz_ssa <- data.frame(name = unique(genderdata::kantrowitz$name),
@@
 
-cf_results <- as.character(cf_kantrowitz_ssa$kantrowitz) ==
-  as.character(cf_kantrowitz_ssa$ssa)
+  cf_kantrowitz_ssa$kantrowitz <- as.character(cf_kantrowitz_ssa$kantrowitz) 
+  cf_kantrowitz_ssa$ssa<- as.character(cf_kantrowitz_ssa$ssa) 
+  cf_kantrowitz_ssa[cf_kantrowitz_ssa == "NA"]  <- NA
+  save(cf_kantrowitz_ssa, file = "data/cf-kantrowitz-ssa.rda", compress = TRUE)
+  }
 
-# Error rate is the number of mismatches to the total number of names
-kantrowitz_error_rate <- 1 -(sum(cf_results) / length(cf_results))
-```
+# Check for mismatches between Kantrowitz and our method
+false_either <- ifelse((cf_kantrowitz_ssa$kantrowitz == "either" &
+                          (cf_kantrowitz_ssa$ssa == "male" |
+                             cf_kantrowitz_ssa$ssa == "female")),
+                     TRUE, FALSE)
 
-```{r}
-# Hardcode the error rate for now. This value is based on running the code above
-# locally.
-kantrowitz_error_rate <- 0.2325766
+false_male <- ifelse(cf_kantrowitz_ssa$kantrowitz == "male" &
+                     cf_kantrowitz_ssa$ssa == "female",
+                     TRUE, FALSE)
+
+false_female <- ifelse(cf_kantrowitz_ssa$kantrowitz == "female" &
+                     cf_kantrowitz_ssa$ssa == "male",
+                     TRUE, FALSE)
+
+false_either_percent <- sum(false_either, na.rm = TRUE) / nrow(cf_kantrowitz_ssa)
+false_male_percent <- sum(false_male, na.rm = TRUE) / nrow(cf_kantrowitz_ssa)
+false_female_percent <- sum(false_female, na.rm = TRUE) / nrow(cf_kantrowitz_ssa)
+
 ```
@@
 There are therefore many names that it cannot identify.
-Assuming for the moment that our method provides more accurate results, we estimate that `r percent(kantrowitz_error_rate)` percent of the names in the Kantrowitz corpus are misclassified or classified as ambigious when a valid prediction could be made.
-**This isn't immediately obvious to me how this error rate was derived.
-Do we know why or how Kantrowitz is classifying things as ambiguous?
-Maybe he's just using a much higher, and theoretically more rigorous, cut-off (say 90%) than our 50% cut-off.**
+Assuming for the moment that our method provides more accurate results, we estimate that `r percent(false_either_percent)` percent of the names in the Kantrowitz corpus are classified ambiguously when a gender could be reasonably predicted from the name, that `r percent(false_male_percent)` percent of the names are classified as male when they should be classified as female, and that `r percent(false_female_percent)` are classified as female when they should be classified as male.
+This error rate is a separate concern from the much smaller size of the Kantrowitz corpuse.^[This error rate was calculated using the Social Security Administration database, assuming a range of years between 1932 and 2012, which is the default setting within the Gender package.
+Of course, the error rate would fluctuate for different time periods and regions.]
 

### 2014-09-22 article.Rmd
Change title to prediction
Typos and minor edits

@@
 ---
-title: "Jane, John … Leslie? A Historical Method for Algorithmic Gender Detection"
+title: "Jane, John … Leslie? A Historical Method for Algorithmic Gender Prediction"
 author: "Cameron Blevins and Lincoln Mullen"
@@
 One of the leading journals in the historical profession is the *American Historical Review*, the flagship journal of the American Historical Association.
-Along with a few dozen articles published in its five issues each year, the AHR publishes roughly one thousand books reviews per year covering (in its own words) "every major field of historical study." The AHR is not only one of the widest-ranging journals in the profession, it is also the oldest; the journal has been publishing continuously since 1895.
-The range, scope, and length of the AHR makes it an ideal source to analyze the historical profession on a large scale.
+Along with a few dozen articles published in its five issues each year, the *AHR* publishes roughly one thousand books reviews per year covering (in its own words) "every major field of historical study." The *AHR* is not only one of the widest-ranging journals in the profession, it is also the oldest; the journal has been publishing continuously since 1895.
+The range, scope, and length of the *AHR* makes it an ideal source to analyze the historical profession on a large scale.
 
-We take the AHR as a rough proxy for gatekeeping within the larger historical profession.
-When a book appears in the AHR it serves as a signal that other historians in positions of power have "approved" it for consideration.
+We take the <i>American Historical Review</i> as a rough proxy for gatekeeping within the larger historical profession.
+When a book appears in the *AHR* it serves as a signal that other historians in positions of power have "approved" it for consideration.
 This signal is not necessarily correlated with the quality of the book.
 Even if it garners a negative review, the fact that it appears in the journal at all is a measure of the fact that the profession's gatekeepers have deemed it important enough to review.
-On the flip side, a book that does not appear in the AHR may have gone on to have a significant impact on the profession.
+On the flip side, a book that does not appear in the *AHR* may have gone on to have a significant impact on the profession.
 It is precisely this professional gatekeeping dimension that makes the AHR useful to study in the context of gender representation.
@@
 The temporal trajectory of gender representations roughly resembles that of history dissertations: women began making inroads in the late 1960s and have made steady gains over the past four decades.
-By the twenty-first century the proportion of female authors reviewed in the AHR had more than tripled.
+By the twenty-first century the proportion of female authors reviewed in the *AHR* had more than tripled.
 
@@
 
-But a closer look shows important differences between the proportion of women writing history dissertations and the proportion of women appearing in the AHR.
-Although both have trended upwards since the 1970s the slope is a lot steeper for dissertations than it is for the AHR.
-Female historians have very nearly closed the gaps in terms of newly completed dissertations, but the glass ceiling remains stubbornly low in terms of what happens from that point onwards.
-In book reviews published in the AHR male authors continue to outnumber female authors by a factor of nearly 2 to 1.
+But a closer look shows important differences between the proportion of women writing history dissertations and the proportion of women appearing in the *AHR*.
+Although both have trended upwards since the 1970s the slope is a lot steeper for dissertations than it is for the *AHR*.
+Female historians have very nearly closed the gap in terms of newly completed dissertations, but the glass ceiling remains stubbornly low in terms of what happens from that point onwards.
+In book reviews published in the *AHR* male authors continue to outnumber female authors by a factor of nearly 2 to 1.
 The long-term gains made by female doctoral students do not carry over in the pages of the historical profession's leading journal.
@@
 
-The path from a dissertation to an AHR book review is not an automatic one.
-Not all dissertations are published as books, not all books are submitted to the American Historical Review, and not all book submissions are accepted by the AHR editors for review.
+The path from a dissertation to a book review in the *American Historical Association* is not an automatic one.
+Not all dissertations are published as books, not all books are submitted to the American Historical Review, and not all book submissions are accepted by the *AHR* editors for review.
 As of August 2014 the journal’s website reminds readers that "the AHR receives over 3,000 books a year; we have the resources to publish at most 1,000 reviews a year (approximately 200 per issue)."^[http://www.indiana.edu/~ahrweb/book_review_guide.html - retrieved August 25 2014].
@@
 This gender gap is more likely to impact the production of a book than a dissertation, as graduate students are generally younger and less likely to be juggling child-rearing responsibilities while writing than their older peers.
-More insidiously, women continue to face insidious gender discrimination.
+More troublingly, women continue to face insidious gender discrimination.
 
-We then used our program to infer the gender of not just book authors reviewed by the AHR, but the gender of the reviewers themselves.
+We then used our program to infer the gender of not just book authors reviewed by the *AHR*, but the gender of the reviewers themselves.
 The story is much the same for reviewers as it is for authors: more than than twice as many men as women appear as reviewers in the journal.
@@
 In short, women are much more likely to write reviews of other women.
-And while men still write the majority of reviews of female-authored books, they tend to gravitate more towards male authors – who are, of course, already over-represented in the AHR.
+And while men still write the majority of reviews of female-authored books, they tend to gravitate more towards male authors – who are, of course, already over-represented in the *AHR*.
 
@@
 In both African history and Ancient/Medieval history female historians made some quite dramatic gains during the late-nineties and aughts.
-The most uneven parties, however, are also the two subject categories that publish the most book reviews in the AHR: Modern/Early Modern Europe and the United States/Canada.
+The most uneven parties, however, are also the two subject categories that publish the most book reviews in the *AHR*: Modern/Early Modern Europe and the United States/Canada.
 Both sub-fields have made progress over the four decades but still hover at around two-thirds male.

### 2014-09-22 article.Rmd
Minor edits; conclusion

@@
 One 2013 analysis of 2,500 recent history Ph.D.'s found that "gender played little role in employment patterns across particular professions and industries."^[http://historians.org/Documents/Many_Careers_of_History_PhDs_Final.pdf] Our own analysis reveals a more complicated reality.
-The flagship journal of professional historians continues to publish twice as many reviews of books by female authors as male authors.
+The flagship journal of professional historians continues to publish twice as many reviews of books by male authors as female authors.
 This disparity is put into even sharper relief when set against the relative parity achieved by women in producing history dissertations.
-Women might be getting hired at comparable rates to men, but discriminatory gatekeeping remains alive and well in the historical profession.
+Women might be getting hired at comparable rates to men, but discriminatory gatekeeping remains alive and well in the profession.
 Our program allows us to quantitatively measure the scope of the problem on a large scale, how it's changed over time, and what remains to be done.
@@
 
-In conclusion, our algorithm for predicting gender from first names combined with historical datasets in the Gender package provides a powerful new method for historians interested in the study of gender.
-The digital turn has made available very large datasets, yet in the midst of abundance historians always have to deal with the problem of scarcity in the sources.
-Our methods uses the abundant datasets to produce more information from scarce datasets.
-The temporal dimension of our work matches a historical question to a historical method, providing much improved over simple, anachronistic lookup methods.
-This method is extensible to any time or place for which there is a suitable dataset of names from which to make the predictions.
+Our algorithm for predicting gender from first names, when combined with historical datasets in the Gender package, provides a powerful new method for researchers interested in the study of gender.
+The digital turn has made available very large datasets, yet in the midst of abundance many humanists continue to deal with the problem of scarcity in their sources.
+Our method uses the abundant existing datasets to infer more information from scarce datasets.
+The temporal dimension of our work matches historical questions to a historical method, providing much improved results over simple, anachronistic lookup methods.
+This method is extensible to any time or place for which there is a suitable dataset of names from which to make predictions about gender and naming practices.
 As we have demonstrated in our study of gatekeeping in the American historical profession, our method can reveal important details about gender even from datasets which do not otherwise include that information, and it can do so over long historical periods.

### 2014-09-26 article.Rmd
Facet reviewers by subfield

@@
 ```{r load-libraries, include=FALSE}
-options(stringsAsFactors = FALSE)
 library(genderdata)
@@
 # I like 
-article_theme <- theme(legend.title = element_blank()) + theme_few()
+article_theme <- theme(legend.title = element_blank(),
+                       strip.background = element_rect(color = "white")) +
+  theme_classic()
 
@@
 
-**Cameron: Do you have the mapping you use from the AHR's 68 categories to the 8 or so you used?
-With that, I just have one or two more lines to facet this plot.
-(It doesn't look right because it's not faceted yet.)**
+```{r reviews-by-category}
+categories <- read.csv("data/ahr-categories.csv", stringsAsFactors = FALSE)
 
-```{r}
 ahr_predicted %>%
-  filter(gender_author == "female") %>%
-  select(year, gender_reviewer, category) %>% 
-  group_by(year, gender_reviewer, category) %>%  
+  left_join(categories, by = "category") %>%
+  filter(gender_author == "female",
+         !our_category %in% c(NA, 
+                              "Latin and South America, Carribean",
+                              "Middle East")) %>%
+  select(year, gender_reviewer, our_category) %>% 
+  group_by(year, gender_reviewer, our_category) %>%  
   summarize(n = n()) %>%
@@
   select(female_proportion, male_proportion, unknown_proportion,
-         year, category) %>% 
-  gather(gender, proportion, -year, -category) %>%
+         year, our_category) %>% 
+  gather(gender, proportion, -year, -our_category) %>%
 ggplot(aes(x = year, y = proportion, fill = gender)) + 
   geom_bar(binwidth = 1, stat = "identity") +
-  ggtitle("Gender of reviewers of female book authors in AHR by category") +
+  facet_wrap(~ our_category) +
+  xlim(1970, 2012) +
+  ggtitle("Gender of reviewers of female AHR authors by category, 1970-2012") +
   xlab(NULL) + ylab(NULL) +
@@
   scale_fill_brewer(palette = article_palette,
-                    labels = c("Female", "Male", "Unknown")) +
+                    labels = c("Female", "Male", "Unknown"),
+                    name = "Gender") +
   article_theme +
-  short_scale
+  theme(strip.background = element_rect(color = "white"))
 ```

### 2014-09-26 article.Rmd
Implement suggestions from writing group

@@
 It outlines a new method for algorithmically inferring the gender of a name from historical data in order to take into account how naming practices change over time and space.
-In contrast to existing methods of predicting gender, this method provides a higher level of accuracy and a precise statement of the likelihood that a name was a given gender.
+In contrast to existing methods of predicting gender, this method provides a higher level of accuracy and a precise statement of the likelihood that a name was a given gender, especially for datasets where names can be associated with dates.
 The article describes the methodology as implemented in the [Gender](https://github.com/ropensci/gender) package for the [R programming language](http://www.r-project.org/).
@@
 [Weingart 2014] Kieran Healy has used network analysis to demonstrate how researchers can infer information from datasets while also touching on the troubling ethical issue of surveillance [Healy 2013].] Many of these datasets deal with lists of people: company payrolls, military rosters, passenger bookings, records of correspondence, or lists of published works.
-Some of these records might contain associated information such a person's address, age, or rank, while others contain little more than a list of personal names.
+Some of these records might contain associated information such as a person's address, age, or rank, while others contain little more than a list of personal names.
+A common question in the history of American religion, for example, is whether more women than men participated in a church or religious group.
+A simple membership roll might contain names, but it is unlikely to identify the gender of the members.
+One might also ask a question about changing patterns in male and female authorship of books.
+Library catalogs contain enormous amounts of data about publication, but they do not record the sex of authors.
+
 Given large but sparse datasets, researchers have to use their ingenuity to derive additional kinds of information.
@@
 We recognize that gender is a fluid concept, both today and in the past, and that this method glosses over the complexities of historical definitions of gender and of the relationship between biological sex and gender [Butler 1990, Butler 1993].
-In some sense this is unavoidable since the large historical datasets which make this kind of analysis possible were created by state agencies.
+In some sense this is unavoidable since the large historical datasets which make this kind of analysis possible were created by state agencies such as the U.S.
+Census Bureau.
 For the category of gender (as for race) states have offered a limited number of boxes to check, and those boxes extert a tremendous power to define state subjects [Brubaker and Cooper 2000, 14-17, Pascoe 2009, Canaday 2011, Landrum 2014].
@@
 The method's real utility stems from being able to process larger datasets than a single name.
-Let's use, say, a list of editors from a college newspaper to illustrate how a researcher might apply it to their own data.
+Let's use, say, a hypothetical list of editors from a college newspaper to illustrate how a researcher might apply it to their own data.
 The package's prediction function allows researchers to choose which reference datasets they would like to use and the range of years for making their predictions.^[See the package vignette for more details: `vignette("predicting-gender", package = "gender")`.]
@@
 Map(gender, editors$name, years = editors$birth_year) %>% 
-  do.call(rbind.data.frame, .)
+  do.call(rbind.data.frame, .) %>% print.data.frame(row.names = FALSE)
 ```
@@
 What follows is a quantitative supplement to Lunbeck's analysis that uses our program to analyze gender representation amongst historians across a much larger scale and a much longer time period.
+We have found that although women have acheived rough equality in terms of the number of dissertations written, there is still a significant gap between male and female authorship of books in history.
 
@@
 This gender gap is more likely to impact the production of a book than a dissertation, as graduate students are generally younger and less likely to be juggling child-rearing responsibilities while writing than their older peers.
-More troublingly, women continue to face insidious gender discrimination.
+More troublingly, women continue to face gender discrimination.
 
@@
 This method is extensible to any time or place for which there is a suitable dataset of names from which to make predictions about gender and naming practices.
+This method is potentially useful to anyone who has a dataset of names, especially if those names can be associated with particular dates.
 As we have demonstrated in our study of gatekeeping in the American historical profession, our method can reveal important details about gender even from datasets which do not otherwise include that information, and it can do so over long historical periods.

### 2014-09-26 article.Rmd
Fixed citation

@@
 
-One 2013 analysis of 2,500 recent history Ph.D.'s found that "gender played little role in employment patterns across particular professions and industries."^[http://historians.org/Documents/Many_Careers_of_History_PhDs_Final.pdf] Our own analysis reveals a more complicated reality.
+One 2013 analysis of 2,500 recent history Ph.D.'s found that "gender played little role in employment patterns across particular professions and industries" [Wood and Townsend 2013].
+Our own analysis reveals a more complicated reality.
 The flagship journal of professional historians continues to publish twice as many reviews of books by male authors as female authors.
@@
 Women might be getting hired at comparable rates to men, but discriminatory gatekeeping remains alive and well in the profession.
-Our program allows us to quantitatively measure the scope of the problem on a large scale, how it's changed over time, and what remains to be done.
+Our program allows us to quantitatively measure the scope of the problem on a large scale and how it has changed over time.
 
@@
 
+[Wood and Townsend 2013] Wood, L.
+Maren, and Robert B.
+Townsend.
+*The Many Careers of History PhDs: A Study of Job Outcomes*.
+American Historical Association, 2013: <http://historians.org/Documents/Many_Careers_of_History_PhDs_Final.pdf>.
+
 [World Bank 2014-] World Bank. "Life expectancy at birth, total (years)." *World Bank Open Data*. Accessed 12 September 2014, http://data.worldbank.org/indicator/SP.DYN.LE00.IN.

### 2014-09-26 article.Rmd
Fixed footnote

@@
 Not all dissertations are published as books, not all books are submitted to the American Historical Review, and not all book submissions are accepted by the *AHR* editors for review.
-As of August 2014 the journal’s website reminds readers that "the AHR receives over 3,000 books a year; we have the resources to publish at most 1,000 reviews a year (approximately 200 per issue)."^[http://www.indiana.edu/~ahrweb/book_review_guide.html - retrieved August 25 2014].
-Fewer and fewer female historians make it through each successive step.
+As of August 2014 the journal's website reminds readers that "the AHR receives over 3,000 books a year; we have the resources to publish at most 1,000 reviews a year (approximately 200 per issue)."^[The website is available at <http://www.indiana.edu/~ahrweb/book_review_guide.html> (retrieved August 25, 2014).] Fewer and fewer female historians make it through each successive step.
 Establishing a single causal reason for this pattern is all but impossible, and instead a murky concoction of interrelated factors likely contributes to the gap.

### 2014-09-29 article.Rmd
Nuanced some langauge
Plot tweak, citations

@@
 plot_names <-  c("madison",
-                 "keanu",
+                 #"keanu", #keanu doesn't seem to actually flip, or at least doesn't show up on the chart?
                  "addison",
@@
 One 2013 analysis of 2,500 recent history Ph.D.'s found that "gender played little role in employment patterns across particular professions and industries" [Wood and Townsend 2013].
-Our own analysis reveals a more complicated reality.
+Our own analysis reveals that the story is different for book publishing than for employment.
 The flagship journal of professional historians continues to publish twice as many reviews of books by male authors as female authors.
@@
 
+[American Historical Review n.d.] American Historical Review.
+"Book Reviewing in the AHR." http://www.historians.org/publications-and-directories/american-historical-review/book-review-guide.
+Accessed 29 September 2014.
+
 [Bird et al. 2009] Bird, Steve, Edward Loper, and Ewan Klein. *Natural Language Processing with Python* (O’Reilly, 2009).
@@
 
-[Genderize.io 2014-] *Genderize.io*. Accessed 15 September 2014, http://genderize.io/.
+[Genderize.io n.d.] *Genderize.io*. http://genderize.io/. Accessed 15 September 2014.
 
@@
 
-[Healy 2013] Healy, Kieran. "Using Metadata to Find Paul Revere." *kieranhealy.org*. 9 June 2013, http://kieranhealy.org/blog/archives/2013/06/09/using-metadata-to-find-paul-revere/
+[Healy 2013] Healy, Kieran. "Using Metadata to Find Paul Revere." *kieranhealy.org*. 9 June 2013, <http://kieranhealy.org/blog/archives/2013/06/09/using-metadata-to-find-paul-revere/>.
 
@@
 [Lunbeck 2005] Lunbeck, Elizabeth, et al.
-The Status of Women in the Historical Profession.
-American Historical Association Committee on Women Historians (2005): http://www.historians.org/Documents/About%20AHA%20and%20Membership/CWH-Report_5.20.05.pdf.
+"The Status of Women in the Historical Profession." American Historical Association Committee on Women Historians.
+May 2005, <http://www.historians.org/Documents/About%20AHA%20and%20Membership/CWH-Report_5.20.05.pdf>.
 
-[Muenchen 2014-] Muenchen, Robert A. "The Popularity of Data Analysis Software," *r4stats.com*. Accessed 12 September 2014, http://r4stats.com/articles/popularity/.
+[Muenchen n.d.] Muenchen, Robert A. "The Popularity of Data Analysis Software," *r4stats.com*. <http://r4stats.com/articles/popularity/>. Accessed 12 September 2014.
 
@@
 "Top Languages for analytics, data mining, data science," *KDN Nuggets*.
-27 August 2013, http://www.kdnuggets.com/2013/08/languages-for-analytics-data-mining-data-science.html.
+27 August 2013, <http://www.kdnuggets.com/2013/08/languages-for-analytics-data-mining-data-science.html>.
 
@@
 Data.gov dataset ID US-GOV-SSA-338.
-Accessed 31 August 2014, http://catalog.data.gov/dataset/baby-names-from-social-security-card-applications-national-level-data
+Accessed 31 August 2014, <http://catalog.data.gov/dataset/baby-names-from-social-security-card-applications-national-level-data>.
 
@@
 "Baby Names from Social Security Card Applications-Data by State and District of Columbia." Data.gov dataset ID US-GOV-SSA-339.
-Accessed 31 August 2014, http://catalog.data.gov/dataset/baby-names-from-social-security-card-applications-data-by-state-and-district-of-
+Accessed 31 August 2014, <http://catalog.data.gov/dataset/baby-names-from-social-security-card-applications-data-by-state-and-district-of->.
 
@@
 
-[Wickham 2014] Wickham, Hadley. "Tidy Data." *Journal of Statistical Software*, 50 (2014): http://www.jstatsoft.org/v59/i10.
+[Wickham 2014] Wickham, Hadley. "Tidy Data." *Journal of Statistical Software*, 50 (2014): <http://www.jstatsoft.org/v59/i10>.
 
@@
 
-[World Bank 2014-] World Bank. "Life expectancy at birth, total (years)." *World Bank Open Data*. Accessed 12 September 2014, http://data.worldbank.org/indicator/SP.DYN.LE00.IN.
+[World Bank 2014-] World Bank. "Life expectancy at birth, total (years)." *World Bank Open Data*. <http://data.worldbank.org/indicator/SP.DYN.LE00.IN>. Accessed 12 September 2014

### 2014-09-29 article.Rmd
Added plot, plot edits, remapped categories

@@
 
-The path from a dissertation to a book review in the *American Historical Association* is not an automatic one.
-Not all dissertations are published as books, not all books are submitted to the American Historical Review, and not all book submissions are accepted by the *AHR* editors for review.
-As of August 2014 the journal's website reminds readers that "the AHR receives over 3,000 books a year; we have the resources to publish at most 1,000 reviews a year (approximately 200 per issue)."^[The website is available at <http://www.indiana.edu/~ahrweb/book_review_guide.html> (retrieved August 25, 2014).] Fewer and fewer female historians make it through each successive step.
+The path from a dissertation to a book review in the *American Historical Association* is not automatic.
+Not all dissertations are published as books, not all books are submitted to the *American Historical Review*, and not all book submissions are accepted by the *AHR* editors for review.
+As of September 2014 the journal's website reminds readers that "the AHR receives over 3,000 books a year; we have the resources to publish at most 1,000 reviews a year (approximately 200 per issue)." [American Historical Review n.d.] Fewer and fewer female historians make it through each successive step.
 Establishing a single causal reason for this pattern is all but impossible, and instead a murky concoction of interrelated factors likely contributes to the gap.
@@
 This gender gap is more likely to impact the production of a book than a dissertation, as graduate students are generally younger and less likely to be juggling child-rearing responsibilities while writing than their older peers.
-More troublingly, women continue to face gender discrimination.
+More troublingly, women continue to face gender discrimination in the professional sphere.
 
-We then used our program to infer the gender of not just book authors reviewed by the *AHR*, but the gender of the reviewers themselves.
+We  used our program to infer the gender of not just book authors reviewed by the *AHR*, but the gender of the reviewers themselves.
 The story is much the same for reviewers as it is for authors: more than than twice as many men as women appear as reviewers in the journal.
@@
 And while men still write the majority of reviews of female-authored books, they tend to gravitate more towards male authors – who are, of course, already over-represented in the *AHR*.
+
+```{r}
+ahr_predicted %>%
+  filter(gender_author == "male") %>%
+  select(year, gender_reviewer) %>%  
+  group_by(year, gender_reviewer) %>% 
+  summarize(n = n()) %>% 
+  spread(gender_reviewer, n, fill = 0) %>%
+  mutate(female_proportion = female / (female + male + unknown),
+         male_proportion = male / (female + male + unknown),
+         unknown_proportion = unknown / (female + male + unknown)) %>%  
+  select(female_proportion, male_proportion, unknown_proportion, year) %>%
+  gather(gender, proportion, -year) %>%
+  ggplot(aes(x = year, y = proportion, fill = gender)) + 
+  geom_bar(binwidth = 1, stat = "identity") +
+  ggtitle("Gender of reviewers of MALE book authors in AHR") +
+  xlab(NULL) + ylab(NULL) +
+  theme(legend.position = "bottom") +
+  geom_hline(yintercept = 0.5) +
+  scale_fill_brewer(palette = article_palette, 
+                    labels = c("Female", "Male", "Unknown")) +
+  article_theme +
+  short_scale
+```
 
@@
   geom_bar(binwidth = 1, stat = "identity") +
-  ggtitle("Gender of reviewers of female book authors in AHR") +
+  ggtitle("Gender of reviewers of FEMALE book authors in AHR") +
   xlab(NULL) + ylab(NULL) +
@@
 
-The American Historical Association also categorizes their reviews under different historical sub-fields.
+The *AHR* also categorizes its reviews under different historical sub-fields.
 Although the journal’s taxonomy has changed quite a bit over time, it did allow us to refine our analysis.
@@
   left_join(categories, by = "category") %>%
-  filter(gender_author == "female",
-         !our_category %in% c(NA, 
+  filter(!our_category %in% c(NA, 
+                              "Articles",
                               "Latin and South America, Carribean",
                               "Middle East")) %>%
-  select(year, gender_reviewer, our_category) %>% 
-  group_by(year, gender_reviewer, our_category) %>%  
+  select(year, gender_author, our_category) %>% 
+  group_by(year, gender_author, our_category) %>%  
   summarize(n = n()) %>%
-  spread(gender_reviewer, n, fill = 0) %>% 
+  spread(gender_author, n, fill = 0) %>% 
   mutate(female_proportion = female / (female + male + unknown),
@@
   xlim(1970, 2012) +
-  ggtitle("Gender of reviewers of female AHR authors by category, 1970-2012") +
+  ggtitle("Gender of AHR authors by book category, 1970-2012") +
   xlab(NULL) + ylab(NULL) +

### 2014-09-29 article.Rmd
Copy-edits

@@
 Whereas before it had been an infrequently used male name, after 1985 it skyrocketed in popularity and by the early 2000s was one of the ten most frequently used names for baby girls.
-Broadly speaking, in the late twentieth century Americans girls became increasingly more likely than American boys to receive a "trendy" name, and once that name was associated with females it was unlikely to be given to male babies.
+Broadly speaking, in the late twentieth century Americans girls became increasingly more likely than American boys to receive a "trendy" name, and once that name was associated with women it was unlikely to be given to male babies.
 These names present a moving target for researchers hoping to use them to infer additional information about the people under study.
@@
 An education policy analyst studying the demographics of different urban school districts, for instance, can turn to a variety of tools that use current databases of names, and these tools will likely identify the gender of Leslie Fay correctly.
-One online service, [Genderize.io](http://genderize.io/), determines the gender of first names "from user profiles across major social networks." Checking the service for the name Leslie returns a prediction that the name is female, along with an estimate of male/female proportions of the uses among members of unnamed social media sites.^[In this example the Genderize.io service has been accessed through an API wrapper provided in our Gender package described below.]
+One online service, [Genderize.io](http://genderize.io/), determines the gender of first names "from user profiles across major social networks." [Genderize.io n.d.] Checking the service for the name Leslie returns a prediction that the name is female, along with an estimate of male/female proportions of the uses among members of unnamed social media sites.^[In this example the Genderize.io service has been accessed through an API wrapper provided in our Gender package described below.]
 
@@
 But researchers studying a longer timespan need to take into account the changing nature of naming practices.
-If that same education researcher wanted to compare the demographics of contemporary urban schools with those from the early twentieth century, contemporary tools like Genderize.io would misidentify Leslie Fay's gender in historical datasets.
+If that same education researcher wanted to compare the demographics of contemporary urban schools with those from the early twentieth century, existing tools like Genderize.io would misidentify Leslie Fay's gender in historical datasets.
 The "Leslie" problem is not just for researchers who think of themselves as historians.
 Anyone studying a period longer than a few years, or anyone studying a group whose demographics do not match the groups used by a contemporary tool like Genderize.io will also encounter this problem.
-As of 2012, average Americans have a lifespan of nearly seventy-nine years---more than enough time for naming practices to change quite dramatically between their births and deaths [World Bank 2014].
+As of 2012, the average American lifespan was nearly seventy-nine years---more than enough time for naming practices to change quite dramatically between their births and deaths [World Bank 2014].
 Predicting gender from first names therefore requires a fundamentally historical method.
@@
 As previously mentioned, the history of how the Social Security Administration collected the data effects its validity.
-Specifically, because the data extends back to 1880 but the first applications were gathered after 1935, the sex ratios in the dataset are skewed, as demonstrated in this plot.
+Specifically, because the data extends back to 1880 but the first applications were gathered after 1935, the sex ratios in the dataset are skewed in the years before 1930.
 
@@
 
-It is extremely improbable that nearly 70 percent of the people born in 1900 were female.
+It is extremely improbable that nearly seventy percent of the people born in 1900 were female.
 Exactly why the dataset has this bias is unclear, though we speculate that it is because the applicants for Social Security in the early years of the program were approaching retirement age, which was set much closer to the average life expectancy in 1935 than it is today.
@@
 The solution to this problem is two-fold.
-First, we recommend that researchers use the IPUMS-USA dataset to make predictions for years from 1790 to 1930 (which avoids the "bubble" in the SSA data) and that they use the SSA dataset for years since 1930.^[The IPUMS-USA dataset includes only apparently random fluctuations in the gender ratio in any given year.] Second, we have built in a correction to the SSA dataset when using the `gender()` function.
+First, we recommend that researchers use the IPUMS-USA dataset to make predictions for years from 1790 to 1930 (which avoids the "bubble" in the SSA data) and that they use the SSA dataset for years after 1930.^[The IPUMS-USA dataset includes only apparently random fluctuations in the gender ratio in any given year.] Second, we have built in a correction to the SSA dataset when using the `gender()` function.
 If we assume that the secondary sex ratio (that is, the ratio of male to female births) in any given year does not deviate from `0.5` (that is, equality), it is possible to calculate a correction factor for each year or range of years to even out the dataset.
@@
 This method is potentially useful to anyone who has a dataset of names, especially if those names can be associated with particular dates.
-As we have demonstrated in our study of gatekeeping in the American historical profession, our method can reveal important details about gender even from datasets which do not otherwise include that information, and it can do so over long historical periods.
+As we have demonstrated in our study of gatekeeping in the American historical profession, our method can reveal important insights about gender even from datasets which do not otherwise include that information, and it can do so over long historical periods.
 

### 2014-09-29 article.Rmd
Deleted faceted chart section for AHR

@@
 
-The path from a dissertation to a book review in the *American Historical Association* is not automatic.
+The path from a dissertation to a monograph to a book review in the *American Historical Association* isn't automatic.
 Not all dissertations are published as books, not all books are submitted to the *American Historical Review*, and not all book submissions are accepted by the *AHR* editors for review.
@@
   short_scale
-```
-
-The *AHR* also categorizes its reviews under different historical sub-fields.
-Although the journal’s taxonomy has changed quite a bit over time, it did allow us to refine our analysis.
-Were there more reviews of female authors in certain sub-fields rather than others, and has that changed over the past forty years?
-Caribbean/Latin American history has had something approaching equal representation for the past decade-and-a-half.
-In both African history and Ancient/Medieval history female historians made some quite dramatic gains during the late-nineties and aughts.
-The most uneven parties, however, are also the two subject categories that publish the most book reviews in the *AHR*: Modern/Early Modern Europe and the United States/Canada.
-Both sub-fields have made progress over the four decades but still hover at around two-thirds male.
-Gender representation is not evenly distributed across the profession.
-
-```{r reviews-by-category}
-categories <- read.csv("data/ahr-categories.csv", stringsAsFactors = FALSE)
-
-ahr_predicted %>%
-  left_join(categories, by = "category") %>%
-  filter(!our_category %in% c(NA, 
-                              "Articles",
-                              "Latin and South America, Carribean",
-                              "Middle East")) %>%
-  select(year, gender_author, our_category) %>% 
-  group_by(year, gender_author, our_category) %>%  
-  summarize(n = n()) %>%
-  spread(gender_author, n, fill = 0) %>% 
-  mutate(female_proportion = female / (female + male + unknown),
-         male_proportion = male / (female + male + unknown),
-         unknown_proportion = unknown / (female + male + unknown)) %>%  
-  select(female_proportion, male_proportion, unknown_proportion,
-         year, our_category) %>% 
-  gather(gender, proportion, -year, -our_category) %>%
-ggplot(aes(x = year, y = proportion, fill = gender)) + 
-  geom_bar(binwidth = 1, stat = "identity") +
-  facet_wrap(~ our_category) +
-  xlim(1970, 2012) +
-  ggtitle("Gender of AHR authors by book category, 1970-2012") +
-  xlab(NULL) + ylab(NULL) +
-  theme(legend.position = "bottom") +
-  geom_hline(yintercept = 0.5) +
-  scale_fill_brewer(palette = article_palette,
-                    labels = c("Female", "Male", "Unknown"),
-                    name = "Gender") +
-  article_theme +
-  theme(strip.background = element_rect(color = "white"))
 ```

### 2014-09-29 article.Rmd
Only use authors with known gender in "gap" chart

--- 2014-09-29 article.Rmd
@@
 In book reviews published in the *AHR* male authors continue to outnumber female authors by a factor of nearly 2 to 1.
-The long-term gains made by female doctoral students do not carry over in the pages of the historical profession's leading journal.
+The long-term gains made by female doctoral students do not carry over in the pages of the historical profession's leading journal.^[The chart below includes only authors whose gender we successfully inferred; it does not include authors of unknown gender as in the previous charts.]
 
 ```{r}
-diss_proportions$type <- "dissertation"
-ahr_proportions$type <- "book"
+ahr_proportions_no_unknown <- ahr_predicted %>%
+  select(year, gender_author) %>%  
+  group_by(year, gender_author) %>% 
+  summarize(n = n()) %>% 
+  spread(gender_author, n, fill = 0) %>%
+  mutate(female_proportion = female / (female + male),
+         male_proportion = male / (female + male)) %>%
+  select(female_proportion, male_proportion, year) %>%
+  gather(gender, proportion, -year)
 
-combined_proportions <- rbind(diss_proportions, ahr_proportions)
+diss_proportions_no_unknown <- diss_predicted %>%
+  select(year, gender_author) %>%
+  filter(year >= 1950) %>%
+  group_by(year, gender_author) %>% 
+  summarize(n = n()) %>% 
+  spread(gender_author, n, fill = 0) %>% 
+  mutate(female_proportion = female / (female + male),
+         male_proportion = male / (female + male)) %>%
+  select(female_proportion, male_proportion, year) %>%
+  gather(gender, proportion, -year)
+
+diss_proportions_no_unknown$type <- "dissertation"
+ahr_proportions_no_unknown$type <- "book"
+
+combined_proportions <- rbind(diss_proportions_no_unknown,
+                              ahr_proportions_no_unknown)
 
 combined_proportions %>%
-  filter(gender != "unknown_proportion") %>%
   ggplot(aes(x = year, y = proportion, color = gender, size = type)) + 

### 2014-09-29 article.Rmd
Note to add later
Order all bar plots female, unknown, male
This permits both male and female to be compared to the 0.5 line.

@@
 diss_predicted %>%
+  mutate(gender_author = factor(gender_author, 
+                                levels = c("female", "unknown", "male"))) %>%
   ggplot(aes(x = year, fill = gender_author)) + 
@@
   scale_fill_brewer(palette = article_palette, name = "",
-                      labels = c("Female", "Male")) +
+                      labels = c("Female", "Unknown", "Male")) +
   article_theme
@@
 
+diss_proportions$gender <- factor(diss_proportions$gender, 
+                                  levels = c("female_proportion", 
+                                             "unknown_proportion", 
+                                             "male_proportion"))
+
 diss_proportions  %>%
-  ggplot(aes(x = year, y = proportion, fill = gender)) +
+  ggplot(aes(x = year, y = proportion, fill = gender, order = gender)) +
   geom_bar(binwidth = 1, stat = "identity") +
@@
   scale_fill_brewer(palette = article_palette,
-                    labels = c("Female", "Male", "Unknown")) +
+                    labels = c("Female", "Unknown", "Male")) +
   article_theme
@@
 
+ahr_proportions$gender <- factor(ahr_proportions$gender, 
+                                 levels = c("female_proportion", 
+                                            "unknown_proportion", 
+                                            "male_proportion"))
+
 ahr_proportions %>%
-  ggplot(aes(x = year, y = proportion, fill = gender)) + 
+  ggplot(aes(x = year, y = proportion, fill = gender, order = gender)) + 
   geom_bar(binwidth = 1, stat = "identity") +
@@
   scale_fill_brewer(palette = article_palette,
-                      labels = c("Female", "Male", "Unknown")) +
+                      labels = c("Female", "Unknown", "Male")) +
   article_theme
@@
   gather(gender, proportion, -year) %>%
-  ggplot(aes(x = year, y = proportion, fill = gender)) + 
+  mutate(gender = factor(gender, levels = c("female_proportion",
+                                            "unknown_proportion",
+                                            "male_proportion"))) %>%
+  ggplot(aes(x = year, y = proportion, fill = gender, order = gender)) + 
   geom_bar(binwidth = 1, stat = "identity") +
@@
   scale_fill_brewer(palette = article_palette, 
-                    labels = c("Female", "Male", "Unknown")) +
+                    labels = c("Female", "Unknown", "Male")) +
   article_theme +

### 2014-09-29 article.Rmd

--- 2014-09-29 article.Rmd
@@
 article_palette <- "Dark2"
+f_color <- "#8E001C"
+u_color <- "#D3CEAA"
+m_color <- "#4C4C4C"
+tri_color <- c(f_color, u_color, m_color) 
 ```
@@
   short_scale + theme(legend.position = "bottom") +
-  scale_fill_brewer(palette = article_palette, name = "",
-                      labels = c("Female", "Male")) +
+  scale_fill_manual(values=tri_color, name = "",
+                    labels = c("Female", "Unknown", "Male")) +
+  guides(fill = guide_legend(reverse = TRUE)) +
+  #scale_fill_brewer(palette = article_palette, name = "",
+  #                    labels = c("Female", "Unknown", "Male")) +
   article_theme
@@
   geom_hline(yintercept = 0.5) +
-  scale_fill_brewer(palette = article_palette,
-                    labels = c("Female", "Male", "Unknown")) +
+  scale_fill_manual(values=tri_color, name = "",
+                    labels = c("Female", "Unknown", "Male")) +
+  guides(fill = guide_legend(reverse = TRUE)) +
+  #scale_fill_brewer(palette = article_palette,
+  #                  labels = c("Female", "Unknown", "Male")) +
   article_theme
@@
   geom_hline(yintercept = 0.5) +
-  scale_fill_brewer(palette = article_palette,
-                      labels = c("Female", "Male", "Unknown")) +
+  scale_fill_manual(values=tri_color, name = "",
+                    labels = c("Female", "Unknown", "Male")) +
+  guides(fill = guide_legend(reverse = TRUE)) +
+  #scale_fill_brewer(palette = article_palette,
+  #                    labels = c("Female", "Unknown", "Male")) +
   article_theme
@@
   geom_line() +
-  short_scale +
+  scale_x_continuous(breaks = seq(1950, 2010, 10),
+                     labels = seq(1950, 2010, 10),
+                     limits = c(1950, 2025)) +
   scale_size_discrete(name = "Type",
                       labels = c("Book", "Dissertation"), range = c(2,1)) +
-  scale_color_brewer(palette = article_palette, name = "Gender",
+  scale_color_manual(values=c(f_color, m_color), name = "Gender",
                     labels = c("Female", "Male")) +
+  guides(color=FALSE, size=FALSE) +
+  #guides(color = guide_legend(reverse = TRUE)) +
+  annotate("text", x = 2017, y = 0.77, label = "Male Authors\n (Books)", size=5, fontface="bold", color = m_color) +
+  annotate("text", x = 2020, y = 0.57, label = "Male Authors\n (Dissertations)", size=4, color = m_color) +
+  annotate("text", x = 2017, y = 0.23, label = "Female Authors\n (Books)", size=5, fontface="bold", color = f_color) +
+  annotate("text", x = 2020, y = 0.43, label = "Female Authors\n (Dissertations)", size=4, color = f_color) +
+  #scale_color_brewer(palette = article_palette, name = "Gender",
+  #                     labels = c("Female", "Male")) +
   xlab(NULL) + ylab("Proportion of authors by gender") +
-  ggtitle("The gap between the genders of book and dissertation authors") + 
+  ggtitle("Gender Gap of Dissertations and of Books Reviewed in the AHR") +
   article_theme +
@@
   geom_hline(yintercept = 0.5) +
-  scale_fill_brewer(palette = article_palette, 
-                    labels = c("Female", "Male", "Unknown")) +
+  scale_fill_manual(values=tri_color, name = "",
+                    labels = c("Female", "Unknown", "Male")) +
+  guides(fill = guide_legend(reverse = TRUE)) +
+  #scale_fill_brewer(palette = article_palette, 
+  #                  labels = c("Female", "Unknown", "Male")) +
   article_theme +
@@
 Women might be getting hired at comparable rates to men, but discriminatory gatekeeping remains alive and well in the profession.
+
 Our program allows us to quantitatively measure the scope of the problem on a large scale and how it has changed over time.
+[to write later: include thoughts how this is even more important as gender discrimination turns away from the out-in-the-open kind of discrimination to more subtle disadvantages and systemic problems.]
 

### 2014-10-06 article.Rmd
Minor edits

@@
 This paper addresses two trends in the humanities: the rise of gender as an analytical tool and the proliferation of large, machine-readable datasets.
-It outlines a new method for algorithmically inferring the gender of a name from historical data in order to take into account how naming practices change over time and space.
-In contrast to existing methods of predicting gender, this method provides a higher level of accuracy and a precise statement of the likelihood that a name was a given gender, especially for datasets where names can be associated with dates.
+It outlines a new method for algorithmically inferring the gender of a name using historical data.
+In contrast to existing methods of predicting gender, this method takes into account how naming practices change over time and space.
+It provides a higher level of accuracy and a precise statement of the likelihood that a name was a given gender, especially for datasets where names can be associated with dates.
 The article describes the methodology as implemented in the [Gender](https://github.com/ropensci/gender) package for the [R programming language](http://www.r-project.org/).
@@
 In general, names that transition from male to female also became increasingly popular over time.
-For example, the name "Madison" changed from male to female quite suddenly after 1985.
+For example, the name Madison changed from male to female quite suddenly after 1985.
 Whereas before it had been an infrequently used male name, after 1985 it skyrocketed in popularity and by the early 2000s was one of the ten most frequently used names for baby girls.
@@
 These names present a moving target for researchers hoping to use them to infer additional information about the people under study.
-In addition, the prevalance of changes in the gender of names in the last half of the twentieth century implies that gendered lists of names drawn from contemporary data are anachronistic and so of little use for earlier periods.^[The few exceptions of names which flip from being female to male, such as "Jan", have a different causal mechanism.
+In addition, the prevalence of changes in the gender of names in the last half of the twentieth century implies that gendered lists of names drawn from contemporary data are anachronistic and so of little use for earlier periods.^[The few exceptions of names which flip from being female to male, such as "Jan", have a different causal mechanism.
 We speculate that Jan as an English name is female but that Jan as a Scandinavian name is male.
@@
 
-For those working on contemporary subjects the "Leslie" problem is not an especially pressing one.
+For those working on contemporary subjects the "Leslie problem" is not an especially pressing one.
 An education policy analyst studying the demographics of different urban school districts, for instance, can turn to a variety of tools that use current databases of names, and these tools will likely identify the gender of Leslie Fay correctly.
@@
 The "Leslie" problem is not just for researchers who think of themselves as historians.
-Anyone studying a period longer than a few years, or anyone studying a group whose demographics do not match the groups used by a contemporary tool like Genderize.io will also encounter this problem.
+Anyone studying a period longer than a few years, or anyone studying a group whose demographics do not match the groups used by a contemporary tool such as Genderize.io will also encounter this problem.
 As of 2012, the average American lifespan was nearly seventy-nine years---more than enough time for naming practices to change quite dramatically between their births and deaths [World Bank 2014].
@@
 
-Our solution to the "Leslie" problem is to create a software package that combines a predictive algorithm with several historical datasets suitable to various times and regions.
+Our solution to the "Leslie problem" is to create a software package that combines a predictive algorithm with several historical datasets suitable to various times and regions.
 The algorithm calculates the proportion of male or female uses of a name in a given birth year or range of years.
@@
 For example, our program predicts that a person named Leslie born in 1950 is female, but only just barely: 52 percent of the babies named Leslie born in the United States in 1950 were girls and 48 percent were boys.
-Including these probability figures allows researchers to conduct more nuanced analysis and to determine for themselves what level of certainty is acceptable for predicting the gender of names.
+Including these probability figures allows researchers to conduct more nuanced analyses and to determine for themselves what level of certainty is acceptable for predicting the gender of names.
 
@@
 Census Bureau.
-For the category of gender (as for race) states have offered a limited number of boxes to check, and those boxes extert a tremendous power to define state subjects [Brubaker and Cooper 2000, 14-17, Pascoe 2009, Canaday 2011, Landrum 2014].
+For the category of gender (as for race) states have offered a limited number of boxes to check, and those boxes exert a tremendous power to define state subjects [Brubaker and Cooper 2000, 14-17, Pascoe 2009, Canaday 2011, Landrum 2014].
 This method does not presuppose that the gender binary imposed by governments and their wider societies reflect how an individual identifies oneself and is neutral as to what the categories "male" or "female" actually mean in any given historical context.
@@
 
-Why do scholars need a historical method for inferring gender?
+Why do scholars need a method for inferring gender that relies on historical data?
 Let's start with a comparison.
@@
 The most significant problem with the Kantrowitz names corpus and thus the NLTK implementation is that it assumes that names are timeless.
-As pointed out above, this makes the corpus problematic for historical purposes.
+As pointed out above, this makes the corpus problematical for historical purposes.
 Furthermore the Kantrowitz corpus includes other oddities which make it less useful for research.
@@
 Assuming for the moment that our method provides more accurate results, we estimate that `r percent(false_either_percent)` percent of the names in the Kantrowitz corpus are classified ambiguously when a gender could be reasonably predicted from the name, that `r percent(false_male_percent)` percent of the names are classified as male when they should be classified as female, and that `r percent(false_female_percent)` are classified as female when they should be classified as male.
-This error rate is a separate concern from the much smaller size of the Kantrowitz corpuse.^[This error rate was calculated using the Social Security Administration database, assuming a range of years between 1932 and 2012, which is the default setting within the Gender package.
+This error rate is a separate concern from the much smaller size of the Kantrowitz corpus.^[This error rate was calculated using the Social Security Administration database, assuming a range of years between 1932 and 2012, which is the default setting within the Gender package.
 Of course, the error rate would fluctuate for different time periods and regions.]
@@
 The Genderize.io API is, for example, a more sophisticated implementation of gender prediction than the NLTK algorithm.
-Besides predicting male or female for gender, it also reports the proportion of male or female names, along with a count of the number of uses of the name that it is basing its prediction on.
+Besides predicting male or female for gender, it also reports the proportion of male and female names, along with a count of the number of uses of the name on which it is basing its prediction.
 Genderize will also permit the user to customize a prediction for different countries, which is an important feature.
@@
 R and Python are two of the most commonly used languages for data analysis [Piatesky 2014, Muenchen 2014-].^[Surveys and analysis of useage point out the growth of R and the continuing popularity of Python for "data science" generally.] Python's existing packages for gender prediction all implement a method similar to the NLTK or Genderize.io.^[These Python libraries include [estimate.gender](https://pypi.python.org/pypi/estimate.gender/0.4), [genderize](https://github.com/SteelPangolin/genderize), and [gender.py](https://github.com/block8437/gender.py)] To our knowledge, [CRAN](http://cran.rstudio.com/) (the central repository for R packages) does not include any packages dedicated to gender prediction.
-Thus two of the most popular data analysis languages used in the digital humanities currently have no satisfactory method for predicting gender from names for historical and humanistic research.
+Thus two of the most popular data analysis languages used in the digital humanities currently have no satisfactory method for predicting gender from names for historical and humanities research.
 
@@
 We have plans, however, to extend the underlying data to six other European and North American countries included in the [North Atlantic Population Project](https://www.nappdata.org/napp/) for the nineteenth century.
-We also welcome contributions of datasets for other regions and times from other scholars.] The first dataset contains names of applicants for Social Security and available from [Data.gov](https://www.data.gov/).
-The second dataset contains names gathered in the decennial censuses and available from the IPUMS-USA (Integrated Public Use Microdata Series) service from the Minnesota Population Center at the University of Minnesota.
+We also welcome contributions of datasets for other regions and times from other scholars.] The first dataset contains names of applicants for Social Security and is available from [Data.gov](https://www.data.gov/).
+The second dataset contains names gathered in the decennial censuses and is available from the IPUMS-USA (Integrated Public Use Microdata Series) service from the Minnesota Population Center at the University of Minnesota.
 
@@
 The description "baby names" provided by the SSA is a serious misnomer.
-When Social Security became available during the New Deal, its first beneficiaries were adults past or near retirment age.
+When Social Security became available during the New Deal, its first beneficiaries were adults past or near retirement age.
 The dataset goes back to 1880, the birth year for a 55 year-old adult when Social Security was enacted.
@@
 This data has been aggregated by IPUMS at the University of Minnesota and is released as a sample of the total census data.
-Unlike the SSA dataset, which includes a 100% sample for every reported to the Social Security Administration and used more than five times, the IPUMS data contains 5% or 10% samples.
+Unlike the SSA dataset, which includes a 100% sample for every name reported to the Social Security Administration and used more than five times, the IPUMS data contains 5% or 10% samples of names from the total census data.
 Because the `gender()` function relies on proportions of uses of names, rather than raw counts of people with the names, the sampling does not diminish the reliability of the function's predictions [Ruggles et al.
@@
 
-These two datasets (and any future datasets to be added to the package) thus contain a simple table of data.^[The format of this data departs from the more generally desirable "tidy data" framework proposed by Hadley Wickham for performance reasons.
+These two SSA and IPUMS datasets (and any future datasets to be added to the package) thus contain a simple table of data.^[The format of this data departs from the more generally desirable "tidy data" framework proposed by Hadley Wickham for performance reasons.
 By keeping the data in a "wider" format instead of a tidy format as defined by Wickham, the function is able to make its predictions faster [Wickham 2014b].] The columns contain the `name`, `year`, and number of `female` and `male` uses of that name in a particular year.
@@
 Let's begin by assuming that we want to predict the gender of someone named Sidney who was born in 1935 using the Social Security Administration dataset.
-Because the dataset contains a list of names for each year, we can simply lookup the row for Sidney in 1935.
+Because the dataset contains a list of names for each year, we can simply look up the row for Sidney in 1935.
 Using the [dplyr](https://github.com/hadley/dplyr) package for R, which provided a grammar for data manipulation, this can be expressed with the action "filter":
@@
 Thus, according to the Social Security Administration, there were 974 boys and 93 girls named Sidney born in 1935.
-We can add another command to calculate the *proportion* of females and males ("mutate" in dplyr's vocabulary) rather than raw numbers.
+We can add another command to calculate the *proportion* of females and males ("mutate" in the [dplyr](https://github.com/hadley/dplyr) package vocabulary) rather than raw numbers.
 
@@
 
-Next we can sum up the male and female columns ("summarize" in dplyr's vocabulary) and calculate the proportions of female and male uses of "Sidney" during that decade.
+Next we can sum up the male and female columns ("summarize" in [dplyr](https://github.com/hadley/dplyr) package vocabulary) and calculate the proportions of female and male uses of "Sidney" during that decade.
 
@@
 
-In other words, for the decade of the 1930s, we can calculate that there is a `r gender("sidney", year = c(1930, 1939))$proportion_male %>% percent()` percent chance that person named Sidney was male.
+In other words, for the decade of the 1930s, we can calculate that there is a `r gender("sidney", year = c(1930, 1939))$proportion_male %>% percent()` percent chance that a person named Sidney was male.
 This is roughly the same as the probability we calculated above for just 1935, but our method also returns the figures it used to calculate those probabilities: 1,067 instances of "Sidney" in 1935 versus 10,110 total instances for the decade as a whole.
@@
 The method's real utility stems from being able to process larger datasets than a single name.
-Let's use, say, a hypothetical list of editors from a college newspaper to illustrate how a researcher might apply it to their own data.
+Let's use, for example, a hypothetical list of editors from a college newspaper to illustrate how a researcher might apply it to their own data.
 The package's prediction function allows researchers to choose which reference datasets they would like to use and the range of years for making their predictions.^[See the package vignette for more details: `vignette("predicting-gender", package = "gender")`.]
@@
 
-This basic call of the `gender()` function would be suitable if we did not know the birth year for each editor. But we can also specify that in the `gender()` function by using this call:
+This basic call of the `gender()` function would be suitable if we did not know the birth year for each editor. But we can also specify that information in the `gender()` function by using this call:
 
@@
 Specifically, because the data extends back to 1880 but the first applications were gathered after 1935, the sex ratios in the dataset are skewed in the years before 1930.
+For example, this dataset implies that thirty percent of the people born in 1900 were male.
 
@@
 What follows is a quantitative supplement to Lunbeck's analysis that uses our program to analyze gender representation amongst historians across a much larger scale and a much longer time period.
-We have found that although women have acheived rough equality in terms of the number of dissertations written, there is still a significant gap between male and female authorship of books in history.
+We have found that although women have achieved rough equality in terms of the number of dissertations written, there is still a significant gap between male and female authorship of books in history.
 
@@
   geom_bar(binwidth = 1, stat = "identity") +
-  ggtitle("Proportion of dissertations about history by gender, 1950-2012") +
+  ggtitle("Proportion of dissertations about history by gender 1950-2012") +
   xlab(NULL) + ylab(NULL) +
@@
 The "coin of the realm" for many historians remains the publication of a scholarly monograph to be read and evaluated by peers.
-This process often takes the form of book reviews published in academic journals.
+These coins are often measured takes the form of book reviews published in academic journals.
 One of the leading journals in the historical profession is the *American Historical Review*, the flagship journal of the American Historical Association.
 Along with a few dozen articles published in its five issues each year, the *AHR* publishes roughly one thousand books reviews per year covering (in its own words) "every major field of historical study." The *AHR* is not only one of the widest-ranging journals in the profession, it is also the oldest; the journal has been publishing continuously since 1895.
-The range, scope, and length of the *AHR* makes it an ideal source to analyze the historical profession on a large scale.
+The range, scope, and longevity of the *AHR* makes it an ideal source to analyze the historical profession on a large scale.
 
@@
 Even if it garners a negative review, the fact that it appears in the journal at all is a measure of the fact that the profession's gatekeepers have deemed it important enough to review.
-On the flip side, a book that does not appear in the *AHR* may have gone on to have a significant impact on the profession.
-It is precisely this professional gatekeeping dimension that makes the AHR useful to study in the context of gender representation.
+To be fair, a book that does not appear in the *AHR* may have gone on to have a significant impact on the profession.
+It is precisely this professional gatekeeping dimension that makes the *AHR* useful for studying gender representation.
 
@@
 In book reviews published in the *AHR* male authors continue to outnumber female authors by a factor of nearly 2 to 1.
-The long-term gains made by female doctoral students do not carry over in the pages of the historical profession's leading journal.^[The chart below includes only authors whose gender we successfully inferred; it does not include authors of unknown gender as in the previous charts.]
+The long-term gains made by female doctoral students do not carry over into the pages of the historical profession's leading journal.^[The chart below includes only authors whose gender we successfully inferred; it does not include authors of unknown gender as in the previous charts.]
 
@@
 Establishing a single causal reason for this pattern is all but impossible, and instead a murky concoction of interrelated factors likely contributes to the gap.
-For one, a 2010 survey by the American Historical Association revealed the female historians dedicated substantially more time to child and elder care than their male colleagues, leaving less time for research.
+For one, a 2010 survey by the American Historical Association revealed that female historians dedicated substantially more time to child and elder care than their male colleagues, leaving less time for research.
 This gender gap is more likely to impact the production of a book than a dissertation, as graduate students are generally younger and less likely to be juggling child-rearing responsibilities while writing than their older peers.
-More troublingly, women continue to face gender discrimination in the professional sphere.
+More troubling, women continue to face gender discrimination in the professional sphere.
 
 We used our program to infer the gender of not just book authors reviewed by the *AHR*, but the gender of the reviewers themselves.
-The story is much the same for reviewers as it is for authors: more than than twice as many men as women appear as reviewers in the journal.
+The story is much the same for reviewers as it is for authors: more than twice as many men as women appear as reviewers in the journal.
 But gender inflects book reviews in less direct ways by shaping who writes reviews of which book authors.
@@
 The temporal dimension of our work matches historical questions to a historical method, providing much improved results over simple, anachronistic lookup methods.
-This method is extensible to any time or place for which there is a suitable dataset of names from which to make predictions about gender and naming practices.
+This method is extensible to any time or place for which there is a suitable dataset of names from which to make predictions about gender.
 This method is potentially useful to anyone who has a dataset of names, especially if those names can be associated with particular dates.

### 2014-10-06 article.Rmd
Adding more context, citations

@@
 
-This paper addresses two trends in the humanities: the rise of gender as an analytical tool and the proliferation of large, machine-readable datasets.
+This paper addresses two trends in the humanities: the use of gender as an analytical framework and the proliferation of large, machine-readable datasets.
 It outlines a new method for algorithmically inferring the gender of a name using historical data.
@@
 It goes on to apply the method to a case study in which we study gender and gatekeeping in the history profession by inferring the gender of authors of history dissertations and of authors and reviewers in the *American Historical Review*.
+The Gender package illustrates how humanities scholars can make concrete methodologial contributions to computer science and related fields.
 
-### Introduction: The "Leslie" Problem
+### Introduction: The "Leslie Problem"
 
@@
 If that same education researcher wanted to compare the demographics of contemporary urban schools with those from the early twentieth century, existing tools like Genderize.io would misidentify Leslie Fay's gender in historical datasets.
-The "Leslie" problem is not just for researchers who think of themselves as historians.
+The "Leslie problem" is not just for researchers who think of themselves as historians.
 Anyone studying a period longer than a few years, or anyone studying a group whose demographics do not match the groups used by a contemporary tool such as Genderize.io will also encounter this problem.
@@
 
-The Social Security Administration (SSA) "[Baby Names](http://catalog.data.gov/dataset/baby-names-from-social-security-card-applications-national-level-data)" dataset was created as a result of the Social Security Act of 1935 during the New Deal.
-^[The Social Security Administration also provides a related dataset which contains the names of applicants broken down by state and year since 1910 [Social Security Administration 2014-b].] This dataset contains a list of first names with the number of male and female uses of the name in every year beginning with 1880 and ending with 2012.
+The Social Security Administration (SSA) [Baby Names](http://catalog.data.gov/dataset/baby-names-from-social-security-card-applications-national-level-data) dataset was created as a result of the Social Security Act of 1935 during the New Deal.
+^[The Social Security Administration also provides a related dataset which contains the names of applicants broken down by state and year since 1910 [Social Security Administration 2014-b].] This dataset contains a list of first names along with how many times each name was assigned to each gender in every year beginning with 1880 and ending with 2012.
 The SSA list includes any name which was used more than five times in a given year, thereby capturing all but the most infrequently used names from each year.
@@
 The "coin of the realm" for many historians remains the publication of a scholarly monograph to be read and evaluated by peers.
-These coins are often measured takes the form of book reviews published in academic journals.
+Once published, the impact of a book is often shaped by its reviews in academic journals.
 One of the leading journals in the historical profession is the *American Historical Review*, the flagship journal of the American Historical Association.
 Along with a few dozen articles published in its five issues each year, the *AHR* publishes roughly one thousand books reviews per year covering (in its own words) "every major field of historical study." The *AHR* is not only one of the widest-ranging journals in the profession, it is also the oldest; the journal has been publishing continuously since 1895.
-The range, scope, and longevity of the *AHR* makes it an ideal source to analyze the historical profession on a large scale.
+The range, scope, longevity, and reputation of the *AHR* makes it an ideal proxy to study gatekeeping within the larger historical profession.
 
-We take the <i>American Historical Review</i> as a rough proxy for gatekeeping within the larger historical profession.
-When a book appears in the *AHR* it serves as a signal that other historians in positions of power have "approved" it for consideration.
-This signal is not necessarily correlated with the quality of the book.
+A few caveats are in order.
+First, a book that never appears in the *AHR* at all may still go on to have a significant impact on the profession.
+Second, a book's appearance in the *AHR* is not necessarily correlated with the *quality* of the book, as the journal prints negative reviews as well as positive ones.
+Nevertheless, when a book appears in the *AHR* it serves as a signal that other historians in positions of power have "approved" it for consideration.
 Even if it garners a negative review, the fact that it appears in the journal at all is a measure of the fact that the profession's gatekeepers have deemed it important enough to review.
-To be fair, a book that does not appear in the *AHR* may have gone on to have a significant impact on the profession.
-It is precisely this professional gatekeeping dimension that makes the *AHR* useful for studying gender representation.
+It is precisely this professional gatekeeping dimension that we can analyze using the Gender package.
 
 Scraping the table-of-contents of every issue of the American Historical Review results in a dataset of close to 60,000 books reviewed by the journal since it began publication in 1895.
-Our program then inferred the gender of the authors of these books, which we could then use to plot the proportion of female and male authors over time.
+Our program then inferred the gender of the authors of these books, which we could in turn use to plot the proportion of female and male authors over time.
 The temporal trajectory of gender representations roughly resembles that of history dissertations: women began making inroads in the late 1960s and have made steady gains over the past four decades.
@@
 Establishing a single causal reason for this pattern is all but impossible, and instead a murky concoction of interrelated factors likely contributes to the gap.
-For one, a 2010 survey by the American Historical Association revealed that female historians dedicated substantially more time to child and elder care than their male colleagues, leaving less time for research.
-This gender gap is more likely to impact the production of a book than a dissertation, as graduate students are generally younger and less likely to be juggling child-rearing responsibilities while writing than their older peers.
-More troubling, women continue to face gender discrimination in the professional sphere.
+For one, a 2011 survey by the American Historical Association revealed that female historians dedicated substantially more time to child and elder care than their male colleagues, leaving less time for research.
+[Townsend 2012] This gender gap is more likely to impact the production of a book than a dissertation, as graduate students are generally younger and less likely to be juggling child-rearing responsibilities while writing than their older peers.
+Their male colleagues, meanwhile, enjoy more time to write and receive more accolades for their writing.
+Since 2004, the Pulitzer Prize for History has been awarded to just one female historian ["The Pulitzer Prizes: History" 2014-], while the Bancroft Prize in American History has been awarded to 24 men and 6 women over that same period ["Bancroft Prize"].
+The *American Historical Review* is, unfortunately, far from alone in its gatekeeping practices.
 
-We used our program to infer the gender of not just book authors reviewed by the *AHR*, but the gender of the reviewers themselves.
+We went on to use our program to infer the gender of not just book authors reviewed by the *AHR*, but the gender of the reviewers themselves.
 The story is much the same for reviewers as it is for authors: more than twice as many men as women appear as reviewers in the journal.
@@
 
+["Bancroft Prize"] "Bancroft Prize." *Wikipedia* <http://en.wikipedia.org/wiki/Bancroft_Prize> Accessed 5 October 2014.
+
 [Bird et al. 2009] Bird, Steve, Edward Loper, and Ewan Klein. *Natural Language Processing with Python* (O’Reilly, 2009).
@@
 
+["The Pulitzer Prizes: History" 2014-] "The Pulitzer Prizes: History." 2014-. <http://www.pulitzer.org/bycat/History> Accessed 5 October 2014.
+
 [R Core Team 2014-] R Core Team. *R: A language and environment for statistical computing*. R Foundation for Statistical Computing, Vienna, Austria, 2014. <http://www.R-project.org/>
@@
 
-[Weingart 2014] Weingart, Scott. "The Moral Role of DH in a Data-Driven World." *the scottbot irregular*. 9 September 2014, http://www.scottbot.net/HIAL/?p=40944
+[Townsend 2012] Townsend, Robert B.
+"What Makes a Successful Academic Career in History?
+A Field Report from the Higher Ranks." *Perspectives on History*.
+December 2012, <http://www.historians.org/publications-and-directories/perspectives-on-history/december-2012/what-makes-a-successful-academic-career-in-history>
+
+[Weingart 2014] Weingart, Scott. "The Moral Role of DH in a Data-Driven World." *the scottbot irregular*. 9 September 2014, <http://www.scottbot.net/HIAL/?p=40944>
 

### 2014-10-06 article.Rmd
Working on conclusion

@@
 Women might be getting hired at comparable rates to men, but discriminatory gatekeeping remains alive and well in the profession.
-
-Our program allows us to quantitatively measure the scope of the problem on a large scale and how it has changed over time.
-[to write later: include thoughts how this is even more important as gender discrimination turns away from the out-in-the-open kind of discrimination to more subtle disadvantages and systemic problems.]
+As explicit instances of sexism fade from view the disadvantages women face are growing more and more opaque.
+But that does not make them any less damaging.
+As discrimination becomes less blatant and more systemic, digital methods like the Gender package are a way to a shine light on these cloaked disadvantages, to reveal the true extent of the problem and the work that remains to be done.
 
@@
 Our algorithm for predicting gender from first names, when combined with historical datasets in the Gender package, provides a powerful new method for researchers interested in the study of gender.
-The digital turn has made available very large datasets, yet in the midst of abundance many humanists continue to deal with the problem of scarcity in their sources.
-Our method uses the abundant existing datasets to infer more information from scarce datasets.
-The temporal dimension of our work matches historical questions to a historical method, providing much improved results over simple, anachronistic lookup methods.
-This method is extensible to any time or place for which there is a suitable dataset of names from which to make predictions about gender.
-This method is potentially useful to anyone who has a dataset of names, especially if those names can be associated with particular dates.
-As we have demonstrated in our study of gatekeeping in the American historical profession, our method can reveal important insights about gender even from datasets which do not otherwise include that information, and it can do so over long historical periods.
+The digital turn has made available very large datasets, yet in the midst of abundance many humanists continue to deal with the problem of incompleteness and scarcity within their sources.
+Our method uses *abundant* existing datasets to infer additional information from *scarce* datasets.
+The temporal dimension of our work matches historical questions to a historical method, providing much improved results over simple, anachronistic, or ahistorical lookup methods.
+This method is also extensible to different times and places, provided a researcher has a similar reference dataset of names and genders from the period or area under study.
+
+The analytical category of gender has transformed research in the humanities for the past half-century.
+More recently, the digitization movement has opened up exciting opportunities to interrogate gender in new ways.
+Projects in the digital humanities have ranged from recovering the voices of early-modern women writers to tracing the linguistic styles of male and female authors [Women Writers Project n.d., Argamon et.
+al.
+2009].
+The Gender package is a new tool to add to this approach.
+It reverses the traditional disciplinary direction bewteen computer science and the humanities.
+Rather than adapting existing computational methodologies, the Gender package uses a humanistic starting point---the historical relationship between gender and naming practices---to contribute a new method to computer science.
 
@@
 Accessed 29 September 2014.
+
+[Argamon et.
+al.
+2009] Argamon, Shlomo, Jean-Baptiste Goulain, Russell Horton, and Mark Olsen.
+"Vive la Différence!
+Text Mining Gender Difference in French Literature" *Digital Humanities Quarterly*.
+3.2 (2009).
 
@@
 
+[Women Writers Project n.d.] Women Writers Project. "About." <http://www.wwp.northeastern.edu/about/>. Accessed 5 October 2014.
+
 [Wood and Townsend 2013] Wood, L.

### 2014-10-06 article.Rmd
More conclusion

@@
 This method is also extensible to different times and places, provided a researcher has a similar reference dataset of names and genders from the period or area under study.
-
+The Gender package reverses the traditional disciplinary direction bewteen computer science and the humanities.
+Rather than adapting existing computational methodologies, it uses a humanistic starting point---the historical relationship between gender and naming practices---to contribute a new method to computer science.
 The analytical category of gender has transformed research in the humanities for the past half-century.
@@
 2009].
-The Gender package is a new tool to add to this approach.
-It reverses the traditional disciplinary direction bewteen computer science and the humanities.
-Rather than adapting existing computational methodologies, the Gender package uses a humanistic starting point---the historical relationship between gender and naming practices---to contribute a new method to computer science.
+The proliferation of machine-readable datasets and digital texts brings a vast trove of material with which to study gender.
+But unlocking its full range of meaning necessitates not just accessing the data that *is* there, but inferring information that *isn't* there.
+The Gender package is a powerful tool to do just that.
 

### 2014-10-07 article.Rmd
Copy edits
Conclusion edits

@@
 One of the leading journals in the historical profession is the *American Historical Review*, the flagship journal of the American Historical Association.
-Along with a few dozen articles published in its five issues each year, the *AHR* publishes roughly one thousand books reviews per year covering (in its own words) "every major field of historical study." The *AHR* is not only one of the widest-ranging journals in the profession, it is also the oldest; the journal has been publishing continuously since 1895.
-The range, scope, longevity, and reputation of the *AHR* makes it an ideal proxy to study gatekeeping within the larger historical profession.
+Along with a few dozen articles published in its five issues each year, the *AHR* publishes roughly one thousand books reviews per year covering (in its own words) "every major field of historical study." [American Historical Review n.d.] The *AHR* is not only one of the widest-ranging journals in the profession, it is also the oldest; the journal has been publishing continuously since 1895.
+The range, scope, longevity, and reputation of the *AHR* makes it an ideal proxy to study gatekeeping within the historical discipline.
 
@@
 Not all dissertations are published as books, not all books are submitted to the *American Historical Review*, and not all book submissions are accepted by the *AHR* editors for review.
-As of September 2014 the journal's website reminds readers that "the AHR receives over 3,000 books a year; we have the resources to publish at most 1,000 reviews a year (approximately 200 per issue)." [American Historical Review n.d.] Fewer and fewer female historians make it through each successive step.
-Establishing a single causal reason for this pattern is all but impossible, and instead a murky concoction of interrelated factors likely contributes to the gap.
-For one, a 2011 survey by the American Historical Association revealed that female historians dedicated substantially more time to child and elder care than their male colleagues, leaving less time for research.
+As of September 2014 the journal's website reminds readers that "the AHR receives over 3,000 books a year; we have the resources to publish at most 1,000 reviews a year (approximately 200 per issue)." [American Historical Review n.d.] But fewer and fewer female historians make it through each successive step.
+Establishing a single causal reason for this pattern is all but impossible, and instead a murky concoction of interrelated factors likely contributes to the successively widening gap.
+A 2011 survey by the American Historical Association revealed that female historians dedicated substantially more time to child and elder care than their male colleagues, leaving them with less time for research.
 [Townsend 2012] This gender gap is more likely to impact the production of a book than a dissertation, as graduate students are generally younger and less likely to be juggling child-rearing responsibilities while writing than their older peers.
-Their male colleagues, meanwhile, enjoy more time to write and receive more accolades for their writing.
-Since 2004, the Pulitzer Prize for History has been awarded to just one female historian ["The Pulitzer Prizes: History" 2014-], while the Bancroft Prize in American History has been awarded to 24 men and 6 women over that same period ["Bancroft Prize"].
-The *American Historical Review* is, unfortunately, far from alone in its gatekeeping practices.
+Their male colleagues, meanwhile, enjoy more time to publish and receive more accolades when they do.
+As of 2014, the Pulitzer Prize for History has been awarded to exactly one female historian in the past nineteen years.
+The Bancroft Prize in American History has been awarded to almost four times as many male historians as female historians over that same period.
+["Bancroft Prize"] The *American Historical Review* is, unfortunately, far from alone in its gatekeeping practices.
 
@@
 The temporal dimension of our work matches historical questions to a historical method, providing much improved results over simple, anachronistic, or ahistorical lookup methods.
-This method is also extensible to different times and places, provided a researcher has a similar reference dataset of names and genders from the period or area under study.
-The Gender package reverses the traditional disciplinary direction bewteen computer science and the humanities.
-Rather than adapting existing computational methodologies, it uses a humanistic starting point---the historical relationship between gender and naming practices---to contribute a new method to computer science.
-The analytical category of gender has transformed research in the humanities for the past half-century.
-More recently, the digitization movement has opened up exciting opportunities to interrogate gender in new ways.
+This method is also extensible to different times and places, provided that a researcher has a suitable reference dataset of names and genders from the period or area under study.
+The Gender package reverses the traditional disciplinary direction between the humanities and computer science.
+Rather than adapting existing computational methodologies, our method uses a humanistic starting point---the historical relationship between gender and naming practices---to contribute a new method to computer science.
+
+The analytical category of gender has transformed humanities research over the past half-century.
+More recently, the digitization movement has opened up opportunities to apply this category to new sources and in new ways.
 Projects in the digital humanities have ranged from recovering the voices of early-modern women writers to tracing the linguistic styles of male and female authors [Women Writers Project n.d., Argamon et.
@@
 2009].
-The proliferation of machine-readable datasets and digital texts brings a vast trove of material with which to study gender.
+The proliferation of machine-readable datasets and digital texts brings with it a vast trove of material with which to study gender.
 But unlocking its full range of meaning necessitates not just accessing the data that *is* there, but inferring information that *isn't* there.
@@
 3.2 (2009).
+<http://www.digitalhumanities.org/dhq/vol/3/2/000042/000042.html>
 
-["Bancroft Prize"] "Bancroft Prize." *Wikipedia* <http://en.wikipedia.org/wiki/Bancroft_Prize> Accessed 5 October 2014.
+["Bancroft Prize"] "Bancroft Prize." *Wikipedia* <http://en.wikipedia.org/wiki/Bancroft_Prize>. Accessed 5 October 2014.
 

### 2014-10-07 article.Rmd
Formatting figures for submission

@@
 author: "Cameron Blevins and Lincoln Mullen"
-output: html_document
+output:
+  html_document: default
 ---
@@
 tri_color <- c(f_color, u_color, m_color)
+
 ```
@@
 
-```{r plot-leslie}
+```{r figure01}
 # We cannot simply use the Social Security figures (even though they probably
@@
 ```
+**Figure 1: Gender of the name Leslie, 1820-2012**
 
@@
 
-```{r plot-names-that-change}
+```{r figure02}
 plot_names <-  c("madison",
@@
 ```
+**Figure 2: Proportion of female uses**
 
@@
 
-```{r}
+```{r figure03}
 genderdata::ssa_national %>%
@@
 ```
+**Figure 3: Percentage of SSA dataset that is female**
 
@@
 
-```{r}
+```{r figure04}
 diss_predicted %>%
@@
 ```
+**Figure 4: Dissertations about history by gender, 1950-2012**
 
@@
 
-```{r}
+```{r figure05}
 diss_proportions <- diss_predicted %>%
@@
 ```
+**Figure 5: Proportion of dissertations about history by gender 1950-2012**
 
@@
 
-```{r}
+```{r figure06}
 ahr_proportions <- ahr_predicted %>%
@@
 ```
+**Figure 6: Authors of books reviewed in the AHR by gender**
 
@@
 
-```{r}
+```{r figure07}
 ahr_proportions_no_unknown <- ahr_predicted %>%
@@
 ```
+**Figure 7: Gender Gap of Dissertations and of Books Reviewed in the *AHR***
 
@@
 
-```{r}
+```{r figure08}
 ahr_predicted %>%
@@
 ```
+**Figure 8: Gender of reviewers of MALE book authors in *AHR***
 
-```{r}
+```{r figure09}
 ahr_predicted %>%
@@
 ```
+**Figure 9: Gender of reviewers of FEMALE book authors in *AHR***
 

### 2014-10-07 article.Rmd
Typos

@@
 The article describes the methodology as implemented in the [Gender](https://github.com/ropensci/gender) package for the [R programming language](http://www.r-project.org/).
-It goes on to apply the method to a case study in which we study gender and gatekeeping in the history profession by inferring the gender of authors of history dissertations and of authors and reviewers in the *American Historical Review*.
+It goes on to apply the method to a case study in which we examine gender and gatekeeping in the history profession by inferring the gender of authors of history dissertations and of authors and reviewers in the *American Historical Review*.
 The Gender package illustrates how humanities scholars can make concrete methodologial contributions to computer science and related fields.
@@
 
-The Social Security Administration (SSA) [Baby Names](http://catalog.data.gov/dataset/baby-names-from-social-security-card-applications-national-level-data) dataset was created as a result of the Social Security Act of 1935 during the New Deal.
-^[The Social Security Administration also provides a related dataset which contains the names of applicants broken down by state and year since 1910 [Social Security Administration 2014-b].] This dataset contains a list of first names along with how many times each name was assigned to each gender in every year beginning with 1880 and ending with 2012.
+The Social Security Administration (SSA) [Baby Names](http://catalog.data.gov/dataset/baby-names-from-social-security-card-applications-national-level-data) dataset was created as a result of the Social Security Act of 1935 during the New Deal.^[The Social Security Administration also provides a related dataset which contains the names of applicants broken down by state and year since 1910 [Social Security Administration 2014-b].] This dataset contains a list of first names along with how many times each name was assigned to each gender in every year beginning with 1880 and ending with 2012.
 The SSA list includes any name which was used more than five times in a given year, thereby capturing all but the most infrequently used names from each year.

### 2015-06-23 article.Rmd
submission files
First crack at revisions
Rewrote abstract, introduction, started to change AHR section

--- 2014-10-07 article.Rmd
@@
   html_document: default
+  word_document: default
 ---
@@
 
-This paper addresses two trends in the humanities: the use of gender as an analytical framework and the proliferation of large, machine-readable datasets.
-It outlines a new method for algorithmically inferring the gender of a name using historical data.
-In contrast to existing methods of predicting gender, this method takes into account how naming practices change over time and space.
-It provides a higher level of accuracy and a precise statement of the likelihood that a name was a given gender, especially for datasets where names can be associated with dates.
-The article describes the methodology as implemented in the [Gender](https://github.com/ropensci/gender) package for the [R programming language](http://www.r-project.org/).
-It goes on to apply the method to a case study in which we examine gender and gatekeeping in the history profession by inferring the gender of authors of history dissertations and of authors and reviewers in the *American Historical Review*.
-The Gender package illustrates how humanities scholars can make concrete methodologial contributions to computer science and related fields.
+This paper outlines a new method for inferring the gender of personal names in large, machine-readable datasets.
+In contrast to existing methods of predicting gender that treat names as static entities, this method uses a humanistic approach to take into account the dynamic nature of naming practices.
+It uses historical data to measure the likelihood that a name was associated with a particular gender based on the time or place under study.
+This approach generates more accurate results for sources that encompass changing periods of time, providing digital humanities scholars with a flexible tool to apply and adapt to analyses of gender.
+The article first describes the methodology as implemented in the [Gender](https://github.com/ropensci/gender) package for the [R programming language](http://www.r-project.org/).
+It goes on to apply the method to a case study in which we examine gender and gatekeeping in the American historical profession over the past half-century.
+The Gender package illustrates the importance of incorporating humanistic approaches to computer science and related fields.
 
@@
 
-As humanities researchers grapple with larger and larger datasets they will increasingly be able to not just extract existing information from that data but to also infer additional information from it [Goldstone and Underwood 2014].^[The availability of large-scale datasets dealing with personal information raises important ethical issues for digital humanists, particularly in light of the revelations about domestic spying at the NSA brought forth by Edward Snowden in 2013.
-[Weingart 2014] Kieran Healy has used network analysis to demonstrate how researchers can infer information from datasets while also touching on the troubling ethical issue of surveillance [Healy 2013].] Many of these datasets deal with lists of people: company payrolls, military rosters, passenger bookings, records of correspondence, or lists of published works.
-Some of these records might contain associated information such as a person's address, age, or rank, while others contain little more than a list of personal names.
-A common question in the history of American religion, for example, is whether more women than men participated in a church or religious group.
-A simple membership roll might contain names, but it is unlikely to identify the gender of the members.
-One might also ask a question about changing patterns in male and female authorship of books.
-Library catalogs contain enormous amounts of data about publication, but they do not record the sex of authors.
+The analytical category of gender has transformed humanities research over the past half-century, and more recently, the digitization movement has opened up opportunities to apply this category to new sources and in new ways.
+Initiatives like the Orlando Project, the Poetess Archive, and the Women Writers Project have done crucial work to recover the voices of female authors and writers [Orlando Project, n.d., Poetess Archive n.d., Women Writers Project n.d.].
+Other scholars have used digital analysis to study linguistic styles of male and female playwrights or representations of gendered bodies in European fairytales [Argamon et.
+al.
+2009, Weingart and Jorgensen 2013].
+The proliferation of machine-readable datasets and digital texts brings with it a vast trove of material with which to study gender.
+Much more, however, can still be done.
+To better equip humanities researchers to take full advantage of digital sources, we have developed a methodology that infers gender from one of the most common features of humanities datasets: personal names.
 
-Given large but sparse datasets, researchers have to use their ingenuity to derive additional kinds of information.
-What can researchers infer from just a name?
-On a basic level, first names often imply a person's gender.
-"Jane Fay" is almost certainly a woman, while her brother "John Fay" is almost certainly a man.
-Given a dataset matching names to genders, a straightforward computer program can easily infer their respective genders.
+As humanities datasets grow larger and larger, researchers have not only been able to access information that is there, but to infer information that isn't there.
+Topic modeling, for example, has been used to uncover scholarship trends over a century of academic articles in literary studies.
+[Goldstone and Underwood 2014].
+Can researchers similarly use personal names to infer information about individual people?
+After all, personal names appear across a wide range of sources in the humanities: literary characters in stories and novels, authorship information in archival metadata, or individuals mentioned by historical newspapers and magazine articles.
+Authors, characters, or historical actors frequently stand at the analytical center of humanities scholarship, and their names provide an intriguing way to infer additional information about them.^[The availability of large-scale datasets dealing with personal information raises important ethical issues for digital humanists, particularly in light of the revelations about domestic spying at the NSA brought forth by Edward Snowden in 2013.
+[Weingart 2014] Kieran Healy has used network analysis to demonstrate how researchers can infer information from datasets while also touching on the troubling ethical issue of surveillance [Healy 2013].]
 
-The problem with inferring gender from names is that the link between gender and naming practices, like language itself, is not static.
+One of the central kinds of information researchers can infer from a personal name is the gender of that individual: "Jane Fay," for instance, is almost certainly a woman, while her brother "John Fay" is almost certainly a man.
+A computer program can infer their respective genders by looking up “Jane” and “John” in an existing dataset that matches names to genders.
+But this approach runs into a problem: the link between gender and naming practices, like language itself, is not static.
 What about Jane and John Fay's sibling, Leslie?
@@
 
-Other names used in the United States such as Madison, Morgan, Sydney, and Kendall, have also flipped their gender over the past century.
-Such changes usually follow set patterns.
+For those working on contemporary subjects the "Leslie problem" is not an especially pressing one.
+A film scholar analyzing male and female characters in modern American television shows, for instance, can turn to a variety of tools that use current databases of names, and these tools will likely identify the gender of Leslie Fay correctly.
+One online service, Genderize.io, determines the gender of first names "from user profiles across major social networks." [Genderize.io n.d.] Checking the service for the name Leslie returns a prediction that the name is female, along with an estimate of male/female proportions of the uses among members of unnamed social media sites.^[In this example the Genderize.io service has been accessed through an API wrapper provided in our Gender package described below.]
+
+**I took out the genderize code here because I think it's not immediately intuitive for the reader, given that we haven't totally introduced our package yet. -CB**
+
+But researchers studying a longer timespan need to take into account the changing nature of naming practices.
+If that same film scholar wanted to compare the characters of modern television shows to characters during the silent film era,  existing tools like Genderize.io would misidentify Leslie Fay's gender in the earlier screenplays.
+Other names used in the United States such as Madison, Morgan, Sydney, and Kendall, have also flipped their gender over the past century and would similarly be misidentified.
+The prevalence of changes in the gender of names in the last half of the twentieth century implies that gendered lists of names drawn from contemporary data are anachronistic and so of limited use for datasets that encompass changing periods of time.^[The few exceptions of names which flip from being female to male, such as "Jan", have a different causal mechanism.
+We speculate that Jan as an English name is female but that Jan as a Scandinavian name is male.
+Because the name has remained relatively uncommon, other causes besides changing gender preferences can explain the change.]
+
+**I propose cutting this part or moving it to another section, as I think we need to move faster to get into the payoff of our method: Such changes usually follow set patterns.
 In general, names tend to shift from being predominantly male to being predominantly female; the shift rarely happens in the other direction.
@@
 Broadly speaking, in the late twentieth century Americans girls became increasingly more likely than American boys to receive a "trendy" name, and once that name was associated with women it was unlikely to be given to male babies.
-These names present a moving target for researchers hoping to use them to infer additional information about the people under study.
-In addition, the prevalence of changes in the gender of names in the last half of the twentieth century implies that gendered lists of names drawn from contemporary data are anachronistic and so of little use for earlier periods.^[The few exceptions of names which flip from being female to male, such as "Jan", have a different causal mechanism.
-We speculate that Jan as an English name is female but that Jan as a Scandinavian name is male.
-Because the name has remained relatively uncommon, other causes besides changing gender preferences can explain the change.]
+These names present a moving target for researchers hoping to use them to infer additional information about the people under study.**
 
@@
 
-For those working on contemporary subjects the "Leslie problem" is not an especially pressing one.
-An education policy analyst studying the demographics of different urban school districts, for instance, can turn to a variety of tools that use current databases of names, and these tools will likely identify the gender of Leslie Fay correctly.
-One online service, [Genderize.io](http://genderize.io/), determines the gender of first names "from user profiles across major social networks." [Genderize.io n.d.] Checking the service for the name Leslie returns a prediction that the name is female, along with an estimate of male/female proportions of the uses among members of unnamed social media sites.^[In this example the Genderize.io service has been accessed through an API wrapper provided in our Gender package described below.]
-
-```{r genderize, echo=TRUE, results='markup'}
-gender("Leslie", method = "genderize")
-```
-
-But researchers studying a longer timespan need to take into account the changing nature of naming practices.
-If that same education researcher wanted to compare the demographics of contemporary urban schools with those from the early twentieth century, existing tools like Genderize.io would misidentify Leslie Fay's gender in historical datasets.
 The "Leslie problem" is not just for researchers who think of themselves as historians.
 Anyone studying a period longer than a few years, or anyone studying a group whose demographics do not match the groups used by a contemporary tool such as Genderize.io will also encounter this problem.
-As of 2012, the average American lifespan was nearly seventy-nine years---more than enough time for naming practices to change quite dramatically between their births and deaths [World Bank 2014].
-Predicting gender from first names therefore requires a fundamentally historical method.
+For example, a literary scholar studying a corpus of poems from the mid-twentieth century might wish to compare the stanza structures of male and female poets.
+Using a tool like Genderize.io based on contemporary data risks misidentifying these writers, many of who were born (and named) more than a century prior to the creation of these modern name datasets.
+This problem will only increase with growing life expectancies: as of 2012, the average American lifespan was nearly seventy-nine years---more than enough time for naming practices to change quite dramatically between a person’s birth and death [World Bank 2014].
+Predicting gender from first names therefore requires a dynamic method that takes into change over time.
 
@@
 For example, our program predicts that a person named Leslie born in 1950 is female, but only just barely: 52 percent of the babies named Leslie born in the United States in 1950 were girls and 48 percent were boys.
-Including these probability figures allows researchers to conduct more nuanced analyses and to determine for themselves what level of certainty is acceptable for predicting the gender of names.
+Researchers can use these probabilities to determine what level of certainty is acceptable for predicting the gender of different names.
 
-We recognize that gender is a fluid concept, both today and in the past, and that this method glosses over the complexities of historical definitions of gender and of the relationship between biological sex and gender [Butler 1990, Butler 1993].
-In some sense this is unavoidable since the large historical datasets which make this kind of analysis possible were created by state agencies such as the U.S.
+It is important to understand our method’s limitations as well as its advantages.
+The approach of inferring gender from personal names is still a blunt way to study a complex subject.
+Gender theorists and feminist scholars have spent decades unpacking the full range of meanings behind gender as an analytical concept.
+On its own, our method has little to say about crucial topics such as the intersectionality of gender with race, class, and power, or the fluid relationship between biological sex and individual identity [Butler 1990, Butler 1993, Crenshaw 1991, McCall 2005].
+The tool itself does not reflect how an individual person identifies oneself and is neutral as to what the categories “male” or “female” actually mean in any given historical and social context.
+The question of gender identity is especially relevant for our method, given its reliance on large datasets created by state agencies such as the U.S.
 Census Bureau.
-For the category of gender (as for race) states have offered a limited number of boxes to check, and those boxes exert a tremendous power to define state subjects [Brubaker and Cooper 2000, 14-17, Pascoe 2009, Canaday 2011, Landrum 2014].
-This method does not presuppose that the gender binary imposed by governments and their wider societies reflect how an individual identifies oneself and is neutral as to what the categories "male" or "female" actually mean in any given historical context.
-Researchers are free---indeed, obligated---to make that determination for themselves, and this package provides access to the underlying dataset so that researchers can interrogate every assumption that comes with it.
-This method does not argue against the complexities of historical understandings of gender, but as with any project it does require researchers who use the method to think through how the data was gathered, what it implies, and how it reflects social and cultural practices.
+Much like the state’s definition of racial categories, governments impose their own vision of a gender binary onto their subjects, which in turn exerts a tremendous definitional power over societies and individuals [Brubaker and Cooper 2000, 14-17, Pascoe 2009, Canaday 2011, Landrum 2014].
+Researchers who use this method need to be fully aware of the ways it replicates this problematic, state-defined gender binary.
+
+As with any tool, the onus is on the researcher to understand how its underlying data was gathered, what it implies, and how it reflects social and cultural practices.
+Imperfect as it may be, this method nevertheless gives digital humanities scholars an important means of studying gender in textual collections.
+We have made the method as transparent and flexible as possible through the inclusion of probability figures - ex.
+52 percent of the babies named Leslie born in the United States in 1950 were girls and 48 percent were boys.
+This makes the underlying data visible to users and allows them the opportunity to interrogate the assumptions that come with it.
+When used thoughtfully, this method can infer additional information about individuals that stand at the heart of humanities scholarship, providing another way to see and interpret sources across large scales of analysis.
 
 The remainder of this article is divided into two sections.
-First, we describe our method in more detail, compare it to existing methods, and explain how to use it.
-Second, we apply the method to a case study of gatekeeping in the historical profession in order to demonstrate its usefulness.
+First, we describe our method in more detail, outline its advantages over existing methods, and explain how to use it.
+Second, we apply the method to a case study of gatekeeping in the historical profession in order to demonstrate its usefulness for humanities scholars.
 
@@
 What follows is a quantitative supplement to Lunbeck's analysis that uses our program to analyze gender representation amongst historians across a much larger scale and a much longer time period.
-We have found that although women have achieved rough equality in terms of the number of dissertations written, there is still a significant gap between male and female authorship of books in history.
+We study one of the bedrocks of the historical profession, scholarly research, and focus on two types of scholarly output: the dissertation and the monograph.
+Although women have achieved rough parity with men in terms of the number of completed dissertations, female historians continue to face an unequal playing field in the monographic landscape.
+Applying our method to the pages of the American Historical Review reveals a persistent and significant gap in the number of male-authored and female-authored books reviewed by one of the field’s premier academic journals.
 
-Our analysis focuses on one of the bedrocks of the historical profession: scholarly research.
-We begin with the history dissertation, often the defining scholarly output of a historian's early career.
+We begin with the history dissertation, often the defining scholarly production of a historian's early career.
 The completion of a dissertation marks a standardized moment of transition out of the "training" phase of the historical profession.
-To study dissertations about history, we used data supplied by ProQuest for roughly eighty thousand PhD dissertations completed in the United States.^[ProQuest's dataset does not permit us to limit dissertations just to dissertations produced in history departments.
+To identify how many women and men completed PhD-level training in history each year, we used data supplied by ProQuest for roughly eighty thousand PhD dissertations completed in the United States between 1950 and 2012.^[ProQuest's dataset does not permit us to limit dissertations just to dissertations produced in history departments.
 These dissertations include dissertations *about* history.
@@
 American Historical Association, *Directory of History Dissertations*, https://secure.historians.org/pubs/dissertations/.
-Data export provided to author by Liz Townsend, March 3, 2014.] Identifying the gender of their authors identifies the approximate number of women and men who are completing PhD-level training in history each year.
-Our program uses the year a historian wrote his or her dissertation to estimate a period for when they might have been born: we have assumed that a historian was between 25 and 45 years old when completing the dissertation.
-Using this temporal information, we are better able to infer their gender and chart how the larger representation of women and men changes year-by-year.
+Data export provided to author by Liz Townsend, March 3, 2014.] Our program used the year a historian wrote his or her dissertation to estimate a period for when they might have been born, assuming that a historian was between 25 and 45 years old when completing their dissertation.
+With this temporal information, we were able to infer their gender and chart how the larger representation of women and men changed year-by-year.
 
@@
 
-Another way to examine this trend is to look at the proportion of male and female dissertation authors over time, which smoothes out changes in the absolute number of dissertations produced.
+Another way to examine this trend is to ignore changes in the *absolute* number of dissertations produced each year and to instead look at changes in the *proportion* of male and female dissertation authors.
 The proportion of dissertations written by women has steadily increased over the past half-century, a change that began in the late 1960s and continued through the early 2000s.
 Since that point, the proportion of dissertations written by women has largely plateaued at a few percentage points below the proportion written by men.
-Female historians have achieved something approaching parity with male historians in terms of how many women and women complete dissertations each year.
+Female historians have achieved something approaching parity with male historians in terms of how many women and men complete dissertations each year.
 
@@
 
+[Crenshaw 1991] Crenshaw, Kimberle. "Mapping the Margins: Intersectionality, Identity Politics, and Violence against Women of Color." *Stanford Law Review*. 43.6 (July 1991): 1241-1299.
+
 [Genderize.io n.d.] *Genderize.io*. http://genderize.io/. Accessed 15 September 2014.
@@
 [Goldstone and Underwood 2014] Goldstone, Andrew, and Ted Underwood.
-"The Quiet Transformations of Literary Studies: What Thirteen Thousand Scholars Could Tell Us." *New Literary History* (Forthcoming 2014).
+"The Quiet Transformations of Literary Studies: What Thirteen Thousand Scholars Could Tell Us." *New Literary History* 45.3 (Summer 2014): 359-384.
 
@@
 
+[McCall 2005] McCall, Leslie. "The Complexity of Intersectionality." *Signs*. 30.3 (Spring 2005): 1771-1800.
+
 [Muenchen n.d.] Muenchen, Robert A. "The Popularity of Data Analysis Software," *r4stats.com*. <http://r4stats.com/articles/popularity/>. Accessed 12 September 2014.
+
+[Orlando Project n.d.] Orlando Project. "About." <http://www.artsrn.ualberta.ca/orlando/?page_id=2>. Accessed 23 June 2015.]
 
@@
 27 August 2013, <http://www.kdnuggets.com/2013/08/languages-for-analytics-data-mining-data-science.html>.
+
+[The Poetess Archive n.d.] The Poetess Archive. "About the Poetess Archive." <http://idhmc.tamu.edu/poetess/about/index.html>. Accessed 23 June 2015.]
 
@@
 
+[Weingart and Jorgensen 2013] Weingart, Scott, and Jeana Jorgensen. "Computational Analysis of the Body in European Fairy Tales." *Literary and Linguistic Computing*, 28.3 (2013): 404-416.
+
 [Wickham 2014] Wickham, Hadley. "Tidy Data." *Journal of Statistical Software*, 50 (2014): <http://www.jstatsoft.org/v59/i10>.

### 2015-06-23 article.Rmd
AHR section revisions

@@
 Although women have achieved rough parity with men in terms of the number of completed dissertations, female historians continue to face an unequal playing field in the monographic landscape.
-Applying our method to the pages of the American Historical Review reveals a persistent and significant gap in the number of male-authored and female-authored books reviewed by one of the field’s premier academic journals.
+Applying our method to the pages of the *American Historical Review* reveals a persistent and significant gap in the number of male-authored and female-authored books reviewed by one of the field’s premier academic journals.
 
@@
 
-But what happens after the dissertation?
-It is, after all, only the first major stage of a historian’s research output.
-The "coin of the realm" for many historians remains the publication of a scholarly monograph to be read and evaluated by peers.
-Once published, the impact of a book is often shaped by its reviews in academic journals.
-One of the leading journals in the historical profession is the *American Historical Review*, the flagship journal of the American Historical Association.
-Along with a few dozen articles published in its five issues each year, the *AHR* publishes roughly one thousand books reviews per year covering (in its own words) "every major field of historical study." [American Historical Review n.d.] The *AHR* is not only one of the widest-ranging journals in the profession, it is also the oldest; the journal has been publishing continuously since 1895.
-The range, scope, longevity, and reputation of the *AHR* makes it an ideal proxy to study gatekeeping within the historical discipline.
+But the dissertation is only the first major piece of scholarship produced by an academic historian.
+The second usually centers on the writing of an academic monograph to be read and evaluated by their peers.
+These monographs remain the "coin of the realm" for many historians, and the value of that coin often depends on it being reviewed in academic journals.
+To study how gender structures the monograph reviews, we turn to one of the leading journals in the historical profession: the *American Historical Review*, the flagship journal of the American Historical Association.
+The *AHR* publishes roughly one thousand total book reviews spread across five issues each year, covering (in its own words) "every major field of historical study." [American Historical Review n.d.] The AHR is not only one of the widest-ranging journals in the profession, it is also the oldest; the journal has been publishing continuously since 1895.
+The range, scope, longevity, and reputation of the *AHR* makes it an ideal proxy to study gendered gatekeeping related to historical monographs.
 
@@
 
-Scraping the table-of-contents of every issue of the American Historical Review results in a dataset of close to 60,000 books reviewed by the journal since it began publication in 1895.
+Scraping the table-of-contents of every issue of the *American Historical Review* results in a dataset of close to 60,000 books reviewed by the journal since it began publication in 1895.
 Our program then inferred the gender of the authors of these books, which we could in turn use to plot the proportion of female and male authors over time.
-The temporal trajectory of gender representations roughly resembles that of history dissertations: women began making inroads in the late 1960s and have made steady gains over the past four decades.
-By the twenty-first century the proportion of female authors reviewed in the *AHR* had more than tripled.
+The temporal trajectory of gender representation in the journal’s reviews roughly resembles that of history dissertations: women began making inroads in the late 1960s and continue to make steady gains over the next four decades.
+By the twenty-first century, the proportion of female authors reviewed in the *AHR* had more than tripled.
 
@@
 
-But a closer look shows important differences between the proportion of women writing history dissertations and the proportion of women appearing in the *AHR*.
-Although both have trended upwards since the 1970s the slope is a lot steeper for dissertations than it is for the *AHR*.
-Female historians have very nearly closed the gap in terms of newly completed dissertations, but the glass ceiling remains stubbornly low in terms of what happens from that point onwards.
-In book reviews published in the *AHR* male authors continue to outnumber female authors by a factor of nearly 2 to 1.
-The long-term gains made by female doctoral students do not carry over into the pages of the historical profession's leading journal.^[The chart below includes only authors whose gender we successfully inferred; it does not include authors of unknown gender as in the previous charts.]
+But a closer look shows substantial differences between the proportion of women writing history dissertations and the proportion of women appearing in the *AHR*.
+Although both have trended upwards since the 1970s, the slope of that line is a lot steeper for dissertations than it is for the *AHR*.
+Female historians have very nearly closed the gap in terms of newly completed dissertations, but the glass ceiling remains stubbornly low in terms of authoring monographs that are then reviewed by the discipline’s leading academic journal.
+In book reviews published in the *AHR*, male authors continue to outnumber female authors by a factor of nearly 2 to 1.
+The long-term gains made by female doctoral students do not carry over into the pages of the *American Historical Review*.^[The chart below includes only authors whose gender we successfully inferred; it does not include authors of unknown gender as in the previous charts.]
 
@@
 
+Not every historian wants or seeks a career path that neatly moves from writing a dissertation to authoring a monograph to being reviewed in the *American Historical Review*.
+But for those that do, each of these subsequent stages represents a winnowing process: not all graduate students produce dissertations, not all PhD’s write books, not all books are submitted to the *American Historical Review*, and not all submissions are accepted by the *AHR* editors for review.
+Nearly equal numbers of female and male historians make it through the initial stage of writing a dissertation.
+But in the final round of gatekeeping – a monograph review in the *AHR* – there are nearly twice as many men as women.
+
+Without more information it is all but impossible to establish a causal reason for this pattern.
+Instead, a murky concoction of factors likely contributes to the different gender gaps between dissertations and monograph reviews in the AHR.
+For one, dissertations and monographs are frequently written not only at different points in a person’s career, but also their family life.
+A 2011 survey by the American Historical Association revealed that female historians dedicated substantially more time to child and elder care than their male colleagues, leaving them with less time for research.
+[Townsend 2012] This gender gap is more likely to impact the production of a book than a dissertation, as graduate students are generally younger and less likely to be juggling child-rearing responsibilities while writing than their older peers.
+Their male colleagues, meanwhile, enjoy more time to publish their research.
+
+***left off revising here***
+
 The path from a dissertation to a monograph to a book review in the *American Historical Association* isn't automatic.
 Not all dissertations are published as books, not all books are submitted to the *American Historical Review*, and not all book submissions are accepted by the *AHR* editors for review.
-As of September 2014 the journal's website reminds readers that "the AHR receives over 3,000 books a year; we have the resources to publish at most 1,000 reviews a year (approximately 200 per issue)." [American Historical Review n.d.] But fewer and fewer female historians make it through each successive step.
+As of September 2014 the journal's website reminds readers that "the *AHR* receives over 3,000 books a year; we have the resources to publish at most 1,000 reviews a year (approximately 200 per issue)." [American Historical Review n.d.] But fewer and fewer female historians make it through each successive step.
 Establishing a single causal reason for this pattern is all but impossible, and instead a murky concoction of interrelated factors likely contributes to the successively widening gap.

### 2015-06-23 article.Rmd
AHR Section, thoughts on conclusion

@@
 
-We begin with the history dissertation, often the defining scholarly production of a historian's early career.
+We begin with the history dissertation, often the defining scholarly output of a historian's early career.
 The completion of a dissertation marks a standardized moment of transition out of the "training" phase of the historical profession.
@@
 These monographs remain the "coin of the realm" for many historians, and the value of that coin often depends on it being reviewed in academic journals.
-To study how gender structures the monograph reviews, we turn to one of the leading journals in the historical profession: the *American Historical Review*, the flagship journal of the American Historical Association.
-The *AHR* publishes roughly one thousand total book reviews spread across five issues each year, covering (in its own words) "every major field of historical study." [American Historical Review n.d.] The AHR is not only one of the widest-ranging journals in the profession, it is also the oldest; the journal has been publishing continuously since 1895.
-The range, scope, longevity, and reputation of the *AHR* makes it an ideal proxy to study gendered gatekeeping related to historical monographs.
+To study the role of gender in monograph reviews, we turn to one of the leading journals in the historical profession: the *American Historical Review*, the flagship journal of the American Historical Association.
+The *AHR* publishes roughly one thousand total book reviews spread across five issues each year, covering (in its own words) "every major field of historical study." [American Historical Review n.d.] The *AHR* is not only one of the widest-ranging journals in the profession, it is also the oldest; the journal has been publishing continuously since 1895.
+The range, scope, longevity, and reputation of the *AHR* makes it an ideal proxy to study how professional gatekeeping operates in relation to gender.
 
 A few caveats are in order.
-First, a book that never appears in the *AHR* at all may still go on to have a significant impact on the profession.
+First, a book that is never reviewed by the *AHR* may still go on to have a significant impact on the profession.
 Second, a book's appearance in the *AHR* is not necessarily correlated with the *quality* of the book, as the journal prints negative reviews as well as positive ones.
@@
 Even if it garners a negative review, the fact that it appears in the journal at all is a measure of the fact that the profession's gatekeepers have deemed it important enough to review.
-It is precisely this professional gatekeeping dimension that we can analyze using the Gender package.
+It is precisely this professional gatekeeping dimension that we can analyze using the Gender package: are women more likely to be barred from this gate than men, and if so, by how much?
 
@@
 Our program then inferred the gender of the authors of these books, which we could in turn use to plot the proportion of female and male authors over time.
-The temporal trajectory of gender representation in the journal’s reviews roughly resembles that of history dissertations: women began making inroads in the late 1960s and continue to make steady gains over the next four decades.
+The temporal trajectory of gender representation in the journal’s reviews roughly resembles that of history dissertations: women began making inroads in the late 1960s and continued to make steady gains over the next four decades.
 By the twenty-first century, the proportion of female authors reviewed in the *AHR* had more than tripled.
@@
 
-But a closer look shows substantial differences between the proportion of women writing history dissertations and the proportion of women appearing in the *AHR*.
-Although both have trended upwards since the 1970s, the slope of that line is a lot steeper for dissertations than it is for the *AHR*.
-Female historians have very nearly closed the gap in terms of newly completed dissertations, but the glass ceiling remains stubbornly low in terms of authoring monographs that are then reviewed by the discipline’s leading academic journal.
-In book reviews published in the *AHR*, male authors continue to outnumber female authors by a factor of nearly 2 to 1.
-The long-term gains made by female doctoral students do not carry over into the pages of the *American Historical Review*.^[The chart below includes only authors whose gender we successfully inferred; it does not include authors of unknown gender as in the previous charts.]
+But a closer look shows substantial differences between the proportion of women as the authors of history dissertations and the proportion of women appearing in the *AHR* as the authors of reviewed books.
+Although both have trended upwards since the 1970s, the slope of that line was a lot steeper for dissertations than it was for the *AHR*.
+Female historians have very nearly closed the gap in terms of newly completed dissertations, but the glass ceiling remains much lower in terms of getting reviewed by the discipline’s leading academic journal.
+The long-term gains made by female doctoral students do not carry over into the pages of the *American Historical Review*, where male monograph authors continue to outnumber female authors by a factor of nearly 2 to 1.^[The chart below includes only authors whose gender we successfully inferred; it does not include authors of unknown gender as in the previous charts.]
 
@@
 Not every historian wants or seeks a career path that neatly moves from writing a dissertation to authoring a monograph to being reviewed in the *American Historical Review*.
-But for those that do, each of these subsequent stages represents a winnowing process: not all graduate students produce dissertations, not all PhD’s write books, not all books are submitted to the *American Historical Review*, and not all submissions are accepted by the *AHR* editors for review.
-Nearly equal numbers of female and male historians make it through the initial stage of writing a dissertation.
-But in the final round of gatekeeping – a monograph review in the *AHR* – there are nearly twice as many men as women.
+For those that do, however, each of these subsequent stages represents a winnowing process: not all graduate students produce dissertations, not all PhD’s write books, not all books are submitted to the *American Historical Review*, and not all submissions are accepted by the *AHR* editors for review.^[As of 2015 the journal's website reminds readers that "the *AHR* receives over 3,000 books a year; we have the resources to publish at most 1,000 reviews a year (approximately 200 per issue)." [American Historical Review n.d.]] Slighty fewer women than men go through the initial stage of writing a history dissertation.
+But by the final round of gatekeeping – a monograph reviewed in the *AHR* – the gender gap grows until there are little more than half as many women as men.
+The *American Historical Review* is, unfortunately, far from alone.
+Many of the profession's most prestigious annual honors go to drastically more men than women.
+As of 2015, the Pulitzer Prize for History has been awarded to exactly two female historians in the past twenty years ["The Pulitzer Prizes: History" 2015-].
+The Bancroft Prize in American History has been awarded to almost five times as many male historians as female historians over that same period ["Bancroft Prize"].
 
-Without more information it is all but impossible to establish a causal reason for this pattern.
-Instead, a murky concoction of factors likely contributes to the different gender gaps between dissertations and monograph reviews in the AHR.
-For one, dissertations and monographs are frequently written not only at different points in a person’s career, but also their family life.
-A 2011 survey by the American Historical Association revealed that female historians dedicated substantially more time to child and elder care than their male colleagues, leaving them with less time for research.
-[Townsend 2012] This gender gap is more likely to impact the production of a book than a dissertation, as graduate students are generally younger and less likely to be juggling child-rearing responsibilities while writing than their older peers.
-Their male colleagues, meanwhile, enjoy more time to publish their research.
+Examples like the Pulitzer and Bancroft Prizes are only the most glaring cases of gender inequity in the historical profession.
+Overt discrimination still exists, but inequity operates in ways that are far more cloaked and far more complex, which frequently makes them difficult to recognize or remedy.
+They take the form of subconscious bias that subtly weakens the assessment of work done by female historians.
+Gender discrimination frequently intersects with other kinds of racial and socioeconomic inequity [Townsend 2008].
+And larger societal structures impede the professional lives of women: a 2011 survey by the American Historical Association revealed that female historians dedicated substantially more time to child and elder care than their male colleagues, leaving them with less time for research [Townsend 2012].
+All of these factors combine and amplify each other to contribute to the gender gap in the historical profession on display in the pages of the *American Historical Review*.
 
-***left off revising here***
-
-The path from a dissertation to a monograph to a book review in the *American Historical Association* isn't automatic.
-Not all dissertations are published as books, not all books are submitted to the *American Historical Review*, and not all book submissions are accepted by the *AHR* editors for review.
-As of September 2014 the journal's website reminds readers that "the *AHR* receives over 3,000 books a year; we have the resources to publish at most 1,000 reviews a year (approximately 200 per issue)." [American Historical Review n.d.] But fewer and fewer female historians make it through each successive step.
-Establishing a single causal reason for this pattern is all but impossible, and instead a murky concoction of interrelated factors likely contributes to the successively widening gap.
-A 2011 survey by the American Historical Association revealed that female historians dedicated substantially more time to child and elder care than their male colleagues, leaving them with less time for research.
-[Townsend 2012] This gender gap is more likely to impact the production of a book than a dissertation, as graduate students are generally younger and less likely to be juggling child-rearing responsibilities while writing than their older peers.
-Their male colleagues, meanwhile, enjoy more time to publish and receive more accolades when they do.
-As of 2014, the Pulitzer Prize for History has been awarded to exactly one female historian in the past nineteen years.
-The Bancroft Prize in American History has been awarded to almost four times as many male historians as female historians over that same period.
-["Bancroft Prize"] The *American Historical Review* is, unfortunately, far from alone in its gatekeeping practices.
-
-We went on to use our program to infer the gender of not just book authors reviewed by the *AHR*, but the gender of the reviewers themselves.
+The hidden, murky, and systemic dimensions of gender inequity makes methods like the Gender package all the more imperative.
+Our program was able to infer the genders of tens of thousands of authors, aggregate and chart their trajectory over time, and compare the pattern to another dataset of dissertation authors.
+Doing so revealed an unmistakable gap in the number of women who were getting reviewed by teh *American Historical Review*.
+But our program was able to look at subtler patterns - the kind that make gender inequity so insidious.
+After inferring the gender of book authors reviewed by the *AHR*, we were also able to infer the gender of the reviewers themselves.
 The story is much the same for reviewers as it is for authors: more than twice as many men as women appear as reviewers in the journal.
-But gender inflects book reviews in less direct ways by shaping who writes reviews of which book authors.
+But gender also inflects book reviews by shaping who writes reviews of which authors.
 About three times as many men write reviews of male-authored books as do women.
-In the case of female-authored books the ratio of female-to-male reviewers is much closer to 50/50.
-In short, women are much more likely to write reviews of other women.
-And while men still write the majority of reviews of female-authored books, they tend to gravitate more towards male authors – who are, of course, already over-represented in the *AHR*.
+In the case of female-authored books the ratio of female-to-male reviewers is closer to 50/50.
+In short, women are far more likely to write reviews of other women, while men tend to gravitate more towards male authors – who are, of course, already over-represented in the *AHR*.
 
@@
 
+Gender inequity is not always an overt case of deciding to award a prestigious prize to a male historian; it can be as subtle as deciding to write a book review of a male author.
+Countless individual actions like these constitute the relationships that knit together the field of academic history.
+And too often these tiny actions benefit men more than women: chairing a panel of mostly male historians, taking on a new male advisee, or assigning mostly male-authored books on a syllabus.
+On their own they might appear benign - one might argue that male historians simply happen to be interested in similar research topics - but in aggregate they contribute to a broader professional landscape that puts female historians at a disadvantage.
+
 One 2013 analysis of 2,500 recent history Ph.D.'s found that "gender played little role in employment patterns across particular professions and industries" [Wood and Townsend 2013].
-Our own analysis reveals that the story is different for book publishing than for employment.
-The flagship journal of professional historians continues to publish twice as many reviews of books by male authors as female authors.
-This disparity is put into even sharper relief when set against the relative parity achieved by women in producing history dissertations.
+Our own analysis reveals a different story in terms of getting a book reviewed in the field's most prestigious journal.
+The *American Historical Review* continues to publish twice as many reviews of male-authored books as female-authored books.
+This disparity is put into even sharper relief when set against the much smaller gender gap in history dissertations.
 Women might be getting hired at comparable rates to men, but discriminatory gatekeeping remains alive and well in the profession.
-As explicit instances of sexism fade from view the disadvantages women face are growing more and more opaque.
-But that does not make them any less damaging.
-As discrimination becomes less blatant and more systemic, digital methods like the Gender package are a way to a shine light on these cloaked disadvantages, to reveal the true extent of the problem and the work that remains to be done.
 
 ### Conclusion
+
+Include in conclusion
+- gender is a tool that humanities scholars can put all of their tremendous skills in contextualizing, reocgnizing subtlety, and putting it in conversation with other sturctural power of race class and sexuality.
+- Point to examples of how much it's downloaded, Ted Underwood, and especially Ben Schmidt's student evaluations, which is exactly the kind of ways it can be used to shed light on hidden patterns of systemic bias and discrimination.
 
@@
 
-The analytical category of gender has transformed humanities research over the past half-century.
-More recently, the digitization movement has opened up opportunities to apply this category to new sources and in new ways.
-Projects in the digital humanities have ranged from recovering the voices of early-modern women writers to tracing the linguistic styles of male and female authors [Women Writers Project n.d., Argamon et.
-al.
-2009].
-The proliferation of machine-readable datasets and digital texts brings with it a vast trove of material with which to study gender.
-But unlocking its full range of meaning necessitates not just accessing the data that *is* there, but inferring information that *isn't* there.
-The Gender package is a powerful tool to do just that.
-
 ### References
@@
 "Book Reviewing in the AHR." http://www.historians.org/publications-and-directories/american-historical-review/book-review-guide.
-Accessed 29 September 2014.
+Accessed 23 June 2015.
 
@@
 
+[Townsend 2008] Townsend, Robert B.
+"The Status of Women and Minorities in the History Profession, 2008." *Perspectives on History*.
+September 2008, <http://www.historians.org/publications-and-directories/perspectives-on-history/september-2008/the-status-of-women-and-minorities-in-the-history-profession-2008>
+
 [Townsend 2012] Townsend, Robert B.

### 2015-07-30 article.Rmd
Mostly AHR section

@@
 For example, a literary scholar studying a corpus of poems from the mid-twentieth century might wish to compare the stanza structures of male and female poets.
-Using a tool like Genderize.io based on contemporary data risks misidentifying these writers, many of who were born (and named) more than a century prior to the creation of these modern name datasets.
+Using a tool like Genderize.io based on contemporary data risks misidentifying these writers, many of whom were born (and named) more than a century prior to the creation of these modern name datasets.
 This problem will only increase with growing life expectancies: as of 2012, the average American lifespan was nearly seventy-nine years---more than enough time for naming practices to change quite dramatically between a person’s birth and death [World Bank 2014].
@@
 Examples like the Pulitzer and Bancroft Prizes are only the most glaring cases of gender inequity in the historical profession.
-Overt discrimination still exists, but inequity operates in ways that are far more cloaked and far more complex, which frequently makes them difficult to recognize or remedy.
-They take the form of subconscious bias that subtly weakens the assessment of work done by female historians.
-Gender discrimination frequently intersects with other kinds of racial and socioeconomic inequity [Townsend 2008].
-And larger societal structures impede the professional lives of women: a 2011 survey by the American Historical Association revealed that female historians dedicated substantially more time to child and elder care than their male colleagues, leaving them with less time for research [Townsend 2012].
+Overt discrimination still exists, but inequity operates in ways that are far more cloaked and far more complex, making them difficult to recognize and remedy.
+They take the form of subconscious bias that subtly weakens the assessment of work done by female historians, whether in a seminar paper or a job application.
+Moreover, gender discrimination frequently intersects with other kinds of racial and socioeconomic inequity that erects even higher barriers in the women's careers [Townsend 2008].
+Gendered family dynamics add another layer of challenges for women: a 2011 survey by the American Historical Association revealed that female historians dedicated substantially more time to child and elder care than their male colleagues, leaving them with less time for the kind of academic research that the profession rewards most highly [Townsend 2012].
 All of these factors combine and amplify each other to contribute to the gender gap in the historical profession on display in the pages of the *American Historical Review*.
+The hidden, murky, and systemic dimensions of gender inequity makes methods like the Gender package all the more imperative for reasearchers.
 
-The hidden, murky, and systemic dimensions of gender inequity makes methods like the Gender package all the more imperative.
-Our program was able to infer the genders of tens of thousands of authors, aggregate and chart their trajectory over time, and compare the pattern to another dataset of dissertation authors.
-Doing so revealed an unmistakable gap in the number of women who were getting reviewed by teh *American Historical Review*.
-But our program was able to look at subtler patterns - the kind that make gender inequity so insidious.
+Finally, we used the Gender package to reveal some of the subtler ways in which gender iniquity operates in the historical profession.
 After inferring the gender of book authors reviewed by the *AHR*, we were also able to infer the gender of the reviewers themselves.
-The story is much the same for reviewers as it is for authors: more than twice as many men as women appear as reviewers in the journal.
-But gender also inflects book reviews by shaping who writes reviews of which authors.
-About three times as many men write reviews of male-authored books as do women.
-In the case of female-authored books the ratio of female-to-male reviewers is closer to 50/50.
+This led us to two findings.
+First, the story is much the same for reviewers as it is for authors: more than twice as many men as women appear as reviewers in the journal.
+This is important because reviewers for the *AHR* garner a certain degree of prestige from their work - perhaps not to the same prestige as winning a Pulitzer Prize, but prestigious nonetheless.
+To become a reviewer, one must possess a Ph.D.
+or equivalent degree and have published at least one book-length manuscript [American Historical Review n.d.].
+*AHR* reviewers are understood to have sufficient experience and command over a given sub-field to evaluate the merits and weaknesses of a new contribution to that literature.
+Being a reviewer for the journal is, in many ways, a proxy for expertise.
+Which is a problem.
+When there are twice as many male "experts" as female "experts" appearing in the *AHR* it imbues our understanding of what constitutes historical expertise.
+Or perhaps more accurately, who is and who isn't an "expert" in the field.
+
+The second finding we arrived at when examining the gender of *AHR* reviewers was a better understanding of who reviews which book authors.
+Once again, the methodology reveals patterns that we might suspect but nonetheless tends to stay hidden from direct sight.
+About three times as many men as women write reviews of male-authored books in the *AHR*.
+In the case of female-authored books, the ratio of female-to-male reviewers is closer to 50/50.
 In short, women are far more likely to write reviews of other women, while men tend to gravitate more towards male authors – who are, of course, already over-represented in the *AHR*.
@@
 
-Gender inequity is not always an overt case of deciding to award a prestigious prize to a male historian; it can be as subtle as deciding to write a book review of a male author.
-Countless individual actions like these constitute the relationships that knit together the field of academic history.
-And too often these tiny actions benefit men more than women: chairing a panel of mostly male historians, taking on a new male advisee, or assigning mostly male-authored books on a syllabus.
+The findings related to *AHR* reviewers help illustrate the ways in which gender inequity is so insidious.
+It is not always an overt case of deciding to award a prestigious prize to a male historian; it can be as seemingly innocuous as deciding to write a book review of a male author's book.
+Countless individual actions like these constitute the core relationships that knit together the profession, and too often these tiny actions benefit men more than women: chairing a panel of mostly male historians, taking on a new male advisee, or assigning mostly male-authored books on a syllabus.
 On their own they might appear benign - one might argue that male historians simply happen to be interested in similar research topics - but in aggregate they contribute to a broader professional landscape that puts female historians at a disadvantage.
@@
 One 2013 analysis of 2,500 recent history Ph.D.'s found that "gender played little role in employment patterns across particular professions and industries" [Wood and Townsend 2013].
-Our own analysis reveals a different story in terms of getting a book reviewed in the field's most prestigious journal.
-The *American Historical Review* continues to publish twice as many reviews of male-authored books as female-authored books.
-This disparity is put into even sharper relief when set against the much smaller gender gap in history dissertations.
+Our own analysis reveals a different story, at least in terms of getting a book reviewed in the field's most prestigious journal.
+The *American Historical Review* continues to publish twice as many reviews of male-authored books as female-authored books, and those same reviews are more than twice as likely to be written by a man than a woman.
+This disparity is put into even sharper relief when set against the much smaller gender gap between female and male historians who complete dissertations.
 Women might be getting hired at comparable rates to men, but discriminatory gatekeeping remains alive and well in the profession.
+The Gender package is one more tool to uncover those practices.
 
@@
 - Point to examples of how much it's downloaded, Ted Underwood, and especially Ben Schmidt's student evaluations, which is exactly the kind of ways it can be used to shed light on hidden patterns of systemic bias and discrimination.
+- Posner's latest blog post on impreative of creating humanistic tools
 

### 2015-08-05 article.Rmd
Who needs MS Word?
Slightly less jargony abstract

@@
   html_document: default
-  word_document: default
 ---
@@
 
-This paper outlines a new method for inferring the gender of personal names in large, machine-readable datasets.
-In contrast to existing methods of predicting gender that treat names as static entities, this method uses a humanistic approach to take into account the dynamic nature of naming practices.
+This article describes a new method for inferring the gender of personal names using large historical datasets.
+In contrast to existing methods of predicting gender that treat names as if they are timelessly associated with gender, this method uses a historical approach that takes into account the dynamic nature of naming practices.
 It uses historical data to measure the likelihood that a name was associated with a particular gender based on the time or place under study.
-This approach generates more accurate results for sources that encompass changing periods of time, providing digital humanities scholars with a flexible tool to apply and adapt to analyses of gender.
+This approach generates more accurate results for sources that encompass changing periods of time, providing digital humanities scholars with a tool to estimate the gender of names.
 The article first describes the methodology as implemented in the [Gender](https://github.com/ropensci/gender) package for the [R programming language](http://www.r-project.org/).
 It goes on to apply the method to a case study in which we examine gender and gatekeeping in the American historical profession over the past half-century.
-The Gender package illustrates the importance of incorporating humanistic approaches to computer science and related fields.
+The Gender package illustrates the importance of incorporating historical approaches to computer science and related fields.
 

### 2015-08-10 article.Rmd
Adds to intro, conclusion

@@
 The approach of inferring gender from personal names is still a blunt way to study a complex subject.
-Gender theorists and feminist scholars have spent decades unpacking the full range of meanings behind gender as an analytical concept.
-On its own, our method has little to say about crucial topics such as the intersectionality of gender with race, class, and power, or the fluid relationship between biological sex and individual identity [Butler 1990, Butler 1993, Crenshaw 1991, McCall 2005].
-The tool itself does not reflect how an individual person identifies oneself and is neutral as to what the categories “male” or “female” actually mean in any given historical and social context.
+Gender theorists and feminist scholars have spent decades unpacking the full range of meanings behind gender as an analytical concept, including the fluid relationship between biological sex and individual identity [Butler 1990, Butler 1993].
+On its own, our method has little to say about crucial topics such as the intersectionality of gender with race, class, and power, or what the categories “male” or “female” actually mean in any given historical and social context [Crenshaw 1991, McCall 2005].
+Perhaps most importantly, the tool itself does not reflect how an individual person self-identifies and how gender shapes their lived experience.
 The question of gender identity is especially relevant for our method, given its reliance on large datasets created by state agencies such as the U.S.
 Census Bureau.
-Much like the state’s definition of racial categories, governments impose their own vision of a gender binary onto their subjects, which in turn exerts a tremendous definitional power over societies and individuals [Brubaker and Cooper 2000, 14-17, Pascoe 2009, Canaday 2011, Landrum 2014].
-Researchers who use this method need to be fully aware of the ways it replicates this problematic, state-defined gender binary.
+Much like the state’s definition of racial categories, governments impose their own vision of a gender binary onto their subjects in ways that are deeply problematic.
+This gender binary reflects existing power structures more than the lived identities of individuals.
+The state's gender categorization exerts a tremendous definitional sway over societies and the people who comprise them, most notably in its systematic exclusion of transgender identities.
+[Brubaker and Cooper 2000, 14-17, Pascoe 2009, Canaday 2011, Landrum 2014].
 
-As with any tool, the onus is on the researcher to understand how its underlying data was gathered, what it implies, and how it reflects social and cultural practices.
+As with any tool, researchers who use this method need to be fully aware of the ways it replicates a problematic, state-defined gender binary.
+The onus is on the user to understand how its underlying data was gathered, what it implies, and how it reflects social and cultural practices.
 Imperfect as it may be, this method nevertheless gives digital humanities scholars an important means of studying gender in textual collections.
@@
 This makes the underlying data visible to users and allows them the opportunity to interrogate the assumptions that come with it.
-When used thoughtfully, this method can infer additional information about individuals that stand at the heart of humanities scholarship, providing another way to see and interpret sources across large scales of analysis.
+When used thoughtfully, this method can infer additional information about individuals who stand at the heart of humanities scholarship, providing another way to see and interpret sources across large scales of analysis.
 
@@
 
-Include in conclusion
-- gender is a tool that humanities scholars can put all of their tremendous skills in contextualizing, reocgnizing subtlety, and putting it in conversation with other sturctural power of race class and sexuality.
-- Point to examples of how much it's downloaded, Ted Underwood, and especially Ben Schmidt's student evaluations, which is exactly the kind of ways it can be used to shed light on hidden patterns of systemic bias and discrimination.
-- Posner's latest blog post on impreative of creating humanistic tools
-
-Our algorithm for predicting gender from first names, when combined with historical datasets in the Gender package, provides a powerful new method for researchers interested in the study of gender.
 The digital turn has made available very large datasets, yet in the midst of abundance many humanists continue to deal with the problem of incompleteness and scarcity within their sources.
 Our method uses *abundant* existing datasets to infer additional information from *scarce* datasets.
-The temporal dimension of our work matches historical questions to a historical method, providing much improved results over simple, anachronistic, or ahistorical lookup methods.
-This method is also extensible to different times and places, provided that a researcher has a suitable reference dataset of names and genders from the period or area under study.
-The Gender package reverses the traditional disciplinary direction between the humanities and computer science.
-Rather than adapting existing computational methodologies, our method uses a humanistic starting point---the historical relationship between gender and naming practices---to contribute a new method to computer science.
+Its major contribution is to incorporate a historical approach to this algorithm, one that takes into account how gendered naming practices have changed over time.
+This temporal methodology provides much improved results over simpler, anachronistic, or ahistorical lookup methods.
+The Gender package is, of course, far from perfect.
+By relying on historical datasets collected by state agencies, it reinforces a gender binary and larger power structure that obscures how people may identify themselves and how they experience the world.
+To borrow a phrase from Miriam Posner, the Gender package does not meet the "radical, unrealized potential of digital humanities." But it might.
+
+The Gender package is entirely open-source, and we have provided detailed and transarent documentation of its code and data.
+Researchers who wish to build upon it can modify the package accordingly.
+To take one imagined scenario: Facebook allows its users to either select from a large spectrum of gender identities or to write in their own prefered identity.
+Because these identities are associated with the user's first name and date of birth, this is exactly the kind of historical dataset that could be incorporated into the Gender package.
+Doing so would replace the package's simplistic categories of "Male" and "Female" with ones like "Female to Male", "Bigender", or "Cisgender Female".
+Although this imagined scenario would create its own set of challenges, it would represent a major improvement in how the Gender package dealt with the complexity and fluidity of gender identity.
+^[This imagined scenario runs into its own set of problems.
+Facebook's business model makes it unlikely to grant researchers open access to its proprietary and closely-guarded datasets.
+And even if it did grant access to this data, Facebook collected it for particular purposes and in particular ways.
+Just as government agencies like the Social Security Administration collect data to serve the purposes of the American state, Facebook's data collection reflects the priorities of a global corporation rather than the lived experience of its users.] By building as much flexibility and transparency into the method as possible, we hope that others can continue to make just these kinds of improvements.
+
+Even as it currently stands, the Gender package meets a pressing methodological need.
+Since its initial release in February 2014, the Gender package has been downloaded more than ________ times [??????] and incorporated as a pedagogical tool into digital humanities courses [Jockers 2014].
+Researchers have used it to uncover gender inequity in a range of different areas, from biases in students' teaching evaluations to disparities between attendees and authors at the annual Digital Humanities Conference [Schmidt 2015, Weingart 2015].
+These implementations help illustrate the value of the method.
+Its major intervention is to algorithmically inferring gender in textual sources in a way that takes into account change over time.
+Rather than trying to shoehorn existing computational methods, the Gender package uses a humanistic starting point---the historical relationship between gender and naming practices---to contribute a new method to computer science.
+This reverses the traditional disciplinary direction between the humanities and computer science.
 
@@
 
+[Jockers 2014] Jockers, Matthew.
+"Experimenting with “gender” package in R." *Matthew L.
+Jockers*.
+25 February 2015, <http://www.matthewjockers.net/2014/02/25/experimenting-with-gender-package-in-r/>.
+Accessed 10 August 2015.
+
 [Landrum 2014] Landrum, Shane.
@@
 
-[The Poetess Archive n.d.] The Poetess Archive. "About the Poetess Archive." <http://idhmc.tamu.edu/poetess/about/index.html>. Accessed 23 June 2015.]
+[The Poetess Archive n.d.] The Poetess Archive. "About the Poetess Archive." <http://idhmc.tamu.edu/poetess/about/index.html>. Accessed 23 June 2015.
+
+[Posner 2015] Posner, Miriam.
+"What’s Next: The Radical, Unrealized Potential of Digital Humanities." *Miriam Posner's Blog*.
+27 July 2015, <http://miriamposner.com/blog/whats-next-the-radical-unrealized-potential-of-digital-humanities/>.
+Accessed 10 August 2015.
 
@@
 Minneapolis, MN: Minnesota Population Center [producer and distributor], 2010.
+
+[Schmidt 2015] Schmidt, Ben. "Gender and Teacher Reviews." February 2015. <http://benschmidt.org/profGender/>.
 
@@
 
-[Weingart 2014] Weingart, Scott. "The Moral Role of DH in a Data-Driven World." *the scottbot irregular*. 9 September 2014, <http://www.scottbot.net/HIAL/?p=40944>
+[Weingart 2014] Weingart, Scott. "The Moral Role of DH in a Data-Driven World." *the scottbot irregular*. 9 September 2014, <http://www.scottbot.net/HIAL/?p=40944>.
+
+[Weingart 2015] Weingart, Scott. "Acceptances to Digital Humanities 2015 (part 3)." *the scottbot irregular*. 27 June 2015, <http://www.scottbot.net/HIAL/?p=41355>.
 

### 2015-08-13 article.Rmd

@@
 It goes on to apply the method to a case study in which we examine gender and gatekeeping in the American historical profession over the past half-century.
-The Gender package illustrates the importance of incorporating historical approaches to computer science and related fields.
+The Gender package illustrates the importance of incorporating historical approaches into computer science and related fields.
 
@@
 
-As humanities datasets grow larger and larger, researchers have not only been able to access information that is there, but to infer information that isn't there.
-Topic modeling, for example, has been used to uncover scholarship trends over a century of academic articles in literary studies.
+As humanities datasets grow larger and larger, researchers have not only been able to access information that *is* there, but to infer information that *isn't* there.
+Topic modeling, for example, has been used to uncover underlying trends across a century of academic articles in literary studies.
 [Goldstone and Underwood 2014].
 Can researchers similarly use personal names to infer information about individual people?
-After all, personal names appear across a wide range of sources in the humanities: literary characters in stories and novels, authorship information in archival metadata, or individuals mentioned by historical newspapers and magazine articles.
-Authors, characters, or historical actors frequently stand at the analytical center of humanities scholarship, and their names provide an intriguing way to infer additional information about them.^[The availability of large-scale datasets dealing with personal information raises important ethical issues for digital humanists, particularly in light of the revelations about domestic spying at the NSA brought forth by Edward Snowden in 2013.
+After all, personal names appear across a wide range of sources in the humanities: literary characters in stories and novels, authorship information in archival metadata, individuals mentioned by historical newspapers and magazine articles.
+Authors, characters, or historical actors are the kind of individuals who stand at the analytical center of humanities scholarship, and their names provide an intriguing way to infer additional information about them.^[The availability of large-scale datasets dealing with personal information raises important ethical issues for digital humanists, particularly in light of the revelations about domestic spying at the NSA brought forth by Edward Snowden in 2013.
 [Weingart 2014] Kieran Healy has used network analysis to demonstrate how researchers can infer information from datasets while also touching on the troubling ethical issue of surveillance [Healy 2013].]
 
-One of the central kinds of information researchers can infer from a personal name is the gender of that individual: "Jane Fay," for instance, is almost certainly a woman, while her brother "John Fay" is almost certainly a man.
+One of the basic pieces of information researchers can infer from a personal name is the gender of that individual: "Jane Fay," for instance, is likely a woman, while her brother "John Fay" is likely a man.
 A computer program can infer their respective genders by looking up “Jane” and “John” in an existing dataset that matches names to genders.
@@
 That is because the conventional gender for the name Leslie switched over the course of one hundred years.
-In 1900 some 92% of the babies born in the United States who were named Leslie were male, while in 2000 about 96% of the Leslies born in that year were female.
+In 1900 some 92% of the babies born in the United States who were named Leslie were classified as male, while in 2000 about 96% of the Leslies born in that year were classified as female.
 
@@
 
-For those working on contemporary subjects the "Leslie problem" is not an especially pressing one.
-A film scholar analyzing male and female characters in modern American television shows, for instance, can turn to a variety of tools that use current databases of names, and these tools will likely identify the gender of Leslie Fay correctly.
-One online service, Genderize.io, determines the gender of first names "from user profiles across major social networks." [Genderize.io n.d.] Checking the service for the name Leslie returns a prediction that the name is female, along with an estimate of male/female proportions of the uses among members of unnamed social media sites.^[In this example the Genderize.io service has been accessed through an API wrapper provided in our Gender package described below.]
+For those working on contemporary subjects, the "Leslie problem" is not an especially pressing one.
+There are a variety of tools that use current databases of names, including Genderize.io, which predicts the gender of a first name from the user profiles of unnamed social media sites [Genderize.io n.d.].
+A film scholar analyzing male and female characters in a database of modern American television shows might use Genderize.io to accurately predict the gender of these characters, including Leslie Fay.
+**do we need this footnote?^[In this example the Genderize.io service has been accessed through an API wrapper provided in our Gender package described below.]** But researchers studying a longer timespan need to take into account the changing nature of naming practices.
+If that same film scholar wanted to compare the characters of modern television shows to characters during the silent film era,  existing tools like Genderize.io would misidentify Leslie Fay's gender in the earlier screenplays.
+Other names used in the United States such as Madison, Morgan, Sydney, and Kendall, have also flipped their gender over the past century and would similarly be misidentified.
+The prevalence of changes in the gender of names in the last half of the twentieth century implies that gendered lists of names drawn from contemporary data are anachronistic and so of limited use for datasets that encompass changing periods of time.^[Most names flip from male to female.
+The few exceptions, such as "Jan", have a different causal mechanism.
+We speculate that Jan as an English name is female but that Jan as a Scandinavian name is male.
+Because the name has remained relatively uncommon, other causes besides changing gender preferences can explain the change.]
 
 **I took out the genderize code here because I think it's not immediately intuitive for the reader, given that we haven't totally introduced our package yet. -CB**
-
-But researchers studying a longer timespan need to take into account the changing nature of naming practices.
-If that same film scholar wanted to compare the characters of modern television shows to characters during the silent film era,  existing tools like Genderize.io would misidentify Leslie Fay's gender in the earlier screenplays.
-Other names used in the United States such as Madison, Morgan, Sydney, and Kendall, have also flipped their gender over the past century and would similarly be misidentified.
-The prevalence of changes in the gender of names in the last half of the twentieth century implies that gendered lists of names drawn from contemporary data are anachronistic and so of limited use for datasets that encompass changing periods of time.^[The few exceptions of names which flip from being female to male, such as "Jan", have a different causal mechanism.
-We speculate that Jan as an English name is female but that Jan as a Scandinavian name is male.
-Because the name has remained relatively uncommon, other causes besides changing gender preferences can explain the change.]
 
@@
 Anyone studying a period longer than a few years, or anyone studying a group whose demographics do not match the groups used by a contemporary tool such as Genderize.io will also encounter this problem.
-For example, a literary scholar studying a corpus of poems from the mid-twentieth century might wish to compare the stanza structures of male and female poets.
-Using a tool like Genderize.io based on contemporary data risks misidentifying these writers, many of whom were born (and named) more than a century prior to the creation of these modern name datasets.
+A literary scholar studying a corpus of poems from the mid-twentieth century might wish to compare the stanza structures of male and female poets.
+Existing tools based on contemporary data risks misidentifying these writers, many of whom were born (and named) more than a century prior to the creation of these modern name datasets.
 This problem will only increase with growing life expectancies: as of 2012, the average American lifespan was nearly seventy-nine years---more than enough time for naming practices to change quite dramatically between a person’s birth and death [World Bank 2014].
@@
 It thus can provide not only a prediction of the gender of a name, but also a measure of that prediction's likelihood.
-For example, our program predicts that a person named Leslie born in 1950 is female, but only just barely: 52 percent of the babies named Leslie born in the United States in 1950 were girls and 48 percent were boys.
+For example, our program predicts that a person named Leslie who was born in 1950 is female, but only just barely: 52 percent of the babies named Leslie born in the United States in 1950 were girls and 48 percent were boys.
 Researchers can use these probabilities to determine what level of certainty is acceptable for predicting the gender of different names.
@@
 
-As with any tool, researchers who use this method need to be fully aware of the ways it replicates a problematic, state-defined gender binary.
+Researchers who use this method need to be fully aware of the ways it replicates a problematic, state-defined gender binary.
 The onus is on the user to understand how its underlying data was gathered, what it implies, and how it reflects social and cultural practices.
@@
 
-In a 2005 report for the American Historical Association, Elizabeth Lunbeck acknowledged "a sea change in the [historical] profession with respect to gender" before going on to describe the painful limitations of this sea change for female historians: ongoing personal discrimination, lower salaries, and barriers to securing high-ranking positions.
+In a 2005 report for the American Historical Association, Elizabeth Lunbeck acknowledged "a sea change in the [historical] profession with respect to gender" before going on to describe the limits to this change for female historians, who face ongoing personal discrimination, lower salaries, and barriers to securing high-ranking positions.
 Lunbeck's report drew in part on a survey of 362 female historians that produced a rich source of responses detailing the deep and multi-faceted challenges facing women in the profession [Lunbeck 2005].
 What follows is a quantitative supplement to Lunbeck's analysis that uses our program to analyze gender representation amongst historians across a much larger scale and a much longer time period.
-We study one of the bedrocks of the historical profession, scholarly research, and focus on two types of scholarly output: the dissertation and the monograph.
+We study one of the bedrocks of the historical profession --- scholarly research --- and focus on two types of output: the dissertation and the monograph.
 Although women have achieved rough parity with men in terms of the number of completed dissertations, female historians continue to face an unequal playing field in the monographic landscape.
@@
 
-The digital turn has made available very large datasets, yet in the midst of abundance many humanists continue to deal with the problem of incompleteness and scarcity within their sources.
+The digital turn has made large datasets widely available, yet in the midst of abundance many humanists continue to deal with the problem of incompleteness and scarcity within their sources.
 Our method uses *abundant* existing datasets to infer additional information from *scarce* datasets.
-Its major contribution is to incorporate a historical approach to this algorithm, one that takes into account how gendered naming practices have changed over time.
-This temporal methodology provides much improved results over simpler, anachronistic, or ahistorical lookup methods.
+Its major contribution is to incorporate a historical method, one that takes into account how gendered naming practices have changed over time.
+This temporal approach provides much improved results over simpler, anachronistic, or ahistorical lookup methods.
 The Gender package is, of course, far from perfect.
 By relying on historical datasets collected by state agencies, it reinforces a gender binary and larger power structure that obscures how people may identify themselves and how they experience the world.
-To borrow a phrase from Miriam Posner, the Gender package does not meet the "radical, unrealized potential of digital humanities." But it might.
+To borrow a phrase from Miriam Posner, the Gender package does not necessarily meet the "radical, unrealized potential of digital humanities" [Posner 2015].
+But it might.
 
-The Gender package is entirely open-source, and we have provided detailed and transarent documentation of its code and data.
-Researchers who wish to build upon it can modify the package accordingly.
-To take one imagined scenario: Facebook allows its users to either select from a large spectrum of gender identities or to write in their own prefered identity.
+The Gender package is entirely open-source, and we have provided detailed documentation for its code and underlying data so that researchers who wish to build upon it can modify the package accordingly.
+To take one imagined scenario: Facebook allows its users to select from a spectrum of gender identities or to write in their own prefered identity.
 Because these identities are associated with the user's first name and date of birth, this is exactly the kind of historical dataset that could be incorporated into the Gender package.
-Doing so would replace the package's simplistic categories of "Male" and "Female" with ones like "Female to Male", "Bigender", or "Cisgender Female".
-Although this imagined scenario would create its own set of challenges, it would represent a major improvement in how the Gender package dealt with the complexity and fluidity of gender identity.
-^[This imagined scenario runs into its own set of problems.
+Doing so would replace the package's simplistic categories of "Male" and "Female" with more detailed categories like "Female to Male", "Bigender", or "Cisgender Female".
+Although this imagined scenario would create its own set of challenges, it would represent a major improvement in how the Gender package dealt with the complexity and fluidity of gender identity.^[This imagined scenario runs into its own set of problems.
 Facebook's business model makes it unlikely to grant researchers open access to its proprietary and closely-guarded datasets.
 And even if it did grant access to this data, Facebook collected it for particular purposes and in particular ways.
-Just as government agencies like the Social Security Administration collect data to serve the purposes of the American state, Facebook's data collection reflects the priorities of a global corporation rather than the lived experience of its users.] By building as much flexibility and transparency into the method as possible, we hope that others can continue to make just these kinds of improvements.
+Just as government agencies like the Social Security Administration collect data to serve the purposes of the American state, Facebook's data collection more accurately reflects the priorities of a global corporation rather than the lived experience of its users.] By building as much flexibility and transparency into the method as possible, we hope that others can continue to make just these kinds of improvements.
 
-Even as it currently stands, the Gender package meets a pressing methodological need.
-Since its initial release in February 2014, the Gender package has been downloaded more than ________ times [??????] and incorporated as a pedagogical tool into digital humanities courses [Jockers 2014].
-Researchers have used it to uncover gender inequity in a range of different areas, from biases in students' teaching evaluations to disparities between attendees and authors at the annual Digital Humanities Conference [Schmidt 2015, Weingart 2015].
+As it currently stands, the Gender package meets a pressing methodological need.
+Since its initial release in February 2014, the Gender package has been downloaded more than **________ times [??????]** and incorporated as a pedagogical tool within digital humanities courses [Jockers 2014].
+Researchers have used it to uncover instances gender inequity in a range of different areas, from biases in students' teaching evaluations to disparities between attendees and authors at the annual Digital Humanities Conference [Schmidt 2015, Weingart 2015].
 These implementations help illustrate the value of the method.
 Its major intervention is to algorithmically inferring gender in textual sources in a way that takes into account change over time.
-Rather than trying to shoehorn existing computational methods, the Gender package uses a humanistic starting point---the historical relationship between gender and naming practices---to contribute a new method to computer science.
-This reverses the traditional disciplinary direction between the humanities and computer science.
+Rather than trying to shoehorn existing computer science methods to fit humanities research, the Gender package reverses this traditional disciplinary relationship.
+It uses a humanistic starting point---the historical relationship between gender and naming practices---to contribute a new method to computer science.
 
@@
 3.2 (2009).
-<http://www.digitalhumanities.org/dhq/vol/3/2/000042/000042.html>
+<http://www.digitalhumanities.org/dhq/vol/3/2/000042/000042.html>.
 
@@
 
-[Orlando Project n.d.] Orlando Project. "About." <http://www.artsrn.ualberta.ca/orlando/?page_id=2>. Accessed 23 June 2015.]
+[Orlando Project n.d.] Orlando Project. "About." <http://www.artsrn.ualberta.ca/orlando/?page_id=2>. Accessed 23 June 2015.
 
@@
 
-[R Core Team 2014-] R Core Team. *R: A language and environment for statistical computing*. R Foundation for Statistical Computing, Vienna, Austria, 2014. <http://www.R-project.org/>
+[R Core Team 2015-] R Core Team. *R: A language and environment for statistical computing*. R Foundation for Statistical Computing, Vienna, Austria, 2015. <http://www.R-project.org/>.
 
@@
 "The Status of Women and Minorities in the History Profession, 2008." *Perspectives on History*.
-September 2008, <http://www.historians.org/publications-and-directories/perspectives-on-history/september-2008/the-status-of-women-and-minorities-in-the-history-profession-2008>
+September 2008, <http://www.historians.org/publications-and-directories/perspectives-on-history/september-2008/the-status-of-women-and-minorities-in-the-history-profession-2008>.
 
@@
 A Field Report from the Higher Ranks." *Perspectives on History*.
-December 2012, <http://www.historians.org/publications-and-directories/perspectives-on-history/december-2012/what-makes-a-successful-academic-career-in-history>
+December 2012, <http://www.historians.org/publications-and-directories/perspectives-on-history/december-2012/what-makes-a-successful-academic-career-in-history>.
 
@@
 
-[World Bank 2014-] World Bank. "Life expectancy at birth, total (years)." *World Bank Open Data*. <http://data.worldbank.org/indicator/SP.DYN.LE00.IN>. Accessed 12 September 2014
+[World Bank 2014-] World Bank. "Life expectancy at birth, total (years)." *World Bank Open Data*. <http://data.worldbank.org/indicator/SP.DYN.LE00.IN>. Accessed 12 September 2014.

### 2015-08-14 article.Rmd
Revise introductory text

@@
 This article describes a new method for inferring the gender of personal names using large historical datasets.
-In contrast to existing methods of predicting gender that treat names as if they are timelessly associated with gender, this method uses a historical approach that takes into account the dynamic nature of naming practices.
+In contrast to existing methods of predicting gender that treat names as if they are timelessly associated with genders, this method uses a historical approach that takes into account changes in naming practices.
 It uses historical data to measure the likelihood that a name was associated with a particular gender based on the time or place under study.
@@
 
-The analytical category of gender has transformed humanities research over the past half-century, and more recently, the digitization movement has opened up opportunities to apply this category to new sources and in new ways.
-Initiatives like the Orlando Project, the Poetess Archive, and the Women Writers Project have done crucial work to recover the voices of female authors and writers [Orlando Project, n.d., Poetess Archive n.d., Women Writers Project n.d.].
+The analytical category of gender has transformed humanities research over the past half-century, and, more recently, widespread digitization has opened up opportunities to apply this category to new sources and in new ways.
+Initiatives like the [Orlando Project](http://www.artsrn.ualberta.ca/orlando/), the [Poetess Archive](http://idhmc.tamu.edu/poetess/), and the [Women Writers Project](http://www.wwp.northeastern.edu/) have done crucial work to recover the voices of female authors and writers [Orlando Project, n.d., Poetess Archive n.d., Women Writers Project n.d.].
 Other scholars have used digital analysis to study linguistic styles of male and female playwrights or representations of gendered bodies in European fairytales [Argamon et.

### 2015-08-14 article.Rmd
Improve the line of fit

@@
   ggplot(aes(x = year, y = prop_female)) +
-  geom_smooth() + 
+  geom_smooth(span = 0.5) + 
   geom_point() +

### 2015-08-14 article.Rmd
Update

@@
 For those working on contemporary subjects, the "Leslie problem" is not an especially pressing one.
-There are a variety of tools that use current databases of names, including Genderize.io, which predicts the gender of a first name from the user profiles of unnamed social media sites [Genderize.io n.d.].
+There are a variety of tools that use current databases of names, including [Genderize.io](https://genderize.io/), which predicts the gender of a first name from the user profiles of unnamed social media sites [Genderize.io n.d.].
 A film scholar analyzing male and female characters in a database of modern American television shows might use Genderize.io to accurately predict the gender of these characters, including Leslie Fay.
-**do we need this footnote?^[In this example the Genderize.io service has been accessed through an API wrapper provided in our Gender package described below.]** But researchers studying a longer timespan need to take into account the changing nature of naming practices.
+But researchers studying a longer timespan need to take into account the changing nature of naming practices.
 If that same film scholar wanted to compare the characters of modern television shows to characters during the silent film era,  existing tools like Genderize.io would misidentify Leslie Fay's gender in the earlier screenplays.
 Other names used in the United States such as Madison, Morgan, Sydney, and Kendall, have also flipped their gender over the past century and would similarly be misidentified.
-The prevalence of changes in the gender of names in the last half of the twentieth century implies that gendered lists of names drawn from contemporary data are anachronistic and so of limited use for datasets that encompass changing periods of time.^[Most names flip from male to female.
+The prevalence of changes in the gender of names in the last half of the twentieth century implies that gendered lists of names drawn from contemporary data are anachronistic and so of limited use for datasets that encompass changing periods of time.^[In general, names tend to shift from being predominantly male to being predominantly female; the shift rarely happens in the other direction.
+In general, names that transition from male to female also became increasingly popular over time.
+For example, the name Madison changed from male to female quite suddenly after 1985.
+Whereas before it had been an infrequently used male name, after 1985 it skyrocketed in popularity and by the early 2000s was one of the ten most frequently used names for baby girls.
+Broadly speaking, in the late twentieth century Americans girls became increasingly more likely than American boys to receive a "trendy" name, and once that name was associated with women it was unlikely to be given to male babies.
 The few exceptions, such as "Jan", have a different causal mechanism.
@@
 
-**I took out the genderize code here because I think it's not immediately intuitive for the reader, given that we haven't totally introduced our package yet. -CB**
-
-**I propose cutting this part or moving it to another section, as I think we need to move faster to get into the payoff of our method: Such changes usually follow set patterns.
-In general, names tend to shift from being predominantly male to being predominantly female; the shift rarely happens in the other direction.
-In general, names that transition from male to female also became increasingly popular over time.
-For example, the name Madison changed from male to female quite suddenly after 1985.
-Whereas before it had been an infrequently used male name, after 1985 it skyrocketed in popularity and by the early 2000s was one of the ten most frequently used names for baby girls.
-Broadly speaking, in the late twentieth century Americans girls became increasingly more likely than American boys to receive a "trendy" name, and once that name was associated with women it was unlikely to be given to male babies.
-These names present a moving target for researchers hoping to use them to infer additional information about the people under study.**
-
 ```{r figure02}
 plot_names <-  c("madison",
-                 #"keanu", #keanu doesn't seem to actually flip, or at least doesn't show up on the chart?
                  "addison",
@@
 The algorithm calculates the proportion of male or female uses of a name in a given birth year or range of years.
-It thus can provide not only a prediction of the gender of a name, but also a measure of that prediction's likelihood.
-For example, our program predicts that a person named Leslie who was born in 1950 is female, but only just barely: 52 percent of the babies named Leslie born in the United States in 1950 were girls and 48 percent were boys.
+It thus can provide not only a prediction of the gender of a name, but also a measure of that prediction's accuracy.
+For example, our program predicts that a person named Leslie who was born in 1950 is female, but reports a low level of certainty in that prediction, since just 52 percent of the babies named Leslie born in the United States in 1950 were girls and 48 percent were boys.
 Researchers can use these probabilities to determine what level of certainty is acceptable for predicting the gender of different names.
 
-It is important to understand our method’s limitations as well as its advantages.
+It is important to understand our method's limitations as well as its advantages.
 The approach of inferring gender from personal names is still a blunt way to study a complex subject.

### 2015-08-15 article.Rmd
Finish revising introduction

@@
 
-It is important to understand our method's limitations as well as its advantages.
-The approach of inferring gender from personal names is still a blunt way to study a complex subject.
-Gender theorists and feminist scholars have spent decades unpacking the full range of meanings behind gender as an analytical concept, including the fluid relationship between biological sex and individual identity [Butler 1990, Butler 1993].
-On its own, our method has little to say about crucial topics such as the intersectionality of gender with race, class, and power, or what the categories “male” or “female” actually mean in any given historical and social context [Crenshaw 1991, McCall 2005].
-Perhaps most importantly, the tool itself does not reflect how an individual person self-identifies and how gender shapes their lived experience.
-The question of gender identity is especially relevant for our method, given its reliance on large datasets created by state agencies such as the U.S.
+Researchers who use this method need to be fully aware of its limitations, in particular its dependency on a state-defined gender binary.
+Inferring gender from personal names is a blunt tool to study a complex subject.
+Gender theorists and feminist scholars have spent decades unpacking the full range of meanings behind gender as an analytical concept, including the fluid relationship between biological sex and gender identity [Butler 1990, Butler 1993].
+On its own, our method has little to say about crucial topics such as the intersectionality of gender with race, class, and power, or what the categories "male" or "female" mean in any given historical and social context [Crenshaw 1991, McCall 2005].
+The tool cannot reflect how an individual person self-identifies and how gender shapes their lived experience, since it relies on large datasets created by state agencies such as the U.S.
 Census Bureau.
-Much like the state’s definition of racial categories, governments impose their own vision of a gender binary onto their subjects in ways that are deeply problematic.
-This gender binary reflects existing power structures more than the lived identities of individuals.
-The state's gender categorization exerts a tremendous definitional sway over societies and the people who comprise them, most notably in its systematic exclusion of transgender identities.
+Much like the state’s definition of racial categories, governments impose a gender binary in their data collection that reflects existing power structures more than the lived identities of individuals.
 [Brubaker and Cooper 2000, 14-17, Pascoe 2009, Canaday 2011, Landrum 2014].
+When using the predictions provided by this method, users must judge for themselves how well they reflect the social and cultural practices of the time and place under study.
 
-Researchers who use this method need to be fully aware of the ways it replicates a problematic, state-defined gender binary.
-The onus is on the user to understand how its underlying data was gathered, what it implies, and how it reflects social and cultural practices.
-Imperfect as it may be, this method nevertheless gives digital humanities scholars an important means of studying gender in textual collections.
-We have made the method as transparent and flexible as possible through the inclusion of probability figures - ex.
-52 percent of the babies named Leslie born in the United States in 1950 were girls and 48 percent were boys.
+Imperfect as it admittedly is, this method nevertheless gives digital humanities scholars an important means of studying gender in textual collections.
+We have made the method as transparent and flexible as possible through the inclusion of probability figures.
 This makes the underlying data visible to users and allows them the opportunity to interrogate the assumptions that come with it.

### 2015-08-15 article.Rmd
Cite R and Python books

@@
 There are therefore many names that it cannot identify.
-Assuming for the moment that our method provides more accurate results, we estimate that `r percent(false_either_percent)` percent of the names in the Kantrowitz corpus are classified ambiguously when a gender could be reasonably predicted from the name, that `r percent(false_male_percent)` percent of the names are classified as male when they should be classified as female, and that `r percent(false_female_percent)` are classified as female when they should be classified as male.
+Assuming for the moment that our method provides more accurate results, we estimate that `r percent(false_either_percent)` percent of the names in the Kantrowitz corpus are classified as ambiguous when a gender could be reasonably predicted from the name, that `r percent(false_male_percent)` percent of the names are classified as male when they should be classified as female, and that `r percent(false_female_percent)` are classified as female when they should be classified as male.
 This error rate is a separate concern from the much smaller size of the Kantrowitz corpus.^[This error rate was calculated using the Social Security Administration database, assuming a range of years between 1932 and 2012, which is the default setting within the Gender package.
@@
 Genderize.io reports that its "database contains 142848 distinct names across 77 countries and 85 languages." Genderize is unsuitable for historical work, however, because it is based only on contemporary data.
-According to the documentation for its API, "It utilizes big datasets of information, from user profiles across major social networks." It would be anachronistic to apply these datasets to the past, and Genderize.io provides no functionality to filter results chronologically as it does geographically.
+According to the documentation for its API, "it utilizes big datasets of information, from user profiles across major social networks." It would be anachronistic to apply these datasets to the past, and Genderize.io provides no functionality to filter results chronologically as it does geographically.
 In addition, Genderize.io does not make clear exactly what comprises the dataset and how it was gathered, which keeps scholars from interrogating the value of the source [Genderize.io 2014-].^[Quotations are taken from the [Genderize.io website](http://genderize.io), with correspondence from the site's developer for confirmation.]
 
-R and Python are two of the most commonly used languages for data analysis [Piatesky 2014, Muenchen 2014-].^[Surveys and analysis of useage point out the growth of R and the continuing popularity of Python for "data science" generally.] Python's existing packages for gender prediction all implement a method similar to the NLTK or Genderize.io.^[These Python libraries include [estimate.gender](https://pypi.python.org/pypi/estimate.gender/0.4), [genderize](https://github.com/SteelPangolin/genderize), and [gender.py](https://github.com/block8437/gender.py)] To our knowledge, [CRAN](http://cran.rstudio.com/) (the central repository for R packages) does not include any packages dedicated to gender prediction.
+R and Python are two of the most commonly used languages for data analysis [Piatesky 2013, Muenchen 2014-].
+Surveys and analysis of useage point out the growth of R and the continuing popularity of Python for data science generally, and scholars are producing guides to using these languages for historical or humanities research [Programming Historian n.d., Arnold and Tilton 2015, Jockers 2014a].
+Yet Python's existing packages for gender prediction all implement a method similar to the NLTK or Genderize.io.^[These Python libraries include [estimate.gender](https://pypi.python.org/pypi/estimate.gender/0.4), [genderize](https://github.com/SteelPangolin/genderize), and [gender.py](https://github.com/block8437/gender.py)] The only other R package for gender prediction,   [genderizeR](https://cran.rstudio.com/web/packages/genderizeR/), uses the Genderize.io API.
 Thus two of the most popular data analysis languages used in the digital humanities currently have no satisfactory method for predicting gender from names for historical and humanities research.
 
-To that end we have created the [Gender](https://github.com/ropensci/gender) package for R which includes both the predictive algorithm and an associated [genderdata](https://github.com/lmullen/gender-data-pkg) package containing various historical datasets.
+To that end we have created the [Gender](https://github.com/ropensci/gender) package for R which includes both a predictive algorithm and an associated [genderdata](https://github.com/lmullen/gender-data-pkg) package containing various historical datasets.
 This R implementation is based on an [earlier Python implementation](https://github.com/cblevins/Gender-ID-By-Time) by [Cameron Blevins](http://www.cameronblevins.org/) and [Bridget Baird](http://www.conncoll.edu/directories/emeritus-faculty/bridget-baird/).
@@
 As it currently stands, the Gender package meets a pressing methodological need.
-Since its initial release in February 2014, the Gender package has been downloaded more than **________ times [??????]** and incorporated as a pedagogical tool within digital humanities courses [Jockers 2014].
+Since its initial release in February 2014, the Gender package has been downloaded more than **________ times [??????]** and incorporated as a pedagogical tool within digital humanities courses [Jockers 2014b].
 Researchers have used it to uncover instances gender inequity in a range of different areas, from biases in students' teaching evaluations to disparities between attendees and authors at the annual Digital Humanities Conference [Schmidt 2015, Weingart 2015].
@@
 <http://www.digitalhumanities.org/dhq/vol/3/2/000042/000042.html>.
+
+[Arnold and Tilton 2015] Arnold, Taylor, and Lauren Tilton. *Humanities Data in R: Exploring Networks, Geospatial Data, and Text*. Springer, USA, 2015.
 
@@
 
-[Jockers 2014] Jockers, Matthew.
+[Jockers 2014a] Jockers, Matthew. *Text Analysis with R for Students of Literature*. Springer, USA, 2014.
+
+[Jockers 2014b] Jockers, Matthew.
 "Experimenting with “gender” package in R." *Matthew L.
 Jockers*.
-25 February 2015, <http://www.matthewjockers.net/2014/02/25/experimenting-with-gender-package-in-r/>.
+25 February 2014, <http://www.matthewjockers.net/2014/02/25/experimenting-with-gender-package-in-r/>.
 Accessed 10 August 2015.
@@
 
-[Muenchen n.d.] Muenchen, Robert A. "The Popularity of Data Analysis Software," *r4stats.com*. <http://r4stats.com/articles/popularity/>. Accessed 12 September 2014.
+[Muenchen 2014-] Muenchen, Robert A. "The Popularity of Data Analysis Software," *r4stats.com*. <http://r4stats.com/articles/popularity/>. Accessed 12 September 2014.
 
@@
 Accessed 10 August 2015.
+
+[Programming Historian n.d.] The Programming Historian. <http://programminghistorian.org/>. Accessed 15 August 2015.
 

### 2015-08-15 article.Rmd
Why R

@@
 This R implementation is based on an [earlier Python implementation](https://github.com/cblevins/Gender-ID-By-Time) by [Cameron Blevins](http://www.cameronblevins.org/) and [Bridget Baird](http://www.conncoll.edu/directories/emeritus-faculty/bridget-baird/).
+
+We have chosen the R programming language for a constellation of reasons.
+The R language is open-source, so it is freely available to scholars and students.
+The language has a strong tradition of being friendly to scholarship.
+It was originally created for statistics and many of its core contributors are academics; it provides facilities citing packages as scholarship.
+[CRAN](https://cran.r-project.org/) (the Comprehensive R Archive Network) offers a central location for publishing R packages for all users.
+These include a rigorous set of standards to ensure the quality of packages contributed.
+R has a number of language features, such as data frames, which permit easy manipulation of data.^[The gender package depends especially on [R Core Team 2014-, Wickham and Francois 2014].] In particular, R permits the publication of packages containing data alongside packages containing code---a crucial feature for the Gender package.
 The Gender package is affiliated with [rOpenSci](http://ropensci.org/), an initiative that supports reproducible research and open data for scientists using R.
-The rOpenSci advisory board has provided code review and publicity for the project.^[The gender package depends especially on [R Core Team 2014-, Wickham and Francois 2014].]
+The rOpenSci advisory board has provided code review, publicity, and an infrastructure to support the project.
+Finally, our method is in principle extensible to other languages, such as Python, should other scholars choose to do so.
 

### 2015-08-15 article.Rmd
Revise description of datasets

@@
 
-The possibilities for inferring gender from names depends on two things.
-First, it needs a suitable (and suitably large) dataset for the time period and region under study.
+Inferring gender from names depends on two things.
+First, it requires a suitable (and suitably large) dataset for the time period and region under study.
 Unsurprisingly such datasets are almost always gathered in the first instance by governments, though their compilation and digitization may be undertaken by scholarly researchers.
@@
 The second has been to provide as many datasets as possible in order to localize the predictions to particular times and places.
-To that end, [Benjamin Schmidt](http://benschmidt.org/) contributed a dataset of US Census Data and accompanying code, as well as provided a significant correction to predictions based on Social Security Administration data.
 
-The Gender package currently uses two datasets which make it suitable for studying the United States from the first federal census in 1790 onwards.^[The authors of the package are both American historians, which explains why the initial release of the package focuses on that country.
-We have plans, however, to extend the underlying data to six other European and North American countries included in the [North Atlantic Population Project](https://www.nappdata.org/napp/) for the nineteenth century.
-We also welcome contributions of datasets for other regions and times from other scholars.] The first dataset contains names of applicants for Social Security and is available from [Data.gov](https://www.data.gov/).
+The Gender package currently uses several datasets which make it suitable for studying the United States from the first federal census in 1790 onwards.] The first dataset contains names of applicants for Social Security and is available from [Data.gov](https://www.data.gov/).
 The second dataset contains names gathered in the decennial censuses and is available from the IPUMS-USA (Integrated Public Use Microdata Series) service from the Minnesota Population Center at the University of Minnesota.
@@
 It is well-known among American historians that the Social Security Act was written to exclude farmworkers as a means of also excluding African Americans.
-In examining the dataset we have not found any reason to believe that this policy noticably skewed the prediction of gender from names.]
+In examining the dataset we have not found any reason to believe that this policy noticeably skewed the prediction of gender from names.]
 
-The IPUMS-USA dataset, contributed by Benjamin Schmidt, contains records from the United States decennial census from 1790 to 1930.
+The IPUMS-USA dataset, contributed by [Benjamin Schmidt](http://benschmidt.org/), contains records from the United States decennial census from 1790 to 1930.
 This dataset includes the birth year and numbers of males and females under the age of 62 for all the years in that range.
-This data has been aggregated by IPUMS at the University of Minnesota and is released as a sample of the total census data.
+This data has been aggregated by [IPUMS](https://usa.ipums.org/usa/) at the University of Minnesota and is released as a sample of the total census data.
 Unlike the SSA dataset, which includes a 100% sample for every name reported to the Social Security Administration and used more than five times, the IPUMS data contains 5% or 10% samples of names from the total census data.
@@
 
-These two SSA and IPUMS datasets (and any future datasets to be added to the package) thus contain a simple table of data.^[The format of this data departs from the more generally desirable "tidy data" framework proposed by Hadley Wickham for performance reasons.
+The next release of the Gender package will include a dataset from the [North Atlantic Population Project](https://www.nappdata.org/napp/).
+This dataset will include names from Canada, Great Britain, Germany, Iceland, Norway, and Sweden from 1801 to 1910.^[The North Atlantic Population Project, which is also project of the Minnesota Population Center, also includes data for the United States of America, but we have excluded it since IPUMS-USA and the SSA dataset is more suitable when focusing on that country alone.] The SSA, IPUMS, and NAPP datasets (and any future datasets to be added to the package) all have the same simple tabular format.^[The format of this data departs from the more generally desirable "tidy data" framework proposed by Hadley Wickham for performance reasons.
 By keeping the data in a "wider" format instead of a tidy format as defined by Wickham, the function is able to make its predictions faster [Wickham 2014b].] The columns contain the `name`, `year`, and number of `female` and `male` uses of that name in a particular year.
+This simple format makes it possible to extend the package to include any place and time period for which there is suitable data.^[We welcome such contributions at the [project's GitHub repository](https://github.com/ropensci/gender).]
 

### 2015-08-15 article.Rmd
Explain why we don't adjust for spellings

@@
 In other words, there is an approximately `r gender("sidney", year = 1935)$proportion_male %>% percent()` percent chance that a person born in 1935 named Sidney was male.
-In 2012, for comparison, there was an approximately `r gender("sidney", year = 2012)$proportion_female %>% percent()` percent chance that a person born named Sidney was female.
+In 2012, for comparison, there was an approximately `r gender("sidney", year = 2012)$proportion_female %>% percent()` percent chance that a person born named Sidney was female.^[The Gender package makes no effort to combine alternative spellings for names, such as Sidney and Sydney, since we think it is better to treat those as separate names.
+For instance, although they are homophones, Geri is predominately a female name, and Gerry is predominately a male name.]
 
@@
 genderdata::ssa_national %>%
-  filter(name == "sidney", year >= 1930 & year <= 1939) 
+  filter(name == "sidney", 
+         year >= 1930,
+         year <= 1939) 
 ```

### 2015-08-20 article.Rmd
Revisions to example

@@
 library(ggplot2)
+library(gridExtra)
 library(mullenMisc)
@@
 Thus, according to the Social Security Administration, there were 974 boys and 93 girls named Sidney born in 1935.
-We can add another command to calculate the *proportion* of females and males ("mutate" in the [dplyr](https://github.com/hadley/dplyr) package vocabulary) rather than raw numbers.
+We can add another command to calculate the *proportion* of females and males ("mutate" in the vocabulary of the [dplyr](https://github.com/hadley/dplyr) package) rather than raw numbers.
 
@@
 genderdata::ssa_national %>%
-  filter(name == "sidney", 
-         year >= 1930,
-         year <= 1939) 
+  filter(name == "sidney", year >= 1930, year <= 1939) 
 ```
@@
 Let's use, for example, a hypothetical list of editors from a college newspaper to illustrate how a researcher might apply it to their own data.
-The package's prediction function allows researchers to choose which reference datasets they would like to use and the range of years for making their predictions.^[See the package vignette for more details: `vignette("predicting-gender", package = "gender")`.]
+The package's prediction function allows researchers to choose which reference datasets they would like to use and the range of years for making their predictions.
 
-```{r}
-editors <- data.frame(name = c("Madison", "Madison", "Morgan",
+```{r, echo=FALSE}
+editors <- data_frame(name = c("Madison", "Madison", "Morgan",
                                   "Morgan", "Jan", "Jan"),
-                         birth_year = c(1934, 1990, 1948, 1970, 1965, 1998),
-                         stringsAsFactors = FALSE) %>% tbl_df()
+                         birth_year = c(1934, 1990, 1948, 1970, 1965, 1998))
 ```
@@
 
-The simplest way to use the package would be to call the `gender()` function, specifying the method, in this case the Social Security Administration dataset since all the names are after 1930, and convert the results to a data frame (one of R's native data structures).
+The simplest way to use the package would be to call the `gender()` function, specifying the method, in this case the Social Security Administration dataset since all the names are after 1930, and specify the range of years in the dataset.
+However, using
 
-```{r, echo=TRUE, results='markup'}
-gender(editors$name, method = "ssa") %>% do.call(rbind.data.frame, .)
-```
+^[See the [package vignette](https://cran.rstudio.com/web/packages/gender/vignettes/predicting-gender.html) for more details about how to use the package effectively, including suggestions about performance: `vignette("predicting-gender", package = "gender")`.]
 

### 2015-08-21 article.Rmd
Revised methods section

@@
 ```{r, echo=FALSE}
-editors <- data_frame(name = c("Madison", "Madison", "Morgan",
-                                  "Morgan", "Jan", "Jan"),
+editors <- data_frame(name = c("Madison", "Madison", "Morgan", "Morgan",
+                               "Jan", "Jan"),
                       birth_year = c(1934, 1990, 1948, 1970, 1965, 1998))
@@
 
-The simplest way to use the package would be to call the `gender()` function, specifying the method, in this case the Social Security Administration dataset since all the names are after 1930, and specify the range of years in the dataset.
-However, using
+The simplest way to use the package would be to call the `gender()` function, specifying the method, in this case the Social Security Administration dataset since all the names are after 1930, and specifying a single range of years to make the predictions.
+However, we can instead use the `gender_df()` function provided by the package.
+By passing it our data frame of people and specifying which columns contain the names and birth years (or a columns containing a minimum and maximum for a range of years), we can efficiently predict the gender for each person.^[See the [package vignette](https://cran.rstudio.com/web/packages/gender/vignettes/predicting-gender.html) for more details about how to use the package effectively, including suggestions about performance: `browseVignettes( "gender")`.]
 
-^[See the [package vignette](https://cran.rstudio.com/web/packages/gender/vignettes/predicting-gender.html) for more details about how to use the package effectively, including suggestions about performance: `vignette("predicting-gender", package = "gender")`.]
-
-This basic call of the `gender()` function would be suitable if we did not know the birth year for each editor. But we can also specify that information in the `gender()` function by using this call:
-
-```{r, echo=TRUE, results='markup'}
-Map(gender, editors$name, years = editors$birth_year) %>% 
-  do.call(rbind.data.frame, .) %>% print.data.frame(row.names = FALSE)
+```{r}
+gender_df(editors, name_col = "name", year_col = "birth_year")
 ```
@@
 We could then calculate a minimum and a maximum year of birth for each, and run the prediction function on that range for each person.
-The exact code to accomplish these types of analysis can be found in the Gender package's vignette.
+The exact code to accomplish these types of analysis can be found in the [vignette for the gender package](https://cran.rstudio.com/web/packages/gender/vignettes/predicting-gender.html).
 
-As previously mentioned, the history of how the Social Security Administration collected the data effects its validity.
+As previously mentioned, the history of how the Social Security Administration collected the data affects its validity.
 Specifically, because the data extends back to 1880 but the first applications were gathered after 1935, the sex ratios in the dataset are skewed in the years before 1930.
@@
 
-It is extremely improbable that nearly seventy percent of the people born in 1900 were female.
-Exactly why the dataset has this bias is unclear, though we speculate that it is because the applicants for Social Security in the early years of the program were approaching retirement age, which was set much closer to the average life expectancy in 1935 than it is today.
+It is extremely improbable that nearly seventy percent of the people born in 1900 were female.^[We are grateful to Ben Schmidt for drawing out attention to this phenomenon.] Exactly why the dataset has this bias is unclear, though we speculate that it is because the applicants for Social Security in the early years of the program were approaching retirement age, which was set much closer to the average life expectancy in 1935 than it is today.
 Since women tend to live longer than men, they were overrepresented in Social Security applications.

### 2015-08-21 article.Rmd
Revisions to AHR section

@@
 What follows is a quantitative supplement to Lunbeck's analysis that uses our program to analyze gender representation amongst historians across a much larger scale and a much longer time period.
-We study one of the bedrocks of the historical profession --- scholarly research --- and focus on two types of output: the dissertation and the monograph.
+We study one of the bedrocks of the historical profession---scholarly research---and focus on two types of output: the dissertation and the monograph.
 Although women have achieved rough parity with men in terms of the number of completed dissertations, female historians continue to face an unequal playing field in the monographic landscape.
@@
 We begin with the history dissertation, often the defining scholarly output of a historian's early career.
-The completion of a dissertation marks a standardized moment of transition out of the "training" phase of the historical profession.
+The completion of a dissertation marks a standardized moment of transition out of the training phase of the historical profession.
 To identify how many women and men completed PhD-level training in history each year, we used data supplied by ProQuest for roughly eighty thousand PhD dissertations completed in the United States between 1950 and 2012.^[ProQuest's dataset does not permit us to limit dissertations just to dissertations produced in history departments.
@@
 Second, a book's appearance in the *AHR* is not necessarily correlated with the *quality* of the book, as the journal prints negative reviews as well as positive ones.
-Nevertheless, when a book appears in the *AHR* it serves as a signal that other historians in positions of power have "approved" it for consideration.
+Nevertheless, when a book appears in the *AHR* it serves as a signal that other historians in positions of power are taking it seriously.
 Even if it garners a negative review, the fact that it appears in the journal at all is a measure of the fact that the profession's gatekeepers have deemed it important enough to review.
-It is precisely this professional gatekeeping dimension that we can analyze using the Gender package: are women more likely to be barred from this gate than men, and if so, by how much?
+It is precisely this professional gatekeeping dimension that we can analyze using the gender package: are books by men more likely to be reviewed then books by women, and if so, how much more likely?
 
@@
 Not every historian wants or seeks a career path that neatly moves from writing a dissertation to authoring a monograph to being reviewed in the *American Historical Review*.
-For those that do, however, each of these subsequent stages represents a winnowing process: not all graduate students produce dissertations, not all PhD’s write books, not all books are submitted to the *American Historical Review*, and not all submissions are accepted by the *AHR* editors for review.^[As of 2015 the journal's website reminds readers that "the *AHR* receives over 3,000 books a year; we have the resources to publish at most 1,000 reviews a year (approximately 200 per issue)." [American Historical Review n.d.]] Slighty fewer women than men go through the initial stage of writing a history dissertation.
-But by the final round of gatekeeping – a monograph reviewed in the *AHR* – the gender gap grows until there are little more than half as many women as men.
+For those that do, however, each of these subsequent stages represents a winnowing process: not all graduate students produce dissertations, not all PhDs write books, not all books are submitted to the *American Historical Review*, and not all submissions are accepted by the *AHR* editors for review.^[As of 2015 the journal's website reminds readers that "the *AHR* receives over 3,000 books a year; we have the resources to publish at most 1,000 reviews a year (approximately 200 per issue)." [American Historical Review n.d.]] Slighty fewer women than men go through the initial stage of writing a history dissertation.
+But by the final round of gatekeeping---a monograph reviewed in the *AHR*---the gender gap grows until there are little more than half as many women as men.
 The *American Historical Review* is, unfortunately, far from alone.
@@
 
-Examples like the Pulitzer and Bancroft Prizes are only the most glaring cases of gender inequity in the historical profession.
+Examples like the Pulitzer and Bancroft Prizes are only the most obvious cases of gender inequity in the historical profession.
 Overt discrimination still exists, but inequity operates in ways that are far more cloaked and far more complex, making them difficult to recognize and remedy.
@@
 All of these factors combine and amplify each other to contribute to the gender gap in the historical profession on display in the pages of the *American Historical Review*.
-The hidden, murky, and systemic dimensions of gender inequity makes methods like the Gender package all the more imperative for reasearchers.
+The hidden, murky, and systemic dimensions of gender inequity makes methods like the gender package all the more imperative for reasearchers.
 
-Finally, we used the Gender package to reveal some of the subtler ways in which gender iniquity operates in the historical profession.
+Finally, we used the gender package to reveal some of the subtler ways in which gender iniquity operates in the historical profession.
 After inferring the gender of book authors reviewed by the *AHR*, we were also able to infer the gender of the reviewers themselves.
@@
 First, the story is much the same for reviewers as it is for authors: more than twice as many men as women appear as reviewers in the journal.
-This is important because reviewers for the *AHR* garner a certain degree of prestige from their work - perhaps not to the same prestige as winning a Pulitzer Prize, but prestigious nonetheless.
-To become a reviewer, one must possess a Ph.D.
-or equivalent degree and have published at least one book-length manuscript [American Historical Review n.d.].
+This is important because reviewers for the *AHR* garner a certain degree of prestige from their work---perhaps not to the same prestige as winning a Pulitzer Prize, but prestige nonetheless.
+To become a reviewer, one must possess a PhD or equivalent degree and have published at least one book-length manuscript [American Historical Review n.d.].
 *AHR* reviewers are understood to have sufficient experience and command over a given sub-field to evaluate the merits and weaknesses of a new contribution to that literature.
 Being a reviewer for the journal is, in many ways, a proxy for expertise.
-Which is a problem.
 When there are twice as many male "experts" as female "experts" appearing in the *AHR* it imbues our understanding of what constitutes historical expertise.
-Or perhaps more accurately, who is and who isn't an "expert" in the field.
 
@@
 In the case of female-authored books, the ratio of female-to-male reviewers is closer to 50/50.
-In short, women are far more likely to write reviews of other women, while men tend to gravitate more towards male authors – who are, of course, already over-represented in the *AHR*.
+In short, women are far more likely to write reviews of other women, while men tend to gravitate more towards male authors---who are, of course, already over-represented in the *AHR*.
 
@@
 
-The findings related to *AHR* reviewers help illustrate the ways in which gender inequity is so insidious.
+The findings related to *AHR* reviewers help illustrate the ways in which gender inequity in the profession is so insidious.
 It is not always an overt case of deciding to award a prestigious prize to a male historian; it can be as seemingly innocuous as deciding to write a book review of a male author's book.
 Countless individual actions like these constitute the core relationships that knit together the profession, and too often these tiny actions benefit men more than women: chairing a panel of mostly male historians, taking on a new male advisee, or assigning mostly male-authored books on a syllabus.
-On their own they might appear benign - one might argue that male historians simply happen to be interested in similar research topics - but in aggregate they contribute to a broader professional landscape that puts female historians at a disadvantage.
+On their own they might appear benign---one might argue that male historians simply happen to be interested in similar research topics---but in aggregate they contribute to a broader professional landscape that puts female historians at a disadvantage.
 
-One 2013 analysis of 2,500 recent history Ph.D.'s found that "gender played little role in employment patterns across particular professions and industries" [Wood and Townsend 2013].
+One 2013 analysis of 2,500 recent history PhDs found that "gender played little role in employment patterns across particular professions and industries" [Wood and Townsend 2013].
 Our own analysis reveals a different story, at least in terms of getting a book reviewed in the field's most prestigious journal.
@@
 Women might be getting hired at comparable rates to men, but discriminatory gatekeeping remains alive and well in the profession.
-The Gender package is one more tool to uncover those practices.
+The gender package is one more tool to uncover those practices.
 
@@
 This temporal approach provides much improved results over simpler, anachronistic, or ahistorical lookup methods.
-The Gender package is, of course, far from perfect.
+The gender package is, of course, far from perfect.
 By relying on historical datasets collected by state agencies, it reinforces a gender binary and larger power structure that obscures how people may identify themselves and how they experience the world.
@@
 
-The Gender package is entirely open-source, and we have provided detailed documentation for its code and underlying data so that researchers who wish to build upon it can modify the package accordingly.
+The gender package is entirely open-source, and we have provided detailed documentation for its code and underlying data so that researchers who wish to build upon it can modify the package accordingly.
 To take one imagined scenario: Facebook allows its users to select from a spectrum of gender identities or to write in their own prefered identity.
-Because these identities are associated with the user's first name and date of birth, this is exactly the kind of historical dataset that could be incorporated into the Gender package.
+Because these identities are associated with the user's first name and date of birth, this is exactly the kind of historical dataset that could be incorporated into the gender package.
 Doing so would replace the package's simplistic categories of "Male" and "Female" with more detailed categories like "Female to Male", "Bigender", or "Cisgender Female".

### 2015-08-21 article.Rmd
Remove hypothetical scenario

@@
 The gender package is entirely open-source, and we have provided detailed documentation for its code and underlying data so that researchers who wish to build upon it can modify the package accordingly.
-To take one imagined scenario: Facebook allows its users to select from a spectrum of gender identities or to write in their own prefered identity.
-Because these identities are associated with the user's first name and date of birth, this is exactly the kind of historical dataset that could be incorporated into the gender package.
-Doing so would replace the package's simplistic categories of "Male" and "Female" with more detailed categories like "Female to Male", "Bigender", or "Cisgender Female".
-Although this imagined scenario would create its own set of challenges, it would represent a major improvement in how the Gender package dealt with the complexity and fluidity of gender identity.^[This imagined scenario runs into its own set of problems.
-Facebook's business model makes it unlikely to grant researchers open access to its proprietary and closely-guarded datasets.
-And even if it did grant access to this data, Facebook collected it for particular purposes and in particular ways.
-Just as government agencies like the Social Security Administration collect data to serve the purposes of the American state, Facebook's data collection more accurately reflects the priorities of a global corporation rather than the lived experience of its users.] By building as much flexibility and transparency into the method as possible, we hope that others can continue to make just these kinds of improvements.
+By building as much flexibility and transparency into the method as possible, we hope that others can continue to make just these kinds of improvements.
 

### 2015-08-21 article.Rmd
Estimate of downloads

@@
 
-As it currently stands, the Gender package meets a pressing methodological need.
-Since its initial release in February 2014, the Gender package has been downloaded more than **________ times [??????]** and incorporated as a pedagogical tool within digital humanities courses [Jockers 2014b].
-Researchers have used it to uncover instances gender inequity in a range of different areas, from biases in students' teaching evaluations to disparities between attendees and authors at the annual Digital Humanities Conference [Schmidt 2015, Weingart 2015].
+As it currently stands, the gender package meets a pressing methodological need.
+As of August 2015, the Gender package has been downloaded more than 21,000 times and incorporated as a pedagogical tool within digital humanities courses [Jockers 2014b].^[The download counts come from the [download logs from RStudio's CRAN mirror](http://cran-logs.rstudio.com/) as tabulated by Gabor Csardi's [cranlogs package](http://cranlogs.r-pkg.org/).
+[Csardi 2015]) Researchers have used it to uncover instances gender inequity in a range of different areas, from biases in students' teaching evaluations to disparities between attendees and authors at the annual Digital Humanities Conference [Schmidt 2015, Weingart 2015].
 These implementations help illustrate the value of the method.
@@
 [Crenshaw 1991] Crenshaw, Kimberle. "Mapping the Margins: Intersectionality, Identity Politics, and Violence against Women of Color." *Stanford Law Review*. 43.6 (July 1991): 1241-1299.
+
+[Csardi 2015]  Csardi, Gabor. cranlogs: Download Logs from the RStudio CRAN Mirror. R package version 2.0.0 (2015): <http://CRAN.R-project.org/package=cranlogs>
 

### 2015-08-21 article.Rmd
Revise conclusion

--- 2015-08-21 article.Rmd
@@
 By relying on historical datasets collected by state agencies, it reinforces a gender binary and larger power structure that obscures how people may identify themselves and how they experience the world.
-To borrow a phrase from Miriam Posner, the Gender package does not necessarily meet the "radical, unrealized potential of digital humanities" [Posner 2015].
-But it might.
+To borrow a phrase from Miriam Posner, the gender package does not necessarily meet the "radical, unrealized potential of digital humanities" [Posner 2015].
 
-The gender package is entirely open-source, and we have provided detailed documentation for its code and underlying data so that researchers who wish to build upon it can modify the package accordingly.
+Nevertheless, the methodology that have described here does take advantage of the datasets available to historians.
+The predictions it provides are not intended to be taken as the entirety of what historians can say about gender in the past.
+But they provide the ability to uncover data about gender where it was previously unknown, thus providing further grist to the mill of more traditional methods.
+Furthermore, the gender package is entirely open-source, and we have provided detailed documentation for its code and underlying data so that researchers who wish to build upon it can modify the package accordingly.
 By building as much flexibility and transparency into the method as possible, we hope that others can continue to make just these kinds of improvements.
 
-As it currently stands, the Gender package meets a pressing methodological need.
-Since its initial release in February 2014, the Gender package has been downloaded more than **________ times [??????]** and incorporated as a pedagogical tool within digital humanities courses [Jockers 2014b].
-Researchers have used it to uncover instances gender inequity in a range of different areas, from biases in students' teaching evaluations to disparities between attendees and authors at the annual Digital Humanities Conference [Schmidt 2015, Weingart 2015].
+As it currently stands, the gender package meets a methodological need.
+As of August 2015, the gender package has been downloaded more than 21,000 times and incorporated as a pedagogical tool within digital humanities courses [Jockers 2014b].^[The download counts come from the [download logs from RStudio's CRAN mirror](http://cran-logs.rstudio.com/) as tabulated by Gabor Csardi's [cranlogs package](http://cranlogs.r-pkg.org/).
+[Csardi 2015] RStudio is the only CRAN mirror to release its logs.) Researchers have used it to uncover instances gender inequity in a range of different areas, from biases in students' teaching evaluations to disparities between attendees and authors at the annual Digital Humanities Conference [Schmidt 2015, Weingart 2015].
 These implementations help illustrate the value of the method.
 Its major intervention is to algorithmically inferring gender in textual sources in a way that takes into account change over time.
-Rather than trying to shoehorn existing computer science methods to fit humanities research, the Gender package reverses this traditional disciplinary relationship.
+Rather than trying to shoehorn existing computer science methods to fit humanities research, the gender package reverses this traditional disciplinary relationship.
 It uses a humanistic starting point---the historical relationship between gender and naming practices---to contribute a new method to computer science.

### 2015-08-21 article.Rmd
Fix footnote

@@
 As of August 2015, the gender package has been downloaded more than 21,000 times and incorporated as a pedagogical tool within digital humanities courses [Jockers 2014b].^[The download counts come from the [download logs from RStudio's CRAN mirror](http://cran-logs.rstudio.com/) as tabulated by Gabor Csardi's [cranlogs package](http://cranlogs.r-pkg.org/).
-[Csardi 2015] RStudio is the only CRAN mirror to release its logs.) Researchers have used it to uncover instances gender inequity in a range of different areas, from biases in students' teaching evaluations to disparities between attendees and authors at the annual Digital Humanities Conference [Schmidt 2015, Weingart 2015].
+[Csardi 2015] RStudio is the only CRAN mirror to release its logs.)] Researchers have used it to uncover instances gender inequity in a range of different areas, from biases in students' teaching evaluations to disparities between attendees and authors at the annual Digital Humanities Conference [Schmidt 2015, Weingart 2015].
 These implementations help illustrate the value of the method.

### 2015-08-21 article.Rmd
Typos

@@
 
-The Kantrowitz corpus provides the list of names.^[The examples below is drawn from our Gender package which implements the Kantrowitz dataset for sake of comparsion, though its use is not recommended for actual research.
+The Kantrowitz corpus provides the list of names.^[The examples below is drawn from our Gender package which implements the Kantrowitz dataset for sake of comparison, though its use is not recommended for actual research.
 The Gender package combines the Kantrowitz corpus's two lists into a single list, identifying names that are ambiguous because they appear on both lists as `"either"`.]
@@
 
-Nevertheless, the methodology that have described here does take advantage of the datasets available to historians.
+Nevertheless, the methodology that we have described here does take advantage of the datasets available to historians.
 The predictions it provides are not intended to be taken as the entirety of what historians can say about gender in the past.

### 2015-08-21 article.Rmd
NAPP in present tense

@@
 
-The next release of the Gender package will include a dataset from the [North Atlantic Population Project](https://www.nappdata.org/napp/).
-This dataset will include names from Canada, Great Britain, Germany, Iceland, Norway, and Sweden from 1801 to 1910.^[The North Atlantic Population Project, which is also project of the Minnesota Population Center, also includes data for the United States of America, but we have excluded it since IPUMS-USA and the SSA dataset is more suitable when focusing on that country alone.] The SSA, IPUMS, and NAPP datasets (and any future datasets to be added to the package) all have the same simple tabular format.^[The format of this data departs from the more generally desirable "tidy data" framework proposed by Hadley Wickham for performance reasons.
+The gender package also features a dataset from the [North Atlantic Population Project](https://www.nappdata.org/napp/).
+This dataset includes names from Canada, Great Britain, Germany, Iceland, Norway, and Sweden from 1801 to 1910.^[The North Atlantic Population Project, which is also project of the Minnesota Population Center, also includes data for the United States of America, but we have excluded it since IPUMS-USA and the SSA dataset is more suitable when focusing on that country alone.] The SSA, IPUMS, and NAPP datasets (and any future datasets to be added to the package) all have the same simple tabular format.^[The format of this data departs from the more generally desirable "tidy data" framework proposed by Hadley Wickham for performance reasons.
 By keeping the data in a "wider" format instead of a tidy format as defined by Wickham, the function is able to make its predictions faster [Wickham 2014b].] The columns contain the `name`, `year`, and number of `female` and `male` uses of that name in a particular year.

### 2015-08-21 article.Rmd
Lowercase gender package

@@
 This approach generates more accurate results for sources that encompass changing periods of time, providing digital humanities scholars with a tool to estimate the gender of names.
-The article first describes the methodology as implemented in the [Gender](https://github.com/ropensci/gender) package for the [R programming language](http://www.r-project.org/).
+The article first describes the methodology as implemented in the [gender](https://github.com/ropensci/gender) package for the [R programming language](http://www.r-project.org/).
 It goes on to apply the method to a case study in which we examine gender and gatekeeping in the American historical profession over the past half-century.
-The Gender package illustrates the importance of incorporating historical approaches into computer science and related fields.
+The gender package illustrates the importance of incorporating historical approaches into computer science and related fields.
 
@@
 
-The Kantrowitz corpus provides the list of names.^[The examples below is drawn from our Gender package which implements the Kantrowitz dataset for sake of comparison, though its use is not recommended for actual research.
-The Gender package combines the Kantrowitz corpus's two lists into a single list, identifying names that are ambiguous because they appear on both lists as `"either"`.]
+The Kantrowitz corpus provides the list of names.^[The examples below is drawn from our gender package which implements the Kantrowitz dataset for sake of comparison, though its use is not recommended for actual research.
+The gender package combines the Kantrowitz corpus's two lists into a single list, identifying names that are ambiguous because they appear on both lists as `"either"`.]
 
@@
 Assuming for the moment that our method provides more accurate results, we estimate that `r percent(false_either_percent)` percent of the names in the Kantrowitz corpus are classified as ambiguous when a gender could be reasonably predicted from the name, that `r percent(false_male_percent)` percent of the names are classified as male when they should be classified as female, and that `r percent(false_female_percent)` are classified as female when they should be classified as male.
-This error rate is a separate concern from the much smaller size of the Kantrowitz corpus.^[This error rate was calculated using the Social Security Administration database, assuming a range of years between 1932 and 2012, which is the default setting within the Gender package.
+This error rate is a separate concern from the much smaller size of the Kantrowitz corpus.^[This error rate was calculated using the Social Security Administration database, assuming a range of years between 1932 and 2012, which is the default setting within the gender package.
 Of course, the error rate would fluctuate for different time periods and regions.]
@@
 
-To that end we have created the [Gender](https://github.com/ropensci/gender) package for R which includes both a predictive algorithm and an associated [genderdata](https://github.com/lmullen/gender-data-pkg) package containing various historical datasets.
+To that end we have created the [gender](https://github.com/ropensci/gender) package for R which includes both a predictive algorithm and an associated [genderdata](https://github.com/lmullen/gender-data-pkg) package containing various historical datasets.
 This R implementation is based on an [earlier Python implementation](https://github.com/cblevins/Gender-ID-By-Time) by [Cameron Blevins](http://www.cameronblevins.org/) and [Bridget Baird](http://www.conncoll.edu/directories/emeritus-faculty/bridget-baird/).
@@
 These include a rigorous set of standards to ensure the quality of packages contributed.
-R has a number of language features, such as data frames, which permit easy manipulation of data.^[The gender package depends especially on [R Core Team 2014-, Wickham and Francois 2014].] In particular, R permits the publication of packages containing data alongside packages containing code---a crucial feature for the Gender package.
-The Gender package is affiliated with [rOpenSci](http://ropensci.org/), an initiative that supports reproducible research and open data for scientists using R.
+R has a number of language features, such as data frames, which permit easy manipulation of data.^[The gender package depends especially on [R Core Team 2014-, Wickham and Francois 2014].] In particular, R permits the publication of packages containing data alongside packages containing code---a crucial feature for the gender package.
+The gender package is affiliated with [rOpenSci](http://ropensci.org/), an initiative that supports reproducible research and open data for scientists using R.
 The rOpenSci advisory board has provided code review, publicity, and an infrastructure to support the project.
@@
 
-The Gender package currently uses several datasets which make it suitable for studying the United States from the first federal census in 1790 onwards.] The first dataset contains names of applicants for Social Security and is available from [Data.gov](https://www.data.gov/).
+The gender package currently uses several datasets which make it suitable for studying the United States from the first federal census in 1790 onwards.] The first dataset contains names of applicants for Social Security and is available from [Data.gov](https://www.data.gov/).
 The second dataset contains names gathered in the decennial censuses and is available from the IPUMS-USA (Integrated Public Use Microdata Series) service from the Minnesota Population Center at the University of Minnesota.
@@
 In other words, there is an approximately `r gender("sidney", year = 1935)$proportion_male %>% percent()` percent chance that a person born in 1935 named Sidney was male.
-In 2012, for comparison, there was an approximately `r gender("sidney", year = 2012)$proportion_female %>% percent()` percent chance that a person born named Sidney was female.^[The Gender package makes no effort to combine alternative spellings for names, such as Sidney and Sydney, since we think it is better to treat those as separate names.
+In 2012, for comparison, there was an approximately `r gender("sidney", year = 2012)$proportion_female %>% percent()` percent chance that a person born named Sidney was female.^[The gender package makes no effort to combine alternative spellings for names, such as Sidney and Sydney, since we think it is better to treat those as separate names.
 For instance, although they are homophones, Geri is predominately a female name, and Gerry is predominately a male name.]

### 2015-08-21 article.Rmd
Additional citations about difficulty of gender binary

@@
 On its own, our method has little to say about crucial topics such as the intersectionality of gender with race, class, and power, or what the categories "male" or "female" mean in any given historical and social context [Crenshaw 1991, McCall 2005].
+The leap from datasets that contain records of sex at birth to gender has rightly been critiqued by scholars [Posner 2015, Westbrook and Saperstein 2015].
 The tool cannot reflect how an individual person self-identifies and how gender shapes their lived experience, since it relies on large datasets created by state agencies such as the U.S.
@@
 
+[Westbrook and Saperstein 2015] Westbrook, Laurel, and Aliya Saperstein.
+"New Categories Are Not Enough: Rethinking the Measurement of Sex and Gender in Social Surveys." *Gender & Society*, 29.4 (2015): 534-560.
+
 [Wickham 2014] Wickham, Hadley. "Tidy Data." *Journal of Statistical Software*, 50 (2014): <http://www.jstatsoft.org/v59/i10>.

### 2015-08-21 article.Rmd
Sharpen do-nothing critique and citation to Eckert

@@
 Nevertheless, the methodology that we have described here does take advantage of the datasets available to historians.
-The predictions it provides are not intended to be taken as the entirety of what historians can say about gender in the past.
-But they provide the ability to uncover data about gender where it was previously unknown, thus providing further grist to the mill of more traditional methods.
+While we agree with the category of critique that we have cited, we do not think that they should deter scholars from making use of our method.
+While such critiques point out genuine problems, they also seldom contain concrete proposals that can actually be implemented in historical research.^[Discussing exactly this problem of gender/sex binaries in large corpora, Penelope Eckert argues that "binary sex-class gets us pretty far" and argues that "at the moment" it is not "whether there's any way we can code large corpora for something subtler about gender than sex
+class" [Eckert 2012].] The predictions that our method provides are by no means intended to be taken as the entirety of what historians can say about gender in the past.
+But they do provide the ability to uncover data about gender where it was previously unknown, thus providing further grist to the mill of more traditional methods.
 Furthermore, the gender package is entirely open-source, and we have provided detailed documentation for its code and underlying data so that researchers who wish to build upon it can modify the package accordingly.
@@
 [Csardi 2015]  Csardi, Gabor. cranlogs: Download Logs from the RStudio CRAN Mirror. R package version 2.0.0 (2015): <http://CRAN.R-project.org/package=cranlogs>
+
+[Eckert 2012] Eckert, Penelope.
+"Coding for Gender and Sexuality." NSF-LDC workshop on coding sociolinguistic corpora.
+LSA Annual Meeting, Portland (2012).
+<https://www.ldc.upenn.edu/sites/www.ldc.upenn.edu/files/eckert-paper.pdf>.
 

### 2015-08-23 article.Rmd
Edits
Minor edits to intro, cleaned up a few citations, edits to AHR section,
reconfigured the conclusion to add in a hypothetical example (but
without specifically referring to Facebook)

--- 2015-08-21 article.Rmd
@@
 This article describes a new method for inferring the gender of personal names using large historical datasets.
-In contrast to existing methods of predicting gender that treat names as if they are timelessly associated with genders, this method uses a historical approach that takes into account changes in naming practices.
+In contrast to existing methods of predicting gender that treat names as if they are timelessly associated with one gender, this method uses a historical approach that takes into account changes in naming practices.
 It uses historical data to measure the likelihood that a name was associated with a particular gender based on the time or place under study.
@@
 Can researchers similarly use personal names to infer information about individual people?
-After all, personal names appear across a wide range of sources in the humanities: literary characters in stories and novels, authorship information in archival metadata, individuals mentioned by historical newspapers and magazine articles.
-Authors, characters, or historical actors are the kind of individuals who stand at the analytical center of humanities scholarship, and their names provide an intriguing way to infer additional information about them.^[The availability of large-scale datasets dealing with personal information raises important ethical issues for digital humanists, particularly in light of the revelations about domestic spying at the NSA brought forth by Edward Snowden in 2013.
+After all, personal names appear across a wide range of sources in the humanities: literary characters in stories and novels, authorship information in archival metadata, or individuals mentioned by historical newspapers and magazine articles.
+Authors, characters, or historical actors are the sorts of people who stand at the analytical center of humanities scholarship, and their names provide an intriguing way to infer additional information about them.^[The availability of large-scale datasets dealing with personal information raises important ethical issues for digital humanists, particularly in light of the revelations about domestic spying at the NSA brought forth by Edward Snowden in 2013.
 [Weingart 2014] Kieran Healy has used network analysis to demonstrate how researchers can infer information from datasets while also touching on the troubling ethical issue of surveillance [Healy 2013].]
@@
 For example, our program predicts that a person named Leslie who was born in 1950 is female, but reports a low level of certainty in that prediction, since just 52 percent of the babies named Leslie born in the United States in 1950 were girls and 48 percent were boys.
-Researchers can use these probabilities to determine what level of certainty is acceptable for predicting the gender of different names.
+These probabilities help a user determine what level of certainty is acceptable for predicting the gender of different names.
 
@@
 Gender theorists and feminist scholars have spent decades unpacking the full range of meanings behind gender as an analytical concept, including the fluid relationship between biological sex and gender identity [Butler 1990, Butler 1993].
-On its own, our method has little to say about crucial topics such as the intersectionality of gender with race, class, and power, or what the categories "male" or "female" mean in any given historical and social context [Crenshaw 1991, McCall 2005].
+On its own, our method has little to say about crucial topics such as the intersectionality of gender with race, class, and power, the lives of transgender people, or what the categories "male" or "female" mean in any given historical and social context [Crenshaw 1991, McCall 2005].
 The tool cannot reflect how an individual person self-identifies and how gender shapes their lived experience, since it relies on large datasets created by state agencies such as the U.S.
 Census Bureau.
-Much like the state’s definition of racial categories, governments impose a gender binary in their data collection that reflects existing power structures more than the lived identities of individuals.
+Much like the state’s definition of racial categories, governments impose a gender binary in their data collection that says more about existing power structures than the ways in which people self-identify.
 [Brubaker and Cooper 2000, 14-17, Pascoe 2009, Canaday 2011, Landrum 2014].
-When using the predictions provided by this method, users must judge for themselves how well they reflect the social and cultural practices of the time and place under study.
+When using the predictions provided by this method, users must judge for themselves how they reflect the social and cultural practices of the time and place under study.
 
-Imperfect as it admittedly is, this method nevertheless gives digital humanities scholars an important means of studying gender in textual collections.
+Imperfect as it is, this method nevertheless gives digital humanities scholars a much-needed tool to study gender in textual collections.
 We have made the method as transparent and flexible as possible through the inclusion of probability figures.
@@
 R and Python are two of the most commonly used languages for data analysis [Piatesky 2013, Muenchen 2014-].
-Surveys and analysis of useage point out the growth of R and the continuing popularity of Python for data science generally, and scholars are producing guides to using these languages for historical or humanities research [Programming Historian n.d., Arnold and Tilton 2015, Jockers 2014a].
+Surveys and analysis of usage point out the growth of R and the continuing popularity of Python for data science generally, and scholars are producing guides to using these languages for historical or humanities research [Programming Historian n.d., Arnold and Tilton 2015, Jockers 2014a].
 Yet Python's existing packages for gender prediction all implement a method similar to the NLTK or Genderize.io.^[These Python libraries include [estimate.gender](https://pypi.python.org/pypi/estimate.gender/0.4), [genderize](https://github.com/SteelPangolin/genderize), and [gender.py](https://github.com/block8437/gender.py)] The only other R package for gender prediction,   [genderizeR](https://cran.rstudio.com/web/packages/genderizeR/), uses the Genderize.io API.
@@
 The language has a strong tradition of being friendly to scholarship.
-It was originally created for statistics and many of its core contributors are academics; it provides facilities citing packages as scholarship.
+It was originally created for statistics and many of its core contributors are academics; it provides facilities for citing packages as scholarship.
 [CRAN](https://cran.r-project.org/) (the Comprehensive R Archive Network) offers a central location for publishing R packages for all users.
@@
 The first is to abstract the predictive algorithm to the simplest possible form so that it is usable for a wide range of historical problems rather than depending on the format of any particular data set.
-The second has been to provide as many datasets as possible in order to localize the predictions to particular times and places.
+The second has been to provide as many datasets as possible in order for users to tailor the algorithm's predictions to particular times and places.
 
-The gender package currently uses several datasets which make it suitable for studying the United States from the first federal census in 1790 onwards.] The first dataset contains names of applicants for Social Security and is available from [Data.gov](https://www.data.gov/).
-The second dataset contains names gathered in the decennial censuses and is available from the IPUMS-USA (Integrated Public Use Microdata Series) service from the Minnesota Population Center at the University of Minnesota.
+The gender package currently uses several datasets which make it suitable for studying the United States from the first federal census in 1790 onwards.
+The first dataset contains names of applicants for Social Security and is available from [Data.gov](https://www.data.gov/) [Social Security Administration 2014a].
+The second dataset contains names gathered in the decennial censuses and is available from the IPUMS-USA (Integrated Public Use Microdata Series) service from the Minnesota Population Center at the University of Minnesota [Ruggles et al.
+2010].
 
-The Social Security Administration (SSA) [Baby Names](http://catalog.data.gov/dataset/baby-names-from-social-security-card-applications-national-level-data) dataset was created as a result of the Social Security Act of 1935 during the New Deal.^[The Social Security Administration also provides a related dataset which contains the names of applicants broken down by state and year since 1910 [Social Security Administration 2014-b].] This dataset contains a list of first names along with how many times each name was assigned to each gender in every year beginning with 1880 and ending with 2012.
+The Social Security Administration (SSA) [Baby Names](http://catalog.data.gov/dataset/baby-names-from-social-security-card-applications-national-level-data) dataset was created as a result of the Social Security Act of 1935 during the New Deal.^[The Social Security Administration also provides a related dataset which contains the names of applicants broken down by state and year since 1910 [Social Security Administration 2014b].] This dataset contains a list of first names along with how many times each name was assigned to each gender in every year beginning with 1880 and ending with 2012.
 The SSA list includes any name which was used more than five times in a given year, thereby capturing all but the most infrequently used names from each year.
@@
 
-The simplest way to use the package would be to call the `gender()` function, specifying the method, in this case the Social Security Administration dataset since all the names are after 1930, and specifying a single range of years to make the predictions.
-However, we can instead use the `gender_df()` function provided by the package.
-By passing it our data frame of people and specifying which columns contain the names and birth years (or a columns containing a minimum and maximum for a range of years), we can efficiently predict the gender for each person.^[See the [package vignette](https://cran.rstudio.com/web/packages/gender/vignettes/predicting-gender.html) for more details about how to use the package effectively, including suggestions about performance: `browseVignettes( "gender")`.]
-
-```{r}
-gender_df(editors, name_col = "name", year_col = "birth_year")
-```
-
 By taking into account the year of birth, we find that four of our six names were likely male, whereas we might otherwise have predicted that all six were female.
@@
 
+A case study that relies on the gender package illustrates how the method can reveal hidden patterns within large textual collections.
 In a 2005 report for the American Historical Association, Elizabeth Lunbeck acknowledged "a sea change in the [historical] profession with respect to gender" before going on to describe the limits to this change for female historians, who face ongoing personal discrimination, lower salaries, and barriers to securing high-ranking positions.
-Lunbeck's report drew in part on a survey of 362 female historians that produced a rich source of responses detailing the deep and multi-faceted challenges facing women in the profession [Lunbeck 2005].
+Lunbeck's report drew in part on a survey of 362 female historians that produced a rich source of responses detailing the entrenched and multi-faceted challenges facing women in the profession [Lunbeck 2005].
 What follows is a quantitative supplement to Lunbeck's analysis that uses our program to analyze gender representation amongst historians across a much larger scale and a much longer time period.
@@
 
-Nevertheless, the methodology that we have described here does take advantage of the datasets available to historians.
-The predictions it provides are not intended to be taken as the entirety of what historians can say about gender in the past.
-But they provide the ability to uncover data about gender where it was previously unknown, thus providing further grist to the mill of more traditional methods.
-Furthermore, the gender package is entirely open-source, and we have provided detailed documentation for its code and underlying data so that researchers who wish to build upon it can modify the package accordingly.
+Nevertheless, the methodology that we have described here does take advantage of the historical datasets available to scholars.
+The predictions it generates help to uncover patterns about gender where they were previously unknown, providing further grist to the mill of more traditional methods.
+Furthermore, the gender package is entirely open-source, and we have provided detailed documentation for its code and underlying data so that those who wish to build upon it can modify the package accordingly.
+For instance, future researchers might compile datasets of names that move beyond the gender binary imposed by government agencies.
+They could then incorporate these datasets into the gender package, replacing the "Male" and "Female" designations with a much wider spectrum of potential gender identities.
 By building as much flexibility and transparency into the method as possible, we hope that others can continue to make just these kinds of improvements.
@@
 As of August 2015, the gender package has been downloaded more than 21,000 times and incorporated as a pedagogical tool within digital humanities courses [Jockers 2014b].^[The download counts come from the [download logs from RStudio's CRAN mirror](http://cran-logs.rstudio.com/) as tabulated by Gabor Csardi's [cranlogs package](http://cranlogs.r-pkg.org/).
-[Csardi 2015] RStudio is the only CRAN mirror to release its logs.)] Researchers have used it to uncover instances gender inequity in a range of different areas, from biases in students' teaching evaluations to disparities between attendees and authors at the annual Digital Humanities Conference [Schmidt 2015, Weingart 2015].
-These implementations help illustrate the value of the method.
-Its major intervention is to algorithmically inferring gender in textual sources in a way that takes into account change over time.
+[Csardi 2015] RStudio is the only CRAN mirror to release its logs.)] Researchers have used it to uncover instances gender inequity in a range of different areas, from biases in students' teaching evaluations to the disparities between attendees and authors at the annual Digital Humanities Conference [Schmidt 2015, Weingart 2015].
+These implementations illustrate the value of the method.
+Its major intervention is to algorithmically infer gender in textual sources in a way that takes into account change over time.
 Rather than trying to shoehorn existing computer science methods to fit humanities research, the gender package reverses this traditional disciplinary relationship.
@@
 
-[Social Security Administration 2014-a] Social Security Administration.
+[Social Security Administration 2014a] Social Security Administration.
 "Baby Names from Social Security Card Applications-National Level Data".
@@
 
-[Social Security Administration 2014-b] Social Security Administration.
+[Social Security Administration 2014b] Social Security Administration.
 "Baby Names from Social Security Card Applications-Data by State and District of Columbia." Data.gov dataset ID US-GOV-SSA-339.

### 2015-08-24 article.Rmd

@@
 On its own, our method has little to say about crucial topics such as the intersectionality of gender with race, class, and power, the lives of transgender people, or what the categories "male" or "female" mean in any given historical and social context [Crenshaw 1991, McCall 2005].
+The leap from datasets that contain records of sex at birth to gender has rightly been critiqued by scholars [Posner 2015, Westbrook and Saperstein 2015].
 The tool cannot reflect how an individual person self-identifies and how gender shapes their lived experience, since it relies on large datasets created by state agencies such as the U.S.
@@
 
-Nevertheless, the methodology that we have described here does take advantage of the historical datasets available to scholars.
-The predictions it generates help to uncover patterns about gender where they were previously unknown, providing further grist to the mill of more traditional methods.
+Nevertheless, the methodology that we have described here does take advantage of the datasets available to historians.
+While we agree with the category of critique that we have cited, we do not think that they should deter scholars from making use of our method.
+While such critiques point out genuine problems, they also seldom contain concrete proposals that can actually be implemented in historical research.^[Discussing exactly this problem of gender/sex binaries in large corpora, Penelope Eckert argues that "binary sex-class gets us pretty far" and that "at the moment" it is not "whether there's any way we can code large corpora for something subtler about gender than sex-class" [Eckert 2012].]  The predictions it generates help to uncover patterns about gender where they were previously unknown, providing further grist to the mill of more traditional methods.
 Furthermore, the gender package is entirely open-source, and we have provided detailed documentation for its code and underlying data so that those who wish to build upon it can modify the package accordingly.
 For instance, future researchers might compile datasets of names that move beyond the gender binary imposed by government agencies.
-They could then incorporate these datasets into the gender package, replacing the "Male" and "Female" designations with a much wider spectrum of potential gender identities.
-By building as much flexibility and transparency into the method as possible, we hope that others can continue to make just these kinds of improvements.
+They could then incorporate these datasets into the gender package, replacing the "male" and "female" designations with a much wider spectrum of potential gender identities.
 
 As it currently stands, the gender package meets a methodological need.
-As of August 2015, the gender package has been downloaded more than 21,000 times and incorporated as a pedagogical tool within digital humanities courses [Jockers 2014b].^[The download counts come from the [download logs from RStudio's CRAN mirror](http://cran-logs.rstudio.com/) as tabulated by Gabor Csardi's [cranlogs package](http://cranlogs.r-pkg.org/).
+As of August 2015, the gender package has been downloaded more than 22,000 times and incorporated as a pedagogical tool within digital humanities courses [Jockers 2014b].^[The download counts come from the [download logs from RStudio's CRAN mirror](http://cran-logs.rstudio.com/) as tabulated by Gabor Csardi's [cranlogs package](http://cranlogs.r-pkg.org/).
 [Csardi 2015] RStudio is the only CRAN mirror to release its logs.)] Researchers have used it to uncover instances gender inequity in a range of different areas, from biases in students' teaching evaluations to the disparities between attendees and authors at the annual Digital Humanities Conference [Schmidt 2015, Weingart 2015].

### 2015-08-24 R/load-data.R
Don't use option() for stringsAsFactors

@@ -1,5 +1,4 @@
 # Load the data and do any cleaning
-options(stringsAsFactors = FALSE)
 library(dplyr)
 library(stringr)
 

### 2015-08-24 R/predict-gender.R

@@ -1,4 +1,3 @@
-options(stringsAsFactors = FALSE)
 library(gender)
 library(dplyr)
 library(tidyr)

### 2015-08-24 article.Rmd
Footnote/link to repository for article

@@
 It goes on to apply the method to a case study in which we examine gender and gatekeeping in the American historical profession over the past half-century.
-The gender package illustrates the importance of incorporating historical approaches into computer science and related fields.
+The gender package illustrates the importance of incorporating historical approaches into computer science and related fields.^[Please see the `lmullen/gender-article` [GitHub repository](https://github.com/lmullen/gender-article/) for the code used to create this article.]
 

### 2015-08-24 article.Rmd
Final copy-editing, conclusion revision

@@
 They take the form of subconscious bias that subtly weakens the assessment of work done by female historians, whether in a seminar paper or a job application.
-Moreover, gender discrimination frequently intersects with other kinds of racial and socioeconomic inequity that erects even higher barriers in the women's careers [Townsend 2008].
+Moreover, gender discrimination frequently intersects with other kinds of racial and socioeconomic inequity that erects even higher barriers in women's careers [Townsend 2008].
 Gendered family dynamics add another layer of challenges for women: a 2011 survey by the American Historical Association revealed that female historians dedicated substantially more time to child and elder care than their male colleagues, leaving them with less time for the kind of academic research that the profession rewards most highly [Townsend 2012].
@@
 Being a reviewer for the journal is, in many ways, a proxy for expertise.
-When there are twice as many male "experts" as female "experts" appearing in the *AHR* it imbues our understanding of what constitutes historical expertise.
+When there are twice as many male "experts" as female "experts" appearing in the *AHR* it imbues our understanding of what constitutes historical expertise and who possesses it.
 
 The second finding we arrived at when examining the gender of *AHR* reviewers was a better understanding of who reviews which book authors.
-Once again, the methodology reveals patterns that we might suspect but nonetheless tends to stay hidden from direct sight.
+Once again, the methodology reveals patterns that we might suspect but tend to stay hidden from direct sight.
 About three times as many men as women write reviews of male-authored books in the *AHR*.
-In the case of female-authored books, the ratio of female-to-male reviewers is closer to 50/50.
+For female-authored books, the ratio of female-to-male reviewers is closer to 50/50.
 In short, women are far more likely to write reviews of other women, while men tend to gravitate more towards male authors---who are, of course, already over-represented in the *AHR*.
@@
 To borrow a phrase from Miriam Posner, the gender package does not necessarily meet the "radical, unrealized potential of digital humanities" [Posner 2015].
+While we agree with this line of critique, we don't think it should deter researchers from using our method.^[While such critiques point out genuine problems, they also seldom contain concrete proposals that can be practically implemented in historically oriented research.
+Discussing the problem of gender/sex binaries in large corpora, Penelope Eckert argues that "binary sex-class gets us pretty far" and that "at the moment" she is pessimistic "whether there's any way we can code large corpora for something subtler about gender than sex-class" [Eckert 2012].
+We thank Heather Froehlich for this citation.]
 
-Nevertheless, the methodology that we have described here does take advantage of the datasets available to historians.
-While we agree with the category of critique that we have cited, we do not think that they should deter scholars from making use of our method.
-While such critiques point out genuine problems, they also seldom contain concrete proposals that can actually be implemented in historical research.^[Discussing exactly this problem of gender/sex binaries in large corpora, Penelope Eckert argues that "binary sex-class gets us pretty far" and that "at the moment" it is not "whether there's any way we can code large corpora for something subtler about gender than sex-class" [Eckert 2012].]  The predictions it generates help to uncover patterns about gender where they were previously unknown, providing further grist to the mill of more traditional methods.
+The gender package makes a pragmatic trade-off between complexity and discovery.
+Its reliance on a male/female binary dampens the complexity of gender identity while simulatenously pointing towards new discoveries about gender within large textual collections.
+Its use of historical datasets helps its users uncover patterns where they were previously unknown, providing further grist to the mill of more traditional methods.
 Furthermore, the gender package is entirely open-source, and we have provided detailed documentation for its code and underlying data so that those who wish to build upon it can modify the package accordingly.
 For instance, future researchers might compile datasets of names that move beyond the gender binary imposed by government agencies.
-They could then incorporate these datasets into the gender package, replacing the "male" and "female" designations with a much wider spectrum of potential gender identities.
+They could then incorporate these datasets into the gender package, replacing "male" and "female" designations with a much wider spectrum of potential gender identities.
 
-As it currently stands, the gender package meets a methodological need.
-As of August 2015, the gender package has been downloaded more than 22,000 times and incorporated as a pedagogical tool within digital humanities courses [Jockers 2014b].^[The download counts come from the [download logs from RStudio's CRAN mirror](http://cran-logs.rstudio.com/) as tabulated by Gabor Csardi's [cranlogs package](http://cranlogs.r-pkg.org/).
+As it currently stands, the gender package meets an important methodological need.
+As of August 2015, the gender package has been downloaded more than 22,000 times and incorporated as a pedagogical tool within digital humanities courses [Jockers 2014b].^[The download counts reflect statistics as of August 20, 2015, and come from the [download logs from RStudio's CRAN mirror](http://cran-logs.rstudio.com/) as tabulated by Gabor Csardi's [cranlogs package](http://cranlogs.r-pkg.org/).
 [Csardi 2015] RStudio is the only CRAN mirror to release its logs.)] Researchers have used it to uncover instances gender inequity in a range of different areas, from biases in students' teaching evaluations to the disparities between attendees and authors at the annual Digital Humanities Conference [Schmidt 2015, Weingart 2015].

### 2015-08-24 article.Rmd
Minor abstract edits

@@
 This article describes a new method for inferring the gender of personal names using large historical datasets.
-In contrast to existing methods of predicting gender that treat names as if they are timelessly associated with one gender, this method uses a historical approach that takes into account changes in naming practices.
+In contrast to existing methods of gender prediction that treat names as if they are timelessly associated with one gender, this method uses a historical approach that takes into account how naming practices change over time.
 It uses historical data to measure the likelihood that a name was associated with a particular gender based on the time or place under study.
-This approach generates more accurate results for sources that encompass changing periods of time, providing digital humanities scholars with a tool to estimate the gender of names.
+This approach generates more accurate results for sources that encompass changing periods of time, providing digital humanities scholars with a tool to estimate the gender of names across large textual collections.
 The article first describes the methodology as implemented in the [gender](https://github.com/ropensci/gender) package for the [R programming language](http://www.r-project.org/).

### 2015-08-24 article.Rmd
Intro copy edits

@@
 
-As humanities datasets grow larger and larger, researchers have not only been able to access information that *is* there, but to infer information that *isn't* there.
+As the digital archive grows larger and larger, researchers have not only been able to access information that *is* there, but to infer information that *isn't* there.
 Topic modeling, for example, has been used to uncover underlying trends across a century of academic articles in literary studies.
@@
 For those working on contemporary subjects, the "Leslie problem" is not an especially pressing one.
-There are a variety of tools that use current databases of names, including [Genderize.io](https://genderize.io/), which predicts the gender of a first name from the user profiles of unnamed social media sites [Genderize.io n.d.].
+There are a variety of tools that use current databases of names.
+[Genderize.io](https://genderize.io/), for instance, predicts the gender of a first name from the user profiles of unnamed social media sites [Genderize.io n.d.].
 A film scholar analyzing male and female characters in a database of modern American television shows might use Genderize.io to accurately predict the gender of these characters, including Leslie Fay.

### 2015-08-24 article.Rmd
Move critique back into text

@@
 To borrow a phrase from Miriam Posner, the gender package does not necessarily meet the "radical, unrealized potential of digital humanities" [Posner 2015].
-While we agree with this line of critique, we don't think it should deter researchers from using our method.^[While such critiques point out genuine problems, they also seldom contain concrete proposals that can be practically implemented in historically oriented research.
-Discussing the problem of gender/sex binaries in large corpora, Penelope Eckert argues that "binary sex-class gets us pretty far" and that "at the moment" she is pessimistic "whether there's any way we can code large corpora for something subtler about gender than sex-class" [Eckert 2012].
+While we agree with this line of critique, we don't think it should deter researchers from using our method.
+While such critiques point out genuine problems, they also seldom contain concrete proposals that can be practically implemented in historically oriented research.^[Discussing the problem of gender/sex binaries in large corpora, Penelope Eckert argues that "binary sex-class gets us pretty far" and that "at the moment" she is pessimistic "whether there's any way we can code large corpora for something subtler about gender than sex-class" [Eckert 2012].
 We thank Heather Froehlich for this citation.]

### 2015-08-24 article.Rmd
Figures as separate files; smart quotes in Word document
Word document? WTF?

@@
 output:
-  html_document: default
-  word_document: default
+  html_document: 
+    self_contained: false
+  word_document:
+    pandoc_args: [
+      "--smart"
+    ]
 ---

### 2015-12-02 article.Rmd
Fix incorrect example of editors

@@
                                "Jan", "Jan"),
-                      birth_year = c(1934, 1990, 1948, 1970, 1965, 1998))
+                      year = c(1934, 1990, 1948, 1970, 1965, 1998))
 ```
@@
 ```{r, echo=TRUE, results='markup'}
-editors
+gender_df(editors, method = "ssa")
 ```
